### 2019-05-02 comparison_of_methods.Rmd
add initial comparison of methods report

---
title: "Comparison of methods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-05-02 index.Rmd
testing dummy markdown
initial commit of R notebook

--- 2019-05-02 comparison_of_methods.Rmd
@@
 ---
-title: "Comparison of methods"
-output: html_document
+title: "Comparison of Methods"
+output:
+  html_document:
+    df_print: paged
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.
+
+Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.
+
+```{r}
+plot(cars)
 ```
 
-## R Markdown
+Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+The preview shows you a rendered HTML copy of the contents of the editor.
+Consequently, unlike *Knit*, *Preview* does not run any R code chunks.
+Instead, the output of the chunk when it was last run in the editor is displayed.

### 2019-05-02 index.Rmd
initital attempt at transferring Marie's latex write-up to R notebook

--- 2019-05-02 comparison_of_methods.Rmd
@@
 ---
-title: "Comparison of methods"
-output: html_document
+title: "Report"
+output:
+  html_document:
+    df_print: paged
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Introduction
 
-## R Markdown
+The classical setup in FDA is to assume that we observe a sample of $n$ independant realizations $X_1,\ldots,X_n$ of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+The first steps of a classical approach to analyse such data are to first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE) and then perform FPCA to reduce the dimension of the data.
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+However, it might very well happen that the function $X$ belongs to $\M \subset L^2([a,b],\R)$ a low-dimensional nonlinear manifold (gives examples and references).
+If it is indeed the case, it is no longer appropriate to use the norm induced by the $L^2$ inner product to calculate the distance between two curves.
+Explain why.
+One should instead use the geodesic distance $d_\M$ induced by $\M$ which takes into account the intrinsic structure of $\M$.
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.
 
-```{r cars}
-summary(cars)
-```
+Now back to functional data living in a low-dimensional manifold embedded in the ambient Hilbert space.
+There exist many nonlinear dimension reduction methods in the literature, also known as manifold learning methods, which are particularly popular in computer vision.
+However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio.
+(is it true?
+ref?) This is a luxury that FDA cannot always afford.
 
-## Including Plots
+Though powerful techniques exist for smoothing/interpolating discretely observed noisy functional data, the recovered functional versions of the data, $\tilde X_1,\ldots,\tilde X_n$, may not lie even close to the functional manifold $\M$.
+This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs.
 
-You can also embed plots, for example:
+% the following two sentences are not related to each other
+% Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
+Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
+With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$ when we have access only to discretely-observed noisy functional observations.
+Say why pairwise distances important, e.g.
+classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression).
+Certainly, pairwise geodesic distances are legitimate objects of interest in themselves but our motivation is more general in nature.
+Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.
+like the Berkeley growth curves, MDS based on pairwise geodesic versus MDS based on pairwise $L_2$.
+The former separates the genders better?
+Something like this, but hopefully a different dataset.???
+
+Briefly outline the methodology: MDS plus density ridge are used in a cool way to estimate (consistently?, let's hope..) pairwise geodesic distances.
+
+Along the way, we introduce a new tool to assess pairwise geodesic distance estimation.
+This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
+
+# Proposed method for estimating geodesic distances
+
+Let $X_1,\ldots,X_n$ be independent realizations of a random function $X\in\M \subset L^2([a,b],\R)$, where $\M$ is an unknown manifold of dimension $d$.
+Suppose that each curve $X_i$ is observed on a grid $T_i=(t_{i1},\ldots,t_{iK})$ and with measurements errors, i.e.
+that we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ are of mean zero and uncorrelated with each other.
+We assume that the grids $T_1,\ldots,T_n$ are dense.
+
+Our method is described by the following steps:
+
+1. Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing:
+$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
+where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
+2. Obtain a $s$-dimensional representation $\tilde X^s_1,\ldots,\tilde X^s_n$ of the functions $\tilde X_1,\ldots,\tilde X_n$ that "preserves" the pairwise $L^2$ distances by using MDS.
+3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
+4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n\times n $ matrix $\hat G$ as
+$$
+\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
+ d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
+ 0 & \textrm{otherwise.}
+  \end{array} \right.
+$$
+
+Since the ridge estimation obtained from noisy measurements of a manifold should approximate well the manifold (ref.
+Genevose and al.
+2014), we expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
+Ridge estimation suffers from the curse of dimensionality, this is why we first reduce the dimension of our data with MDS and then apply the shift-mean algorithm to estimate the ridge.
+
+# Comparison to other methods
+
+Susan: why are we describing Isomap as the first sentence in this Simulation section?
+Shouldn't we begin by describing the goal.
+Something like: We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances for discretely-observed noisy functional data, compared to various alternative approaches.
+
+MH : Sure, the introduction sentence was missing and we can also put the description of IsoMAP somewhere else.
+
+Susan question: Also, isn't the real name for IsoGeo Floyd-Warshall or Dijkstra’s? These are the two options in python scikitlearn's isomap routine, where the default is actually automatic selection.
+
+MH : Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, p-isomap is one of these).
+I wanted to introduce that IsoGeo to avoid confusion but maybe it was not clear, I changed a bit the description.
+
+Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
+The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
+In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
+are estimated based on $G$.
+The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, and can be calculated easily with Floyd-Warshall or Dijkstra’s algorithms.
+The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
+
+%In the sequel, we use the first step of Isomap to estimate pairwise geodesic distances and we call this procedure IsoGeo.
+In the sequel, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
+
+We compare our method to the following ones:
+* [\bf RD] Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
+* [\bf SS] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
+* [\bf pI] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller).
+Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
+ [\bf RP] same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
+
+# TODOs
+
+* Clean up the writing in this notebook. Add code and graphs where helpful.
+* Related to tuning parameter
+    + Choose the number of neighbors $N$ in IsoGeo.
+    + Choose projection dimension for avoiding curse of dimensionality in performing density ridge scms routine
+    + Choose the bandwidth $h$ in scms.
+* Related to spartanSim.R simulations results
+    + Why do some Monte Carlo simulations fail?
+    + Is P-isomap always choosing zero penalty automatically?
+    + Shouldn’t the smoothing method be doing better? Thought the smooth was close to the true curve.
+* Next time we submit spartanSim.R, make sure to hold dataset constant for different projection dimensions
+* Read Marie's suggested paper on robust ISOMAP
+* Revisit Mueller manifold examples. Need to write down some differential equations to solve the geodesic minimisation problem?

### 2019-05-09 index.Rmd
a few comments were updated

@@
 The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
+```{r}
+3 + 5
+```
 
@@
 * Next time we submit spartanSim.R, make sure to hold dataset constant for different projection dimensions
-* Read Marie's suggested paper on robust ISOMAP
+* Read Marie's suggested paper on [robust ISOMAP](https://arxiv.org/abs/1306.3373)
 * Revisit Mueller manifold examples. Need to write down some differential equations to solve the geodesic minimisation problem?

### 2019-05-09 Geo_estimation.R
updated comments

@@ -1,16 +1,15 @@
 # TODO Marie: this function should call assess_goodness_estimation
 # TODO Marie: this function should actually be renamed to reflect that it is an oracle estimation procedure
-# TODO Marie/Susan: work on tuning parameters
 # TODO Marie/Susan: if it turns out random projection method g is still worth considering, implement parallel for loops
 ##### Estimation of geodesic distances
 
 
 # We estimate the geodesic matrix obtained via
 # a) Floyd's Algorithm on true data (just to make sure that the theoretical geo. is calculated correctly)
-# b) Floyd's Algorithm on raw data
+# b) Floyd's Algorithm on noisy data
 # c) Floyd's Algorithm on smooth data (only for FD)
 # d) p-isomap of Muller on smooth data
-# e) mds of smooth data + scms + Floyd's Algorithm (only for FD) 
+# e) mds of smooth data + scms + Floyd's Algorithm (only for FD)
 # f) scms + Floyd's Algorithm (only for ED)
 # g) random proj. of smooth data + scms + Floyd's Algorithm (only for FD)
 
@@ -54,7 +53,7 @@ py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
 scms = import_from_path("scms",path='.')
 
 Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,nb_proj,grid,reg_grid,common_grid_true=0){
-  
+
   samplesize = nrow(true_data)
   K = ncol(true_data)
   norm_true_geo = sqrt(sum(true_geo^2))
@@ -62,14 +61,14 @@ Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,
   list2_to_return=list()
   list3_to_return=list()
   list4_to_return=list()
-  
+
   ## Apply method a) and b) for FD observed on a common grid or Euclidean data
-  if(!FD_true||common_grid_true==1){ 
-    
-    ### a) Estimation from the true data 
-    
+  if(!FD_true||common_grid_true==1){
+
+    ### a) Estimation from the true data
+
     print("Floyd's Algorithm on true data")
-    
+
     if(FD_true){
       a = reg_grid[1]
       b = reg_grid[K]
[…]

### 2019-05-09 functions.R
clean up

@@ -1,156 +1,0 @@
-################################################
-## calculates the local covariance at (x0,y0)
-loc.cov = function(obj,x0,h,shift_type,kernel){
-
-  data = data.matrix(obj$data)
-
-  weights = exp( -((data[,1]-x0[1])^2+(data[,2]-x0[2])^2)/h^2 ) / nrow(data)
-
-  if (shift_type == "ms") {
-
-    local.mean.x = sum(weights * data[,1]) / sum(weights)
-    local.mean.y = sum(weights * data[,2]) / sum(weights)
-
-  } else if(shift_type == "scms") {
-
-    kernel_sigma = h
-    kernelscms = scms$make_isotropic_gaussian_kernel(kernel_sigma)
-    denoised = scms$subspace_constrained_mean_shift_update(x0, data, kernelscms, kernel_sigma)
-    local.mean.x = x0[1]+denoised[1]
-    local.mean.y = x0[2]+denoised[2]
-
-  } else if(shift_type == "none") {
-    local.mean.x = x0[1]; local.mean.y = x0[2];
-  }
-
-  if (kernel == "double-peak") {
-    # double-peak kernel
-    m = matrix(c(sum(weights*(data[,1]-local.mean.x)^2) ,
-                 sum(weights*(data[,1]-local.mean.x)*(data[,2]-local.mean.y)) ,
-                 sum(weights*(data[,2]-local.mean.y)*(data[,1]-local.mean.x)) ,
-                 sum(weights*(data[,2]-local.mean.y)^2  ) ) ,nrow=2)
-  } else if (kernel == "butterfly") {
-    # butterfly kernel
-    m = matrix(c(sum(weights*(data[,1]-local.mean.x)^2 / ( (data[,1]-local.mean.x)^2+(data[,2]-local.mean.y)^2 ) ) ,
-                 sum(weights*(data[,1]-local.mean.x)*(data[,2]-local.mean.y) / ( (data[,1]-local.mean.x)^2+(data[,2]-local.mean.y)^2 ) ) ,
-                 sum(weights*(data[,2]-local.mean.y)*(data[,1]-local.mean.x) / ( (data[,1]-local.mean.x)^2+(data[,2]-local.mean.y)^2 ) ) ,
-                 sum(weights*(data[,2]-local.mean.y)^2 / ( (data[,1]-local.mean.x)^2+(data[,2]-local.mean.y)^2 ) )) ,nrow=2)
-  }
-  return(m)
-}
-
-
-# plot eigenvector field
[…]

### 2019-05-09 runSCMS.R

@@ -1,13 +1,0 @@
-library(reticulate)
-scms = import_from_path("scms",path='.')
-py_run_file("demo.py")
-
-blah <- import("demo",convert=TRUE)
-scmspath = blah$denoised
-plot(scmspath[,1],scmspath[,2])
-
-obj = EuclideanExamples("bananas",100,0)
-data = obj$data
-plot(data)
-denoised = scms$scms(data,0.2)
-points(denoised,col="red")

### 2019-05-09 Geo_estimation.R
more housekeeping

@@ -301,3 +301,26 @@ Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,
 
   return(c(list1_to_return,list2_to_return,list3_to_return,list4_to_return))
 }
+
+##### Smooth data one curve at the time with bspline and penatlty on second derivative
+
+smooth_FD_bspline <- function(discrete_data,grid,reg_grid){
+  
+  K = ncol(discrete_data)
+  nbasis <- K+2
+  basis_obj <- create.bspline.basis(c(reg_grid[1],reg_grid[K]),nbasis)
+  loglambda <- seq(-10,-1,by=0.5)
+  gcv<-rep(0,length(loglambda))
+  for(i in 1:length(loglambda)){
+    fd_par_obj <- fdPar(fdobj=basis_obj,Lfdobj=2,lambda=10^loglambda[i])
+    smooth_res <- smooth.basis(t(grid),t(discrete_data),fd_par_obj)
+    gcv[i]=mean(smooth_res$gcv)
+  }
+  ind_m=which(gcv==min(gcv))
+  fd_par_obj <- fdPar(fdobj=basis_obj,Lfdobj=2,lambda=10^loglambda[ind_m])
+  smooth_res <- smooth.basis(t(grid),t(discrete_data),fd_par_obj)
+  smooth_curve_tmp <- smooth_res$fd
+  smooth_curve <- eval.fd(reg_grid,smooth_curve_tmp)
+  return(t(smooth_curve))
+  
+}

### 2019-05-09 Isomap.R

@@ -1,3 +1,26 @@
+## Computes the pairwise Euclidean distances between all data samples
+##
+## input:
+##       data: NxD matrix (N samples, D features)
+## output:
+##	NxN matrix with the pairwise Euclidean distances
+
+
+pairwiseDistances = function(data){
+  
+  num_samples = nrow(data)
+  
+  d = t(rowSums(data^2))
+  d = d[rep(1, each = num_samples),]
+  d = d + t(d) - 2*data%*%t(data)
+  diag(d) = 0
+  d = sqrt(d)
+  
+  return(d)
+  
+}
+
+
 ## Computes the Isomap embedding by Tenenbaum, de Silva and Langford (2000).
 ## It is based on the Matlab implementation by Tenenbaum.
 ## This version uses Floyd's Algorithm to compute the shortest path tree.

### 2019-05-09 pairwiseDistances.R

@@ -1,21 +1,0 @@
-## Computes the pairwise Euclidean distances between all data samples
-##
-## input:
-##       data: NxD matrix (N samples, D features)
-## output:
-##	NxN matrix with the pairwise Euclidean distances
-
-
-pairwiseDistances = function(data){
-
-        num_samples = nrow(data)
-
-        d = t(rowSums(data^2))
-        d = d[rep(1, each = num_samples),]
-        d = d + t(d) - 2*data%*%t(data)
-        diag(d) = 0
-        d = sqrt(d)
-
-	return(d)
-
-}

### 2019-05-09 smooth_FD_bspline.R

@@ -1,22 +1,0 @@
-##### Smooth data one curve at the time with bspline and penatlty on second derivative
-
-smooth_FD_bspline <- function(discrete_data,grid,reg_grid){
-  
-  K = ncol(discrete_data)
-  nbasis <- K+2
-  basis_obj <- create.bspline.basis(c(reg_grid[1],reg_grid[K]),nbasis)
-  loglambda <- seq(-10,-1,by=0.5)
-  gcv<-rep(0,length(loglambda))
-  for(i in 1:length(loglambda)){
-    fd_par_obj <- fdPar(fdobj=basis_obj,Lfdobj=2,lambda=10^loglambda[i])
-    smooth_res <- smooth.basis(t(grid),t(discrete_data),fd_par_obj)
-    gcv[i]=mean(smooth_res$gcv)
-  }
-  ind_m=which(gcv==min(gcv))
-  fd_par_obj <- fdPar(fdobj=basis_obj,Lfdobj=2,lambda=10^loglambda[ind_m])
-  smooth_res <- smooth.basis(t(grid),t(discrete_data),fd_par_obj)
-  smooth_curve_tmp <- smooth_res$fd
-  smooth_curve <- eval.fd(reg_grid,smooth_curve_tmp)
-  return(t(smooth_curve))
-  
-}

### 2019-05-09 test_analytic_geodesic_distance.Rmd
cleaned up file dependencies and wrote a notebook for testing whether analytic calculations are correct

---
title: Test if analytic derivation of geodesic distance for curves from sim_functional_data.R
  is correct
output:
  html_document:
    df_print: paged
---

```{r}
source('sim_functional_data.R')
```

The function sim_functional_data.R allows specification of several simulatin scenarios. The samplesize is purposefully set very high for reasons explained below.

```{r}
# Generate data
sim <- sim_functional_data(sce=2,samplesize=1000)
```

The output actually contains more than just the simulated functional data. In particular, it contains the in-sample pairwise geodesic distance matrix (analytic_geo) via analytic derivation. The purpose of this notebook is to see if our analytic derivations match numerical results.

```{r}
names(sim)
```

We would like to compare sim$analytic_geo to geodesic distance estimation via Floyd's algorithm for the noiseless (discretised) curve. This algorithm is guaranteed to do well if we sample the manifold densely enough? (That's why we set the samplesize to be quite high initially.)

```{r}
library(reticulate)
# sometimes reticulate requires specific instruction on where to find your python
# first call Sys.which("python") and then copy that path into 
use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')
getIsomapGdist = import_from_path("getIsomapGdist",path='.')

attach(sim)
a = reg_grid[1]
b = tail(reg_grid,1)
K = length(reg_grid)
samplesize = dim(noiseless_data)[1]
noiseless_data_tmp = (sqrt((b-a)/K))*noiseless_data

# Find a grid of possible values for the number of neigbors
num_neigh_min=get_min_num_neighbors$get_min_num_neighbors(noiseless_data_tmp)
num_neigh_true=seq(num_neigh_min,samplesize/2,by=2)

# Calculate the geo matrix for each number of neighbors and keep the one that gives the minimal error
norm_analytic_geo = sqrt(sum(analytic_geo^2))
Error_true_mani_K= rep(0,length(num_neigh_true))
for(j in 1:length(num_neigh_true)){
  IsomapGdist = getIsomapGdist$getIsomapGdist(noiseless_data_tmp,num_neigh_true[j])
  Error_true_mani_K[j]=sqrt(sum((IsomapGdist - analytic_geo )^2))/norm_analytic_geo
}
ind_op_true=min(which(Error_true_mani_K==min(Error_true_mani_K)))
estim_geo_noiseless_data = getIsomapGdist$getIsomapGdist(noiseless_data_tmp,num_neigh_true[ind_op_true])
```
If the analytic derviation is correct, then estim_geo_noiseless_data should be extremely similar to analytic_geo. We can assess their closeness using assess_goodness_estimation.R. Running this notebook for sce=1 gives rmse on the order of e-16.
```{r}
library(MESS)
source('assess_goodness_estimation.R')
assess_goodness_estimation(estim_geo_noiseless_data,analytic_geo)
```

### 2019-05-09 assess_goodness_estimation.R

@@ -1,5 +1,7 @@
 ##### Function that assess how good a particular estimation of a geodesic matrix is good
 
+library(MESS) # for auc function
+
 # for the moment we just use the relative error || hat A - A ||_F / || A ||_F
 
 assess_goodness_estimation <- function(estim_mat,true_geo){

### 2019-05-09 sim_functional_data.R

@@ -1,8 +1,8 @@
 ##### Dataset generators
 
-# Scenario 1 : random function from the manifolg M={X(t)=u(h_alpha(t)},with h_alpha(t)=t-alpha and u(t)=t
-# TODO Marie: the true geodesic distance calculation is incorrect for scenario 2
-# Scenario 2 : random function from the manifolg M={X(t)=density N(alpha,1)}
+# Scenario 1 : random function from the manifold M={X(t)=u(h_alpha(t)},with h_alpha(t)=t-alpha and u(t)=t
+# TODO Marie: the analytic geodesic distance calculation is incorrect for scenario 2
+# Scenario 2 : random function from the manifold M={X(t)=density N(alpha,1)}
 
 # TODO Marie/Susan: we need more functional scenarios!
 
@@ -13,19 +13,22 @@
 # reg_sampling : if 0 points on the manifold are draw from a norm dist. and if 1 regular sampling 
 # a,b : functions X(t) are defined for t in [a,b]
 # com_grid : if 1 each curve is observed on the same grid and if 0 the grid for each curve is randomly genarated from unif[a,b]
-# plot_true : if 1 plot the true data, the observed the data and the true geodesic matrix
+# plot_true : if 1 plot the true data, the observed data and the true geodesic matrix
 
 ## Output
-# true_data : samplesize x K matrix containing the original data (no noise)
-# discrete_data : samplesize x K matrix containing the observed data
-# true_geo : samplesize x samplesize matrix containing true pairwise geo disctance
+# noiseless_data : samplesize x K matrix containing the original data (no noise)
+# noisy_data : samplesize x K matrix containing the observed data
+# analytic_geo : samplesize x samplesize matrix containing analytic pairwise geo disctance
 # grid : samplesize x K matrix containing the grid on which each data is observed
 # reg_grid : vector of dim K containing a common grid to use for smoothing
 
 #Example of call
 # data<- sim_functional_data(1,20,15,-3,3,0.5,1,1,1)
 
-sim_functional_data<-function(sce,samplesize,K,a,b,SNR,reg_sampling,com_grid,plot_true){
[…]

### 2019-05-14 sim_functional_data.R
testing build of rmarkdown webpage

@@ -1,15 +1,21 @@
 #' ## This function generates functional data living in a submanifold of a Hilbert space
 
-#' Scenario 1: Considered is the manifold $$M=\{X(t) = t-\alpha, -0.9 \le \alpha \le 3\} \subset L^2([a,b])$$ with inherited metric from the ambient $L^2[a,b]$ space.
+#' # Scenario 1: 
+#' Consider the manifold $$M=\{X(t) = t-\alpha, -0.9 \le \alpha \le 3\} \subset L^2([a,b])$$ with inherited metric from the ambient $L^2[a,b]$ space.
 #' The shortest path $\gamma: [0,1] \to M$ between two functions $X_1 = (t-\alpha_1)$ and $X_2 = (t-\alpha_2)$ in $M$ is clearly given by $\gamma(t) = t X_1 + (1-t) X_2$.
 #' The geodesic distance between $X_1$ and $X_2$ is $$L(\gamma) := \int_0^1 || \dot \gamma(t) ||_{L^2} \,dt = ||X_1 - X_2||_{L_2} = (\alpha_1-\alpha_2)\sqrt{b-a}.$$ 
-#' Scenario 2 : Considered is the manifold $$ M = \{ \text{probability density function} N(\alpha,\sigma^2): \alpha \in \mathbb R, \sigma > 0 \} \subset L^2([-\infty,\infty])$$ with inherited metric from the ambient $L^2[-\infty,\infty]$ space. 
-#' Let $X_1$ be the pdf of $N(\alpha_1,\sigma_1^2)$ and $X_2$ be the pdf of $N(\alpha_2,\sigma_2^2)$. 
-#' Again, since $$ b\gamma(t) := t X_1 + (1-t) X_2$$ belongs to $M$ for all $t$, it must be the shortest path between $X_1$ and $X_2$. The geodesic distance between $X_1$ and $X_2$ is
-#' $$\int_0^1 || \dot \gamma(t) ||_{L^2} \,dt = ||X_1 - X_2||_{L_2}$$ which has no easy analytic solution.
+
+#' # Scenario 2: 
 #' Marie's previous example where $$ M = \{ \text{probability density function } N(\alpha,1): \alpha \in \mathbb R \} $$ 
 #' is actually much harder to work with because the "straight" line connecting $X_1$ and $X_2$ in $M$ does not always stay inside of $M$. 
[…]

### 2019-05-14 index.Rmd
updated comments
included primer on differential geometry and added more description of simulation scenarios that should be implemented eventually

@@
 ---
-title: "Report"
+title: "Project Description"
 output:
@@
 
-# Introduction
+\newcommand {\To}{\rightarrow}
+\newcommand {\TO}{\Rightarrow}
+\newcommand {\R}{\mathbb{R}}
+\newcommand {\Prob}{\mathbb{P}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand {\cov}{\textrm{Cov}}
+\newcommand {\var}{\textrm{Var}}
+\newcommand {\1}{\textrm{\textbf{1}}}
+\newcommand{\M}{\mathcal{M}}
 
-The classical setup in FDA is to assume that we observe a sample of $n$ independant realizations $X_1,\ldots,X_n$ of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-The first steps of a classical approach to analyse such data are to first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE) and then perform FPCA to reduce the dimension of the data.
+## Introduction
+
+The classical setup in FDA assumes we observe a sample of $n$ independant realizations $X_1,\ldots,X_n$ of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+In the classical approach to analysing such data, we first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE), next perform FPCA to reduce the dimension of the data, and then proceed by employing multivariate techniques.
 
 However, it might very well happen that the function $X$ belongs to $\M \subset L^2([a,b],\R)$ a low-dimensional nonlinear manifold (gives examples and references).
-If it is indeed the case, it is no longer appropriate to use the norm induced by the $L^2$ inner product to calculate the distance between two curves.
-Explain why.
-One should instead use the geodesic distance $d_\M$ induced by $\M$ which takes into account the intrinsic structure of $\M$.
+If it is indeed the case, it may no longer be appropriate to use $L^2$ distance between two curves as a measure of distance.
+Scenario 2 in [sim_functional_data](sim_functional_data.html) is one such example.
+One may want to consider using some geodesic distance which takes into account the intrinsic structure of $\M$.
 
@@
 
-Now back to functional data living in a low-dimensional manifold embedded in the ambient Hilbert space.
-There exist many nonlinear dimension reduction methods in the literature, also known as manifold learning methods, which are particularly popular in computer vision.
+There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
 However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio.
 (is it true?
-ref?) This is a luxury that FDA cannot always afford.
-
+ref?) This is a luxury that FDA cannot always afford because functional data are usually observed with noise.
+The noise can be such that the observed functional data no longer live on a manifold.
 Though powerful techniques exist for smoothing/interpolating discretely observed noisy functional data, the recovered functional versions of the data, $\tilde X_1,\ldots,\tilde X_n$, may not lie even close to the functional manifold $\M$.
@@
 
-% the following two sentences are not related to each other
-% Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
+Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
 Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.
 
-With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$ when we have access only to discretely-observed noisy functional observations.
-Say why pairwise distances important, e.g.
+With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$ when we have access only to discretely-observed noisy functional observations that possibly live off the true manifold.
+Pairwise distances are important for many downstream tasks, e.g.
 classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression).
@@
 
-Along the way, we introduce a new tool to assess pairwise geodesic distance estimation.
-This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
+## Preliminaries
 
-# Proposed method for estimating geodesic distances
+We first rigorously define the concepts of geodesic and geodesic distance. We will also describe an important helper algorithm which we call IsoGeo.
 
-Let $X_1,\ldots,X_n$ be independent realizations of a random function $X\in\M \subset L^2([a,b],\R)$, where $\M$ is an unknown manifold of dimension $d$.
-Suppose that each curve $X_i$ is observed on a grid $T_i=(t_{i1},\ldots,t_{iK})$ and with measurements errors, i.e.
-that we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ are of mean zero and uncorrelated with each other.
+### Basic differential geometry
+This is take from a paper on [metric learning](http://proceedings.mlr.press/v32/linb14.pdf).
+Let $(\M,g)$ be a $d$-dimensional Riemannian manifold, where $g$ is a Riemannian metric tensor on $\M$.
+This metric tensor can be used to assign a metric on the manifold.
+For each point $p$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_p$ on the tangent space $T_p \M$.
+The norm of a tangent vector $v \in T_p \M$ is defined as $$||v|| = \sqrt{g_p(v,v)}$$
+If $\gamma: [a,b] \subset \R \to \M$ is a smooth curve, its length is then defined as
+$$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt $$
+The geodesic distance between two points $p,q$ on the manifold $\M$, based on this metric tensor $g$, is defined as
+$$ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.$$
+
+### IsoGeo
+
+Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+* The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
+The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
+* In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
+are estimated based on $G$.
+The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, and can be calculated easily with Floyd-Warshall or Dijkstra’s algorithms.
+* The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
+
+In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
+Note that Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, p-isomap is one of these).
+
+## Proposed method for estimating geodesic distances
+
+Let $X_1,\ldots,X_n$ be independent realizations of a random function $X\in\M \subset L^2([a,b],\R)$, where $\M$ is an unknown manifold.
+Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
+we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ are of mean zero and uncorrelated with each other.
 We assume that the grids $T_1,\ldots,T_n$ are dense.
@@
 3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
-4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n\times n $ matrix $\hat G$ as
+4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n \times n $ matrix $\hat G$ as
 $$
@@
 
-# Comparison to other methods
+## Comparison to other methods
 
-Susan: why are we describing Isomap as the first sentence in this Simulation section?
-Shouldn't we begin by describing the goal.
-Something like: We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances for discretely-observed noisy functional data, compared to various alternative approaches.
-
-MH : Sure, the introduction sentence was missing and we can also put the description of IsoMAP somewhere else.
-
-Susan question: Also, isn't the real name for IsoGeo Floyd-Warshall or Dijkstra’s? These are the two options in python scikitlearn's isomap routine, where the default is actually automatic selection.
-
-MH : Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, p-isomap is one of these).
-I wanted to introduce that IsoGeo to avoid confusion but maybe it was not clear, I changed a bit the description.
-
-Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
-The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, and can be calculated easily with Floyd-Warshall or Dijkstra’s algorithms.
-The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
-```{r}
-3 + 5
-```
-
-%In the sequel, we use the first step of Isomap to estimate pairwise geodesic distances and we call this procedure IsoGeo.
-In the sequel, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
+We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances for discretely-observed noisy functional data.
+???Describe simulation scenarios in sim_functional_data.R???
 
 We compare our method to the following ones:
-* [\bf RD] Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
-* [\bf SS] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+
+* **RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
+* **SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
-* [\bf pI] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+* **pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller).
 Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
- [\bf RP] same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
+* **RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
 
-# TODOs
+# Assessment of geodesic distance estimates
+We introduce a new tool to assess pairwise geodesic distance estimation.
+This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
 
-* Clean up the writing in this notebook. Add code and graphs where helpful.
+## TODOs
+
 * Related to tuning parameter

### 2019-05-14 sim_functional_data.R

@@ -1,21 +1,42 @@
-#' ## This function generates functional data living in a submanifold of a Hilbert space
+#' ---
+#' title: Functional manifold data examples
+#' ---
 
-#' # Scenario 1: 
-#' Consider the manifold $$M=\{X(t) = t-\alpha, -0.9 \le \alpha \le 3\} \subset L^2([a,b])$$ with inherited metric from the ambient $L^2[a,b]$ space.
-#' The shortest path $\gamma: [0,1] \to M$ between two functions $X_1 = (t-\alpha_1)$ and $X_2 = (t-\alpha_2)$ in $M$ is clearly given by $\gamma(t) = t X_1 + (1-t) X_2$.
-#' The geodesic distance between $X_1$ and $X_2$ is $$L(\gamma) := \int_0^1 || \dot \gamma(t) ||_{L^2} \,dt = ||X_1 - X_2||_{L_2} = (\alpha_1-\alpha_2)\sqrt{b-a}.$$ 
-
-#' # Scenario 2: 
-#' Marie's previous example where $$ M = \{ \text{probability density function } N(\alpha,1): \alpha \in \mathbb R \} $$ 
-#' is actually much harder to work with because the "straight" line connecting $X_1$ and $X_2$ in $M$ does not always stay inside of $M$. 
-#' To see this, note that $t N(\alpha_1,1) + (1-t) N(\alpha_2,1)$ no longer has variance $1$ for all $t \in [0,1]$. 
-
-#' # Scenario 3:
-#' Considered is the manifold $$ M = \{ \text{probability density function} N(\alpha,\sigma^2): \alpha \in \mathbb R, \sigma > 0 \} \subset L^2([-\infty,\infty])$$ with inherited metric from the ambient $L^2[-\infty,\infty]$ space. 
+#' \newcommand {\To}{\rightarrow}
+#'\newcommand {\TO}{\Rightarrow}
+#'\newcommand {\R}{\mathbb{R}}
+#'\newcommand {\Prob}{\mathbb{P}}
+#'\newcommand{\E}{\mathbb{E}}
+#'\newcommand {\cov}{\textrm{Cov}}
+#'\newcommand {\var}{\textrm{Var}}
+#'\newcommand {\1}{\textrm{\textbf{1}}}
+#'\newcommand{\M}{\mathcal{M}}
+#'
+#' ##Scenario 1: 
[…]

### 2019-05-14 sim_functional_data.R
add explanation scenaio 2 and adjust the initial value for a and b

@@ -19,11 +19,19 @@
 #'
 #'
 #' ##Scenario 2: 
-#' This is based on Manifold 2 in Chen and Muller 2012 where we fix the variance to be $1$. The paper claims that in this case the resulting manifold $$\M = \{ \text{probability density function } N(\alpha,1): \alpha \in \mathbb R \}$$ is isometric.
+#' This is based on Manifold 2 in Chen and Muller 2012 where we fix the variance to be $1$. The paper claims that in this case the resulting manifold $$\M =  \left \{ X_\beta \in L^2([-4,4]) : X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}, \beta \in \R\right \}.$$ is isometric.
 #' Chen and Muller are also working with the $L_2$ inner product as their metric tensor of $\M$. 
 #' Isometric means the geodesic distance between $X_1$ and $X_2$ in $\M$ is the Euclidan distance between the $\alpha$'s up to some scaling factor. ???Why is there a scaling factor???
 #' Note that the "straight" line connecting $X_1$ and $X_2$ in $\M$ does not always stay inside of $\M$ since $t N(\alpha_1,1) + (1-t) N(\alpha_2,1)$ no longer has variance $1$ for all $t \in [0,1]$. 
-#'
+#'  The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is 
+#'\begin{eqnarray*}
+#'d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
+#'&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
+#'&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
+#'&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
[…]

### 2019-05-14 test_analytic_geodesic_distance.Rmd
testing build of rmarkdown webpage
change html

@@
 output:
+  pdf_document: default
   html_document:
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache = TRUE)
+```
 
@@
 # Generate data
-sim <- sim_functional_data(sce=2,samplesize=1000)
+sim <- sim_functional_data(sce=2,samplesize=100)
 ```
@@
 # first call Sys.which("python") and then copy that path into 
-use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+#use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+use_python('/anaconda3/bin/python',required=TRUE)
 get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')
@@
 estim_geo_noiseless_data = getIsomapGdist$getIsomapGdist(noiseless_data_tmp,num_neigh_true[ind_op_true])
+image.plot(estim_geo_noiseless_data,main='geo estimation')
 ```
+
 If the analytic derviation is correct, then estim_geo_noiseless_data should be extremely similar to analytic_geo.
@@
 source('assess_goodness_estimation.R')
-assess_goodness_estimation(estim_geo_noiseless_data,analytic_geo)
+assess_goodness_estimation(estim_geo_noiseless_data,sim$analytic_geo)
 ```

### 2019-05-14 sim_functional_data.R
missing accolade

@@ -143,3 +143,5 @@ sim_functional_data<-function(sce,samplesize=100,K=30,a=-4,b=4,SNR=1,reg_samplin
   }
     
   return(list('noiseless_data'=noiseless_data,'noisy_data'=noisy_data,'analytic_geo'=analytic_geo,'grid'=grid,'reg_grid'=reg_grid))
+
+  

### 2019-05-14 sim_functional_data.R
problem of accolade

@@ -46,6 +46,7 @@
 #' $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$. The geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply
 #' $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>$$.
 
+
 ## Input 
 # samplesize
 # K = number of grid points
@@ -71,7 +72,6 @@ library(fields)
 source('full_geo_from_adj_geo.R')
 
 sim_functional_data<-function(sce,samplesize=100,K=30,a=-4,b=4,SNR=1,reg_sampling=1,com_grid=1,plot_true=1){
-  
   if(sce == 1){
     if(reg_sampling==0){
       Z <- rnorm(samplesize,0,0.3)
@@ -141,7 +141,8 @@ sim_functional_data<-function(sce,samplesize=100,K=30,a=-4,b=4,SNR=1,reg_samplin
     matplot(t(grid),t(noisy_data),main="Observed data",type='l', col=rainbow(samplesize))
     image.plot(analytic_geo,main='analytic geodesic')
   }
-    
-  return(list('noiseless_data'=noiseless_data,'noisy_data'=noisy_data,'analytic_geo'=analytic_geo,'grid'=grid,'reg_grid'=reg_grid))
+    return(list('noiseless_data'=noiseless_data,'noisy_data'=noisy_data,'analytic_geo'=analytic_geo,'grid'=grid,'reg_grid'=reg_grid))
+}
 
-  
+#'bla bla
+#'

### 2019-05-16 test_analytic_geodesic_distance.Rmd
html update
added abstract/title from IMS china writeup. took out caching in test_analytic_geodesic_distance markdown

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE)
-```
+# ```{r setup, include=FALSE}
+# #knitr::opts_chunk$set(cache = TRUE)
+# ```
 
@@
 # first call Sys.which("python") and then copy that path into 
-#use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
-use_python('/anaconda3/bin/python',required=TRUE)
+use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+#use_python('/anaconda3/bin/python',required=TRUE)
 get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')

### 2019-05-16 pairwise_geo_estimation.R
Creation of a new function to estimate pairwise geo only for FD and where one can specify the methods he wants to use.
I also changed the name of the method to be coherent with the report

@@ -1,0 +1,315 @@
+# TODO Marie: this function should call assess_goodness_estimation
+# TODO Marie: this function should actually be renamed to reflect that it is an oracle estimation procedure
+# TODO Marie/Susan: if it turns out random projection method g is still worth considering, implement parallel for loops
+##### Estimation of geodesic distances
+
+
+# We estimate the geodesic matrix obtained via
+# NN Floyd's Algorithm on noiseless data (just to make sure that the theoretical geo. is calculated correctly)
+# RD Floyd's Algorithm on noisy data
+# SS Floyd's Algorithm on smooth data 
+# pI p-isomap of Muller on smooth data
+# OUR mds of smooth data + scms + Floyd's Algorithm (only for FD)
+# RP : random proj. of smooth data + scms + Floyd's Algorithm (only for FD)
+
+
+## Note that a) and b) can only be used if the data are observed on a common grid
+## Note that c) can only be used for functional data
+## In the case of Eucledian data, d), e) and f) use directly the raw data
+
+## Input
+# method : list of dimension 6 indicating which estimation methods should be run. Warning : NN and RD only works if common_grid_true=1
+# true_data : samplesize x K matrix containing the original data (no noise)
+# discrete_data : samplesize x K matrix containing the observed data
+# true_geo : samplesize x samplesize matrix containing true pairwise geo disctance
+# s : dimension use for mds and random projections
+# plot_true : if TRUE plot of geo estimation for each method
+# FD_true : if TRUE functional data, otherwise euclidean data
+# grid : samplesize x K matrix containing the grid on which each data is observed (required only if FD_true =1)
+# common_grid_true : if 1 grid is common for every curve, if 0 different grid for every curve
+
+## Output
+# list of estimated geodesic distance :
+# estim_geo_true_data = method NN
+# estim_geo_noisy_data = method RD
+# estim_geo_smooth_data = method SS
+# estim_geo_penalized_isomap = method pI
[…]

### 2019-05-16 Isomap.R
added information on what return the function

@@ -38,8 +38,8 @@ pairwiseDistances = function(data){
 ##	verbose:	show a summary of the embedding procedure at the end (default TRUE)	 
 ##
 ## output: 
-##	Nxdim matrix (N samples, dim features) with the reduced input data (list of several matrices if more than one dimension specified)
-
+##	low_mani :Nxdim matrix (N samples, dim features) with the reduced input data (list of several matrices if more than one dimension specified)
+##  geod : NxN pairwise geo distance matrix
 
 Isomap = function(data, dims=2, k, mod=FALSE){
 #Isomap = function(data,  k, mod=FALSE, plotResiduals=FALSE, verbose=FALSE){
@@ -132,9 +132,6 @@ Isomap = function(data, dims=2, k, mod=FALSE){
 		i = i+1
 	}
 	
-	rep <- list(all_Y,d)
-	names(rep) <- c("bla","geod")
-
-	return(rep)
+	return(list("low_mani" = all_Y,"geod" = d))
 
 }

### 2019-05-16 main.R
Updated the main to use the function pairwise_geo_estimation for FD. It it now possible to do a plot where the different methods and measures of efficiency are compared.

@@ -1,16 +1,10 @@
 rm(list = ls())
 
-# setwd("/Users/marie_uqam/Dropbox/Marie-Moi/Susan_project/manifold learning/Geo_dist_calculation")
-
 source('EuclideanExamples.R')
-source('functions.R')
-source('Isomap.R')
-source('Modif_Isomap.R')
-source('pairwiseDistances.R')
 source('full_geo_from_adj_geo.R')
-source('smooth_FD_bspline.R')
 source('sim_functional_data.R')
 source('Geo_estimation.R')
+source('pairwise_geo_estimation.R')
 source('sim_Euclidean_data.R')
 source('assess_goodness_estimation.R')
 
@@ -22,11 +16,50 @@ library(matlabr)
 
 # TODO: it doesn't make ansy sense to define these outside of Geo_estimation?
 # use right version of python
-# use_python('/anaconda3/bin/python',required=TRUE)
+use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
 # use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
 pyIso = import_from_path("getIsomapGdist",path='.')
 py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
 scms = import_from_path("scms",path='.')
+
+
+#### Analysis of Functional data
+
+FD_true =TRUE
+
+# set up parameters
+sce =3 # 1, 2 or 3 (for more info see sim_functional_data.R)
+K = 30 # number of grid points (each curve is observed on K points on [a,b])
+samplesize = 100 # number of points on the manifold
+SNR = 0.1 # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
+reg_sampling=TRUE # regular sampling of the point on the manifold or uniformly random
+plotTrue= TRUE 
+com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
+nb_proj = 20 # number of random projection
+meth<- list("NN" = TRUE,"RD" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"RP" = TRUE ) # see pairwise_geo_estimation for more info
+
+# Generate data
+data<- sim_functional_data(sce,samplesize,K,SNR,reg_sampling,com_grid,plotTrue)
+
+# Estimation of geodesic distances with different methods
[…]

### 2019-05-16 sim_functional_data.R
Added scenario 3 and incorportate the domain (a,b) in the function instead of being parameters (to avoid confusion).

@@ -33,7 +33,7 @@
 #'\end{eqnarray*}
 #'where the approximation comes from the fact that we are integrating on $[-4,4]$ and not on $\R$. Moreover the mean $\beta$ has to be close to 0 for the approximation to make sense.
 #'
-#' #Scenario 3: ???This needs to be implemented eventually???
+#' #Scenario 3: 
 #' Consider the manifold $$ \M = \{ \text{probability density function } N(\alpha,\sigma^2): \alpha \in \mathbb R, \sigma > 0 \} $$ with the $L_2$ inner product as the metric tensor.
 #' Let $X_1$ be the pdf of $N(\alpha_1,\sigma_1^2)$ and $X_2$ be the pdf of $N(\alpha_2,\sigma_2^2)$. 
 #' Again, since $$ \gamma(t) := t X_1 + (1-t) X_2$$ belongs to $M$ for all $t$, it must be the shortest path between $X_1$ and $X_2$. The geodesic distance between $X_1$ and $X_2$ is
@@ -71,8 +71,10 @@
 library(fields)
 source('full_geo_from_adj_geo.R')
 
-sim_functional_data<-function(sce,samplesize=100,K=30,a=-4,b=4,SNR=1,reg_sampling=1,com_grid=1,plot_true=1){
+sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_grid=1,plot_true=1){
   if(sce == 1){
+    a<- -4
+    b<- 4
     if(reg_sampling==0){
       Z <- rnorm(samplesize,0,0.3)
       alpha <- apply(cbind(rep(-1,samplesize),Z),1,max)
@@ -91,10 +93,10 @@ sim_functional_data<-function(sce,samplesize=100,K=30,a=-4,b=4,SNR=1,reg_samplin
     analytic_geo <- full_geo(adja_geo,samplesize)
     
   } else if(sce ==2){
-    
+    a<- -4
+    b<- 4
     if(reg_sampling==0){
       Z <- rnorm(samplesize,0,0.3)
-      alpha <- apply(cbind(rep(-1,samplesize),Z),1,max)
       alpha=sort(alpha)
     } else if(reg_sampling==1){
       alpha <- seq(-1,1,length.out=samplesize)
[…]

### 2019-05-16 test_analytic_geodesic_distance.Rmd
updated webpage

--- 2019-05-14 test_analytic_geodesic_distance.Rmd
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE)
-```
-
 ```{r}
@@
 # Generate data
-sim <- sim_functional_data(sce=2,samplesize=100)
+sim <- sim_functional_data(sce=3,samplesize=100)
 ```
@@
 #use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
-use_python('/anaconda3/bin/python',required=TRUE)
+use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
 get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')

### 2019-05-16 Geo_estimation.R

@@ -53,7 +53,7 @@ py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
 scms = import_from_path("scms",path='.')
 
 Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,nb_proj,grid,reg_grid,common_grid_true=0){
-
+  
   samplesize = nrow(true_data)
   K = ncol(true_data)
   norm_true_geo = sqrt(sum(true_geo^2))
@@ -61,14 +61,14 @@ Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,
   list2_to_return=list()
   list3_to_return=list()
   list4_to_return=list()
-
+  
   ## Apply method a) and b) for FD observed on a common grid or Euclidean data
   if(!FD_true||common_grid_true==1){
-
+    
     ### a) Estimation from the true data
-
+    
     print("Floyd's Algorithm on true data")
-
+    
     if(FD_true){
       a = reg_grid[1]
       b = reg_grid[K]
@@ -76,11 +76,11 @@ Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,
     } else{
       true_data_tmp = true_data
     }
-
+    
     # Find a grid of possible values for the number of neigbors
     num_neigh_min=py_min_neigh$get_min_num_neighbors(true_data_tmp)
     num_neigh_true=seq(num_neigh_min,samplesize/2,by=2)
-
+    
     # Calculate the geo matrix for each number of neighbors and keep the one that gives the minimal error
     Error_true_mani_K= rep(0,length(num_neigh_true))
     for(j in 1:length(num_neigh_true)){
@@ -89,11 +89,11 @@ Geo_estimation <- function(true_data,discrete_data,true_geo,plot_true,FD_true,s,
     }
     ind_op_true=min(which(Error_true_mani_K==min(Error_true_mani_K)))
     estim_geo_true_data = pyIso$getIsomapGdist(true_data_tmp,num_neigh_true[ind_op_true])
-
+    
     ### b) Direct estimation from the noisy data
-
+    
     print("Floyd's Algorithm on raw data")
-
+    
     if(FD_true){
       a = reg_grid[1]
       b = reg_grid[K]
[…]

### 2019-05-16 index.Rmd

@@
 \newcommand{\M}{\mathcal{M}}
+
+## Abstract
+
+Title: Geodesic distance estimation for functional data residing near a nonlinear manifold.
+
+Abstract: Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example.
+Since these spaces are not vector spaces, basic operations such as addition and subtraction require special consideration and treatment.
+In this work we set ourselves the task of calculating a geodesic distance between two functions living near a non-linear manifold due to noise.
+This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
+Good estimation of the geodesic distance has implications for many downstream tasks such as clustering, classification, and nonparametric functional regression.
+
+*I wrote this for the IMS China 2019 conference.*
 

### 2019-05-16 test_analytic_geodesic_distance.Rmd
fix webpage

@@
 # first call Sys.which("python") and then copy that path into
+
 #use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
 use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+
 get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')

### 2019-05-22 test_analytic_geodesic_distance.Rmd
updated compare_methods to work with pairwise_geoestimation

@@
 output:
-  pdf_document: default
   html_document:
     df_print: paged
+  pdf_document: default
 ---
+
+```{r}
+library(reticulate)
+# sometimes reticulate requires specific instruction on where to find your python
+# first call Sys.which("python") and then copy that path into
+
+use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+#use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+```
 
@@
 ```{r}
-library(reticulate)
-# sometimes reticulate requires specific instruction on where to find your python
-# first call Sys.which("python") and then copy that path into
-
-#use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
-use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
-
 get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')

### 2019-05-22 compare_methods.R

@@ -10,67 +10,60 @@ print(slurm_arrayid)
 # parameters that will be fixed during the sim study
 K = 30 # number of grid points (each curve is observed on K points on [a,b])
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
-nb_proj = 100 # number of random projection
 plotTrue = FALSE
-a = -2
-b = 3
 FD_true = TRUE
-sce = 1 # there is something wrong with the true geodesic distance calculation for scenario 2, so fix this at scenario 1 for now
+meth<- list("NN" = TRUE,"RD" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"RP" = FALSE)
 
 # parameters under study
-# sce = {1,2} (for more info see sim_functional_data.R),K,a,b,SNR,reg_sampling,com_grid,plot_true
+# scenario
 # samplesize # number of points on the manifold
 # SNR # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
 # reg_sampling {True,False} # regular sampling of the point on the manifold or uniformly random
-# s {2,3,4} # reduced dimension use for mds and random projection
 
+sces = 1:3
 samplesizes = c(100,250)
 SNRs = c(0.1,0.5)
 reg_samplings = c(TRUE,FALSE)
-ss = c(2,3,4)
 mcs = 1:100
 
-unravel=arrayInd(slurm_arrayid,c(length(samplesizes), length(SNRs), length(reg_samplings), length(ss), length(mcs)))
+unravel=arrayInd(slurm_arrayid,c(length(sces),length(samplesizes), length(SNRs), length(reg_samplings), length(mcs)))
 
 # actual parameters for this run
-samplesize = samplesizes[unravel[1,1]]
-SNR = SNRs[unravel[1,2]]
-reg_sampling = reg_samplings[unravel[1,3]]
-s = ss[unravel[1,4]]
+sce = sces[unravel[1,1]]
+samplesize = samplesizes[unravel[1,2]]
+SNR = SNRs[unravel[1,3]]
+reg_sampling = reg_samplings[unravel[1,4]]
 mc = mcs[unravel[1,5]]
 
+
 source('EuclideanExamples.R')
-source('functions.R')
-source('Isomap.R')
-source('Modif_Isomap.R')
-source('pairwiseDistances.R')
[…]

### 2019-05-22 compare_methods.sh

@@ -25,9 +25,9 @@
 #SBATCH --mail-user=susan.wei@unimelb.edu.au
 
 # The maximum running time of the job in days-hours:mins:sec
-#SBATCH --time=0-2:00:00
+#SBATCH --time=1-0:00:00
 
-#SBATCH --array=1-2400%60         ### Array index | %50: number of simultaneously tasks
+#SBATCH --array=1-2400         ### Array index | %50: number of simultaneously tasks
 
 # Get Array ID
 i=${SLURM_ARRAY_TASK_ID}
@@ -41,8 +41,8 @@ fi
 # Run the job from the directory where it was launched (default)
 
 # The modules to load:
-module load MATLAB
+# module load MATLAB
 
 # The job command(s):
 # R --vanilla < spartanSim.R ${i} #this isn't necessary since we get the environment
-R --vanilla < spartanSim.R
+R --vanilla < compare_methods.R

### 2019-05-22 main.R

@@ -1,3 +1,8 @@
+library(reticulate)
+# use right version of python
+# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
 rm(list = ls())
 
 source('EuclideanExamples.R')
@@ -10,14 +15,10 @@ source('assess_goodness_estimation.R')
 
 library(DescTools)
 library(fields)
-library(reticulate)
 library(fda)
 library(matlabr)
 
-# TODO: it doesn't make ansy sense to define these outside of Geo_estimation?
-# use right version of python
-use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
-# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
 pyIso = import_from_path("getIsomapGdist",path='.')
 py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
 scms = import_from_path("scms",path='.')

### 2019-06-20 index.Rmd
scms now uses silverman's rule but keeps isotropic kernel bandwidth matrix using sigma_min of cosmic reconstruction paper.
added scenario 4

--- 2019-05-14 index.Rmd
@@
 \newcommand{\M}{\mathcal{M}}
+
+## Abstract
+
+Title: Pairwise geodesic distance estimation for functional data residing near a nonlinear manifold.
+
+Abstract: Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example.
+Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration and treatment.
+In this work we set ourselves the task of calculating the geodesic distance between two functions living near a non-linear manifold due to noise.
+This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
+Good estimation of the geodesic distance has implications for many downstream tasks in functional data analysis, including functional clustering, and nonparametric functional regression.
 
@@
 
-## Comparison to other methods
+## Simulaton Study
 
@@
 
-We compare our method to the following ones:
+# Comparison to other methods
 
-* **RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
-* **SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+* **(Raw Data) RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
+* **(Spline Smoothing) SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
-* **pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+* **(Penalized-Isomap) pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller).
 Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
-* **RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
+* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
 
 # Assessment of geodesic distance estimates
-We introduce a new tool to assess pairwise geodesic distance estimation.
+We introduce a new metric for assessing the quality of a pairwise geodesic distance estimator.
 This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.

### 2019-06-20 test_analytic_geodesic_distance.Rmd
scms now uses silverman's rule but keeps isotropic kernel bandwidth matrix using sigma_min of cosmic reconstruction paper.
added scenario 4

@@
 library(reticulate)
-# sometimes reticulate requires specific instruction on where to find your python
-# first call Sys.which("python") and then copy that path into
 
-use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
-#use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+use_python('/usr/local/bin/python3.7',required=TRUE)
+# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
+```
+
+Check with py_config()
+
+```{r}
+py_config()
 ```
@@
 
-The function sim_functional_data.R allows specification of several simulatin scenarios. The samplesize is purposefully set very high for reasons explained below.
+The function sim_functional_data.R allows specification of several simulatin scenarios.
+The samplesize should be set very high so that Floyd's algorithm works well on the noiseless data.
+The number of time grid points K also should be high so that the vector data really represents the functional curve.
 
@@
 # Generate data
-sim <- sim_functional_data(sce=3,samplesize=100)
+sim <- sim_functional_data(sce=4,samplesize=1500,K=150,plot_true=1)
 ```
@@
 We can assess their closeness using assess_goodness_estimation.R.
-Running this notebook for sce=1 gives rmse on the order of e-16.
+Three assessment metrics are returned.
+If the calculation is correct, rmse should be close to zero and epsilon_isometry_auc and pearson_corr should both be close to 1.
+Note that running this notebook for sce=1 gives rmse on the order of e-16.
 ```{r}

### 2019-06-20 main.R

@@ -1,7 +1,8 @@
 library(reticulate)
 # use right version of python
+use_python('/usr/local/bin/python3.7',required=TRUE)
 # use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
-use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
 
 rm(list = ls())
 

### 2019-06-20 scms.py

@@ -1,13 +1,26 @@
 import numpy as np
 
-def make_isotropic_gaussian_kernel(sigma):
+
+def make_anisotropic_gaussian_kernel(Hdiag):
     """
     Makes an isotropic gaussian kernel function with bandwidth sigma.
     """
+
     # the kernel function is a function of ||x-y||^2, the squared Euclidean
     # distance between x and y
-    def kernel(d_squared):
-        return 1/np.sqrt(2*np.pi*sigma**2)*np.exp(-1/2.*d_squared/sigma**2)
+
+    # hardcode bandwidth matrix H according to Silverman's
+    H = np.diag(Hdiag)
+
+    def kernel(diff):
+        # diff should be dim by sample size
+        diff = np.transpose(diff)
+        d = diff.shape[0]
+        prod = np.matmul(np.transpose(diff), np.linalg.inv(H))
+        prod = np.matmul(prod, diff)
+        weight_matrix = (2 * np.pi) ** (-d / 2)*np.linalg.det(H)**(-1/2) * np.exp( (-1/2)* prod)
+        weights = np.diag(weight_matrix)
+        return weights
 
     return kernel
 
@@ -18,63 +31,61 @@ def kernel_density_estimate(x, data, kernel):
     This obviously does not use data-dependent kernels.
     """
     # first, we calculate the interpoint distances
-    d_squared = np.sum((data - x)**2, axis=1)
+    d_squared = np.sum((data - x) ** 2, axis=1)
 
     # and evaluate the kernel at each distance
-    weights = kernel(d_squared)
+    weights = kernel(data-x)
 
     # return the kernel density estimate
     return np.mean(weights)
 
 
-def gradient(x, data, kernel, sigma):
+def gradient(x, data, kernel, Hdiag):
     """
     Computes the gradient at x of the kde of the data using isotropic Guassian
     kernels with bandwidth sigma.
     """
     # the number of data points and the dimensionality
-    n,d = data.shape
+    n, d = data.shape
 
     # first, we form the u_i
-    u = 1./(sigma**2)*(data - x)
+    u = np.matmul((data - x),np.diag(1./Hdiag))
 
     # now we form the c_i
-    d_squared = np.sum((data - x)**2, axis=1)
-    c = kernel(d_squared)
+    d_squared = np.sum((data - x) ** 2, axis=1)
+    c = kernel(data-x)
 
[…]

### 2019-06-20 sim_functional_data.R

@@ -13,17 +13,17 @@
 #'\newcommand{\M}{\mathcal{M}}
 #'
 #' ##Scenario 1: 
-#' Consider the manifold $$ \M = \{X(t) = t-\alpha, -0.9 \le \alpha \le 3, t \in [a,b]\} $$ with the $L_2$ inner product as the metric tensor.
-#' The shortest path $\gamma: [0,1] \to \M$ between two functions $X_1 = (t-\alpha_1)$ and $X_2 = (t-\alpha_2)$ in $\M$ is clearly given by $\gamma(t) = t X_1 + (1-t) X_2$.
-#' The geodesic distance between $X_1$ and $X_2$ is $$L(\gamma) = \int_0^1 || \dot \gamma(t) ||_{L^2} \,dt = ||X_1 - X_2||_{L_2} = (\alpha_1-\alpha_2)\sqrt{b-a}.$$ 
+#' Consider the manifold $$ \M = \{X_\alpha: -1 \le \alpha \le 1\} $$ with the $L_2$ inner product as the metric tensor,
+#' where $X_\alpha: [a,b] \to \R$ is given by $X_\alpha(t) = t-\alpha$. Below, we set $a=-4$ and $b=4$.
+#' The shortest path $\gamma: [0,1] \to \M$ between two functions $X_{\alpha_1} = (t-\alpha_1)$ and $X_{\alpha_2} = (t-\alpha_2)$ in $\M$ is clearly given by $\gamma(t) = t X_{\alpha_1} + (1-t) X_{\alpha_2}$.
+#' The geodesic distance between $X_{\alpha_1}$ and $X_{\alpha_2}$ is $$L(\gamma) = \int_0^1 || \dot \gamma(t) ||_{L^2} \,dt = ||X_{\alpha_1} - X_{\alpha_2}||_{L_2} = (\alpha_1-\alpha_2)\sqrt{b-a}.$$ 
 #'
 #'
 #' ##Scenario 2: 
-#' This is based on Manifold 2 in Chen and Muller 2012 where we fix the variance to be $1$. The paper claims that in this case the resulting manifold $$\M =  \left \{ X_\beta \in L^2([-4,4]) : X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}, \beta \in \R\right \}.$$ is isometric.
-#' Chen and Muller are also working with the $L_2$ inner product as their metric tensor of $\M$. 
[…]

### 2019-06-20 main.R
Implementation of robust Isomap which does not require the number of neighbors in Floyd algo.

@@ -1,7 +1,8 @@
 library(reticulate)
 # use right version of python
-# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
-use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+#use_python('/usr/bin/python',required=TRUE)
+use_python('/anaconda3/envs/r-reticulate/bin/python',required=TRUE)
+#use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
 
 rm(list = ls())
 
@@ -12,12 +13,13 @@ source('Geo_estimation.R')
 source('pairwise_geo_estimation.R')
 source('sim_Euclidean_data.R')
 source('assess_goodness_estimation.R')
+source('robust_isomap.R')
 
 library(DescTools)
 library(fields)
 library(fda)
 library(matlabr)
-
+library(igraph)
 
 pyIso = import_from_path("getIsomapGdist",path='.')
 py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
@@ -29,7 +31,7 @@ scms = import_from_path("scms",path='.')
 FD_true =TRUE
 
 # set up parameters
-sce =3 # 1, 2 or 3 (for more info see sim_functional_data.R)
+sce =2 # 1, 2 or 3 (for more info see sim_functional_data.R)
 K = 30 # number of grid points (each curve is observed on K points on [a,b])
 samplesize = 100 # number of points on the manifold
 SNR = 0.1 # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
@@ -37,7 +39,7 @@ reg_sampling=TRUE # regular sampling of the point on the manifold or uniformly r
 plotTrue= TRUE 
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 nb_proj = 20 # number of random projection
-meth<- list("NN" = TRUE,"RD" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"RP" = TRUE ) # see pairwise_geo_estimation for more info
+meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"RI"=TRUE,"OUR" = TRUE,"OUR2" = TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info
 
 # Generate data
 data<- sim_functional_data(sce,samplesize,K,SNR,reg_sampling,com_grid,plotTrue)

### 2019-06-20 pairwise_geo_estimation.R

@@ -6,19 +6,21 @@
 
 # We estimate the geodesic matrix obtained via
 # NN Floyd's Algorithm on noiseless data (just to make sure that the theoretical geo. is calculated correctly)
-# RD Floyd's Algorithm on noisy data
-# SS Floyd's Algorithm on smooth data 
+# RD_o Floyd's Algorithm on noisy data (oracle for num neigh)
+# RD Floyd's Algorithm on noisy data (no oracle)
+# SS_o Floyd's Algorithm on smooth data (oracle for num neigh)
+# SS Floyd's Algorithm on smooth data (no oracle)
 # pI p-isomap of Muller on smooth data
-# OUR mds of smooth data + scms + Floyd's Algorithm (only for FD)
-# RP : random proj. of smooth data + scms + Floyd's Algorithm (only for FD)
+# RI robust isomap of Dimeglio on smooth data (by construction, no oracle needed for num neigh)
+# OUR mds of smooth data + scms + Floyd's Algorithm (oracle for num neigh)
+# OUR2 mds of smooth data + scms + robust isomap (no oracle)
+# RP : random proj. of smooth data + scms + Floyd's Algorithm 
 
 
-## Note that a) and b) can only be used if the data are observed on a common grid
-## Note that c) can only be used for functional data
-## In the case of Eucledian data, d), e) and f) use directly the raw data
+## Note that NN and RD can only be used if the data are observed on a common grid
 
 ## Input
-# method : list of dimension 6 indicating which estimation methods should be run. Warning : NN and RD only works if common_grid_true=1
+# method : list of dimension 8 indicating which estimation methods should be run. Warning : NN and RD only works if common_grid_true=1
 # true_data : samplesize x K matrix containing the original data (no noise)
 # discrete_data : samplesize x K matrix containing the observed data
 # true_geo : samplesize x samplesize matrix containing true pairwise geo disctance
@@ -31,16 +33,21 @@
 ## Output
 # list of estimated geodesic distance :
 # estim_geo_true_data = method NN
+# estim_geo_noisy_data_o = method RD_o
[…]

### 2019-06-20 robust_isomap.R

@@ -1,0 +1,47 @@
+
+
+robust_iso<- function(X){
+  n <- dim(X)[1]
+  D <- as.matrix(dist(X))
+  A <- manif(X)
+  DA <- D*A
+  G <- graph.adjacency(DA, "undirected", weighted = TRUE)
+  GD <- shortest.paths(G)
+}
+
+
+mst <- function(X) {
+  D <- as.matrix(dist(X))
+  G <- graph.adjacency(D, "undirected", weighted = TRUE)
+  T <- minimum.spanning.tree(G)
+  A <- get.adjacency(T)
+  return(A)
+}
+
+
+# Function to obtain an extended graph
+manif <- function(X) {
+  D <- as.matrix(dist(X))
+  A <- mst(X)
+  DA <- D*A
+  R <- apply(DA, 1, max)
+  n <- dim(X)[1]
+  d <- dim(X)[2]
+  A <- matrix(0, nrow=n, ncol=n)
+  L <- 100
+  f <- function(l,k) {
+    (1-l)*X[i,k]+l*X[j,k]
+  }
+  for (i in 1:(n-1)) {
+    for (j in (i+1):n) {
+      P <- outer(1:L/(L+1), 1:d, f)
+      E <- as.matrix(dist(rbind(P, X)))
+      F <- E[(L+1):(L+n), 1:L]
+      if (all(apply(F<R, 2, any))) A[i,j]=1
+    }
+  }
+  A <- A + t(A)
+  return(A)
+}
+
+

### 2019-06-20 compare_methods.R
picking right parameters for compare_methods submission

@@ -12,7 +12,7 @@ K = 30 # number of grid points (each curve is observed on K points on [a,b])
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 plotTrue = FALSE
 FD_true = TRUE
-meth<- list("NN" = TRUE,"RD" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"RP" = FALSE)
+meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )
 
 # parameters under study
 # scenario
@@ -20,7 +20,7 @@ meth<- list("NN" = TRUE,"RD" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"RP" =
 # SNR # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
 # reg_sampling {True,False} # regular sampling of the point on the manifold or uniformly random
 
-sces = 1:3
+sces = c(1,2,4)
 samplesizes = c(100,250)
 SNRs = c(0.1,0.5)
 reg_samplings = c(TRUE,FALSE)

### 2019-06-24 index.Rmd
modularizing real data analysis

@@
 ---
-title: "Project Description"
+title: "The hidden manifold distance for functional manifold data"
 output:
@@
     df_print: paged
+    number_sections: true
 ---
@@
 
-## Abstract
+# Abstract
 
-Title: Pairwise geodesic distance estimation for functional data residing near a nonlinear manifold.
-
-Abstract: Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example.
+Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example.
 Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration and treatment.
@@
 
-## Introduction
+# Introduction
 
 The classical setup in FDA assumes we observe a sample of $n$ independant realizations $X_1,\ldots,X_n$ of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-In the classical approach to analysing such data, we first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE), next perform FPCA to reduce the dimension of the data, and then proceed by employing multivariate techniques.
-
 However, it might very well happen that the function $X$ belongs to $\M \subset L^2([a,b],\R)$ a low-dimensional nonlinear manifold (gives examples and references).
 If it is indeed the case, it may no longer be appropriate to use $L^2$ distance between two curves as a measure of distance.
-Scenario 2 in [sim_functional_data](sim_functional_data.html) is one such example.
+<!-- Scenario 2 in [sim_functional_data](sim_functional_data.html) is one such example.  -->
 One may want to consider using some geodesic distance which takes into account the intrinsic structure of $\M$.
@@
 
-## Preliminaries
+## Setting
+Describe preprocessing of discretely observed functional data: In the classical approach to analysing such data, we first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE), next perform FPCA to reduce the dimension of the data, and then proceed by employing multivariate techniques.
+
+What type of noise structures are allowed/considered? What type of time sampling frequency is assumed? Do not need time-aligned curves!
+
+# Preliminaries
 
@@
 
-### Basic differential geometry
+## Basic differential geometry
 This is take from a paper on [metric learning](http://proceedings.mlr.press/v32/linb14.pdf).
@@
 
-### IsoGeo
+## IsoGeo
 
@@
 
-## Proposed method for estimating geodesic distances
+# Proposed method for estimating geodesic distances
 
@@
 
-## Simulaton Study
+## Selection of tuning parameters
+
+# Simulation study
 
 We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances for discretely-observed noisy functional data.
-???Describe simulation scenarios in sim_functional_data.R???
 
-# Comparison to other methods
+## Alternative estimators of geodesic distance
 
@@
 
-# Assessment of geodesic distance estimates
+## Simulation scenarios
+
+## Performance metrics
 We introduce a new metric for assessing the quality of a pairwise geodesic distance estimator.
@@
 
-## TODOs
+## Results
 
-* Related to tuning parameter
-    + Choose the number of neighbors $N$ in IsoGeo.
-    + Choose projection dimension for avoiding curse of dimensionality in performing density ridge scms routine
-    + Choose the bandwidth $h$ in scms.
-* Related to spartanSim.R simulations results
-    + Why do some Monte Carlo simulations fail?
-    + Is P-isomap always choosing zero penalty automatically?
-    + Shouldn’t the smoothing method be doing better? Thought the smooth was close to the true curve.
-* Next time we submit spartanSim.R, make sure to hold dataset constant for different projection dimensions
-* Read Marie's suggested paper on [robust ISOMAP](https://arxiv.org/abs/1306.3373)
-* Revisit Mueller manifold examples. Need to write down some differential equations to solve the geodesic minimisation problem?
+# Study on "misspecification"
+
+Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance.
+Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance.
+
+# Downstream tasks
+
+In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task.
+There are many tasks we could consider here, but we will focus on distanced-based functional clustering and distance-based functional classification.
+It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.
+
+In both tasks below, we compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.
+
+## Distance-based functional clustering
+
+Describe the $k$-medoids clustering method.
+
+### Simulated data
+
+###  Real data
+
+## Distance-based functional classification
+
+For simplicity, assume the task is bianry classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
+Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
+$$
+\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
+$$
+
+### Simulated data
+
+###  Real data
+A few examples of the datasets used by functional classification papers
+
+* Wheat, rainfall and phoneme in "Achieving near-perfect classification for functional data"
+* Berkeley growth curves in Chen et. al 2014, Biometrics.
+* Tecator and phoneme in technometrics paper.

### 2019-06-24 analyse_growth_curves.R

@@ -1,47 +1,92 @@
 #### Classification of growth curves dataset based on orginal data and velocity
 ## We estimate the geodesic distance matrix with 4 methods
 
-
-
 library(reticulate)
 library(fda)
 library(scatterplot3d)
 library(kmed)
 library(fields)
 library(igraph)
+library(ROCR)
+library(caret)
 source('robust_isomap.R')
 source('pairwise_geo_estimation.R')
 
+# input: lists of discrete_data, reg_grid, and true_group
+# discrete_datas_list
+# reg_grid_list
+# true_group_list
 
-### Reading the data
+# call processRealData which returns return(list("discrete_datas_list" = discrete_datas_list, "reg_grid_list" = reg_grid_list, "true_group_list"=true_group_list))
 
-growth_data <- t(cbind(growth$hgtm,growth$hgtf))
-growth_grid <- growth$age
-true_group<- c(rep(0,39),rep(1,54)) # 0 = boy and 1 = girl
+data = processRealData()
+discrete_datas_list = data[[1]] # etc
 
 # We apply our method to obtain the pairwise geodesic distance (and floyd on smooth data)
 meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = TRUE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )
-Estim<- pairwise_geo_estimation(meth,NA,growth_data,NA,FALSE,TRUE,NA,matrix(rep(growth_grid,93),nrow=93,byrow = TRUE),growth_grid,1,FALSE)
 
-class_err<-rep(0,500)
-for(split in 1:500){
-  train_ind<- sort(sample(1:93,50))
-  test_ind_temp<- 1:93 
-  test_ind<- test_ind_temp[-train_ind]
+estimate_distance <- function(discrete_data,reg_grid){
   
-  # We classify the data using functional version of the Nadaraya–Watson kernel estimator for the class membership probabilities
-  h<- 10
-  deno <- colSums(dnorm(h^(-1)*Estim$estim_geo_mds_scms_RI_h_heuri_s3[train_ind,test_ind]))
-  num_prob_gars<- colSums(dnorm(h^(-1)*Estim$estim_geo_mds_scms_RI_h_heuri_s3[train_ind[which(train_ind<40)],test_ind]))
   
[…]

### 2019-06-24 pairwise_geo_estimation.R

@@ -1,35 +1,32 @@
-# TODO Marie: this function should call assess_goodness_estimation
-# TODO Marie: this function should actually be renamed to reflect that it is an oracle estimation procedure
-# TODO Marie/Susan: if it turns out random projection method g is still worth considering, implement parallel for loops
-##### Estimation of geodesic distances
-
+##### Estimation of pairwise geodesic distances
 
 # We estimate the geodesic matrix obtained via
-# NN Floyd's Algorithm on noiseless data (just to make sure that the theoretical geo. is calculated correctly)
-# RD_o Floyd's Algorithm on noisy data (oracle for num neigh)
-# RD Floyd's Algorithm on noisy data (no oracle)
-# SS_o Floyd's Algorithm on smooth data (oracle for num neigh)
-# SS Floyd's Algorithm on smooth data (no oracle)
-# pI p-isomap of Muller on smooth data
-# OUR mds of smooth data + scms + Floyd's Algorithm (oracle for h and num neigh)
-# OUR2 mds of smooth data + scms + robust isomap (oracle for h)
-# OUR3 mds of smooth data + scms h heuri + robust isomap (no oracle)
+# NN: Floyd's Algorithm on noiseless data (just to make sure that the theoretical geo. is calculated correctly)
+# RD_o: Floyd's Algorithm on noisy data (oracle for num neigh)
+# RD: Floyd's Algorithm on noisy data (no oracle)
+# SS_o: Floyd's Algorithm on smooth data (oracle for num neigh)
+# SS: Floyd's Algorithm on smooth data (no oracle)
+# pI: p-isomap of Muller on smooth data
+# OUR: mds of smooth data + scms + Floyd's Algorithm (oracle for h and num neigh)
+# OUR2: mds of smooth data + scms + robust isomap (oracle for h)
+# OUR3: mds of smooth data + scms h heuri + robust isomap (no oracle)
 # RP : random proj. of smooth data + scms + Floyd's Algorithm 
 
-
-## Note that NN and RD can only be used if the data are observed on a common grid
+# TODO: what if data are not observed on a common grid and NN, RD, RD_o are called? Are there warnings?
+## Note that NN, RD, RD_o can only be used if the data are observed on a common grid
 
 ## Input
 # method : list of dimension 8 indicating which estimation methods should be run. Warning : NN and RD only works if common_grid_true=1
[…]

### 2019-06-24 processRealData.R
processRealData for growth curves

@@ -1,0 +1,46 @@
+library("rlist")
+processRealData = function(){
+  
+  list_of_lists = list()
+  
+  ### GROWTH CURVES from fda package
+  # A list containing the heights of 39 boys and 54 girls from age 1 to 18 and the ages at which they were collected.
+  
+  #  --- Smooth, penalizing the Nth derivative (Lfdobj)  --
+  #  A standard non-monotone smoothing using function smooth.basis
+  #      with a penalty on the Nth derivative (Lfdobj) in order to estimate a
+  #      smooth acceleration function
+  #  This gives a smoother estimate of the acceleration functions
+  #  Smoothing code follows https://github.com/cran/fda/blob/master/demo/growth.R with modification to simultaneous smoothing of both genders
+  
+  attach(growth)
+  true_group<- c(rep(TRUE,39),rep(FALSE,54)) # TRUE = boy, FALSE = girl
+  (ageRng <- range(age))
+  agefine <- seq(ageRng[1],ageRng[2],length=101)
+  knots  <- age
+  norder <- 6
+  nbasis <- length(knots) + norder - 2
+  hgtbasis <- create.bspline.basis(range(knots), nbasis, norder, knots)
+  
+  list_index = 1
+  for (Lfdobj in 1:4) {
+    lambda <- 1e-2
+    growfdPar <- fdPar(hgtbasis, Lfdobj, lambda)
+    
+    # Need 'hgtm', 'hgtf', e.g., from attach(growth)
+    hgtfd <- smooth.basis(growth$age, cbind(growth$hgtm,growth$hgtf), growfdPar)$fd
+    growthhat <- eval.fd(agefine, hgtfd) # growth 
+    velmhat <- eval.fd(agefine, hgtfd, 1) # velocity
+    accmhat <- eval.fd(agefine, hgtfd, 2) # acceleration
+    
+    growthhat_list = list(data = growthhat, reg_grid = agefine, true_group = true_group, name = sprintf("growthhat_der%d",Lfdobj))
+    list_of_lists[[list_index]] = growthhat_list
+    velmhat_list = list(data = velmhat, reg_grid = agefine, true_group = true_group, name = sprintf("velmhat_der%d",Lfdobj))
+    list_of_lists[[list_index+1]] = velmhat_list
+    accmhat_list = list(data = accmhat, reg_grid = agefine, true_group = true_group, name = sprintf("accmhat_der%d",Lfdobj))
[…]

### 2019-06-24 index.Rmd
added (3!) versions of tecator dataset

@@
 ###  Real data
-A few examples of the datasets used by functional classification papers
+Datasets used by functional classification papers
 
-* Wheat, rainfall and phoneme in "Achieving near-perfect classification for functional data"
+* Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
 * Berkeley growth curves in Chen et. al 2014, Biometrics.
-* Tecator and phoneme in technometrics paper.
+* Tecator and phoneme in Mahalanobis technometrics paper.
+
+Datasets used in functional manifold papers
+* Berkeley growth, yeast cell cycle gene expression in Chen and Muller
+* Tecator in Lin and Yao's contamination paper
+* Berkeley growth, gait cycle in Jean-Michel Loubes' robust isomap paper

### 2019-06-24 processRealData.R

@@ -1,4 +1,8 @@
-library("rlist")
+library(fda) # for growth dataset
+library("fda.usc") # for tecator dataset
+
+# returns a list of lists where every element describes a different dataset. all datasets are smoothed
+
 processRealData = function(){
   
   list_of_lists = list()
@@ -41,6 +45,55 @@ processRealData = function(){
     list_of_lists[[list_index+2]]= accmhat_list
     list_index = list_index + 3
   }
+  detach(growth)
+  
+  ### Tecator from fda package
+  # This is from Mahalanobis technometrics paper
+  # "The classification problem here is to separate meat samples with a high fat content(more than 20%) from samples with a low fat content (less than20%) based on absorbance. 
+  # Among the 215 samples, 77 have high fat content and 138 have low fat content."
+  
+  
+  # Warning: there are some weird boundary effects here!
+  # Mahalanobis paper regarding smoothing
+  # "second-order derivatives of the observed functions produces lower misclassification rates. 
+  # Therefore, the analysis of the original data and their second-order derivatives are carried out. 
+  # In both cases, the original discrete observations and their second differences are converted to 
+  # functional observations using a B- spline basis of order 6 with 20 and 40 basis functions, respectively."
+  
+  data(tecator)
+  absorp <- tecator$absorp.fdata
+  Fat20 <- ifelse(tecator$y$Fat < 20, 0, 1) # class membership
+  
+  # bspline parameters
+  norder <- 6
+  lambda <- 1e-2
+  
+  
+  for (nbasis in c(20,40)) {
+    basis <- create.bspline.basis(tecator[["absorp.fdata"]]$rangeval, nbasis, norder)
+    growfdPar <- fdPar(basis, 0, lambda)
+    
+    tecator_fd <- smooth.basis(tecator[["absorp.fdata"]]$argvals, t(absorp$data), growfdPar)$fd
+    grid = absorp$argvals
+    # grid_chopped = grid[-c(1,length(grid))] #TODO: is this cheating? we have to do it to avoid boundary effects
+    if (nbasis==20){
+      # Evaluate on original time grid
+      list_of_lists[[list_index]] = list(data = eval.fd(grid, tecator_fd), reg_grid = grid, true_group = Fat20, name = "tecator")
+      list_index = list_index+1
+    } else{
+      # Evaluate on original time grid
[…]

### 2019-06-24 main.R
weighted L2

@@ -4,6 +4,7 @@ library(fields)
 library(fda)
 library(matlabr)
 library(igraph)
+library(vows)
 
 # this may need to be set at the beginning of the session 
 # Susan's
@@ -26,14 +27,13 @@ source('pairwise_geo_estimation.R')
 source('sim_Euclidean_data.R')
 source('assess_goodness_estimation.R')
 source('robust_isomap.R')
+source('weight_L2.R')
 
 
 #### Analysis of Functional data
 
-FD_true =TRUE
-
 # set up parameters
-sce =2 # 1, 2 or 3 (for more info see sim_functional_data.R)
+sce =1 # 1, 2 or 3 (for more info see sim_functional_data.R)
 K = 30 # number of grid points (each curve is observed on K points on [a,b])
 samplesize = 100 # number of points on the manifold
 SNR = 0.1 # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
@@ -41,14 +41,14 @@ reg_sampling=TRUE # regular sampling of the point on the manifold or uniformly r
 plotTrue= TRUE 
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 nb_proj = 20 # number of random projection
-meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info
+meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = FALSE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=FALSE,"RP" = FALSE, "L2"= TRUE, "w_L2"=TRUE)# see pairwise_geo_estimation for more info
 
 
 # Generate data
 data<- sim_functional_data(sce,samplesize,K,SNR,reg_sampling,com_grid,plotTrue)
 
 # Estimation of geodesic distances with different methods
-Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,plotTrue,FD_true,nb_proj,data$grid,data$reg_grid,com_grid)
[…]

### 2019-06-24 pairwise_geo_estimation.R

@@ -1,4 +1,4 @@
-##### Estimation of pairwise geodesic distances
+##### Estimation of pairwise geodesic distances for functional data
 
 # We estimate the geodesic matrix obtained via
 # NN: Floyd's Algorithm on noiseless data (just to make sure that the theoretical geo. is calculated correctly)
@@ -11,6 +11,9 @@
 # OUR2: mds of smooth data + scms + robust isomap (oracle for h)
 # OUR3: mds of smooth data + scms h heuri + robust isomap (no oracle)
 # RP : random proj. of smooth data + scms + Floyd's Algorithm 
+# L2 : pairwise L2 distances on smooth data
+# w_L2 : weigthed L2 defined in Chen et al (Biometrics)
+
 
 # TODO: what if data are not observed on a common grid and NN, RD, RD_o are called? Are there warnings?
 ## Note that NN, RD, RD_o can only be used if the data are observed on a common grid
@@ -22,56 +25,74 @@
 # true_geo [optional, bypass with NA]: samplesize x samplesize matrix containing true pairwise geo disctance
 # s : dimension use for mds and random projections
 # plot_true : if TRUE plot of geo estimation for each method
-# FD_true : if TRUE functional data, otherwise euclidean data
-# grid : samplesize x K matrix containing the grid on which each data is observed (required only if FD_true =1)
+# nb_proj : number of projection used for RP method
+# grid : samplesize x K matrix containing the grid on which each data is observed 
+# reg_grid : vector of dim K containing a regular grid of K points on [a,b] 
 # common_grid_true : if 1 grid is common for every curve, if 0 different grid for every curve
+# Analytic_geo_available : indicate if we have access to the true pairwise geodesic distances
+# is_data_smoothed : TRUE data has already been smoothed, otherwise false
 
-# TODO: all of these "estim_geo_XXX" variable names should be changed, e.g. estim_geo_true_data should be renamed estim_geo_NN
+# TODO MH : add the possibility of a denser grid?
[…]

### 2019-06-24 processRealData.R

@@ -1,4 +1,3 @@
-library("rlist")
 processRealData = function(){
   
   list_of_lists = list()

### 2019-06-24 sim_functional_data.R

@@ -78,7 +78,7 @@
 
 
 #Example of call
-# data<- sim_functional_data(1,20,15,-3,3,0.5,1,1,1)
+# data<- sim_functional_data(1,100,15,0.5,1,1,1)
 
 library(fields)
 source('full_geo_from_adj_geo.R')

### 2019-06-24 index.Rmd
update cluster_results

@@
 ## Performance metrics
-We introduce a new metric for assessing the quality of a pairwise geodesic distance estimator.
+We describe three different metrics we use to assess the quality of a pairwise geodesic distance estimator.
 This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
@@
 Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance.
-Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance.
+Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?
 
@@
 * Tecator and phoneme in Mahalanobis technometrics paper.
+* yeast cell cycle gene expression (can't find this publicly) in Leng and Muller's "Classification using functional data analysis for temporal gene expression data"
 
 Datasets used in functional manifold papers
-* Berkeley growth, yeast cell cycle gene expression in Chen and Muller
+* Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in Chen and Muller
 * Tecator in Lin and Yao's contamination paper

### 2019-06-24 cluster_results.R

@@ -45,8 +45,9 @@ create_boxplot<- function(sce_details,mat_ind){
   }
 }
 
-combination_res_assess<- cbind(rep(1:6,3),rep(1:3,each=6)) # 18 combinations of methods and assesment measures
-combination_para <- expand.grid(1:3,c(100,250),c(0.1,0.5),c(0,1)) # 24 combinations of the parameters
+# TODO: a bit annoying that these are hardcoded, have to customise them according to compare_methods.R
+combination_res_assess<- cbind(rep(1:7,3),rep(1:3,each=7)) # 7*3 combinations of methods and assesment measures
+combination_para <- expand.grid(c(1,2,4),c(100,250),c(0.1,0.5),c(0,1)) # 24 combinations of the parameters
 
 apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess)
 

### 2019-06-24 cluster_results.R
making cluster_results more flexible

@@ -35,9 +35,16 @@ create_boxplot<- function(sce_details,mat_ind){
   par(mfrow=c(1,3))
   par(oma=c(2,2,2,2))
 
-  boxplot(results[,1:6],main="relative MSE",names=c("NN","RD","SS","OurS1","OurS2","OurS3"))
-  boxplot(results[,7:12],main="AUC of entrywise epsilon-isometry",names=c("NN","RD","SS","OurS1","OurS2","OurS3"))
-  boxplot(results[,13:18],main="Pearson correlation",names=c("NN","RD","SS","OurS1","OurS2","OurS3"))
+  meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info
+  
+  # TODO: this needs to be passed into the script from compare_methods.R?
+  method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3_S1","Our3_S2","Our3_S3")
+  nr_methods = length(method_names)
+  
+  # TODO: VERY annoying that these are hardcoded, have to customise them according to compare_methods.
+  boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
+  boxplot(results[,(1+nr_methods):(2*nr_methods)],main="AUC of entrywise epsilon-isometry",names=method_names)
+  boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation",names=method_names)
   mtext(paste("sce=",sce_details[1],"_samplesize=",sce_details[2],
               "_SNR=",sce_details[3],
               "_reg_sampling=",sce_details[4],sep=""),outer=TRUE,cex=1.5)
@@ -46,8 +53,8 @@ create_boxplot<- function(sce_details,mat_ind){
 }
 
 # TODO: a bit annoying that these are hardcoded, have to customise them according to compare_methods.R
-combination_res_assess<- cbind(rep(1:7,3),rep(1:3,each=7)) # 7*3 combinations of methods and assesment measures
-combination_para <- expand.grid(c(1,2,4),c(100,250),c(0.1,0.5),c(0,1)) # 24 combinations of the parameters
[…]

### 2019-06-24 cluster_results.R
fixed mistake

@@ -8,7 +8,7 @@ choose_el2<-function(ind_vec,list_in1){
   lapply(list_in1,choose_el,ind1=ind_vec[1],ind2=ind_vec[2])
 }
 
-create_boxplot<- function(sce_details,mat_ind){
+create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   
   data_sce = list.files(pattern=paste("_sce=",sce_details[1],
                                       "_samplesize=",sce_details[2],
@@ -35,12 +35,6 @@ create_boxplot<- function(sce_details,mat_ind){
   par(mfrow=c(1,3))
   par(oma=c(2,2,2,2))
 
-  meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info
-  
-  # TODO: this needs to be passed into the script from compare_methods.R?
-  method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3_S1","Our3_S2","Our3_S3")
-  nr_methods = length(method_names)
-  
   # TODO: VERY annoying that these are hardcoded, have to customise them according to compare_methods.
   boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
   boxplot(results[,(1+nr_methods):(2*nr_methods)],main="AUC of entrywise epsilon-isometry",names=method_names)
@@ -52,11 +46,17 @@ create_boxplot<- function(sce_details,mat_ind){
   }
 }
 
+
+# TODO: this needs to be passed into the script from compare_methods.R?
+method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3_S1","Our3_S2","Our3_S3")
+nr_methods = length(method_names)
+
+
 # TODO: a bit annoying that these are hardcoded, have to customise them according to compare_methods.R
 combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # 7*3 combinations of methods and assesment measures
 combination_para <- expand.grid(c(1,2,4),c(100,250),c(0.1,0.5),c(0,1)) # combinations of the parameters
 
-apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess)
[…]

### 2019-06-24 cluster_results.R
fixed a bit of hardcoding

@@ -23,7 +23,7 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   
   temp_res<-apply(mat_ind,1,choose_el2,list_in1=raw_results)
   temp_res2<- lapply(temp_res,unlist,use.names=FALSE)
-  results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=18)
+  results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=3*nr_methods)
   
   ### We plot one boxplot per method and per assesment measure
   pdf(paste("sce=",sce_details[1],
@@ -33,7 +33,7 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
             ".pdf",sep=""),width=15,height=5)
   
   par(mfrow=c(1,3))
-  par(oma=c(2,2,2,2))
+  par(oma=c(3,3,3,3))
 
   # TODO: VERY annoying that these are hardcoded, have to customise them according to compare_methods.
   boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
@@ -48,7 +48,7 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
 
 
 # TODO: this needs to be passed into the script from compare_methods.R?
-method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3_S1","Our3_S2","Our3_S3")
+method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3S1","Our3S2","Our3S3")
 nr_methods = length(method_names)
 
 

### 2019-06-25 test_analytic_geodesic_distance.Rmd
fixed reg_sampling to mean concentrated sampling

@@
 
+TODO: The reg_sampling parameter in sim_functional_data is somewhat of a misnomer.
+In either case when it's 0 or 1, it does NOT mean that the manifold can be sampled evenly.
+Recall swiss roll euclidean example.
+A simple sampling of the one-dimensional manifold parameter does not mean that entire swiss roll is visited, some corners are seldom sampled.
+This will affect downstream Floyd's algorithm.
+Indeed, Diaconis et.
+al in "Sampling from a Manifold" realize the drawbacks of naive sampling of Euclidean manifolds.
+
 ```{r}
 # Generate data
-sim <- sim_functional_data(sce=4,samplesize=1500,K=150,plot_true=1)
+sim <- sim_functional_data(sce=4,samplesize=200,K=100,plot_true=1,reg_sampling=0)
 ```

### 2019-06-25 compare_methods.R

@@ -29,6 +29,7 @@ K = 30 # number of grid points (each curve is observed on K points on [a,b])
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 plotTrue = FALSE
 FD_true = TRUE
+reg_sampling = TRUE
 
 # parameters under study
 # scenario
@@ -39,21 +40,19 @@ FD_true = TRUE
 sces = c(1,2,4)
 samplesizes = c(100,250)
 SNRs = c(0.1,0.5)
-reg_samplings = c(TRUE,FALSE)
 mcs = 1:100
 
-unravel=arrayInd(slurm_arrayid,c(length(sces),length(samplesizes), length(SNRs), length(reg_samplings), length(mcs)))
+unravel=arrayInd(slurm_arrayid,c(length(sces),length(samplesizes), length(SNRs), length(mcs)))
 
 # actual parameters for this run
 sce = sces[unravel[1,1]]
 samplesize = samplesizes[unravel[1,2]]
 SNR = SNRs[unravel[1,3]]
-reg_sampling = reg_samplings[unravel[1,4]]
-mc = mcs[unravel[1,5]]
+mc = mcs[unravel[1,4]]
 
 
 # Generate data
-data<- sim_functional_data(sce, samplesize, K, SNR, reg_sampling, com_grid, plotTrue)
+data<- sim_functional_data(sce, samplesize, K, SNR, com_grid, plotTrue)
 
 # Estimation of geodesic distances with different methods
 meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info

### 2019-06-25 sim_functional_data.R

@@ -2,7 +2,7 @@
 #' title: Functional manifold data examples
 #' ---
 
-#' \newcommand {\To}{\rightarrow}
+#'\newcommand {\To}{\rightarrow}
 #'\newcommand {\TO}{\Rightarrow}
 #'\newcommand {\R}{\mathbb{R}}
 #'\newcommand {\Prob}{\mathbb{P}}
@@ -64,7 +64,7 @@
 # samplesize: number of functional data 
 # K: number of time grid points
 # SNR (signal to noise ratio)
-# reg_sampling : 0 if manifold parameters are drawn randomly and 1 if deterministically
+# reg_sampling : 0 if manifold parameters are drawn uniformly and 1 if manifold parameters are drawn from a concentrated measure. SHOULD ALWAYS SET THIS TO 1
 # com_grid : if 1 each curve is observed on the same grid and if 0 the grid for each curve is randomly genarated from unif[a,b]
 # plot_true : if 1 plot the true data, the observed data and the true geodesic matrix
 
@@ -91,7 +91,9 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
       alpha<- runif(samplesize,-1,1)
       alpha=sort(alpha)
     } else if(reg_sampling==1){
-      alpha <- seq(-1,1,length.out=samplesize)
+      # alpha <- seq(-1,1,length.out=samplesize)
+      alpha<- rnorm(samplesize,0,0.2)
+      alpha=sort(alpha)      
     }
     
     mu_t <- function(t,al){
@@ -110,7 +112,9 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
       alpha<- runif(samplesize,-1,1)
       alpha=sort(alpha)
     } else if(reg_sampling==1){
-      alpha <- seq(-1,1,length.out=samplesize)
+      # alpha <- seq(-1,1,length.out=samplesize)
+      alpha<- rnorm(samplesize,0,0.1)
+      alpha=sort(alpha)
     }
     
     mu_t <- function(t,al){
@@ -128,22 +132,25 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
     nb_alpha<- 10
     nb_beta<- samplesize/10
[…]

### 2019-06-25 analyse_growth_curves.R
real data analysis

@@ -1,61 +1,87 @@
-#### Classification of growth curves dataset based on orginal data and velocity
-## We estimate the geodesic distance matrix with 4 methods
+#### Classification of real dataset preprocessed by processRealData.R
+# We compare results with SS, OUR3, L2 and weight L2 
 
 library(reticulate)
 library(fda)
 library(scatterplot3d)
-library(kmed)
-library(fields)
 library(igraph)
 library(ROCR)
 library(caret)
+library(vows)
+
+# this may need to be set at the beginning of the session 
+# Susan's
+# use_python('/usr/bin/python',required=TRUE)
+# use_python('/usr/local/bin/python3.7',required=TRUE)
+# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
+# Marie's
+# use_python('/anaconda3/envs/r-reticulate/bin/python',required=TRUE)
+# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+
+
 source('robust_isomap.R')
 source('pairwise_geo_estimation.R')
+source("weight_L2.R")
+source("processRealData.R")
 
-# input: lists of discrete_data, reg_grid, and true_group
-# discrete_datas_list
-# reg_grid_list
-# true_group_list
 
-# call processRealData which returns return(list("discrete_datas_list" = discrete_datas_list, "reg_grid_list" = reg_grid_list, "true_group_list"=true_group_list))
 
-data = processRealData()
-discrete_datas_list = data[[1]] # etc
+# call processRealData which each element is a list containing "discrete_data" , "reg_grid"  "true_group"
+dataset = processRealData()
 
-# We apply our method to obtain the pairwise geodesic distance (and floyd on smooth data)
-meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = TRUE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )
[…]

### 2019-06-25 processRealData.R

@@ -1,11 +1,9 @@
-<<<<<<< HEAD
-=======
+
 library(fda) # for growth dataset
 library("fda.usc") # for tecator dataset
 
 # returns a list of lists where every element describes a different dataset. all datasets are smoothed
 
->>>>>>> 53321e987c82d6c0191b763d8d190cf0421a41c5
 processRealData = function(){
   
   list_of_lists = list()

### 2019-06-25 test_analytic_geodesic_distance.Rmd
added more intersting warping scenario 5 and updated compare_methods

@@
 The function sim_functional_data.R allows specification of several simulatin scenarios. 
-The samplesize should be set very high so that Floyd's algorithm works well on the noiseless data.
+# The samplesize should be set very high so that Floyd's algorithm works well on the noiseless data.
 The number of time grid points K also should be high so that the vector data really represents the functional curve.
@@
 # Generate data
-sim <- sim_functional_data(sce=4,samplesize=200,K=100,plot_true=1,reg_sampling=0)
+sim <- sim_functional_data(sce=5,samplesize=200,K=30,plot_true=1,reg_sampling=1)
 ```
 
-The output actually contains more than just the simulated functional data.
-In particular, it contains the in-sample pairwise geodesic distance matrix (analytic_geo) via analytic derivation.
-The purpose of this notebook is to see if our analytic derivations match numerical results.
+The output contains the in-sample pairwise geodesic distance matrix (analytic_geo) via analytic derivation. The purpose of this notebook is to see if our analytic derivations match numerical results.
 
@@
 
-We would like to compare sim$analytic_geo to geodesic distance estimation via Floyd's algorithm for the noiseless (discretised) curve.
+Specifically, we would like to compare sim$analytic_geo to geodesic distance estimation via Floyd's algorithm for the noiseless (discretised) curve.
 This algorithm is guaranteed to do well if we sample the manifold densely enough?
-(That's why we set the samplesize to be quite high initially.)
+Setting the samplesize very high is not enough for this, a concentrated sampling of the manifold is actually needed.
+More on this in sim_functiona_data.R documentation.
+The number of grid points should not be too low since then the L2 distance between the discretized curves won't be well estimated.
+K=30 seems to suffice in general.
 

### 2019-06-25 analyse_growth_curves.R

@@ -61,7 +61,7 @@ for(ind_data in 1:length(dataset)){
   AUC_res <- matrix(unlist(temp,use.names=FALSE),ncol=length(geo_estim))
   boxplot(AUC_res,names=names(geo_estim),main=dataset[[ind_data]]$name,las=2,ylab="AUC")
   
-  AUC_full_res[[dataset[[ind_data]]$name]]<-  AUC_res
+  AUC_full_res[[dataset[[ind_data]]$name]]<-AUC_res
 }
 
 dev.off()

### 2019-06-25 cluster_results.R

@@ -48,7 +48,8 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
 
 
 # TODO: this needs to be passed into the script from compare_methods.R?
-method_names=c("NN","RD_o", "RD", "SS_o","SS","OurS1","OurS2","OurS3","Our3S1","Our3S2","Our3S3")
+method_names=c("SS","Our3S1","Our3S2","Our3S3","L2")
+
 nr_methods = length(method_names)
 
 

### 2019-06-25 compare_methods.R

@@ -25,40 +25,33 @@ slurm_arrayid = as.numeric(slurm_arrayid)
 print(slurm_arrayid)
 
 # parameters that will be fixed during the sim study
-K = 30 # number of grid points (each curve is observed on K points on [a,b])
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 plotTrue = FALSE
 FD_true = TRUE
 reg_sampling = TRUE
+samplesize = 100
+
+unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks), length(mcs)))
 
 # parameters under study
-# scenario
-# samplesize # number of points on the manifold
-# SNR # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
-# reg_sampling {True,False} # regular sampling of the point on the manifold or uniformly random
-
-sces = c(1,2,4)
-samplesizes = c(100,250)
+sces = c(5,2,4)
 SNRs = c(0.1,0.5)
+Ks = c(100,30)
 mcs = 1:100
-
-unravel=arrayInd(slurm_arrayid,c(length(sces),length(samplesizes), length(SNRs), length(mcs)))
 
 # actual parameters for this run
 sce = sces[unravel[1,1]]
-samplesize = samplesizes[unravel[1,2]]
-SNR = SNRs[unravel[1,3]]
+SNR = SNRs[unravel[1,2]]
+K = Ks[unravel[1,3]]
 mc = mcs[unravel[1,4]]
 
 
 # Generate data
 data<- sim_functional_data(sce, samplesize, K, SNR, com_grid, plotTrue)
-
+  
 # Estimation of geodesic distances with different methods
-meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE )# see pairwise_geo_estimation for more info
-Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,plotTrue,FD_true,nb_proj,data$grid,data$reg_grid,com_grid)
-# TODO: decide if we need to save Estim?
[…]

### 2019-06-25 compare_methods.sh

@@ -27,7 +27,7 @@
 # The maximum running time of the job in days-hours:mins:sec
 #SBATCH --time=1-0:00:00
 
-#SBATCH --array=1-2400         ### Array index | %50: number of simultaneously tasks
+#SBATCH --array=1-1200         ### Array index | %50: number of simultaneously tasks
 
 # Get Array ID
 i=${SLURM_ARRAY_TASK_ID}

### 2019-06-25 sim_functional_data.R

@@ -36,6 +36,7 @@
 #' Note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$, so we cannot employ the calculation technique of Scenario 1. 
 #'
 #' ##Scenario 3: 
+#' THIS DOESN'T WORK RIGHT NOW. A CONVEX COMBINATION OF GAUSSIAN DENSITIES IS NOT GAUSSIAN.
 #' Fix $\mu_1,\sigma_1^2, \mu_2, \sigma_2^2$. Let $f_1$ be the normal density with mean $\mu_1$ and variance $\sigma_1^2$. Define $f_2$ analogously.
 #' Consider the manifold $$ \M = \{ X_c: 0 \le c \le 1 \} $$ with the $L_2$ inner product as the metric tensor where
 #' $X_c: [a,b] \to \R$ is given by $X_c(t) = c f_1(t) + (1-c) f_2(t)$.
@@ -57,7 +58,19 @@
 #' We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$. That is, 
 #' $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$
 #' where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$.
-
+#' 
+#' ##Scenario 5: 
+#' This is based on Equations (17) and (18) of https://statistics.uni-bonn.de/fileadmin/Fachbereich_Wirtschaft/Einrichtungen/Statistik/WS0910/Topics_in_Econometrics_and_Statistics/Register2PCA.pdf
+#' but with $z_{i1}, z_{i2}$ set to $1$. (Equation 17 has a typo where the exponentials are missing negative signs).
+#' Let $X_\alpha(t) = \mu(h_\alpha(t))$ defined on $[-3,3]$ where
+#' $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$
+#' and 
+#' $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$
+#' and $h_\alpha(t) = t$ if $\alpha = 0$. 
+#' Consider the manifold
[…]

### 2019-06-25 compare_methods.R
stupid error with unravel

@@ -31,13 +31,14 @@ FD_true = TRUE
 reg_sampling = TRUE
 samplesize = 100
 
-unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks), length(mcs)))
 
 # parameters under study
 sces = c(5,2,4)
 SNRs = c(0.1,0.5)
 Ks = c(100,30)
 mcs = 1:100
+
+unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks), length(mcs)))
 
 # actual parameters for this run
 sce = sces[unravel[1,1]]

### 2019-06-26 cluster_results.R
stupid hardcoding

@@ -54,8 +54,8 @@ nr_methods = length(method_names)
 
 
 # TODO: a bit annoying that these are hardcoded, have to customise them according to compare_methods.R
-combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # 7*3 combinations of methods and assesment measures
-combination_para <- expand.grid(c(1,2,4),c(100,250),c(0.1,0.5),c(0,1)) # combinations of the parameters
+combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
+combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(100,30)) # combinations of the parameters
 
 apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)
 

### 2019-06-26 index.Rmd
fixed reg_sampling to mean concentrated sampling
more paper writing

@@
 ---
-title: "The hidden manifold distance for functional manifold data"
+title: "The hidden manifold distance for functional data"
 output:
@@
     df_print: paged
-    number_sections: true
+    number_sections: yes
+  pdf_document: default
 ---
@@
 3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
-4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n \times n $ matrix $\hat G$ as
+4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n$ by $n$ matrix 
 $$
@@
 
-## Simulation scenarios
-
 ## Performance metrics
@@
 
-## Results
+## Simulation scenarios
+Acknowledge that we are sampling on a very concentrated measure to make sure functional manifold is well sampled.
+Cite Diaconis et.
+al's "Sampling from a manifold".
+How to sample properly from a functional manifold could be interesting future work.
 
-# Study on "misspecification"
+* warping manifold (scenario 5 in sim_functional_data.R)
+* pdf manifold (scenario 2 in sim_functional_data.R)
+* square root pdf manifold (scenario 4 )
+### Results
+
+Submit compare_methods.sh as slurm job using sbatch.
+```{r, engine = 'bash', eval = FALSE}
+sbatch compare_methods.sh
+```
+
+Alternatively, we can run the local version
+```{r, eval=FALSE}
+R compare_methods_local.R
+```
+
+Results are visualized by invoking cluster_results.R
+```{r, eval=FALSE}
+R cluster_results.R
+```
+
+## The cost of employing proposed method when manifold is flat
 
@@
 
-# Downstream tasks
+# Distance-based functional classification
 
 In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task.
-There are many tasks we could consider here, but we will focus on distanced-based functional clustering and distance-based functional classification.
+There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification.
 It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.
-
-In both tasks below, we compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.
-
-## Distance-based functional clustering
-
-Describe the $k$-medoids clustering method.
-
-### Simulated data
-
-###  Real data
-
-## Distance-based functional classification
 
@@
 $$
+We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.
 
-### Simulated data
-
-###  Real data
 Datasets used by functional classification papers
@@
 * Berkeley growth, gait cycle in Jean-Michel Loubes' robust isomap paper
+
+<!-- ## Distance-based functional clustering -->
+
+<!-- Describe the $k$-medoids clustering method. -->
+
+<!-- ### Simulated data -->
+
+<!-- ###  Real data -->

### 2019-06-26 cluster_results.R

@@ -10,11 +10,11 @@ choose_el2<-function(ind_vec,list_in1){
 
 create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   
+  # TOOD: how can we avoid hardcoding this string?
   data_sce = list.files(pattern=paste("_sce=",sce_details[1],
-                                      "_samplesize=",sce_details[2],
-                                      "_SNR=",sce_details[3],
-                                      "_reg_sampling=",sce_details[4]
-                                      ,"_mc=*",sep=""))
+                                      "SNR=",sce_details[2],
+                                      "K=",sce_details[3],
+                                      "_mc=*",sep=""))
   if (length(data_sce) == 0){
     return()
   }
@@ -26,10 +26,10 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=3*nr_methods)
   
   ### We plot one boxplot per method and per assesment measure
+  # TOOD: how can we avoid hardcoding this string?
   pdf(paste("sce=",sce_details[1],
-            "_samplesize=",sce_details[2],
-            "_SNR=",sce_details[3],
-            "_reg_sampling=",sce_details[4],
+            "SNR=",sce_details[2],
+            "_K=",sce_details[3],
             ".pdf",sep=""),width=15,height=5)
   
   par(mfrow=c(1,3))
@@ -39,23 +39,22 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
   boxplot(results[,(1+nr_methods):(2*nr_methods)],main="AUC of entrywise epsilon-isometry",names=method_names)
   boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation",names=method_names)
-  mtext(paste("sce=",sce_details[1],"_samplesize=",sce_details[2],
-              "_SNR=",sce_details[3],
-              "_reg_sampling=",sce_details[4],sep=""),outer=TRUE,cex=1.5)
+  mtext(paste("sce=",sce_details[1],
+              "SNR=",sce_details[2],
+              "K=",sce_details[3],
+              sep=""),outer=TRUE,cex=1.5)
   dev.off()
   }
 }
 
 
-# TODO: this needs to be passed into the script from compare_methods.R?
[…]

### 2019-06-26 compare_methods.R

@@ -1,5 +1,5 @@
 # Title     : compare_methods.R
-# Objective : This is modeled after Marie's main.R but designed for submission to Spartan HPC. Performs sweep of certain parameters.
+# Objective : compare multiple estimators of geodesic distance on simulated data. this is called by compare_methods.sh!
 # Created by: suswei
 # Created on: 3/12/18
 
@@ -27,17 +27,14 @@ print(slurm_arrayid)
 # parameters that will be fixed during the sim study
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 plotTrue = FALSE
-FD_true = TRUE
 reg_sampling = TRUE
 samplesize = 100
-
 
 # parameters under study
 sces = c(5,2,4)
 SNRs = c(0.1,0.5)
 Ks = c(100,30)
 mcs = 1:100
-
 unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks), length(mcs)))
 
 # actual parameters for this run
@@ -51,14 +48,31 @@ mc = mcs[unravel[1,4]]
 data<- sim_functional_data(sce, samplesize, K, SNR, com_grid, plotTrue)
   
 # Estimation of geodesic distances with different methods
-meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = TRUE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE,"L2" = TRUE, "w_L2" = FALSE )# see pairwise_geo_estimation for more info
-Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,plotTrue,nb_proj,data$grid,data$reg_grid,com_grid)
+meth <- list("NN" = FALSE,
+             "RD_o" = FALSE,
+             "RD" = FALSE,
+             "SS_o" = FALSE,
+             "SS" = TRUE,
+             "pI" = FALSE,
+             "OUR" = FALSE,
+             "OUR2" = FALSE,
+             "OUR3"=TRUE,
+             "RP" = FALSE,
+             "L2" = TRUE, 
+             "w_L2" = FALSE )# see pairwise_geo_estimation for more info
+  
+Estim<- pairwise_geo_estimation(method=meth,
+                                true_data = data$noiseless_data,
[…]

### 2019-06-26 pairwise_geo_estimation.R

@@ -62,7 +62,17 @@ pyIso = import_from_path("getIsomapGdist",path='.') # TODO: consider replacing p
 py_min_neigh = import_from_path("get_min_num_neighbors",path='.')
 scms = import_from_path("scms",path='.') # TODO: consider replacing python scms with R version, see https://sites.google.com/site/yenchicr/algorithm
 
-pairwise_geo_estimation <- function(method,true_data,discrete_data,true_geo,plot_true,nb_proj,grid,reg_grid,common_grid_true,Analytic_geo_available=TRUE,is_data_smoothed = FALSE){
+pairwise_geo_estimation <- function(method,
+                                    true_data,
+                                    discrete_data,
+                                    true_geo,
+                                    plot_true,
+                                    nb_proj=NULL,
+                                    grid,
+                                    reg_grid,
+                                    common_grid_true,
+                                    Analytic_geo_available=TRUE,
+                                    is_data_smoothed = FALSE){
   
   samplesize = nrow(discrete_data)
   K = ncol(discrete_data)

### 2019-06-26 main.R
add option K_dense

@@ -33,22 +33,23 @@ source('weight_L2.R')
 #### Analysis of Functional data
 
 # set up parameters
-sce =1 # 1, 2 or 3 (for more info see sim_functional_data.R)
+sce =2 # 1, 2 or 3 (for more info see sim_functional_data.R)
 K = 30 # number of grid points (each curve is observed on K points on [a,b])
+K_dense = 100 # number of grid points on which smoothed curves are observed
 samplesize = 100 # number of points on the manifold
 SNR = 0.1 # signal to noise ratio (in Chen and Muller is 0.1 or 0.5)
 reg_sampling=TRUE # regular sampling of the point on the manifold or uniformly random
 plotTrue= TRUE 
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 nb_proj = 20 # number of random projection
-meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = FALSE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=FALSE,"RP" = FALSE, "L2"= TRUE, "w_L2"=TRUE)# see pairwise_geo_estimation for more info
+meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = FALSE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE, "L2"= TRUE, "w_L2"=TRUE)# see pairwise_geo_estimation for more info
 
 
 # Generate data
 data<- sim_functional_data(sce,samplesize,K,SNR,reg_sampling,com_grid,plotTrue)
 
 # Estimation of geodesic distances with different methods
-Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,plotTrue,nb_proj,data$grid,data$reg_grid,com_grid,FALSE,FALSE)
+Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,plotTrue,nb_proj,data$grid,data$reg_grid,K_dense,com_grid,FALSE,FALSE)
 
 # Comparision of the different methods
 nom_methode <- rep(names(Estim),rep(3,length(Estim)))

### 2019-06-26 pairwise_geo_estimation.R

@@ -28,12 +28,11 @@
 # nb_proj : number of projection used for RP method
 # grid : samplesize x K matrix containing the grid on which each data is observed 
 # reg_grid : vector of dim K containing a regular grid of K points on [a,b] 
+# K_dense : size of the regular grid on which smoothed curve are evaluated 
 # common_grid_true : if 1 grid is common for every curve, if 0 different grid for every curve
 # Analytic_geo_available : indicate if we have access to the true pairwise geodesic distances
 # is_data_smoothed : TRUE data has already been smoothed, otherwise false
 
-# TODO MH : add the possibility of a denser grid?
-# K_dense : if 0 the smooth curves are observe on reg_grid, otherwise they are observed on a regular grid of K_dense points
 
 
 ## Output
@@ -55,7 +54,7 @@
 # Example of call for functional data
 # data<- sim_functional_data(sce=2,samplesize=100)
 # meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE, "L2" = TRUE, "w_L2" = TRUE )
-# Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,TRUE,TRUE,20,data$grid,data$reg_grid,1,FALSE,FALSE)
+# Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,TRUE,TRUE,20,data$grid,data$reg_grid,100,1,FALSE,FALSE)
 
 
 pyIso = import_from_path("getIsomapGdist",path='.') # TODO: consider replacing python Isomap with R implementation of Floyd's algorithm? https://en.wikipedia.org/wiki/Floyd%E2%80%93Warshall_algorithm
@@ -70,6 +69,7 @@ pairwise_geo_estimation <- function(method,
                                     nb_proj=NULL,
                                     grid,
                                     reg_grid,
+                                    K_dense,
[…]

### 2019-06-26 index.Rmd
prepping compare_methods

@@
 output:
+  pdf_document: 
+    number_sections: yes
   html_document:
@@
     number_sections: yes
-  pdf_document: default
 ---
@@
 
-# Introduction
+# Introduction [Susan]
 
@@
 
-## Setting
+## Setting [Susan]
 Describe preprocessing of discretely observed functional data: In the classical approach to analysing such data, we first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE), next perform FPCA to reduce the dimension of the data, and then proceed by employing multivariate techniques.
@@
 
-# Preliminaries
+# Preliminaries [Susan]
 
@@
 
-## Basic differential geometry
+## Basic differential geometry [Susan]
 This is take from a paper on [metric learning](http://proceedings.mlr.press/v32/linb14.pdf).
@@
 
-## IsoGeo
+## IsoGeo [Marie]
 
@@
 
-# Proposed method for estimating geodesic distances
+# Proposed method for estimating geodesic distances [Marie]
 
@@
 
-## Selection of tuning parameters
+## Selection of tuning parameters [Marie]
+Describe heuristics.
 
@@
 
-## Alternative estimators of geodesic distance
+## Alternative estimators of geodesic distance [Marie]
 
@@
 
-## Performance metrics
+## Performance metrics [Susan]
 We describe three different metrics we use to assess the quality of a pairwise geodesic distance estimator.
@@
 
-## Simulation scenarios
+## Simulation scenarios [Susan]
 Acknowledge that we are sampling on a very concentrated measure to make sure functional manifold is well sampled.
@@
 
-## The cost of employing proposed method when manifold is flat
+## The cost of employing proposed method when manifold is flat [Susan/Marie?]
 
@@
 
-# Distance-based functional classification
+# Distance-based functional classification [Marie]
 
@@
 $$
-We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished.
+We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.
+
+The bandwidth in the classifier should be tuned individually for each method.
+Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.
 
@@
 Datasets used in functional manifold papers
+
 * Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in Chen and Muller

### 2019-06-26 cluster_results.R

@@ -13,7 +13,8 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   # TOOD: how can we avoid hardcoding this string?
   data_sce = list.files(pattern=paste("_sce=",sce_details[1],
                                       "SNR=",sce_details[2],
-                                      "K=",sce_details[3],
+                                      "K_dense=",sce_details[3],
+                                      "regsamp=",sce_details[4],
                                       "_mc=*",sep=""))
   if (length(data_sce) == 0){
     return()
@@ -29,7 +30,8 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   # TOOD: how can we avoid hardcoding this string?
   pdf(paste("sce=",sce_details[1],
             "SNR=",sce_details[2],
-            "_K=",sce_details[3],
+            "K_dense=",sce_details[3],
+            "regsamp=",sce_details[4],
             ".pdf",sep=""),width=15,height=5)
   
   par(mfrow=c(1,3))
@@ -54,7 +56,7 @@ nr_methods = length(method_names)
 
 combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
 # TODO: a bit annoying that combination_para is hardcoded, have to customise them according to compare_methods.R
-combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(100,30)) # combinations of the parameters, copied from compare_methods.R
+combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(TRUE,FALSE),c(100,30)) # combinations of the parameters, copied from compare_methods.R
 
 apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)
 

### 2019-06-26 compare_methods.R

@@ -19,60 +19,83 @@ source('sim_Euclidean_data.R')
 source('assess_goodness_estimation.R')
 source('robust_isomap.R')
 
-# get sweeping id
-slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
-slurm_arrayid = as.numeric(slurm_arrayid)
-print(slurm_arrayid)
+
 
 # parameters that will be fixed during the sim study
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
 plotTrue = FALSE
-reg_sampling = TRUE
 samplesize = 100
 
-# parameters under study
+## STUDY THE EFFECT OF THE FOLLOWING PARAMETERS 
+# parameters in sim_functional_data
 sces = c(5,2,4)
 SNRs = c(0.1,0.5)
-Ks = c(100,30)
+reg_samplings = c(TRUE,FALSE)
+Ks_obs = c(100,30)
+# parameters in pairwise_geo_estimation
+Ks_smooth = c(100,30)
+
 mcs = 1:100
-unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks), length(mcs)))
+total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(Ks_smooth)*length(mcs)
 
-# actual parameters for this run
-sce = sces[unravel[1,1]]
-SNR = SNRs[unravel[1,2]]
-K = Ks[unravel[1,3]]
-mc = mcs[unravel[1,4]]
+# get sweeping id, only use this block if submitting to cluster
+# slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
+# slurm_arrayid = as.numeric(slurm_arrayid)
+# print(slurm_arrayid)
 
-
-# Generate data
-data<- sim_functional_data(sce, samplesize, K, SNR, com_grid, plotTrue)
+for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
   
-# Estimation of geodesic distances with different methods
-meth <- list("NN" = FALSE,
-             "RD_o" = FALSE,
-             "RD" = FALSE,
-             "SS_o" = FALSE,
-             "SS" = TRUE,
-             "pI" = FALSE,
-             "OUR" = FALSE,
-             "OUR2" = FALSE,
-             "OUR3"=TRUE,
-             "RP" = FALSE,
-             "L2" = TRUE, 
[…]

### 2019-06-26 sim_functional_data.R

@@ -93,52 +93,61 @@
 # data<- sim_functional_data(1,100,15,0.5,1,1,1)
 
 library(fields)
+library(truncnorm)
 source('full_geo_from_adj_geo.R')
 
 sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_grid=1,plot_true=1){
   
   if(sce == 1){
+    
     a<- -4
     b<- 4
+    
+    alphamin = -1
+    alphamax = 1
+    
     if(reg_sampling==0){
-      alpha<- runif(samplesize,-1,1)
-      alpha=sort(alpha)
+      alpha <- runif(samplesize,alphamin,alphamax)
+      alpha = sort(alpha)
     } else if(reg_sampling==1){
-      # alpha <- seq(-1,1,length.out=samplesize)
-      alpha<- rnorm(samplesize,0,0.2)
-      alpha=sort(alpha)      
+      alpha <- rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)
+      alpha = sort(alpha)      
     }
     
     mu_t <- function(t,al){
       h_t <- t-al
     }
     
-    # TODO: Add mathematical derviation for adja_geo
     adja_geo <- (alpha[-1]- alpha[-samplesize])*sqrt(b-a)
     ### Calculate the analytic geodesic matrix
     analytic_geo <- full_geo(adja_geo,samplesize)
     
   } else if(sce ==2){
+    
     a<- -4
     b<- 4
+    
+    alphamin = -1
+    alphamax = 1
+    
     if(reg_sampling==0){
-      alpha<- runif(samplesize,-1,1)
-      alpha=sort(alpha)
+      alpha <- runif(samplesize,alphamin,alphamax)
+      alpha = sort(alpha)
     } else if(reg_sampling==1){
-      # alpha <- seq(-1,1,length.out=samplesize)
-      alpha<- rnorm(samplesize,0,0.1)
-      alpha=sort(alpha)
+      alpha <- rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1) # doesn't follow Manifold 2 of Chen and Muller. It's called beta in the paper and beta ~ N(0,1). this affects our test_analytic_geodesic_distance resutls
+      alpha = sort(alpha)
     }
     
     mu_t <- function(t,al){
[…]

### 2019-06-26 test_analytic_geodesic_distance.Rmd
prepping compare_methods
fix problems in processRealData

@@
 
-use_python('/usr/local/bin/python3.7',required=TRUE)
+# use_python('/usr/local/bin/python3.7',required=TRUE)
 # use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
@@
 ```
-
-Check with py_config()
 
@@
 # Generate data
-sim <- sim_functional_data(sce=5,samplesize=200,K=30,plot_true=1,reg_sampling=1)
+sim <- sim_functional_data(sce=2,samplesize=100,K=30,plot_true=1,reg_sampling=1)
 ```
@@
 The number of grid points should not be too low since then the L2 distance between the discretized curves won't be well estimated.
-K=30 seems to suffice in general.
 

### 2019-06-26 analyse_real_dataset.R
change name script

@@ -35,7 +35,7 @@ meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = TRUE
 AUC_full_res<- list()
 
 
-for(ind_data in 1:length(dataset)){
+for(ind_data in 15:length(dataset)){
    
   # estimation of geo distance with methods TRUE in meth
   geo_estim = pairwise_geo_estimation(method=meth,
@@ -44,8 +44,9 @@ for(ind_data in 1:length(dataset)){
                                       true_geo=NA,
                                       plot_true=FALSE,
                                       nb_proj=NA,
-                                      grid=matrix(rep(dataset[[ind_data]]$reg_grid,length(dataset[[ind_data]]$reg_grid)),nrow=dataset[[ind_data]]$reg_grid,byrow = TRUE),
+                                      grid=matrix(rep(dataset[[ind_data]]$reg_grid,length(dataset[[ind_data]]$reg_grid)),nrow=length(dataset[[ind_data]]$reg_grid),byrow = TRUE),
                                       reg_grid=dataset[[ind_data]]$reg_grid,
+                                      K_dense =100,
                                       common_grid_true=TRUE,
                                       Analytic_geo_available=FALSE,
                                       is_data_smoothed=TRUE)
@@ -57,14 +58,14 @@ for(ind_data in 1:length(dataset)){
   # train.rows[,i] contains the training indices i-th split
   train.rows = createDataPartition(dataset[[ind_data]]$true_group, times = nr_train_test_splits, p = 0.53, list = FALSE)
   
-  temp<- lapply(geo_estim,mean_auc,train.rows=train.rows,true_group=dataset[[ind_data]]$true_group,h=10)
+  temp<- lapply(geo_estim,mean_auc,train.rows=train.rows,true_group=dataset[[ind_data]]$true_group,h=2)
   AUC_res <- matrix(unlist(temp,use.names=FALSE),ncol=length(geo_estim))
   boxplot(AUC_res,names=names(geo_estim),main=dataset[[ind_data]]$name,las=2,ylab="AUC")
   
   AUC_full_res[[dataset[[ind_data]]$name]]<-AUC_res
 }
 
-dev.off()
+
 
 
 pdf("classification_results_grotwh_curve1.pdf",width=12,height=5)

### 2019-06-26 index.Rmd
fix problems in processRealData
added bib references

@@
   pdf_document:
+    keep_tex: yes
     number_sections: yes
+  word_document: default
   html_document:
@@
     number_sections: yes
+bibliography: references.bib
 ---
@@
 # Abstract
+Functional data analysis is the statistical analysis of smooth infinite-dimensional curves. 
+The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that, since observations are smooth over their domain, the dimensionality of the curves is only artifically high.
+Less studied in functional data analysis is the idea that the curves may actually live in a submanifold with low intrinsic dimension.
+Interesting classes of functional manifold data include the class of probability density functions and the class of warped curves of a common template function. 
+Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration. 
+In this work we set ourselves the task of calculating the pairwise geodesic distance between functional observations that, due to noise, live near a non-linear manifold rather than exactly on the manifold.
+<!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
+We will show that good estimation of the geodesic distance cannot be accomplished by naive methods such as smoothing followed by a shortest-path algorithm. 
+Instead, we propose a dedicated technique for pairwise geodesic distance estimation by first projecting the functional data to the underlying hidden manifold and then perofrming operations on this hidden manifold.
+We conclude the work by showing the implications for many downstream tasks in functional data analysis, in particular distanced-based functional classification.
 
-Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example.
-Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration and treatment.
-In this work we set ourselves the task of calculating the geodesic distance between two functions living near a non-linear manifold due to noise.
-This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
-Good estimation of the geodesic distance has implications for many downstream tasks in functional data analysis, including functional clustering, and nonparametric functional regression.
+<!-- ##############  -->
 
@@
 
-The classical setup in FDA assumes we observe a sample of $n$ independant realizations $X_1,\ldots,X_n$ of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-However, it might very well happen that the function $X$ belongs to $\M \subset L^2([a,b],\R)$ a low-dimensional nonlinear manifold (gives examples and references).
-If it is indeed the case, it may no longer be appropriate to use $L^2$ distance between two curves as a measure of distance.
-<!-- Scenario 2 in [sim_functional_data](sim_functional_data.html) is one such example.  -->
-One may want to consider using some geodesic distance which takes into account the intrinsic structure of $\M$.
+Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+We additionally assume that the function $X$ belongs to a low-dimensional nonlinear manifold $\M \subset L^2([a,b],\R)$, in which case a geodesic distance taking into account the intrinsic structure of $\M$ rather than the $L^2$ distance should be used as a measure of distance between two functions in $\M$.
+As many statistical methods depend on some measure of distance, downstream benefits of using the geodesic distance can be expected.
 
-Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.
+???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.???
 
-There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
+<!-- (gives examples and references for functional manifolds). -->
+<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->
+
+<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
 However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio.
 (is it true?
-ref?) This is a luxury that FDA cannot always afford because functional data are usually observed with noise.
-The noise can be such that the observed functional data no longer live on a manifold.
-Though powerful techniques exist for smoothing/interpolating discretely observed noisy functional data, the recovered functional versions of the data, $\tilde X_1,\ldots,\tilde X_n$, may not lie even close to the functional manifold $\M$.
-This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs.
+ref?)  -->
 
-Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
+Given perfect functional observations, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance can be accomplished straightforwardly using algorithms such as Floyd's.
+However, functional data are observed with noise in such a way that the observed functional data no longer live on a manifold.
+Though powerful techniques exist for smoothing discretely observed noisy functional data, the smoothing is designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
+<!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
+
+<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
 Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.
+-->
 
-With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$ when we have access only to discretely-observed noisy functional observations that possibly live off the true manifold.
-Pairwise distances are important for many downstream tasks, e.g.
+With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$, to be defined  when we have access only to discretely-observed noisy functional observations that possibly live off the true manifold.
+Briefly outline the methodology: MDS plus density ridge are used in a cool way to estimate (consistently?, let's hope..) pairwise geodesic distances.
+
+<!-- Pairwise distances are important for many downstream tasks, e.g.
 classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression).
@@
 Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.
+-->
 
-???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.
-like the Berkeley growth curves, MDS based on pairwise geodesic versus MDS based on pairwise $L_2$.
-The former separates the genders better?
-Something like this, but hopefully a different dataset.???
+## Related Work [Susan]
+[@ChenMuller2012]
 
-Briefly outline the methodology: MDS plus density ridge are used in a cool way to estimate (consistently?, let's hope..) pairwise geodesic distances.
+[@LinYao2017]
+
+[@Dimeglio2014]
 
 ## Setting [Susan]
-Describe preprocessing of discretely observed functional data: In the classical approach to analysing such data, we first obtain a functional version $\tilde X_1,\ldots,\tilde X_n$ of the raw data (by either smoothing or using PACE), next perform FPCA to reduce the dimension of the data, and then proceed by employing multivariate techniques.
-
-What type of noise structures are allowed/considered? What type of time sampling frequency is assumed? Do not need time-aligned curves!
-
-# Preliminaries [Susan]
-
-We first rigorously define the concepts of geodesic and geodesic distance. We will also describe an important helper algorithm which we call IsoGeo.
-
-## Basic differential geometry [Susan]
-This is take from a paper on [metric learning](http://proceedings.mlr.press/v32/linb14.pdf).
-Let $(\M,g)$ be a $d$-dimensional Riemannian manifold, where $g$ is a Riemannian metric tensor on $\M$.
-This metric tensor can be used to assign a metric on the manifold.
-For each point $p$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_p$ on the tangent space $T_p \M$.
-The norm of a tangent vector $v \in T_p \M$ is defined as $$||v|| = \sqrt{g_p(v,v)}$$
-If $\gamma: [a,b] \subset \R \to \M$ is a smooth curve, its length is then defined as
-$$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt $$
-The geodesic distance between two points $p,q$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.$$
-
-## IsoGeo [Marie]
-
-Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
-* The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-* In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, and can be calculated easily with Floyd-Warshall or Dijkstra’s algorithms.
-* The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
-
-In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
-Note that Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, p-isomap is one of these).
-
-# Proposed method for estimating geodesic distances [Marie]
-
 Let $X_1,\ldots,X_n$ be independent realizations of a random function $X\in\M \subset L^2([a,b],\R)$, where $\M$ is an unknown manifold.
@@
 
+Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data by some smoothing method.
+Our proposed methodology is unspecific to the smoothing method employed.
+For simplicity however, let's say we employ spline smoothing to recover the functional versions of the raw data, i.e.\
+$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
+where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
+
+<!-- ##############  -->
+
+## Relevant differential geometry concepts [Susan]
+
+We first rigorously define the concepts of geodesic and geodesic distance.
+Following [@Lin2014], let $(\M,g)$ be a $d$-dimensional Riemannian manifold, where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold.
+Specifically, for each point $p$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_p$ on the tangent space $T_p \M$.
+The norm of a tangent vector $v \in T_p \M$ is defined as $$||v|| = \sqrt{g_p(v,v)}.$$
+If $\gamma: [a,b] \subset \R \to \M$ is a smooth curve, its length is then defined as
+$$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt .$$
+The geodesic distance between two points $p,q$ on the manifold $\M$, based on this metric tensor $g$, is defined as
+$$ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.$$
+<!-- ##############  -->
+
+# Proposed method for estimating geodesic distances [Marie]
+
+We first describe an important helper algorithm which we call IsoGeo. It is a subroutine of the classic Isomap procedure.
+Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+
+\begin{enumerate}
+\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
+The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
+\item In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
+are estimated based on $G$.
+The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Warshall1962]
+<!-- or Dijkstra’s  which one do we actually use in our implementation, the FW or D? -->
+\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
+\end{enumerate}
+
+In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
+Note that Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these).
+
 Our method is described by the following steps:
 
-1. Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing:
-$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
-where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
+<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
+<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
+<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
 2. Obtain a $s$-dimensional representation $\tilde X^s_1,\ldots,\tilde X^s_n$ of the functions $\tilde X_1,\ldots,\tilde X_n$ that "preserves" the pairwise $L^2$ distances by using MDS.
-3. Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with a mean shift algorithm.
+3.
+Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with the subspace constrained mean-shift algorithm [@Ozertem2011].
 4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n$ by $n$ matrix 
@@
 
-Since the ridge estimation obtained from noisy measurements of a manifold should approximate well the manifold (ref.
-Genevose and al.
-2014), we expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
+Since the ridge estimation obtained from noisy measurements of a manifold should approximate well the manifold [@Genovese2014], we expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
 Ridge estimation suffers from the curse of dimensionality, this is why we first reduce the dimension of our data with MDS and then apply the shift-mean algorithm to estimate the ridge.
@@
 ## Selection of tuning parameters [Marie]
-Describe heuristics.
+Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of [@ChenHo2015].
+
+<!-- ##############  -->
 
@@
 Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
-* **(Penalized-Isomap) pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+* **(P-ISOMAP) pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller).
@@
 * **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
+
+We ran Chen and Muller's P-ISOMAP separately in Matlab with their automatic penalty selection and discovered that ???.
 
@@
 * square root pdf manifold (scenario 4 )
+
 ### Results
@@
 
-Alternatively, we can run the local version
+Alternatively, we can run the R script compare_methods.R locally by removing the slurm_arrayid chunk
 ```{r, eval=FALSE}
-R compare_methods_local.R
+R compare_methods.R
 ```
 
-Results are visualized by invoking cluster_results.R
+Simulation results are then visualized by invoking cluster_results.R. There's some unfortunate hardcoding still in cluster_results.R
 ```{r, eval=FALSE}
@@
 Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?
+
+<!-- ##############  -->
 
@@
 Datasets used by functional classification papers
-
-* Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
-* Berkeley growth curves in Chen et. al 2014, Biometrics.
-* Tecator and phoneme in Mahalanobis technometrics paper.
-* yeast cell cycle gene expression (can't find this publicly) in Leng and Muller's "Classification using functional data analysis for temporal gene expression data"
+\begin{itemize}
+\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
+\item Berkeley growth curves in [@ChenReiss2014].
+\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
+\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
+\end{itemize}
 
 Datasets used in functional manifold papers
-
-* Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in Chen and Muller
-* Tecator in Lin and Yao's contamination paper
-* Berkeley growth, gait cycle in Jean-Michel Loubes' robust isomap paper
+\begin{itemize}
+\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
+\item Tecator in [@LinYao2017] contamination paper
+\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
+\end{itemize}
 
@@
 <!-- ###  Real data -->
+
+# References

### 2019-06-26 cluster_results.R
implemented try catch error for integrate maximum subdivisions problem

@@ -9,56 +9,45 @@ choose_el2<-function(ind_vec,list_in1){
 }
 
 create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
-  
-  # TOOD: how can we avoid hardcoding this string?
-  data_sce = list.files(pattern=paste("_sce=",sce_details[1],
-                                      "SNR=",sce_details[2],
-                                      "K_dense=",sce_details[3],
-                                      "regsamp=",sce_details[4],
-                                      "_mc=*",sep=""))
+
+  # TODO: fix hardcoding
+  base_pattern = do.call(sprintf, c(list("sce=%d_SNR=%.01f_regsamp=%d_Kobs=%d_Ksmooth=%d"), sce_details))
+
+  data_sce = list.files(pattern=paste("_", base_pattern,"_mc=*",sep=""))
   if (length(data_sce) == 0){
     return()
   }
   else{
-  raw_results<-lapply(data_sce,readRDS)
-  
-  temp_res<-apply(mat_ind,1,choose_el2,list_in1=raw_results)
-  temp_res2<- lapply(temp_res,unlist,use.names=FALSE)
-  results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=3*nr_methods)
-  
-  ### We plot one boxplot per method and per assesment measure
-  # TOOD: how can we avoid hardcoding this string?
-  pdf(paste("sce=",sce_details[1],
-            "SNR=",sce_details[2],
-            "K_dense=",sce_details[3],
-            "regsamp=",sce_details[4],
-            ".pdf",sep=""),width=15,height=5)
-  
-  par(mfrow=c(1,3))
-  par(oma=c(3,3,3,3))
-
-  # TODO: VERY annoying that these are hardcoded, have to customise them according to compare_methods.
-  boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
-  boxplot(results[,(1+nr_methods):(2*nr_methods)],main="AUC of entrywise epsilon-isometry",names=method_names)
-  boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation",names=method_names)
-  mtext(paste("sce=",sce_details[1],
-              "SNR=",sce_details[2],
-              "K=",sce_details[3],
-              sep=""),outer=TRUE,cex=1.5)
-  dev.off()
+    raw_results<-lapply(data_sce,readRDS)
+    
[…]

### 2019-06-26 compare_methods.R

@@ -27,6 +27,7 @@ plotTrue = FALSE
 samplesize = 100
 
 ## STUDY THE EFFECT OF THE FOLLOWING PARAMETERS 
+mcs = 1:100
 # parameters in sim_functional_data
 sces = c(5,2,4)
 SNRs = c(0.1,0.5)
@@ -35,15 +36,14 @@ Ks_obs = c(100,30)
 # parameters in pairwise_geo_estimation
 Ks_smooth = c(100,30)
 
-mcs = 1:100
 total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(Ks_smooth)*length(mcs)
 
 # get sweeping id, only use this block if submitting to cluster
-# slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
-# slurm_arrayid = as.numeric(slurm_arrayid)
-# print(slurm_arrayid)
+slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
+slurm_arrayid = as.numeric(slurm_arrayid)
+print(slurm_arrayid)
 
-for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
+#for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
   
   # Hardcode
   unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(Ks_smooth),length(mcs)))
@@ -66,7 +66,7 @@ for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
                              SNR = SNR, 
                              reg_sampling = reg_sampling,
                              K = K_obs)
-  )
+  
   
   # Estimation of geodesic distances with different methods
   meth <- list("NN" = FALSE,
@@ -96,6 +96,7 @@ for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
   Rel_errs = lapply(Estim, assess_goodness_estimation, true_geo = data$analytic_geo)
   print(Rel_errs)
   # hardcode 
[…]

### 2019-06-26 sim_functional_data.R

@@ -223,7 +223,7 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
           sqrt(dbeta(x,alpha_beta[comb1,1],alpha_beta[comb1,2])) * sqrt(dbeta(x,alpha_beta[comb2,1],alpha_beta[comb2,2]))
         }
         
-        inprod = integrate(f,lower=a,upper=b)$value
+        inprod = integrate(Vectorize(f),lower=a,upper=b)$value
         analytic_geo[comb1,comb2]<-  
           acos(pmin(pmax(inprod,-1.0),1.0))
       }
@@ -293,7 +293,18 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
       integrand <- function(s){
         X(s)^2
       }
-      return(sqrt(integrate(integrand,lower=lower,upper=upper)$value))
+      
+      out <- tryCatch(
+        {
+          sqrt( integrate(Vectorize(integrand),lower=lower,upper=upper)$value )
+        },
+        error=function(cond) {
+          message(cond)
+          # Choose a return value in case of error
+          sqrt( integrate(Vectorize(integrand),lower=lower,upper=upper,subdivisions=1000)$value )
+        }
+      )    
+      return(out)
     }
     
     # returns dX_alpha/dalpha as a function

### 2019-07-28 compare_methods.sh
prep for new sim

@@ -27,7 +27,7 @@
 # The maximum running time of the job in days-hours:mins:sec
 #SBATCH --time=1-0:00:00
 
-#SBATCH --array=1-1200         ### Array index | %50: number of simultaneously tasks
+#SBATCH --array=1-2400         ### Array index | %50: number of simultaneously tasks
 
 # Get Array ID
 i=${SLURM_ARRAY_TASK_ID}
@@ -44,5 +44,4 @@ fi
 # module load MATLAB
 
 # The job command(s):
-# R --vanilla < spartanSim.R ${i} #this isn't necessary since we get the environment
 R --vanilla < compare_methods.R

### 2019-07-28 pairwise_geo_estimation.R

@@ -236,26 +236,29 @@ pairwise_geo_estimation <- function(method,
     #Define candidate values for delta
     delta_can=c(0,0.02,0.05,0.1,0.2)
   
-    # TODO: need to make sure these files are not overwriting each other when submitted in parallel.
+    # need to make sure intermediate text files are not overwriting each other when submitted in parallel.
     string <- format(Sys.time(), format = "%Y-%j-%H%M%S") 
-    write.table(delta_can,file=paste0('./possible_delta',string,'.txt'),col.names = FALSE, row.names = FALSE)
-    write.table(smooth_data,file=paste0('./discrete_data',string,'.txt'),col.names = FALSE, row.names = FALSE)
-    write.table(num_neigh,file=paste0('./possible_K',string,'.txt'),col.names = FALSE, row.names = FALSE)
-    write.table(c(samplesize,K),file=paste0('./n_K',string,'.txt'),col.names = FALSE, row.names = FALSE)
-    write.table(true_geo,file=paste0('./true_geo',string,'.txt'),col.names = FALSE, row.names = FALSE)
+    dir.create(string)
+    write.table(delta_can,file=paste0('./',string,'/possible_delta.txt'),col.names = FALSE, row.names = FALSE)
+    write.table(smooth_data,file=paste0('./',string,'/discrete_data.txt'),col.names = FALSE, row.names = FALSE)
+    write.table(num_neigh,file=paste0('./',string,'/possible_K.txt'),col.names = FALSE, row.names = FALSE)
+    write.table(c(samplesize,K),file=paste0('./',string,'/n_K.txt'),col.names = FALSE, row.names = FALSE)
+    write.table(true_geo,file=paste0('./',string,'/true_geo.txt'),col.names = FALSE, row.names = FALSE)
   
-    # Run penalized-isomap.m in matlab
-    run_matlab_script("./run_isomap_p.m")
+    file.copy("./run_isomap_p.m",paste0('./',string,'/'))
+    run_matlab_script(paste0('./',string,'/run_isomap_p.m'))
   
-    temp=as.vector(read.table(file='./manidis.txt',header=FALSE))
+    temp=as.vector(read.table(file=paste0('./',string,'/manidis.txt'),header=FALSE))
     estim_geo_penalized_isomap=matrix(temp[,1],ncol=samplesize)
-    temp2=read.table(file='./err_delta_neigh.txt',header=FALSE)
[…]

### 2019-07-28 compare_methods.R
new sim where pisomap runs correctly

@@ -33,69 +33,64 @@ sces = c(5,2,4)
 SNRs = c(0.1,0.5)
 reg_samplings = c(TRUE,FALSE)
 Ks_obs = c(100,30)
-# parameters in pairwise_geo_estimation
 
-total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(Ks_smooth)*length(mcs)
+total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(mcs)
 
 # get sweeping id, only use this block if submitting to cluster
 slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
 slurm_arrayid = as.numeric(slurm_arrayid)
 print(slurm_arrayid)
 
-#for (slurm_arrayid in 1:total_tasks){ #use this for loop if running locally
-  
-  # Hardcode
-  unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(Ks_smooth),length(mcs)))
-  
-  # actual parameters for this run
-  sce = sces[unravel[1,1]]
-  SNR = SNRs[unravel[1,2]]
-  reg_sampling = reg_samplings[unravel[1,3]]
-  K_obs = Ks_obs[unravel[1,4]]
-  K_smooth = Ks_smooth[unravel[1,5]]
-  mc = mcs[unravel[1,6]]
-  
-  
-  # Generate data
-  data<- sim_functional_data(samplesize = samplesize, 
-                             com_grid = com_grid, 
-                             plot_true = plotTrue,
-                             ### parameters under study
-                             sce = sce, 
-                             SNR = SNR, 
-                             reg_sampling = reg_sampling,
-                             K = K_obs)
-  
-  
-  # Estimation of geodesic distances with different methods
-  meth <- list("NN" = FALSE,
-               "RD_o" = FALSE,
-               "RD" = FALSE,
-               "SS_o" = FALSE,
-               "SS" = TRUE,
-               "pI" = TRUE,
-               "OUR" = FALSE,
-               "OUR2" = FALSE,
-               "OUR3"=TRUE,
-               "RP" = FALSE,
-               "L2" = TRUE, 
-               "w_L2" = FALSE )# see pairwise_geo_estimation for more info
-  
-  Estim<- pairwise_geo_estimation(method=meth,
-                                  true_data = data$noiseless_data,
-                                  discrete_data = data$noisy_data,data$analytic_geo,
[…]

### 2019-07-28 index.Rmd
some writing of sim scenario section

@@
 In this work we set ourselves the task of calculating the pairwise geodesic distance between functional observations that, due to noise, live near a non-linear manifold rather than exactly on the manifold.
+Good estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
-We will show that good estimation of the geodesic distance cannot be accomplished by naive methods such as smoothing followed by a shortest-path algorithm. 
-Instead, we propose a dedicated technique for pairwise geodesic distance estimation by first projecting the functional data to the underlying hidden manifold and then perofrming operations on this hidden manifold.
-We conclude the work by showing the implications for many downstream tasks in functional data analysis, in particular distanced-based functional classification.
+<!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
+The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
+We conclude the work by using our pairwise geodesic distance estimator in distanced-based functional classification.
 
@@
 Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-We additionally assume that the function $X$ belongs to a low-dimensional nonlinear manifold $\M \subset L^2([a,b],\R)$, in which case a geodesic distance taking into account the intrinsic structure of $\M$ rather than the $L^2$ distance should be used as a measure of distance between two functions in $\M$.
+We additionally assume that the function $X$ belongs to a low-dimensional nonlinear manifold $\M \subset L^2([a,b],\R)$.
+In this case a geodesic distance taking into account the intrinsic structure of $\M$ should be used as a measure of distance between two functions in $\M$, rather than the $L^2$ distance.
 As many statistical methods depend on some measure of distance, downstream benefits of using the geodesic distance can be expected.
 
-???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.???
+<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->
 
@@
 Given perfect functional observations, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance can be accomplished straightforwardly using algorithms such as Floyd's.
-However, functional data are observed with noise in such a way that the observed functional data no longer live on a manifold.
+However, when functional data are observed with noise, this approach is likely to fail.
+This is because we assume the error is such that the noisy functional manifold data no longer live on a manifold.
 Though powerful techniques exist for smoothing discretely observed noisy functional data, the smoothing is designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
@@
 ## Simulation scenarios [Susan]
-Acknowledge that we are sampling on a very concentrated measure to make sure functional manifold is well sampled.
-Cite Diaconis et.
-al's "Sampling from a manifold".
+
+### Isometric functional manifold of normal density functions
+<!-- Scenario 5 -->
+This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$.
+We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
+We set $a=-4$ and $b=4$.The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
+\begin{eqnarray*}
+d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
+&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
+&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
+&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
+&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
+\end{eqnarray*}
+where the approximation comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
+We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is the Euclidan distance between the $\beta$'s, up to some scaling factor.
+Note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$, so we cannot employ the calculation technique of Scenario 1.
+
+### Functional manifold of square root velocity functions
+It was shown in [@Joshi2007] that the square root representation of probability density functions has a nice closed form geodesic.
+They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
+Note that this concides with the $L_2[0,1]$ inner product.
+[@Joshi2007] showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$.
+That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$.
+
+### Functional manifold of warping functions
+<!-- Scenario 5 -->
+This is based on Equations (17) and (18) of [@Kneip2008] but with $z_{i1}, z_{i2}$ set to $1$.
+(Equation 17 has a typo where the exponentials are missing negative signs).Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$ and $h_\alpha(t) = t$ if $\alpha = 0$.
+Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\} $$.
+The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$
+
+### Sampling from a functional manifold
+As of yet, there is little work as to how to sample from a functional manifold.
+Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
+To safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
+For both the manifold of normal densities and the manifold of warping functions, we sample $\alpha$ according to ???rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)???
+For the manifold of square root beta densities, we sample as follows       
+alpha<- rtruncnorm(nb_alpha,alphamin,alphamax,(alphamax+alphamin)/2,0.3)
+beta<- rtruncnorm(nb_beta,betamin,betamax,(betamax+betamin)/2,0.3)
+
 How to sample properly from a functional manifold could be interesting future work.
-
-* warping manifold (scenario 5 in sim_functional_data.R)
-* pdf manifold (scenario 2 in sim_functional_data.R)
-* square root pdf manifold (scenario 4 )
-
 ### Results
@@
 
-For simplicity, assume the task is bianry classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
+For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
 Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:

### 2019-07-28 sim_functional_data.R

@@ -77,7 +77,7 @@
 # samplesize: number of functional data 
 # K: number of time grid points
 # SNR (signal to noise ratio)
-# reg_sampling : 0 if manifold parameters are drawn uniformly and 1 if manifold parameters are drawn from a concentrated measure. SHOULD ALWAYS SET THIS TO 1
+# reg_sampling : 0 if manifold parameters are drawn uniformly and 1 if manifold parameters are drawn from a concentrated measure.
 # com_grid : if 1 each curve is observed on the same grid and if 0 the grid for each curve is randomly genarated from unif[a,b]
 # plot_true : if 1 plot the true data, the observed data and the true geodesic matrix
 

### 2019-07-29 index.Rmd
more writing of intro and sim

@@
 Functional data analysis is the statistical analysis of smooth infinite-dimensional curves. 
-The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that, since observations are smooth over their domain, the dimensionality of the curves is only artifically high.
-Less studied in functional data analysis is the idea that the curves may actually live in a submanifold with low intrinsic dimension.
-Interesting classes of functional manifold data include the class of probability density functions and the class of warped curves of a common template function. 
+The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the dimensionality of curves is only artifically high if they are smooth.
+There is also another sense in which the dimensionality of the curves under study may be low.
+This is the much less explored idea that curves may actually live in a submanifold with low intrinsic dimension.
+Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
 Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration. 
-In this work we set ourselves the task of calculating the pairwise geodesic distance between functional observations that, due to noise, live near a non-linear manifold rather than exactly on the manifold.
-Good estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis.
+In this work we set ourselves the task of estimation the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near a non-linear manifold rather than exactly on the manifold.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
@@
 The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
-We conclude the work by using our pairwise geodesic distance estimator in distanced-based functional classification.
+Good estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
 
-<!-- ##############  -->
+# Introduction 
+<!-- Susan  -->
+Many statistical methods depend on a measure of distance.
+Methods in functional data analysis, the study of infinite-dimensional smooth curves, is no exception.
+In situations where the manifold hypothesis might be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
+The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.
 
-# Introduction [Susan]
-
-Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-We additionally assume that the function $X$ belongs to a low-dimensional nonlinear manifold $\M \subset L^2([a,b],\R)$.
-In this case a geodesic distance taking into account the intrinsic structure of $\M$ should be used as a measure of distance between two functions in $\M$, rather than the $L^2$ distance.
-As many statistical methods depend on some measure of distance, downstream benefits of using the geodesic distance can be expected.
+Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+Suppose additionally that the function $X$ belongs to a low-dimensional nonlinear Riemannian manifold $\M \subset L^2([a,b],\R)$ where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold as follows [@Lin2014].
+For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
+The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
+$$ d_g(X_i,X_j):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = X_i, \gamma(b) = X_j]  \}$$
+where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
+is the length of a smooth curve $\gamma: [a,b] \subset \R \to \M$.
 
@@
 
-Given perfect functional observations, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance can be accomplished straightforwardly using algorithms such as Floyd's.
-However, when functional data are observed with noise, this approach is likely to fail.
-This is because we assume the error is such that the noisy functional manifold data no longer live on a manifold.
-Though powerful techniques exist for smoothing discretely observed noisy functional data, the smoothing is designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
+Given perfect functional observations $X_i$ and $X_j$, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance $d_g(X_i,X_j)$ can be accomplished straightforwardly using a shortest-path algorithm, e.g.
+the Floyd-Warhsall algorithm [@Floyd1962].
+However, when functional data are observed with noise, this approach is likely to fail. This is because the error may be such that the noisy functional manifold data no longer live on a manifold. 
+A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
+This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
 <!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
@@
 
-With this challenge in mind, we put forth a technique for the specific task of estimating pairwise geodesic distances $\{ d_\M (X_i,X_j\}_{i>j}$, to be defined  when we have access only to discretely-observed noisy functional observations that possibly live off the true manifold.
-Briefly outline the methodology: MDS plus density ridge are used in a cool way to estimate (consistently?, let's hope..) pairwise geodesic distances.
+In this work, we put forth a technique for the specific task of recovering pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ when we have, in lieu of $X_i$ and $X_j$, access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
+This is accomplished by first projecting the functional data on to the underlying hidden manifold and then performing operations on this hidden manifold.
 
@@
 
-## Related Work [Susan]
+## Related Work
+<!-- Susan  -->
 [@ChenMuller2012]
@@
 
-## Setting [Susan]
-Let $X_1,\ldots,X_n$ be independent realizations of a random function $X\in\M \subset L^2([a,b],\R)$, where $\M$ is an unknown manifold.
+# Proposed method for estimating geodesic distances [Marie]
+
 Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
@@
 
-<!-- ##############  -->
-
-## Relevant differential geometry concepts [Susan]
-
-We first rigorously define the concepts of geodesic and geodesic distance.
-Following [@Lin2014], let $(\M,g)$ be a $d$-dimensional Riemannian manifold, where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold.
-Specifically, for each point $p$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_p$ on the tangent space $T_p \M$.
-The norm of a tangent vector $v \in T_p \M$ is defined as $$||v|| = \sqrt{g_p(v,v)}.$$
-If $\gamma: [a,b] \subset \R \to \M$ is a smooth curve, its length is then defined as
-$$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt .$$
-The geodesic distance between two points $p,q$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.$$
-<!-- ##############  -->
-
-# Proposed method for estimating geodesic distances [Marie]
-
 We first describe an important helper algorithm which we call IsoGeo. It is a subroutine of the classic Isomap procedure.
@@
 are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Warshall1962]
-<!-- or Dijkstra’s  which one do we actually use in our implementation, the FW or D? -->
+The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Floyd1962].
+<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
 \item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
@@
 
-We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances for discretely-observed noisy functional data.
+We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances when one only has access to discretely-observed noisy functional data.
+Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.
+
+* Near-$\epsilon$ isometry: area under the curve of the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
+* Relative Frobenius: let $d$ and $\hat d$ be the true and estimated $n \times n$ geodesic distance matrices, respectively. 
+$$
+||d-\hat d||/||d||
+$$
+* Pearson correlation: take the upper diagonal of $d$ and $\hat d$, calculate their Pearson correlation coefficient.
 
@@
 
-## Performance metrics [Susan]
-We describe three different metrics we use to assess the quality of a pairwise geodesic distance estimator.
-This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
-
-## Simulation scenarios [Susan]
+## Simulation scenarios 
+<!-- Susan  -->
 
@@
 
-### Sampling from a functional manifold
+## Sampling from a functional manifold
 As of yet, there is little work as to how to sample from a functional manifold.
@@
 How to sample properly from a functional manifold could be interesting future work.
-### Results
 
+## Results
 Submit compare_methods.sh as slurm job using sbatch.
@@
 ```
+<!--
 
@@
 
-<!-- ##############  -->
+-->
 

### 2019-07-29 index.Rmd
update index

--- 2019-06-26 index.Rmd
@@
 output:
+  html_document:
+    df_print: paged
+    number_sections: yes
+  word_document: default
   pdf_document:
     keep_tex: yes
-    number_sections: yes
-  word_document: default
-  html_document:
-    df_print: paged
     number_sections: yes

### 2019-07-29 index.tex
update index

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The hidden manifold distance for functional data},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{The hidden manifold distance for functional data}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

\newcommand {\To}{\rightarrow}
\newcommand {\TO}{\Rightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\Prob}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand {\cov}{\textrm{Cov}}
\newcommand {\var}{\textrm{Var}}
\newcommand {\1}{\textrm{\textbf{1}}}
\newcommand{\M}{\mathcal{M}}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Functional data analysis is the statistical analysis of smooth
infinite-dimensional curves. The challenge of analyzing
infinite-dimensional objects is ameliorated by the fact that, since
observations are smooth over their domain, the dimensionality of the
curves is only artifically high. Less studied in functional data
analysis is the idea that the curves may actually live in a submanifold
with low intrinsic dimension. Interesting classes of functional manifold
data include the class of probability density functions and the class of
warped curves of a common template function. Since these spaces are not
vector spaces, even basic operations such as addition and subtraction
require special consideration. In this work we set ourselves the task of
calculating the pairwise geodesic distance between functional
observations that, due to noise, live near a non-linear manifold rather
than exactly on the manifold. We will show that good estimation of the
geodesic distance cannot be accomplished by naive methods such as
smoothing followed by a shortest-path algorithm. Instead, we propose a
dedicated technique for pairwise geodesic distance estimation by first
projecting the functional data to the underlying hidden manifold and
then perofrming operations on this hidden manifold. We conclude the work
by showing the implications for many downstream tasks in functional data
analysis, in particular distanced-based functional classification.

\hypertarget{introduction-susan}{%
\section{Introduction {[}Susan{]}}\label{introduction-susan}}

Let \(X_1,\ldots,X_n\) be a sample of \(n\) independant realizations of
a random variable \(X\) that takes value in the Hilbert space
\(L^2([a,b],\mathbb{R})\). We additionally assume that the function
\(X\) belongs to a low-dimensional nonlinear manifold
\(\mathcal{M}\subset L^2([a,b],\mathbb{R})\), in which case a geodesic
distance taking into account the intrinsic structure of \(\mathcal{M}\)
rather than the \(L^2\) distance should be used as a measure of distance
between two functions in \(\mathcal{M}\). As many statistical methods
depend on some measure of distance, downstream benefits of using the
geodesic distance can be expected.

???insert example that shows pairwise geodesic distances give visually
more interesting results than pairwise \(L_2\) distance.???

Given perfect functional observations, i.e.~with no measurement error
and on a very dense domain grid, the estimation of the geodesic distance
can be accomplished straightforwardly using algorithms such as Floyd's.
However, functional data are observed with noise in such a way that the
observed functional data no longer live on a manifold. Though powerful
techniques exist for smoothing discretely observed noisy functional
data, the smoothing is designed for the \(L_2\) recovery of the original
curve, not to ensure the recovered functional versions of the data lie
on or close to the functional manifold \(\mathcal{M}\).

With this challenge in mind, we put forth a technique for the specific
task of estimating pairwise geodesic distances
\(\{ d_\mathcal{M}(X_i,X_j\}_{i>j}\), to be defined when we have access
only to discretely-observed noisy functional observations that possibly
live off the true manifold. Briefly outline the methodology: MDS plus
density ridge are used in a cool way to estimate (consistently?, let's
hope..) pairwise geodesic distances.

\hypertarget{related-work-susan}{%
\subsection{Related Work {[}Susan{]}}\label{related-work-susan}}

(Chen and Muller 2012)

(Lin and Yao 2017)

(Dimeglio et al. 2014)

\hypertarget{setting-susan}{%
\subsection{Setting {[}Susan{]}}\label{setting-susan}}

Let \(X_1,\ldots,X_n\) be independent realizations of a random function
\(X\in\mathcal{M}\subset L^2([a,b],\mathbb{R})\), where \(\mathcal{M}\)
is an unknown manifold. Suppose that each curve \(X_i\) is observed with
measurements errors on a grid \(T_i=(t_{i1},\ldots,t_{iK})\), i.e.~we
observe a sample of \(K\)-dimensional vectors \(Y_1,\ldots,Y_n\) with
\(Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}\), where the random variables
\(\epsilon_{ij}\) are of mean zero and uncorrelated with each other. We
assume that the grids \(T_1,\ldots,T_n\) are dense.

Let \(\tilde X_1,\ldots,\tilde X_n\) denote the functional versions of
the raw data by some smoothing method. Our proposed methodology is
unspecific to the smoothing method employed. For simplicity however,
let's say we employ spline smoothing to recover the functional versions
of the raw data, i.e.\\
\[ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}\]
where \(\lambda>0\) is a tuning parameter controlling the smoothness of
\(\tilde X_i\).

\hypertarget{relevant-differential-geometry-concepts-susan}{%
\subsection{Relevant differential geometry concepts
{[}Susan{]}}\label{relevant-differential-geometry-concepts-susan}}

We first rigorously define the concepts of geodesic and geodesic
distance. Following (Lin et al. 2014), let \((\mathcal{M},g)\) be a
\(d\)-dimensional Riemannian manifold, where \(g\) is a Riemannian
metric tensor on \(\mathcal{M}\) which can be used to assign a metric on
the manifold. Specifically, for each point \(p\) on the manifold, the
Riemannian metric tensor \(g\) has an inner product \(g_p\) on the
tangent space \(T_p \mathcal{M}\). The norm of a tangent vector
\(v \in T_p \mathcal{M}\) is defined as \[||v|| = \sqrt{g_p(v,v)}.\] If
\(\gamma: [a,b] \subset \mathbb{R}\to \mathcal{M}\) is a smooth curve,
its length is then defined as
\[ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt .\] The
geodesic distance between two points \(p,q\) on the manifold
\(\mathcal{M}\), based on this metric tensor \(g\), is defined as
\[ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \mathcal{M}\text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.\]

\hypertarget{proposed-method-for-estimating-geodesic-distances-marie}{%
\section{Proposed method for estimating geodesic distances
{[}Marie{]}}\label{proposed-method-for-estimating-geodesic-distances-marie}}

We first describe an important helper algorithm which we call IsoGeo. It
is a subroutine of the classic Isomap procedure. Isomap is a three steps
procedure that takes as input a set of points
\(x_1,\ldots,x_n\in \mathbb{R}^D\) and produces an embedding of the
input data in the space \(\mathbb{R}^d\) with \(d<D\), that preserves
pairwise geodesic distances.

\begin{enumerate}
\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\mathbb{R}^D}$. The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
\item In the second step, the pairwise geodesic distances $d_\mathcal{M}(x_i,x_j)$
are estimated based on $G$. The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Warshall1962]
<!-- or Dijkstra’s  which one do we actually use in our implementation, the FW or D? -->
\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\mathbb{R}^d$.
\end{enumerate}

In what follows, we call IsoGeo the procedure of estimating pairwise
geodesic distances with the two first steps of Isomap. Note that Floyd
or Dijkstra are methods to find the smallest path given a weigthed graph
but they are not related to how to calculate that graph (which is the
difficult part, depend on number of neighbors, many method out there to
calculate that graph in a robust way, P-ISOMAP is one of these).

Our method is described by the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Obtain a \(s\)-dimensional representation
  \(\tilde X^s_1,\ldots,\tilde X^s_n\) of the functions
  \(\tilde X_1,\ldots,\tilde X_n\) that ``preserves'' the pairwise
  \(L^2\) distances by using MDS.
\item
  Obtain
  \(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
  a ``projection'' of \(\tilde X^s_1,\ldots,\tilde X^s_n\) onto a ridge
  \(\hat \mathcal{M}\) which is computed with the subspace constrained
  mean-shift algorithm (Ozertem and Erdogmus 2011).
\item
  Use IsoGeo to approximate the pairwise geodesic distances
  \(\{d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j)\}_{i>j}\)
  and define the \(n\) by \(n\) matrix \[
  \hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
   d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j) & \textrm{if $i\neq j$,}\\
   0 & \textrm{otherwise.}
    \end{array} \right.
  \]
\end{enumerate}

Since the ridge estimation obtained from noisy measurements of a
manifold should approximate well the manifold (Genovese et al. 2014), we
expect the points
\(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
to lie close to the real manifold \(\mathcal{M}\) and then
\(d_{\hat \mathcal{M}}\) to be close to \(d_\mathcal{M}\). Ridge
estimation suffers from the curse of dimensionality, this is why we
first reduce the dimension of our data with MDS and then apply the
shift-mean algorithm to estimate the ridge.

\hypertarget{selection-of-tuning-parameters-marie}{%
\subsection{Selection of tuning parameters
{[}Marie{]}}\label{selection-of-tuning-parameters-marie}}

Describe heuristics. We pick the bandwidth \(h\) in the subspace
constrained mean-shift using Equation (A1) of (Chen et al. 2015).

\hypertarget{simulation-study}{%
\section{Simulation study}\label{simulation-study}}

We perform a simulation study to ascertain the efficacy of our method
for estimating pairwise geodesic distances for discretely-observed noisy
functional data.

\hypertarget{alternative-estimators-of-geodesic-distance-marie}{%
\subsection{Alternative estimators of geodesic distance
{[}Marie{]}}\label{alternative-estimators-of-geodesic-distance-marie}}

\begin{itemize}
\tightlist
\item
  \textbf{(Raw Data) RD} Apply IsoGeo on the raw data \(Y_1,\ldots,Y_n\)
  to obtain an estimator \(\hat G_{\textrm{RD}}\). Note that this
  procedure can only be used if the grid \(T_i\) is the same for each
  \(i\).
\item
  \textbf{(Spline Smoothing) SS} Transform each vector \(Y_i\) into a
  function \(\tilde X_i\) by spline smoothing. Apply IsoGeo on the
  vectors \(\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n\), where
  \(t_1,\ldots,t_K\) is a regular grid of \([a,b]\), to obtain an
  estimator \(\hat G_{\textrm{SS}}\).
\item
  \textbf{(P-ISOMAP) pI} Transform each vector \(Y_i\) into a function
  \(\tilde X_i\) by spline smoothing. Calculate the weighted graph of
  step I of Isomap with a penalty (describe the penalty of Chen and
  Muller). Apply step II of Isomap to obtain \(\hat G_{\textrm{pI}}\).
\item
  \textbf{(Random Projection) RP} same method as our but change step 2
  by : obtain \(s\)-dimensional representation by random projection and
  setp 4 by obtain \(\hat G\) using a ensemble method.
\end{itemize}

We ran Chen and Muller's P-ISOMAP separately in Matlab with their
automatic penalty selection and discovered that ???.

\hypertarget{performance-metrics-susan}{%
\subsection{Performance metrics
{[}Susan{]}}\label{performance-metrics-susan}}

We describe three different metrics we use to assess the quality of a
pairwise geodesic distance estimator. This is the ROC curve with
\(\epsilon\) on the \(x\)-axis and degree to which near-\(\epsilon\)
isometry holds, i.e.~the percentage of estimated pairwise distances
between \(1-\epsilon\) and \(1+\epsilon\) of the truth pairwise
distance.

\hypertarget{simulation-scenarios-susan}{%
\subsection{Simulation scenarios
{[}Susan{]}}\label{simulation-scenarios-susan}}

Acknowledge that we are sampling on a very concentrated measure to make
sure functional manifold is well sampled. Cite Diaconis et. al's
``Sampling from a manifold''. How to sample properly from a functional
manifold could be interesting future work.

\begin{itemize}
\tightlist
\item
  warping manifold (scenario 5 in sim\_functional\_data.R)
\item
  pdf manifold (scenario 2 in sim\_functional\_data.R)
\item
  square root pdf manifold (scenario 4 )
\end{itemize}

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

Submit compare\_methods.sh as slurm job using sbatch.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{sbatch}\NormalTok{ compare_methods.sh}
\end{Highlighting}
\end{Shaded}

Alternatively, we can run the R script compare\_methods.R locally by
removing the slurm\_arrayid chunk

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{R compare_methods.R}
\end{Highlighting}
\end{Shaded}

Simulation results are then visualized by invoking cluster\_results.R.
There's some unfortunate hardcoding still in cluster\_results.R

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{R cluster_results.R}
\end{Highlighting}
\end{Shaded}

\hypertarget{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}{%
\subsection{The cost of employing proposed method when manifold is flat
{[}Susan/Marie?{]}}\label{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}}

Suppose we take scenario 1 where the geodesic distance coincides with
the \(L_2\) distance. Then what is the cost of employing our distance
estimator compared to smoothing the data and doing numerical integration
to find the \(L_2\) distance?

\hypertarget{distance-based-functional-classification-marie}{%
\section{Distance-based functional classification
{[}Marie{]}}\label{distance-based-functional-classification-marie}}

In this section, we explore whether our geodesic distance estimator has
benefits for downstream analysis task. There are many tasks we could
consider here such as distance-based nonparametric regression and
distanced-based functional clustering, but we will focus on
distance-based functional classification. It must be noted that while
curve alignment, also known as curve registration, is necessarily
performed as a preprocessing technique prior to clustering and
classification, our geodesic distance estimator allows one to forsake
this step.

For simplicity, assume the task is binary classification. Associated to
each functional object \(x\) is a binary \(y\) indicating class
membership. Consider the classifier proposed in Ferraty and Vieu
(2003,2006) which is a functional version of the Nadaraya-Watson kernel
estimator of class membership probabilities: \[
\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
\] We shall compare our method to using \(L_2\) distance, possibly
weighted, and with curve registration already accomplished. Describe
alternative methods in detail.

The bandwidth in the classifier should be tuned individually for each
method. Also we might need to tune MDS dimension \(s\) since in real
data, the dimension of the manifold might be much higher than
encountered in the simulation scenarios where it never goes above 2.

Datasets used by functional classification papers

\begin{itemize}
\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
\item Berkeley growth curves in [@ChenReiss2014].
\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
\end{itemize}

Datasets used in functional manifold papers

\begin{itemize}
\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
\item Tecator in [@LinYao2017] contamination paper
\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ChenMuller2012}{}%
Chen, Dong, and Hans-Georg Muller. 2012. ``NONLINEAR Manifold
Representations for Functional Data.'' \emph{The Annals of Statistics}
40 (1). Institute of Mathematical Statistics: 1--29.
\url{http://www.jstor.org/stable/41713625}.

\leavevmode\hypertarget{ref-ChenHo2015}{}%
Chen, Y., S. Ho, P. E. Freeman, C. R. Genovese, and L. Wasserman. 2015.
``Cosmic Web Reconstruction Through Density Ridges: Method and
Algorithm.'' \emph{Monthly Notices of the Royal Astronomical Society}
454 (1): 1140--56. \url{https://doi.org/10.1093/mnras/stv1996}.

\leavevmode\hypertarget{ref-Dimeglio2014}{}%
Dimeglio, Chloe, Santiago Gallon, Jean-Michel Loubes, and Elie Maza.
2014. ``A Robust Algorithm for Template Curve Estimation Based on
Manifold Embedding.'' \emph{Comput. Stat. Data Anal.} 70 (February).
Amsterdam, The Netherlands, The Netherlands: Elsevier Science Publishers
B. V.: 373--86. \url{https://doi.org/10.1016/j.csda.2013.09.030}.

\leavevmode\hypertarget{ref-Genovese2014}{}%
Genovese, Christopher R., Marco Perone-Pacifico, Isabella Verdinelli,
and Larry Wasserman. 2014. ``NONPARAMETRIC Ridge Estimation.'' \emph{The
Annals of Statistics} 42 (4). Institute of Mathematical Statistics:
1511--45. \url{http://www.jstor.org/stable/43556332}.

\leavevmode\hypertarget{ref-Lin2014}{}%
Lin, Binbin, Ji Yang, Xiaofei He, and Jieping Ye. 2014. ``Geodesic
Distance Function Learning via Heat Flow on Vector Fields.'' In
\emph{Proceedings of the 31st International Conference on International
Conference on Machine Learning - Volume 32}, II--145--II--153. ICML'14.
Beijing, China: JMLR.org.
\url{http://dl.acm.org/citation.cfm?id=3044805.3044909}.

\leavevmode\hypertarget{ref-LinYao2017}{}%
Lin, Zhenhua, and Fang Yao. 2017. ``Functional Regression on Manifold
with Contamination.'' \emph{arXiv E-Prints}, April, arXiv:1704.03005.

\leavevmode\hypertarget{ref-Ozertem2011}{}%
Ozertem, Umut, and Deniz Erdogmus. 2011. ``Locally Defined Principal
Curves and Surfaces.'' \emph{J. Mach. Learn. Res.} 12 (July). JMLR.org:
1249--86. \url{http://dl.acm.org/citation.cfm?id=1953048.2021041}.

\end{document}

### 2019-07-29 index.Rmd
write a bit in proposed method

@@
 output:
-  html_document:
-    df_print: paged
-    number_sections: yes
   word_document: default
@@
     keep_tex: yes
+    number_sections: yes
+  html_document:
+    df_print: paged
     number_sections: yes
@@
 
-Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data by some smoothing method.
+Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
 Our proposed methodology is unspecific to the smoothing method employed.
 For simplicity however, let's say we employ spline smoothing to recover the functional versions of the raw data, i.e.\
-$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
+$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},$$
 where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
 
-We first describe an important helper algorithm which we call IsoGeo. It is a subroutine of the classic Isomap procedure.
+We first describe an important helper algorithm which we call IsoGeo and that has been proposed in @Dimeglio2014.
+It is a modified version of a subroutine of the classic Isomap procedure [@Tenenbaum2000].
 Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+The three steps are as follow:
 
-\begin{enumerate}
-\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-\item In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Floyd1962].
-<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
-\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
-\end{enumerate}
+1.
+Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ if the distance $d_{ij}=||x_i-x_j ||_{2}$ is smaller than a given value $\epsilon$, and the weight associated to an edge $e_{ij}$ is $d_{ij}$.
+2.
+Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$.
+The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, i.e.
+the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's?
+[Marie: In robust_isomap Dijkstra is used but both should give the same results no?
+I'm not sure we need to specify it]-->
+3. Use Multidimensional Scaling (MDS) algorithm on the matrix of pairwise geodesic distances obtained in the second step in order to obtain an embedding in $\R^d$.
 
-In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
-Note that Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these).
+Note that in step 2 Floyd or Dijkstra algorithms are methods to find the shortest path given a weigthed graph but they are not related to how to calculate that graph in step 1 which is usually the difficult part.
+Indeed it depends heavily on the choice of the tuning parameter $\epsilon$, and there exist many methods out there to calculate that graph in a robust way; p-ISOMAP is one of these as well as IsoGeo.
+In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap but where the method to construct the graph $G$ in the first step has been modified as follow in order to avoid using the tuning parameter $\epsilon$:
+
+1'.
+Construct the complete weigthed graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Obtain the minimal spanning tree $G_s$ associated with $G_c$ and denote its set of edges by $E_s$.
+The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
+$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
+where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
+
+For more details on computations involved in IsoGeo, one should refer to Algorithm 1 of @Dimeglio2014.
 

### 2019-07-30 index.Rmd
more writing of methodology section

--- 2019-07-29 index.Rmd
@@
 output:
+  pdf_document:
+    keep_tex: yes
+    number_sections: yes
+  word_document: default
   html_document:
     df_print: paged
-    number_sections: yes
-  word_document: default
-  pdf_document:
-    keep_tex: yes
     number_sections: yes
@@
 There is also another sense in which the dimensionality of the curves under study may be low. 
-This is the much less explored idea that curves may actually live in a submanifold with low intrinsic dimension.
+This is the much less explored idea that the curves may actually live in a submanifold with low intrinsic dimension.
 Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
-Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration. 
-In this work we set ourselves the task of estimation the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near a non-linear manifold rather than exactly on the manifold.
+<!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
+In this work we address the estimation of the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near, rather than exactly on, the manifold.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
@@
 # Introduction 
-<!-- Susan  -->
 Many statistical methods depend on a measure of distance.
-Methods in functional data analysis, the study of infinite-dimensional smooth curves, is no exception.
+Methods in functional data analysis, the study of infinite-dimensional smooth curves, are no exception.
+Certain classes of functional data may naturally live on manifold, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
 In situations where the manifold hypothesis might be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
@@
 
-In this work, we put forth a technique for the specific task of recovering pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ when we have, in lieu of $X_i$ and $X_j$, access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
+In this work, we put forth a technique for the specific task of recovering the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
+$$
+G(i,j)=G(j,i) = \left\{ \begin{array}{ll}
+ \{ d_g (X_i,X_j)\}_{i>j} & \textrm{if $i\neq j$,}\\
+ 0 & \textrm{otherwise.}
+  \end{array} \} \right.
+$$ 
+when we have, in lieu of $X_i$ and $X_j$, access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
 This is accomplished by first projecting the functional data on to the underlying hidden manifold and then performing operations on this hidden manifold.
@@
 ## Related Work
-<!-- Susan  -->
-[@ChenMuller2012]
+The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data.
+A notion of the mean and variation of functional manifold data was introduced.
+Of particular relevance to our work is the modified ISOMAP procedure, called P-ISOMAP, introduced in [@ChenMuller2012], which adds a data-adaptive penalty to allow for noisy functional observations whereas in the classic ISOMAP algorithm, observations are assumed to lie exactly on the manifold.
+Another existing work for estimating the pairwise geodesic distance for functional manifold data can be found in [@Dimeglio2014].
+The estimator is called Robust-ISOMAP for the fact that it is motivated by modifying the ISOMAP algorithm to be less sensitive to outliers.
+We will discuss both P-ISOMAP and robust ISOMAP in more details in the simulation section.
 
-[@LinYao2017]
+<!-- [@LinYao2017]  -->
 
-[@Dimeglio2014]
-
-# Proposed method for estimating geodesic distances [Marie]
+# Proposed method for estimating geodesic distances
 
 Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
-we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ are of mean zero and uncorrelated with each other.
-We assume that the grids $T_1,\ldots,T_n$ are dense.
+we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
+We assume that each observation grid $T_1,\ldots,T_n$ is dense.
 
+We begin by converting the discrete functional observations into continuous ones.
 Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data by some smoothing method.
-Our proposed methodology is unspecific to the smoothing method employed.
-For simplicity however, let's say we employ spline smoothing to recover the functional versions of the raw data, i.e.\
+Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
+For example we may employ spline smoothing to recover the functional versions of the raw data, i.e.\
 $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
@@
 
-We first describe an important helper algorithm which we call IsoGeo. It is a subroutine of the classic Isomap procedure.
-Isomap is a three steps procedure that takes as input a set of points $x_1,\ldots,x_n\in \R^D$ and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+Before introducing our procedure, let us first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
 
 \begin{enumerate}
-\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
+\item The first step constructs a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
 The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-\item In the second step, the pairwise geodesic distances $d_\M(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Floyd1962].
+\item In the second step, the pairwise geodesic distances $d_g(x_i,x_j)$ are estimated based on $G$ using shortest-path algorithms.
+Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, which we may calculate for instance with the Floyd-Warshall algorithm [@Floyd1962].
 <!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
-\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\R^d$.
+\item The last step consists in using multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
 \end{enumerate}
 
-In what follows, we call IsoGeo the procedure of estimating pairwise geodesic distances with the two first steps of Isomap.
-Note that Floyd or Dijkstra are methods to find the smallest path given a weigthed graph but they are not related to how to calculate that graph (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these).
+In what follows, we call IsoGeo the procedure which performs the first two steps of ISOMAP.Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves.
+In fact, P-ISOMAP and robust ISOMAP are both modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
+<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
 
-Our method is described by the following steps:
+Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm [@Ozertem2011].
+Theoretical justificaton for this can be found in [@Genovese2014].
+and then performing operations on this hidden manifold.
+We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
+is described by the following steps:
 
@@
 <!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
-2. Obtain a $s$-dimensional representation $\tilde X^s_1,\ldots,\tilde X^s_n$ of the functions $\tilde X_1,\ldots,\tilde X_n$ that "preserves" the pairwise $L^2$ distances by using MDS.
-3.
-Obtain $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ a "projection" of $\tilde X^s_1,\ldots,\tilde X^s_n$ onto a ridge $\hat \M$ which is computed with the subspace constrained mean-shift algorithm [@Ozertem2011].
-4. Use  IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$ and define the $n$ by $n$ matrix 
+\begin{enumerate}
+\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a common grid $\tilde T=(t_{1},\ldots,t_{K})$.
+Let $s$ be a positive integer, smaller than $K$.
+Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^K$ Euclidean distances on the discretised functions $\tilde X_i$ are preserved.
+\item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain  $\tilde X^{s,\hat \M}_i$.
+\item Use IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
+\end{enumerate}
+Thus our geodesic distance estimators is the $n{\times}n$ matrix $\hat G$ whose elements are given by
 $$
@@
 $$
+Since ridge estimation suffers from the curse of dimensionality, we first reduce the dimension of our data with multidimensional scaling in Step 1 before applying the mean-shift algorithm.
 
-Since the ridge estimation obtained from noisy measurements of a manifold should approximate well the manifold [@Genovese2014], we expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
-Ridge estimation suffers from the curse of dimensionality, this is why we first reduce the dimension of our data with MDS and then apply the shift-mean algorithm to estimate the ridge.
-
-## Selection of tuning parameters [Marie]
+## Selection of tuning parameters
 Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of [@ChenHo2015].
@@
 
-* Near-$\epsilon$ isometry: area under the curve of the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
-* Relative Frobenius: let $d$ and $\hat d$ be the true and estimated $n \times n$ geodesic distance matrices, respectively. 
-$$
-||d-\hat d||/||d||
-$$
-* Pearson correlation: take the upper diagonal of $d$ and $\hat d$, calculate their Pearson correlation coefficient.
+* the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis.
+Near-$\epsilon$ isometry is a relaxation of isometry measured by the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
+* the relative Frobenius metric: this is given by
+$||d-\hat d||/||d||$
+where $d$ and $\hat d$ are the true and estimated $n{\times}n$ geodesic distance matrices, respectively. 
+* the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
 

### 2019-07-30 index.Rmd

@@
 output:
+  word_document: default
   pdf_document:
@@
     number_sections: yes
-  word_document: default
   html_document:

### 2019-07-30 index.Rmd
put Marie's changes back into methodolgoy section

--- 2019-07-30 index.Rmd
@@
 We begin by converting the discrete functional observations into continuous ones.
-Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data by some smoothing method.
+Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
 Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
 For example we may employ spline smoothing to recover the functional versions of the raw data, i.e.\
-$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$
+$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},$$
 where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
 
-Before introducing our procedure, let us first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$, that preserves pairwise geodesic distances.
+Before introducing our procedure, let us first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+\begin{enumerate}
+\item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ if the distance $d_{ij}=||x_i-x_j ||_{2}$ is smaller than a given value $\epsilon$, and the weight associated to an edge $e_{ij}$ is $d_{ij}$.
+\item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms.
+Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e./ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
+\item Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
+\end{enumerate}
+Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves.
+Formulating the graph in step 1 of ISOMAP is in fact the most challenging asepct of the algorithm since it depends heavily on the choice of the tuning parameter $\epsilon$.
+Both P-ISOMAP and robust ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
+<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
 
-\begin{enumerate}
-\item The first step constructs a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\R^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-\item In the second step, the pairwise geodesic distances $d_g(x_i,x_j)$ are estimated based on $G$ using shortest-path algorithms.
-Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, which we may calculate for instance with the Floyd-Warshall algorithm [@Floyd1962].
-<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
-\item The last step consists in using multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
-\end{enumerate}
-
-In what follows, we call IsoGeo the procedure which performs the first two steps of ISOMAP.Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves.
-In fact, P-ISOMAP and robust ISOMAP are both modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
-<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
+In what follows, we call IsoGeo the procedure which performs a modified version of the first step of ISOMAP, proposed in @Dimeglio2014, followed by the original second step of ISOMAP.
+The first step of IsoGeo modifies Step 1 of ISOMAP so that the constructed graph $G$ is independent of the tuning parameter $\epsilon$:
+1'.
+Construct the complete weigthed graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Obtain the minimal spanning tree $G_s$ associated with $G_c$ and denote its set of edges by $E_s$.
+The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
+$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
+where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
+Note that IsoGeo has no tuning parameters.
 
@@
 Theoretical justificaton for this can be found in [@Genovese2014].
-and then performing operations on this hidden manifold.
 We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
-is described by the following steps:
+Our procedure is described by the following steps:
 
@@
 \end{enumerate}
-Thus our geodesic distance estimators is the $n{\times}n$ matrix $\hat G$ whose elements are given by
+Thus our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
 $$

### 2019-08-02 compare_methods.R
changed spartan partition, fixed spelling error in compare methods

@@ -36,11 +36,10 @@ Ks_obs = c(100,30)
 
 total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(mcs)
 
-# get sweeping id, only use this block if submitting to cluster
-slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
-slurm_arrayid = as.numeric(slurm_arrayid)
-print(slurm_arrayid)
-
+# slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
+# slurm_arrayid = as.numeric(slurm_arrayid)
+# print(slurm_arrayid)
+slurm_arrayid = 1
 
 # Hardcode
 unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(mcs)))
@@ -85,7 +84,7 @@ Estim<- pairwise_geo_estimation(method=meth,
                                 grid = data$grid,
                                 reg_grid = data$reg_grid,
                                 common_grid_true =com_grid,
-                                K_dense = K_smooth)
+                                K_dense = Ks_smooth)
 
 # Assessment 
 Rel_errs = lapply(Estim, assess_goodness_estimation, true_geo = data$analytic_geo)

### 2019-08-02 compare_methods.sh

@@ -3,8 +3,7 @@
 # example from https://stackoverflow.com/questions/50653937/how-to-run-a-job-array-in-r-using-the-rscript-command-from-the-command-line
 
 # Partition for the job:
-#SBATCH --partition=cloud
-
+#SBATCH -p cloud,physical
 
 # The name of the job:
 #SBATCH --job-name="manifoldFDAgeodesic"
@@ -25,7 +24,7 @@
 #SBATCH --mail-user=susan.wei@unimelb.edu.au
 
 # The maximum running time of the job in days-hours:mins:sec
-#SBATCH --time=1-0:00:00
+#SBATCH --time=0-3:00:00
 
 #SBATCH --array=1-2400         ### Array index | %50: number of simultaneously tasks
 

### 2019-08-02 index.Rmd

@@
 For example we may employ spline smoothing to recover the functional versions of the raw data, i.e.\
-$$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},$$
+\begin{equation}\label{eq_spline_smoothing} 
+\tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
+\end{equation}
 where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
@@
 ## Alternative estimators of geodesic distance [Marie]
+We compare the performance of our method to the one of the three following methods:
 
-* **(Raw Data) RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
-* **(Spline Smoothing) SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
-Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
-* **(P-ISOMAP) pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
-Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller).
-Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
-* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.
+1.
+**Raw Data (RD)** The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+<!-- to obtain an estimator $\hat G_{\textrm{RD}}$.--> Note that this procedure only makes sense if all the grids $T_1,\ldots,T_n$ are the same.
+2.
+**Spline Smoothing (SS)** The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
+<!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
+3.
+**P-ISOMAP (pI)** The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
+<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->
 

### 2019-08-02 paperAAAI.tex

--- 2019-08-02 index.Rmd
@@
----
-title: "The hidden manifold distance for functional data"
-output:
-  html_document:
-    df_print: paged
-    number_sections: yes
-  word_document: default
-  pdf_document:
-    keep_tex: yes
-    number_sections: yes
-bibliography: references.bib
----
+\def\year{2020}\relax
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
+\usepackage{aaai20}  % DO NOT CHANGE THIS
+\usepackage{times}  % DO NOT CHANGE THIS
+\usepackage{helvet} % DO NOT CHANGE THIS
+\usepackage{courier}  % DO NOT CHANGE THIS
+\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
+\usepackage{graphicx} % DO NOT CHANGE THIS
+\urlstyle{rm} % DO NOT CHANGE THIS
+\def\UrlFont{\rm}  % DO NOT CHANGE THIS
+\usepackage{graphicx}  % DO NOT CHANGE THIS
+\frenchspacing  % DO NOT CHANGE THIS
+\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
+\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
+
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{latexsym}
+\usepackage{amssymb}
+\usepackage{amsbsy}
+\usepackage{graphics}
+\usepackage{epsfig}
+\usepackage[usenames]{color}
+\usepackage{graphicx}
 
@@
 \newcommand{\M}{\mathcal{M}}
+%\nocopyright
+%PDF Info Is REQUIRED.
+% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
+% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
+ \pdfinfo{
+/Title (The hidden manifold distance for functional data)
+/Author (Susan Wei and Marie-H\'el\`{e}ne Descary)
+} %Leave this	
+% /Title ()
+% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
+% Leave the space between \Title and the beginning parenthesis alone
+% /Author ()
+% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case. 
+% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands, 
+% remove them.
 
-# Abstract
+% DISALLOWED PACKAGES
+% \usepackage{authblk} -- This package is specifically forbidden
+% \usepackage{balance} -- This package is specifically forbidden
+% \usepackage{caption} -- This package is specifically forbidden
+% \usepackage{color (if used in text)
+% \usepackage{CJK} -- This package is specifically forbidden
+% \usepackage{float} -- This package is specifically forbidden
+% \usepackage{flushend} -- This package is specifically forbidden
+% \usepackage{fontenc} -- This package is specifically forbidden
+% \usepackage{fullpage} -- This package is specifically forbidden
+% \usepackage{geometry} -- This package is specifically forbidden
+% \usepackage{grffile} -- This package is specifically forbidden
+% \usepackage{hyperref} -- This package is specifically forbidden
+% \usepackage{navigator} -- This package is specifically forbidden
+% (or any other package that embeds links such as navigator or hyperref)
+% \indentfirst} -- This package is specifically forbidden
+% \layout} -- This package is specifically forbidden
+% \multicol} -- This package is specifically forbidden
+% \nameref} -- This package is specifically forbidden
+% \natbib} -- This package is specifically forbidden -- use the following workaround:
+% \usepackage{savetrees} -- This package is specifically forbidden
+% \usepackage{setspace} -- This package is specifically forbidden
+% \usepackage{stfloats} -- This package is specifically forbidden
+% \usepackage{tabu} -- This package is specifically forbidden
+% \usepackage{titlesec} -- This package is specifically forbidden
+% \usepackage{tocbibind} -- This package is specifically forbidden
+% \usepackage{ulem} -- This package is specifically forbidden
+% \usepackage{wrapfig} -- This package is specifically forbidden
+% DISALLOWED COMMANDS
+% \nocopyright -- Your paper will not be published if you use this command
+% \addtolength -- This command may not be used
+% \balance -- This command may not be used
+% \baselinestretch -- Your paper will not be published if you use this command
+% \clearpage -- No page breaks of any kind may be used for the final version of your paper
+% \columnsep -- This command may not be used
+% \newpage -- No page breaks of any kind may be used for the final version of your paper
+% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
+% \pagestyle -- This command may not be used
+% \tiny -- This is not an acceptable font size.
+% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
+% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
+
+\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
+
+% The file aaai20.sty is the style file for AAAI Press 
+% proceedings, working notes, and technical reports.
+%
+\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
+% command to correct it. You may not alter the value below 2.5 in
+\title{The hidden manifold distance for functional data}
+%Your title must be in mixed case, not sentence case. 
+% That means all verbs (including short verbs like be, is, using,and go), 
+% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
+% articles, conjunctions, and prepositions are lower case unless they
+% directly follow a colon or long dash
+\author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be in the same font size and format.
+Use \Large and \textbf to achieve this result when breaking a line
+\textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ %If you have multiple authors and multiple affiliations
+% use superscripts in text and roman font to identify them.
+For example, Sunil Issar,\textsuperscript{\rm 2} J.
+Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}.
+Note that the comma should be placed BEFORE the superscript for optimum readability
+ % email address must be in roman text type, not monospace or sans serif
+}
+ \begin{document}
+
+\maketitle
+
+\begin{abstract}
 Functional data analysis is the statistical analysis of smooth infinite-dimensional curves. 
@@
 Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
-<!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
+%Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
 In this work we address the estimation of the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near, rather than exactly on, the manifold.
-<!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
-<!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
+% This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
+%We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
 The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
@@
 
-# Introduction 
+\end{abstract}
+
+\section{Introduction}
+
 Many statistical methods depend on a measure of distance.
@@
 Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a low-dimensional nonlinear Riemannian manifold $\M \subset L^2([a,b],\R)$ where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold as follows [@Lin2014].
+Suppose additionally that the function $X$ belongs to a low-dimensional nonlinear Riemannian manifold $\M \subset L^2([a,b],\R)$ where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold as follows \cite{Lin2014}.
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
-The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_g(X_i,X_j):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = X_i, \gamma(b) = X_j]  \}$$
-where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
+The norm of a tangent vector $V \in T_X \M$ is defined as $$\|V\| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
+\begin{eqnarray*} 
+d_g(X_i,X_j)&:=&\inf \big\{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth},\\
+&&\quad \quad  \gamma(a) = X_i, \gamma(b) = X_j] \big \},
+\end{eqnarray*}
+where $$ l(\gamma) := \int_a^b \left\| \frac{\,d\gamma}{\,d t}(t) \right\| \,dt$$
 is the length of a smooth curve $\gamma: [a,b] \subset \R \to \M$.
 
-<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->
+%<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->
 
-<!-- (gives examples and references for functional manifolds). -->
-<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->
+%<!-- (gives examples and references for functional manifolds). -->
+%<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->
 
-<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
+%<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
 However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio.
@@
 Given perfect functional observations $X_i$ and $X_j$, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance $d_g(X_i,X_j)$ can be accomplished straightforwardly using a shortest-path algorithm, e.g.
-the Floyd-Warhsall algorithm [@Floyd1962].
+the Floyd-Warhsall algorithm \cite{Floyd1962}.
 However, when functional data are observed with noise, this approach is likely to fail. This is because the error may be such that the noisy functional manifold data no longer live on a manifold. 
@@
 This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
-<!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
+%<!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
 
-<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
+%<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
 Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.
@@
 G(i,j)=G(j,i) = \left\{ \begin{array}{ll}
- \{ d_g (X_i,X_j)\}_{i>j} & \textrm{if $i\neq j$,}\\
- 0 & \textrm{otherwise.}
-  \end{array} \} \right.
+ d_g (X_i,X_j) & \textrm{if $i\neq j$,}\\
+ 0 & \textrm{otherwise,}
+  \end{array} \right.
 $$ 
@@
 
-<!-- Pairwise distances are important for many downstream tasks, e.g.
+% Pairwise distances are important for many downstream tasks, e.g.
 classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression).
@@
 
-## Related Work
-The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data.
+\subsection{Related work}
+
+The work of Chen and Muller \shortcite{ChenMuller2012} was among the first in functional data analysis to consider the manifold hypothesis for functional data.
 A notion of the mean and variation of functional manifold data was introduced.
-Of particular relevance to our work is the modified ISOMAP procedure, called P-ISOMAP, introduced in [@ChenMuller2012], which adds a data-adaptive penalty to allow for noisy functional observations whereas in the classic ISOMAP algorithm, observations are assumed to lie exactly on the manifold.
-Another existing work for estimating the pairwise geodesic distance for functional manifold data can be found in [@Dimeglio2014].
+Of particular relevance to our work is the modified ISOMAP procedure, called P-ISOMAP, introduced in Chen and Muller \shortcite{ChenMuller2012}, which adds a data-adaptive penalty to allow for noisy functional observations whereas in the classic ISOMAP algorithm, observations are assumed to lie exactly on the manifold.
+Another existing work for estimating the pairwise geodesic distance for functional manifold data can be found in Dimeglio et al.
+\shortcite{Dimeglio2014}.
 The estimator is called Robust-ISOMAP for the fact that it is motivated by modifying the ISOMAP algorithm to be less sensitive to outliers.
@@
 
-<!-- [@LinYao2017]  -->
+%<!-- [@LinYao2017]  -->
 
-# Proposed method for estimating geodesic distances
+\section{Proposed method for estimating geodesic distances}
 
@@
 
-Before introducing our procedure, let us first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+Before introducing our procedure, let us first review the ISOMAP algorithm \cite{Tenenbaum2000}, a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
 \begin{enumerate}
@@
 \item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms.
-Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e./ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
-<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
+Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e./ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm \cite{Floyd1962} or with the Dijkstra algorithm \cite{Dijkstra59anote}.
+%<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
 \item Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
@@
 Both P-ISOMAP and robust ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
-<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
+%<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
 
@@
 
-Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm [@Ozertem2011].
-Theoretical justificaton for this can be found in [@Genovese2014].
+Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm \cite{Ozertem2011}.
+Theoretical justificaton for this can be found in Genovese et al.
+\shortcite{Genovese2014}.
 We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
@@
 
-<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
-<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
-<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
+%<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
+%<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
+%<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
 \begin{enumerate}
@@
 Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^K$ Euclidean distances on the discretised functions $\tilde X_i$ are preserved.
-\item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain  $\tilde X^{s,\hat \M}_i$.
+\item Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011} to each of $\tilde X^s_i$ to obtain  $\tilde X^{s,\hat \M}_i$.
 \item Use IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
@@
 
-## Selection of tuning parameters
-Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of [@ChenHo2015].
+\subsection{Selection of tuning parameters}
+Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of \cite{ChenHo2015}.
 
-<!-- ##############  -->
-
-# Simulation study
+\section{Simulation study}
 
@@
 
-* the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis.
+\begin{itemize}
+\item the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis.
 Near-$\epsilon$ isometry is a relaxation of isometry measured by the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
-* the relative Frobenius metric: this is given by
+\item  the relative Frobenius metric: this is given by
 $||d-\hat d||/||d||$
 where $d$ and $\hat d$ are the true and estimated $n{\times}n$ geodesic distance matrices, respectively. 
-* the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
+\item  the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
+\end{itemize}
 
-## Alternative estimators of geodesic distance [Marie]
+\subsection{Alternative estimators of geodesic distance [Marie]}
+
 We compare the performance of our method to the one of the three following methods:
-
-1.
-**Raw Data (RD)** The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+\begin{enumerate}
+\item \emph{Raw Data (RD)} The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
 <!-- to obtain an estimator $\hat G_{\textrm{RD}}$.--> Note that this procedure only makes sense if all the grids $T_1,\ldots,T_n$ are the same.
-2.
-**Spline Smoothing (SS)** The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
-<!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
+\item   \emph{Spline Smoothing (SS)} The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
+%<!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
-3.
-**P-ISOMAP (pI)** The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
-<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->
-
+\item  \emph{P-ISOMAP (pI)} The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
+%<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->
+\end{enumerate}
 We ran Chen and Muller's P-ISOMAP separately in Matlab with their automatic penalty selection and discovered that ???.
 
-## Simulation scenarios 
-<!-- Susan  -->
+\subsection{Simulation scenarios }
+%<!-- Susan  -->
 
-### Isometric functional manifold of normal density functions
-<!-- Scenario 5 -->
-This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$.
+\subsubsection{Isometric functional manifold of normal density functions}
+%<!-- Scenario 5 -->
+This scenario is modified from what is referred to as Manifold 2 in \cite{ChenMuller2012} by fixing the variance of the normal density to be $1$.
 We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
@@
 
-### Functional manifold of square root velocity functions
-It was shown in [@Joshi2007] that the square root representation of probability density functions has a nice closed form geodesic.
+\subsubsection{Functional manifold of square root velocity functions}
+It was shown in \cite{Joshi2007} that the square root representation of probability density functions has a nice closed form geodesic.
 They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
 Note that this concides with the $L_2[0,1]$ inner product.
-[@Joshi2007] showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+\cite{Joshi2007} showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
 We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$.
@@
 
-### Functional manifold of warping functions
-<!-- Scenario 5 -->
-This is based on Equations (17) and (18) of [@Kneip2008] but with $z_{i1}, z_{i2}$ set to $1$.
+\subsubsection{Functional manifold of warping functions}
+%<!-- Scenario 5 -->
+This is based on Equations (17) and (18) of\cite{Kneip2008} but with $z_{i1}, z_{i2}$ set to $1$.
 (Equation 17 has a typo where the exponentials are missing negative signs).Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$ and $h_\alpha(t) = t$ if $\alpha = 0$.
@@
 
-## Sampling from a functional manifold
+\subsection{Sampling from a functional manifold}
 As of yet, there is little work as to how to sample from a functional manifold.
-Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
+Even in the Euclidean case, it is not obvious how sampling should be done \cite{Diaconis2013}.
 To safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
-For both the manifold of normal densities and the manifold of warping functions, we sample $\alpha$ according to ???rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)???
+For both the manifold of normal densities and the manifold of warping functions, we sample $\alpha$ according to ???%rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)???
 For the manifold of square root beta densities, we sample as follows       
-alpha<- rtruncnorm(nb_alpha,alphamin,alphamax,(alphamax+alphamin)/2,0.3)
-beta<- rtruncnorm(nb_beta,betamin,betamax,(betamax+betamin)/2,0.3)
-
+%alpha<-$ rtruncnorm(nb_alpha,alphamin,alphamax,(alphamax+alphamin)/2,0.3)$
+%beta<- $rtruncnorm(nb_beta,betamin,betamax,(betamax+betamin)/2,0.3)$
 How to sample properly from a functional manifold could be interesting future work.
 
-## Results
-Submit compare_methods.sh as slurm job using sbatch.
-```{r, engine = 'bash', eval = FALSE}
-sbatch compare_methods.sh
-```
+\subsection{Results}
 
-Alternatively, we can run the R script compare_methods.R locally by removing the slurm_arrayid chunk
-```{r, eval=FALSE}
-R compare_methods.R
-```
+\subsection{The cost of employing proposed method when manifold is flat}
 
-Simulation results are then visualized by invoking cluster_results.R. There's some unfortunate hardcoding still in cluster_results.R
-```{r, eval=FALSE}
-R cluster_results.R
-```
-<!--
-
-## The cost of employing proposed method when manifold is flat [Susan/Marie?]
-
-Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance.
-Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?
-
--->
-
-# Distance-based functional classification [Marie]
+\section{Distance-based functional classification [Marie]}
 
@@
 It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.
-
 For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
@@
 \item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
-\item Berkeley growth curves in [@ChenReiss2014].
-\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
-\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
+\item Berkeley growth curves in \cite{ChenReiss2014}.
+\item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper.
+\item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data"
 \end{itemize}
@@
 \begin{itemize}
-\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
-\item Tecator in [@LinYao2017] contamination paper
-\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
+\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012}
+\item Tecator in \cite{LinYao2017} contamination paper
+\item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper
 \end{itemize}
 
-<!-- ## Distance-based functional clustering -->
+\bibliographystyle{aaai} 
+\bibliography{references}
 
-<!-- Describe the $k$-medoids clustering method. -->
-
-<!-- ### Simulated data -->
-
-<!-- ###  Real data -->
-
-# References
+\end{document}

### 2019-08-02 compare_methods.R
fixed matlabr package placement?

@@ -8,6 +8,7 @@ library(reticulate)
 library(fields)
 library(fda)
 library(igraph)
+library(matlabr)
 
 source('EuclideanExamples.R')
 source('full_geo_from_adj_geo.R')

### 2019-08-02 pairwise_geo_estimation.R

@@ -56,8 +56,6 @@
 # meth <- list("NN" = TRUE,"RD_o" = TRUE,"RD" = TRUE,"SS_o" = TRUE,"SS" = TRUE,"pI" = FALSE,"OUR" = TRUE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE, "L2" = TRUE, "w_L2" = TRUE )
 # Estim<- pairwise_geo_estimation(meth,data$noiseless_data,data$noisy_data,data$analytic_geo,TRUE,TRUE,20,data$grid,data$reg_grid,100,1,FALSE,FALSE)
 
-library(matlabr)
-
 
 pyIso = import_from_path("getIsomapGdist",path='.') # TODO: consider replacing python Isomap with R implementation of Floyd's algorithm? https://en.wikipedia.org/wiki/Floyd%E2%80%93Warshall_algorithm
 py_min_neigh = import_from_path("get_min_num_neighbors",path='.')

### 2019-08-08 index.Rmd
created workaround to run pisomap in matlab on our sim_functional_data

@@
 # Abstract
-Functional data analysis is the statistical analysis of smooth infinite-dimensional curves. 
+Clustering and classification of data that are curves or surfaces is a common machine learning task.
+Functional data analysis is a branch in statistics dedicated to analyzing objects varying over a continuum, in particular smooth infinite-dimensional curves.
 The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the dimensionality of curves is only artifically high if they are smooth.
@@
 <!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
-In this work we address the estimation of the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near, rather than exactly on, the manifold.
+In this work we address the estimation of the pairwise geodesic distance between noisy observations from a functional manifold.
+The noise is both in the sense of measurement error and that the fact that observations may live near, rather than on, the manifold.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
@@
 The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
-Good estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
+Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
 
 # Introduction 
-Many statistical methods depend on a measure of distance.
-Methods in functional data analysis, the study of infinite-dimensional smooth curves, are no exception.
-Certain classes of functional data may naturally live on manifold, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
-In situations where the manifold hypothesis might be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
+Many statistical and machine learning methods rely on some measure of distance.
+Functional data analysis, the study of infinite-dimensional smooth curves, is no exception.
+Certain classes of functional data naturally live on manifold, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
+In situations where the manifold hypothesis may be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
 The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.
@@
 
-We ran Chen and Muller's P-ISOMAP separately in Matlab with their automatic penalty selection and discovered that ???.
+We ran Chen and Muller's P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric.
+We found that the penalty parameter chosen in this way was always zero, thus reducing hte P-ISOMAP procedure to standard ISOMAP.
+Given this, we did not make further comparison to P-ISOMAP in our simulation study.
 
 ## Simulation scenarios 
-<!-- Susan  -->
+We shall consider three functional manifolds in our simulation study.
+As of yet, there is little work as to how to sample from a functional manifold.
+Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
+How to properly sample from a functional manifold could be interesting future work.
+For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
 
 ### Isometric functional manifold of normal density functions
-<!-- Scenario 5 -->
+<!-- Scenario 2 in sim_functional_data -->
 This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$.
 We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
+We sample $\beta$ according to a truncated standard normal with support on $[-1,1]$.
 We set $a=-4$ and $b=4$.The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
@@
 That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$.
+We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$.
+We sample $\beta$ according to a truncated normal with mean 3.5 and variance 0.09 with support on $[2,5]$.
 
@@
 Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\} $$. 
+We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
 The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$
-
-## Sampling from a functional manifold
-As of yet, there is little work as to how to sample from a functional manifold.
-Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
-To safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
-For both the manifold of normal densities and the manifold of warping functions, we sample $\alpha$ according to ???rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)???
-For the manifold of square root beta densities, we sample as follows       
-alpha<- rtruncnorm(nb_alpha,alphamin,alphamax,(alphamax+alphamin)/2,0.3)
-beta<- rtruncnorm(nb_beta,betamin,betamax,(betamax+betamin)/2,0.3)
-
-How to sample properly from a functional manifold could be interesting future work.
 
@@
 sbatch compare_methods.sh
-```
-
-Alternatively, we can run the R script compare_methods.R locally by removing the slurm_arrayid chunk
-```{r, eval=FALSE}
-R compare_methods.R
 ```

### 2019-08-08 compare_methods.R

@@ -5,6 +5,7 @@
 
 # library(DescTools)
 library(reticulate)
+use_python("/Users/suswei/anaconda3/bin/python",required=TRUE)
 library(fields)
 library(fda)
 library(igraph)
@@ -18,8 +19,6 @@ source('pairwise_geo_estimation.R')
 source('sim_Euclidean_data.R')
 source('assess_goodness_estimation.R')
 source('robust_isomap.R')
-
-
 
 # parameters that will be fixed during the sim study
 com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
@@ -37,10 +36,9 @@ Ks_obs = c(100,30)
 
 total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(mcs)
 
-# slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
-# slurm_arrayid = as.numeric(slurm_arrayid)
-# print(slurm_arrayid)
-slurm_arrayid = 1
+slurm_arrayid <- Sys.getenv('SLURM_ARRAY_TASK_ID')
+slurm_arrayid = as.numeric(slurm_arrayid)
+print(slurm_arrayid)
 
 # Hardcode
 unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(mcs)))
@@ -69,11 +67,11 @@ meth <- list("NN" = FALSE,
              "RD_o" = FALSE,
              "RD" = FALSE,
              "SS_o" = FALSE,
-             "SS" = TRUE,
-             "pI" = TRUE,
+             "SS" = FALSE,
+             "pI" = FALSE,
              "OUR" = FALSE,
              "OUR2" = FALSE,
-             "OUR3"=TRUE,
+             "OUR3"=FALSE,
              "RP" = FALSE,
              "L2" = TRUE, 
              "w_L2" = FALSE )# see pairwise_geo_estimation for more info

### 2019-08-08 compare_methods.sh

@@ -40,7 +40,7 @@ fi
 # Run the job from the directory where it was launched (default)
 
 # The modules to load:
-# module load MATLAB
+module load MATLAB
 
 # The job command(s):
 R --vanilla < compare_methods.R

### 2019-08-08 createSimTextFiles_pisomap.R

@@ -1,0 +1,88 @@
+# Title     : compare_methods.R
+# Objective : compare multiple estimators of geodesic distance on simulated data. this is called by compare_methods.sh!
+# Created by: suswei
+# Created on: 3/12/18
+
+# library(DescTools)
+library(reticulate)
+use_python("/Users/suswei/anaconda3/bin/python",required=TRUE)
+library(fields)
+library(fda)
+library(igraph)
+library(matlabr)
+
+source('EuclideanExamples.R')
+source('full_geo_from_adj_geo.R')
+source('sim_functional_data.R')
+source('Geo_estimation.R')
+source('pairwise_geo_estimation.R')
+source('sim_Euclidean_data.R')
+source('assess_goodness_estimation.R')
+source('robust_isomap.R')
+
+
+
+# parameters that will be fixed during the sim study
+com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a common grid
+plotTrue = FALSE
+samplesize = 100
+Ks_smooth = 100
+
+## STUDY THE EFFECT OF THE FOLLOWING PARAMETERS 
+mcs = 1:10
+# parameters in sim_functional_data
+sces = c(5,2,4)
+SNRs = c(0.1,0.5)
+reg_samplings = c(TRUE,FALSE)
+Ks_obs = c(100)
+
+total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(mcs)
+
+
+for(slurm_arrayid in 1:total_tasks){
+  
+  # Hardcode
+  unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(mcs)))
+  
+  # actual parameters for this run
+  sce = sces[unravel[1,1]]
+  SNR = SNRs[unravel[1,2]]
+  reg_sampling = reg_samplings[unravel[1,3]]
+  K_obs = Ks_obs[unravel[1,4]]
+  mc = mcs[unravel[1,5]]
+  
+  
+  # Generate data
+  data<- sim_functional_data(samplesize = samplesize, 
+                             com_grid = com_grid, 
+                             plot_true = plotTrue,
+                             ### parameters under study
+                             sce = sce, 
+                             SNR = SNR, 
+                             reg_sampling = reg_sampling,
+                             K = K_obs)
+  
+  
+  # Estimation of geodesic distances with different methods
+  meth <- list("NN" = FALSE,
[…]

### 2019-08-08 pairwise_geo_estimation.R

@@ -7,6 +7,8 @@
 # SS_o: Floyd's Algorithm on smooth data (oracle for num neigh)
 # SS: Floyd's Algorithm on smooth data (no oracle)
 # pI: p-isomap of Muller on smooth data
+  # TODO: pI as implemented below still doesn't work with the txt files writing and reading. this is mainly due to the fact that matlabr doesn't allow you to pass in function arguments
+  # Susan made a work around where we first run creatSimTextFiles_pisomap.R in R and then run "run_isomap_p.m" in Matlab. Most delta's, chosen in the oracle way, are chosen to be zero.
 # OUR: mds of smooth data + scms + Floyd's Algorithm (oracle for h and num neigh)
 # OUR2: mds of smooth data + scms + robust isomap (oracle for h)
 # OUR3: mds of smooth data + scms h heuri + robust isomap (no oracle)
@@ -242,25 +244,25 @@ pairwise_geo_estimation <- function(method,
     write.table(num_neigh,file=paste0('./',string,'/possible_K.txt'),col.names = FALSE, row.names = FALSE)
     write.table(c(samplesize,K),file=paste0('./',string,'/n_K.txt'),col.names = FALSE, row.names = FALSE)
     write.table(true_geo,file=paste0('./',string,'/true_geo.txt'),col.names = FALSE, row.names = FALSE)
-  
-    file.copy("./run_isomap_p.m",paste0('./',string,'/'))
-    run_matlab_script(paste0('./',string,'/run_isomap_p.m'))
-  
-    temp=as.vector(read.table(file=paste0('./',string,'/manidis.txt'),header=FALSE))
-    estim_geo_penalized_isomap=matrix(temp[,1],ncol=samplesize)
-    temp2=read.table(file=paste0('./',string,'/err_delta_neigh.txt'),header=FALSE)
-    err_delta_neigh=matrix(temp2[,1],ncol=length(num_neigh))
-    delta_min_tmp=which(err_delta_neigh==min(err_delta_neigh),arr.ind = TRUE)
-    delta_min=delta_can[delta_min_tmp[1,1]]
-  
-    list_to_return[["estim_geo_pI"]]<- estim_geo_penalized_isomap
-    list_to_return[["delta_min"]]<- delta_min
-    
-    unlink(string,recursive=TRUE)
-    
-    if(plot_true){
[…]

### 2019-08-08 penalized_Isomap.m

@@ -1,3 +1,4 @@
+% Marie adpated this from maniMethods.m in PACE package
 function [Manidis] = penalized_Isomap (Dis,K,delta)
     N = size(Dis,1);    
     % initial step: connect K nearest points

### 2019-08-08 run_isomap_p.m

@@ -1,49 +1,66 @@
+% takes all sim_functional_data outputs and runs automatic tuning for
+% pisomap. records how many of the chosen deltas = 0
+% first run createSimTextFiles_pisomap.R to create data directories, than
+% run_isomap_p in matlab
 
+S=dir('2019*')
+for k = 1:numel(S)
+    try
+        fnm = fullfile(S(k).name,'n_K.txt');  
+        file = fopen(fnm,'r') ;
+        temp = fscanf(file,'%f');
+        fclose(file);
+        n=temp(1);
+        K=temp(2);
 
-file = fopen('./n_K.txt','r') ;
-temp = fscanf(file,'%f');
-fclose(file)
-n=temp(1);
-K=temp(2);
+        fnm = fullfile(S(k).name,'possible_K.txt');  
+        file3 = fopen(fnm,'r') ;
+        num_neigh = fscanf(file3,'%f');
+        fclose(file3);
 
-file3 = fopen('./possible_K.txt','r') ;
-num_neigh = fscanf(file3,'%f');
-fclose(file3)
+        fnm = fullfile(S(k).name,'true_geo.txt');  
+        file2 = fopen(fnm,'r') ;
+        geo_temp = fscanf(file2,'%f');
+        fclose(file2);
+        true_geo=reshape(geo_temp,n,n);
 
-file2 = fopen('./true_geo.txt','r') ;
-geo_temp = fscanf(file2,'%f');
-fclose(file2)
-true_geo=reshape(geo_temp,n,n);
+        fnm = fullfile(S(k).name,'discrete_data.txt');  
+        file = fopen(fnm,'r') ;
+        X_temp = fscanf(file,'%f');
+        fclose(file);
+        data = reshape(X_temp,K,n)';
 
-file = fopen('./discrete_data.txt','r') ;
-X_temp = fscanf(file,'%f');
-fclose(file)
-data = reshape(X_temp,K,n)';
+        fnm = fullfile(S(k).name,'possible_delta.txt');  
+        file = fopen(fnm,'r') ;
+        delta_can = fscanf(file,'%f');
+        fclose(file);
 
-file = fopen('./possible_delta.txt','r') ;
-delta_can = fscanf(file,'%f');
-fclose(file)
+        Dis= L2_distance(data',data',1); %*sqrt(range(grid)/(K-1));
+        norm_true_geo = norm(true_geo,'fro');
 
-Dis= L2_distance(data',data',1); %*sqrt(range(grid)/(K-1));
[…]

### 2019-08-09 index.Rmd
more writing

@@
 # Abstract
-Clustering and classification of data that are curves or surfaces is a common machine learning task.
-Functional data analysis is a branch in statistics dedicated to analyzing objects varying over a continuum, in particular smooth infinite-dimensional curves.
-The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the dimensionality of curves is only artifically high if they are smooth.
-There is also another sense in which the dimensionality of the curves under study may be low. 
-This is the much less explored idea that the curves may actually live in a submanifold with low intrinsic dimension.
+The clustering and classification of discretely-sampled data varying over a continuu, e.g.\ time or space, is an important machine learning task. Data of this nature are also known as functional data.
+While one can treat each data point as a vector in Euclidean space, the successful development of functional data analysis demonstrates that it worthwhile to take the functional perspective.
+Specifically, functional data analysis is a branch of statistics dedicated to the analysis of infinite-dimensional \textit{smooth} curves.
+The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the curves are smooth since the dimensionality is then only artifically high.
+This insight underpins functional data analysis.
+There is also another, much less explored, sense in which the dimensionality of the curves under study may be low -- the curves may actually live on a submanifold with low intrinsic dimension.
 Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
 <!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
-In this work we address the estimation of the pairwise geodesic distance between noisy observations from a functional manifold.
-The noise is both in the sense of measurement error and that the fact that observations may live near, rather than on, the manifold.
+This work addresses the estimation of the pairwise geodesic distance between functional manifold data, observed with noise due to both measurement error and the fact that observations may live near, rather than on, the manifold.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
 <!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
-The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
+The proposed methodology first projects the functional data onto the hidden manifold, estimated using subspace-constrained mean-shift.
+Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold.
 Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
@@
 Many statistical and machine learning methods rely on some measure of distance. 
-Functional data analysis, the study of infinite-dimensional smooth curves, is no exception.
-Certain classes of functional data naturally live on manifold, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
-In situations where the manifold hypothesis may be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
+When the observations are functions, it is especially important to use an appropriate measure. 
+Functional data analysis, the study of smooth infinite-dimensional curves, is no exception.
+Certain classes of functional data naturally live on manifolds, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
+In situations where the manifold hypothesis may be plausible, one might consider using a geodesic distance rather than the $L^2$ distance.
 The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.
@@
 Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a low-dimensional nonlinear Riemannian manifold $\M \subset L^2([a,b],\R)$ where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold as follows [@Lin2014].
+Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014].
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
 The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_g(X_i,X_j):=\inf \{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth}, \gamma(a) = X_i, \gamma(b) = X_j]  \}$$
+$$ d_{\M}(X_i,X_j):=\inf \{l(\gamma): \gamma \text{ is piecewise-smooth function from } [a,b] \text{ to } \M, \gamma(a) = X_i, \gamma(b) = X_j \}$$
 where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
-is the length of a smooth curve $\gamma: [a,b] \subset \R \to \M$.
+is the length of the curve $\gamma$.
 
@@
 
-Given perfect functional observations $X_i$ and $X_j$, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance $d_g(X_i,X_j)$ can be accomplished straightforwardly using a shortest-path algorithm, e.g.
-the Floyd-Warhsall algorithm [@Floyd1962].
-However, when functional data are observed with noise, this approach is likely to fail. This is because the error may be such that the noisy functional manifold data no longer live on a manifold. 
-A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
+Given perfect functional observations $X_i$ and $X_j$, i.e.\ observed on a very dense domain grid with no measurement error, the estimation of the geodesic distance $d_g(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
+However, this approach criticaly assumes observations live on the manifold and is thus likely to fail for functional data which are almost certainly observed with noise.
+<!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
 This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
-<!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
-
+This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs.
+-->
 <!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
@@
 -->
-
 In this work, we put forth a technique for the specific task of recovering the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
 $$
-G(i,j)=G(j,i) = \left\{ \begin{array}{ll}
- \{ d_g (X_i,X_j)\}_{i>j} & \textrm{if $i\neq j$,}\\
+G(i,j)=G(j,i) = \begin{cases} 
+      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
       0 & \textrm{otherwise.}
-  \end{array} \} \right.
+   \end{cases}
 $$ 
-when we have, in lieu of $X_i$ and $X_j$, access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
-This is accomplished by first projecting the functional data on to the underlying hidden manifold and then performing operations on this hidden manifold.
+In lieu of $X_i$ and $X_j$, we have access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
 
@@
 
-## Related Work
 The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data.
-A notion of the mean and variation of functional manifold data was introduced.
-Of particular relevance to our work is the modified ISOMAP procedure, called P-ISOMAP, introduced in [@ChenMuller2012], which adds a data-adaptive penalty to allow for noisy functional observations whereas in the classic ISOMAP algorithm, observations are assumed to lie exactly on the manifold.
-Another existing work for estimating the pairwise geodesic distance for functional manifold data can be found in [@Dimeglio2014].
-The estimator is called Robust-ISOMAP for the fact that it is motivated by modifying the ISOMAP algorithm to be less sensitive to outliers.
-We will discuss both P-ISOMAP and robust ISOMAP in more details in the simulation section.
+A notion of the mean and variation of functional manifold data was introduced there.
+Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a data-adaptive penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold.
+This very drawback of ISOMAP led [@Dimeglio2014] to propose the robust-ISOMAP, a version of ISOMAP that is less sensitive to outliers.
+We will discuss P-ISOMAP and robust-ISOMAP in more details in the simulation section.
 
@@
 # Proposed method for estimating geodesic distances
-
 Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
 we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
-We assume that each observation grid $T_1,\ldots,T_n$ is dense.
+We assume that each observation grid $T_1,\ldots,T_n$ is dense ???add reference for dense???.
 
@@
 Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
-For example we may employ spline smoothing to recover the functional versions of the raw data, i.e.\
+For example we may employ spline smoothing ???add reference for spline smoothing???
+to recover the functional versions of the raw data, i.e.\
 \begin{equation}\label{eq_spline_smoothing} 
-\tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
+\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
 \end{equation}
@@
 
-Before introducing our procedure, let us first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+Before introducing our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+The procedure is as follows.
 \begin{enumerate}
 \item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
-Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ if the distance $d_{ij}=||x_i-x_j ||_{2}$ is smaller than a given value $\epsilon$, and the weight associated to an edge $e_{ij}$ is $d_{ij}$.
+Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
 \item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms.
-Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e./ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
 <!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
-\item Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
+\item Use multidimensional scaling ???add reference??? to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
 \end{enumerate}

### 2019-08-11 index.Rmd
more writing, trying to get raw sim outputs for final sim configuration

@@
 # Abstract
-The clustering and classification of discretely-sampled data varying over a continuu, e.g.\ time or space, is an important machine learning task. Data of this nature are also known as functional data.
-While one can treat each data point as a vector in Euclidean space, the successful development of functional data analysis demonstrates that it worthwhile to take the functional perspective.
-Specifically, functional data analysis is a branch of statistics dedicated to the analysis of infinite-dimensional \textit{smooth} curves.
-The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the curves are smooth since the dimensionality is then only artifically high.
-This insight underpins functional data analysis.
-There is also another, much less explored, sense in which the dimensionality of the curves under study may be low -- the curves may actually live on a submanifold with low intrinsic dimension.
-Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
+The clustering and classification of functional data, discretely-sampled data varying over a continuum, is frequently encountered in machine learning.
+While one can naively treat the data as vectors in Euclidean space, one may also regard each data point as a function.
+The challenge of analyzing infinite-dimensional functions can be ameliorated if the curves are smooth since the dimensionality is then only artificially high. 
+This insight underpins functional data analysis, a subfield of statistics that studies functional data.
+For certain types of functional data such as classes of probability density functions and classes of warped curves of a common template function, there is further structure to exploit.
+Namely, when functional data lie on a manifold, the dimensionality at play might be even lower if the manifold is itself of low intrinsic dimension.
 <!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
-This work addresses the estimation of the pairwise geodesic distance between functional manifold data, observed with noise due to both measurement error and the fact that observations may live near, rather than on, the manifold.
+This work addresses the estimation of pairwise geodesic distances between functional manifold data that are observed with noise and which may live near, rather than on, a manifold.
 <!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
 <!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
-The proposed methodology first projects the functional data onto the hidden manifold, estimated using subspace-constrained mean-shift.
+The proposed methodology first sends the observed functional data to the hidden manifold, estimated using subspace-constrained mean-shift. 
 Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold. 
-Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
+Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distance-based functional classification.
 
@@
 Many statistical and machine learning methods rely on some measure of distance.
-When the observations are functions, it is especially important to use an appropriate measure. 
-Functional data analysis, the study of smooth infinite-dimensional curves, is no exception.
-Certain classes of functional data naturally live on manifolds, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
-In situations where the manifold hypothesis may be plausible, one might consider using a geodesic distance rather than the $L^2$ distance.
-The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.
+For example, clustering and classification of functional data, discretely-sampled data varying over a continuum, is a common machine learning task in which distance plays a crucial role.
+The successful development of functional data analysis, a branch of statistics dedicated to the analysis of infinite-dimensional smooth curves, demonstrates that employing the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between functional observations treated as vectors.
+In this work, we advocate that for functional manifold data, i.e.
+functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification of functional manifold data.
+In other words, when the manifold hypothesis is plausible, e.g.\ classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L_2$ distance.
 
-Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
 Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014].
@@
 where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
-is the length of the curve $\gamma$.
+is the length of the curve $\gamma$. ???Is it clear how $l(\gamma)$ depends on $g$???
 
@@
 
-Given perfect functional observations $X_i$ and $X_j$, i.e.\ observed on a very dense domain grid with no measurement error, the estimation of the geodesic distance $d_g(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
-However, this approach criticaly assumes observations live on the manifold and is thus likely to fail for functional data which are almost certainly observed with noise.
+Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error.
+Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
+However, most shortest-path algorithms critically assume observations live on the manifold ???add reference???
+and are thus likely to fail for functional data which are almost certainly observed with noise.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
@@
 -->
-In this work, we put forth a technique for the specific task of recovering the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
+In this work, we put forth a technique for the recovery of the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
 $$
@@
 $$ 
-In lieu of $X_i$ and $X_j$, we have access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
+In lieu of $X_i$ and $X_j$, we have access only to discretely-sampled noisy functional observations $Y_i$ and $Y_j$ that possibly lie off the true manifold.
 
@@
 A notion of the mean and variation of functional manifold data was introduced there.
-Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a data-adaptive penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold.
-This very drawback of ISOMAP led [@Dimeglio2014] to propose the robust-ISOMAP, a version of ISOMAP that is less sensitive to outliers.
+Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold.
+[@Dimeglio2014] also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP that is less sensitive to outliers.
 We will discuss P-ISOMAP and robust-ISOMAP in more details in the simulation section.
@@
 # Proposed method for estimating geodesic distances
-Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
-we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
-We assume that each observation grid $T_1,\ldots,T_n$ is dense ???add reference for dense???.
+Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
+We assume that each observation grid $T_1,\ldots,T_n$ is dense, i.e.\ $K$ is large.
 
-We begin by converting the discrete functional observations into continuous ones.
+We begin by converting the discretly-sampled functional observations into continuous ones.
 Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
-Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
-For example we may employ spline smoothing ???add reference for spline smoothing???
+For example, we may employ spline smoothing ???add reference for spline smoothing???
 to recover the functional versions of the raw data, i.e.\
@@
 where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
+The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
 
-Before introducing our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm [@Ozertem2011].
+Let $\hat M \subset L^2([a,b],\mathbb R)$ denote the collection of basins of attraction of the estimated probability density function of $X$.
+The SCMS algorithm sends each point $\tilde X_i \in L^2([a,b],\mathbb R)$ to a point in $\hat M$.
+This destination is unique and we denote it $\tilde X_i^{\hat \M}$.
+We expect the points $\tilde X^{\hat \M}_1,\ldots,\tilde X^{\hat \M}_n \in \hat M$ to lie close to the real manifold $\M$.
+Theoretical justificaton that $\hat M$ as estimated by SCMS is a reasonable surrogate for the true manifold $\M$ can be found in [@Genovese2014].
+
+Before presenting our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold $\M$ of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
 The procedure is as follows.
@@
 Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
-\item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms.
+\item Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
 Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
-<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
 \item Use multidimensional scaling ???add reference??? to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
 \end{enumerate}
-Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves.
-Formulating the graph in step 1 of ISOMAP is in fact the most challenging asepct of the algorithm since it depends heavily on the choice of the tuning parameter $\epsilon$.
-Both P-ISOMAP and robust ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
+Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weigthed graph but they do not produce the weighted graphs themselves.
+Both P-ISOMAP and robust-ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in step 1.
 <!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
 
-In what follows, we call IsoGeo the procedure which performs a modified version of the first step of ISOMAP, proposed in @Dimeglio2014, followed by the original second step of ISOMAP.
-The first step of IsoGeo modifies Step 1 of ISOMAP so that the constructed graph $G$ is independent of the tuning parameter $\epsilon$:
-1'.
-Construct the complete weigthed graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
-Obtain the minimal spanning tree $G_s$ associated with $G_c$ and denote its set of edges by $E_s$.
+In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP.
+The first step of IsoGeo constructs the weighted graph $G$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$.
+Specifically, we begin by constructing the complete ???explain what complete means???
+weighted graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$.
 The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
@@
 where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
+Let the distance estimated by IsoGeo be denoted $\text{IsoGeo}(x_i,x_j)$.
 Note that IsoGeo has no tuning parameters.
 
-Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm [@Ozertem2011].
-Theoretical justificaton for this can be found in [@Genovese2014].
-We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
-Our procedure is described by the following steps:
+We are now ready to describe our procedure.
+Since SCMS is based on kernel density estimation, we first reduce the dimension of our data with multidimensional scaling before applying SCMS to avoid the curse of dimensionality.
 
@@
 \begin{enumerate}
-\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a common grid $\tilde T=(t_{1},\ldots,t_{K})$.
-Let $s$ be a positive integer, smaller than $K$.
-Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^K$ Euclidean distances on the discretised functions $\tilde X_i$ are preserved.
+\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$.
+Let $s$ be a positive integer, much smaller than $D$.
+Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ are preserved.
 \item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
-\item Use IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
+\item Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
 \end{enumerate}
-Thus our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
+Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
 $$
 \hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
- d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
+ \text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
  0 & \textrm{otherwise.}
@@
 $$
-Since ridge estimation suffers from the curse of dimensionality, we first reduce the dimension of our data with multidimensional scaling in Step 1 before applying the mean-shift algorithm.
 
-## Selection of tuning parameters
-Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of [@ChenHo2015].
+There are two tuning parameters to our procedure.
+First, there is the dimension $s$ in step 1.
+In our simulations we try $s \in \{1,2,3\}$.
+The other parameter is the bandwidth $h$ in the subspace constrained mean-shift.
+In our simulations we set $h$ according to the heuristic in Equation (A1) of [@ChenHo2015].
+In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.
 
@@
 
-We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances when one only has access to discretely-observed noisy functional data.
+We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances.
 Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.
 
-* the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis.
-Near-$\epsilon$ isometry is a relaxation of isometry measured by the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
-* the relative Frobenius metric: this is given by
-$||d-\hat d||/||d||$
-where $d$ and $\hat d$ are the true and estimated $n{\times}n$ geodesic distance matrices, respectively. 
-* the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
+The first metric assesses near-$\epsilon$ isometry, which we define as the percentage of estimated pairwise geodesic distances between $1-\epsilon$ and $1+\epsilon$ of the corresponding true pairwise geodesic distance.
+We form the receiver operating curve with $\epsilon$ on the $x$-axis and near-$\epsilon$ isometry on the $y$-axis.
+The metric is the area under the receiver operating curve.
+
+The second metric $||d-\hat d||/||d||$ is the Frobenius norm of the estimation error matrix $d-\hat d$, relative to the Frobenius norm of the true distance matrix $d$.
+The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
 
 ## Alternative estimators of geodesic distance [Marie]
-We compare the performance of our method to the one of the three following methods:
+We compare the performance of our method to each of the following four methods.
+The first method, denoted RD for raw data, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+Note that this procedure is only sensible if all the grids $T_1,\ldots,T_n$ are the same.
+In the second method, we first smooth the discretely-sampled noisy functional data and then estimate their pairwise distance using the $L^2$ distance.
 
-1.
-**Raw Data (RD)** The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
-<!-- to obtain an estimator $\hat G_{\textrm{RD}}$.--> Note that this procedure only makes sense if all the grids $T_1,\ldots,T_n$ are the same.
-2.
-**Spline Smoothing (SS)** The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
+The third alternative, denoted SS for spline smoothing, applies IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
 <!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
 Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
-3.
-**P-ISOMAP (pI)** The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
+The final alternative we consider is P-ISOMAP, denoted pI, which is the two-step procedure developed in @ChenMuller2012 designed to handle noisy functional data.
+In the first step, a penalty is incorporated to robustify the construction of the weighted graph.
+This is followed by the original second step of ISOMAP.
 <!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->
@@
 
-### Isometric functional manifold of normal density functions
+\textbf{Isometric functional manifold of normal density functions}
 <!-- Scenario 2 in sim_functional_data -->
@@
 We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
-We sample $\beta$ according to a truncated standard normal with support on $[-1,1]$.
-We set $a=-4$ and $b=4$.The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
+First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$.
+Thus the geodesic distance will not coincide with the $L^2$ distance.
+
+We set $a=-4$ and $b=4$ and sample $\beta$ according to a truncated standard normal with support on $[-1,1]$.   The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
 \begin{eqnarray*}
@@
 where the approximation comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
-We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is the Euclidan distance between the $\beta$'s, up to some scaling factor.
-Note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$, so we cannot employ the calculation technique of Scenario 1.
+We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$, the Euclidean distance between the $\beta$'s.
 
-### Functional manifold of square root velocity functions
-It was shown in [@Joshi2007] that the square root representation of probability density functions has a nice closed form geodesic.
+\textbf{Functional manifold of square root velocity functions}
+<!-- Scenario 4 in sim_functional_data -->
+It was shown in @Joshi2007 that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
 They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
 Note that this concides with the $L_2[0,1]$ inner product.
-[@Joshi2007] showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+@Joshi2007 showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
 We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$.
@@
 
-### Functional manifold of warping functions
-<!-- Scenario 5 -->
-This is based on Equations (17) and (18) of [@Kneip2008] but with $z_{i1}, z_{i2}$ set to $1$.
-(Equation 17 has a typo where the exponentials are missing negative signs).Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$ and $h_\alpha(t) = t$ if $\alpha = 0$.
-Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\} $$. 
+\textbf{Functional manifold of warping functions}
+<!-- Scenario 5 in sim_functional_data -->
+This is based on Equations 17 and 18 in @Kneip2008 but with $z_{i1}, z_{i2}$ both set to $1$.
+(Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
+
+Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\}.$$ 
 We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
-The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$
+The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha$$ which is estimated using numerical integration.
 
 ## Results
-Submit compare_methods.sh as slurm job using sbatch.
+???Still need to describe parameters under study: SNR, denseness of observations ($K_{observed}$).
+Describe parameters that are fixed:  $K_smooth$ and com_grid.
+Describe tuning parameters for our method: $s$ and $h$.
+
 ```{r, engine = 'bash', eval = FALSE}
@@
 
-Simulation results are then visualized by invoking cluster_results.R. There's some unfortunate hardcoding still in cluster_results.R
 ```{r, eval=FALSE}

### 2019-08-11 compare_methods.R

@@ -25,13 +25,13 @@ com_grid = 1 # 1 or 0 to indicate if yes or no each curve is observed on a commo
 plotTrue = FALSE
 samplesize = 100
 Ks_smooth = 100
+reg_samplings = TRUE
 
 ## STUDY THE EFFECT OF THE FOLLOWING PARAMETERS 
 mcs = 1:100
 # parameters in sim_functional_data
 sces = c(5,2,4)
 SNRs = c(0.1,0.5)
-reg_samplings = c(TRUE,FALSE)
 Ks_obs = c(100,30)
 
 total_tasks = length(sces)*length(SNRs)*length(reg_samplings)*length(Ks_obs)*length(mcs)
@@ -41,24 +41,23 @@ slurm_arrayid = as.numeric(slurm_arrayid)
 print(slurm_arrayid)
 
 # Hardcode
-unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(reg_samplings), length(Ks_obs),length(mcs)))
+unravel=arrayInd(slurm_arrayid,c(length(sces), length(SNRs), length(Ks_obs),length(mcs)))
 
 # actual parameters for this run
 sce = sces[unravel[1,1]]
 SNR = SNRs[unravel[1,2]]
-reg_sampling = reg_samplings[unravel[1,3]]
-K_obs = Ks_obs[unravel[1,4]]
-mc = mcs[unravel[1,5]]
+K_obs = Ks_obs[unravel[1,3]]
+mc = mcs[unravel[1,4]]
 
 
 # Generate data
 data<- sim_functional_data(samplesize = samplesize, 
                            com_grid = com_grid, 
                            plot_true = plotTrue,
+                           reg_sampling = reg_sampling,
                            ### parameters under study
                            sce = sce, 
                            SNR = SNR, 
-                           reg_sampling = reg_sampling,
                            K = K_obs)
 
 
@@ -67,11 +66,11 @@ meth <- list("NN" = FALSE,
              "RD_o" = FALSE,
              "RD" = FALSE,
              "SS_o" = FALSE,
-             "SS" = FALSE,
+             "SS" = TRUE,
              "pI" = FALSE,
              "OUR" = FALSE,
-             "OUR2" = FALSE,
-             "OUR3"=FALSE,
+             "OUR2" = TRUE,
[…]

### 2019-08-11 compare_methods.sh

@@ -24,9 +24,9 @@
 #SBATCH --mail-user=susan.wei@unimelb.edu.au
 
 # The maximum running time of the job in days-hours:mins:sec
-#SBATCH --time=0-3:00:00
+#SBATCH --time=0-0:05:00
 
-#SBATCH --array=1-2400         ### Array index | %50: number of simultaneously tasks
+#SBATCH --array=1-1200         ### Array index | %50: number of simultaneously tasks
 
 # Get Array ID
 i=${SLURM_ARRAY_TASK_ID}
@@ -40,7 +40,7 @@ fi
 # Run the job from the directory where it was launched (default)
 
 # The modules to load:
-module load MATLAB
+# module load MATLAB
 
 # The job command(s):
 R --vanilla < compare_methods.R

### 2019-08-22 index.Rmd
rmarkdown outputs to body

@@
   html_document:
-    df_print: paged
-    number_sections: yes
-  word_document: default
-  pdf_document:
-    keep_tex: yes
-    number_sections: yes
+    fig_caption: true
 bibliography: references.bib
 ---
+
+```{r, include=FALSE}
+source("cluster_results.R",echo=FALSE)
+```
 
@@
 ## Results
+
 ???Still need to describe parameters under study: SNR, denseness of observations ($K_{observed}$).
@@
 
-```{r, engine = 'bash', eval = FALSE}
+```{r, engine = 'bash', eval = FALSE, include=FALSE}
 sbatch compare_methods.sh
+scp susanwei@spartan.hpc.unimelb.edu.au:/home/susanwei/functional_manifold_data/taskid* /Documents/Githuib/functional_manifold_data/
 ```
 
-```{r, eval=FALSE}
-R cluster_results.R
+```{r, echo=FALSE, eval=TRUE, results='hide'}
+
+setwd("./sim_results")
+
+method_names=c("SS","s=1","s=2","s=3","L2")
+nr_methods = length(method_names)
+
+combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
+# TODO: fix hardcoding
+combination_para <- expand.grid(c(5,2,4),c(0.5),c(30)) # combinations of the parameters, copied from compare_methods.R
+
+apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)
+
+setwd("../")
 ```
+
+\begin{figure}
+\includegraphics{sim_results/sce=5_SNR=high_Kobs=30}
+\caption{blah}
+\end{figure}
+
 <!--

### 2019-08-22 index.tex
rmarkdown outputs to body

@@
-\documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={The hidden manifold distance for functional data},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{The hidden manifold distance for functional data}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
-\begin{document}
-\maketitle
-
 \newcommand {\To}{\rightarrow}
@@
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\section{Abstract}\label{abstract}
 
-Functional data analysis is the statistical analysis of smooth
-infinite-dimensional curves. The challenge of analyzing
-infinite-dimensional objects is ameliorated by the fact that, since
-observations are smooth over their domain, the dimensionality of the
-curves is only artifically high. Less studied in functional data
-analysis is the idea that the curves may actually live in a submanifold
-with low intrinsic dimension. Interesting classes of functional manifold
-data include the class of probability density functions and the class of
-warped curves of a common template function. Since these spaces are not
-vector spaces, even basic operations such as addition and subtraction
-require special consideration. In this work we set ourselves the task of
-calculating the pairwise geodesic distance between functional
-observations that, due to noise, live near a non-linear manifold rather
-than exactly on the manifold. We will show that good estimation of the
-geodesic distance cannot be accomplished by naive methods such as
-smoothing followed by a shortest-path algorithm. Instead, we propose a
-dedicated technique for pairwise geodesic distance estimation by first
-projecting the functional data to the underlying hidden manifold and
-then perofrming operations on this hidden manifold. We conclude the work
-by showing the implications for many downstream tasks in functional data
-analysis, in particular distanced-based functional classification.
+The clustering and classification of functional data, discretely-sampled
+data varying over a continuum, is frequently encountered in machine
+learning. While one can naively treat the data as vectors in Euclidean
+space, one may also regard each data point as a function. The challenge
+of analyzing infinite-dimensional functions can be ameliorated if the
+curves are smooth since the dimensionality is then only artificially
+high. This insight underpins functional data analysis, a subfield of
+statistics that studies functional data. For certain types of functional
+data such as classes of probability density functions and classes of
+warped curves of a common template function, there is further structure
+to exploit. Namely, when functional data lie on a manifold, the
+dimensionality at play might be even lower if the manifold is itself of
+low intrinsic dimension. This work addresses the estimation of pairwise
+geodesic distances between functional manifold data that are observed
+with noise and which may live near, rather than on, a manifold. The
+proposed methodology first sends the observed functional data to the
+hidden manifold, estimated using subspace-constrained mean-shift.
+Geodesic distances are subsequently calculated by employing
+shortest-path algorithms on this estimated manifold. Improved estimation
+of the pairwise geodesic distance has beneficial implications for many
+downstream tasks in functional data analysis which we illustrate in the
+case of distance-based functional classification.
 
-\hypertarget{introduction-susan}{%
-\section{Introduction {[}Susan{]}}\label{introduction-susan}}
+\section{Introduction}\label{introduction}
+
+Many statistical and machine learning methods rely on some measure of
+distance. For example, clustering and classification of functional data,
+discretely-sampled data varying over a continuum, is a common machine
+learning task in which distance plays a crucial role. The successful
+development of functional data analysis, a branch of statistics
+dedicated to the analysis of infinite-dimensional smooth curves,
+demonstrates that employing the \(L^2\) distance between the smoothed
+functional observations has advantages over the Euclidean distance
+between functional observations treated as vectors. In this work, we
+advocate that for functional manifold data, i.e.~functional data that
+lie on a manifold, the geodesic distance is potentially even better than
+the \(L^2\) distance, with benefits realized in downstream tasks such as
+distanced-based clustering and classification of functional manifold
+data. In other words, when the manifold hypothesis is plausible,
+e.g.~classes of probability density functions and classes of warped
+curves of a common template function, clustering and classification
+using the goedesic distance may give better results than using the
+\(L_2\) distance.
 
@@
 a random variable \(X\) that takes value in the Hilbert space
-\(L^2([a,b],\mathbb{R})\). We additionally assume that the function
-\(X\) belongs to a low-dimensional nonlinear manifold
-\(\mathcal{M}\subset L^2([a,b],\mathbb{R})\), in which case a geodesic
-distance taking into account the intrinsic structure of \(\mathcal{M}\)
-rather than the \(L^2\) distance should be used as a measure of distance
-between two functions in \(\mathcal{M}\). As many statistical methods
-depend on some measure of distance, downstream benefits of using the
-geodesic distance can be expected.
+\(L^2([a,b],\R)\). Suppose additionally that the function \(X\) belongs
+to a Riemannian manifold \(\M \subset L^2([a,b],\R)\) with Riemannian
+metric tensor \(g\) which can be used to assign a metric on the manifold
+as follows (Lin et al. 2014). For each point \(X\) on the manifold, the
+Riemannian metric tensor \(g\) has an inner product \(g_X\) on the
+tangent space \(T_X \M\). The norm of a tangent vector \(V \in T_X \M\)
+is defined as \[||V|| = \sqrt{g_X(V,V)}.\] The geodesic distance between
+two functions \(X_i,X_j\) on the manifold \(\M\), based on this metric
+tensor \(g\), is defined as
+\[ d_{\M}(X_i,X_j):=\inf \{l(\gamma): \gamma \text{ is piecewise-smooth function from } [a,b] \text{ to } \M, \gamma(a) = X_i, \gamma(b) = X_j \}\]
+where \[ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt\]
+is the length of the curve \(\gamma\). ???Is it clear how \(l(\gamma)\)
+depends on \(g\)???
 
-???insert example that shows pairwise geodesic distances give visually
-more interesting results than pairwise \(L_2\) distance.???
+Ideally, the functions \(X_i\) and \(X_j\) are observed on a very dense
+domain grid with no measurement error. Then the estimation of the
+geodesic distance \(d_\M(X_i,X_j)\) is straightfoward, e.g.~a
+shortest-path algorithm such as Floyd-Warhsall (Floyd 1962) may be used.
+However, most shortest-path algorithms critically assume observations
+live on the manifold ???add reference??? and are thus likely to fail for
+functional data which are almost certainly observed with noise. In this
+work, we put forth a technique for the recovery of the \(n{\times}n\)
+matrix \(G\) of pairwise geodesic distances, i.e.\\
+\[
+G(i,j)=G(j,i) = \begin{cases} 
+      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
+      0 & \textrm{otherwise.}
+   \end{cases}
+\] In lieu of \(X_i\) and \(X_j\), we have access only to
+discretely-sampled noisy functional observations \(Y_i\) and \(Y_j\)
+that possibly lie off the true manifold.
 
-Given perfect functional observations, i.e.~with no measurement error
-and on a very dense domain grid, the estimation of the geodesic distance
-can be accomplished straightforwardly using algorithms such as Floyd's.
-However, functional data are observed with noise in such a way that the
-observed functional data no longer live on a manifold. Though powerful
-techniques exist for smoothing discretely observed noisy functional
-data, the smoothing is designed for the \(L_2\) recovery of the original
-curve, not to ensure the recovered functional versions of the data lie
-on or close to the functional manifold \(\mathcal{M}\).
+The work of (Chen and Muller 2012) was among the first in functional
+data analysis to consider the manifold hypothesis for functional data. A
+notion of the mean and variation of functional manifold data was
+introduced there. Of particular relevance to this work is their proposed
+P-ISOMAP procedure which adds a penalty to allow for noisy functional
+observations whereas the classic ISOMAP algorithm (Tenenbaum, Silva, and
+Langford 2000) assumes observations lie exactly on the manifold.
+(Dimeglio et al. 2014) also realized this drawback to ISOMAP and
+proposed a procedure we will call robust-ISOMAP that is less sensitive
+to outliers. We will discuss P-ISOMAP and robust-ISOMAP in more details
+in the simulation section.
 
-With this challenge in mind, we put forth a technique for the specific
-task of estimating pairwise geodesic distances
-\(\{ d_\mathcal{M}(X_i,X_j\}_{i>j}\), to be defined when we have access
-only to discretely-observed noisy functional observations that possibly
-live off the true manifold. Briefly outline the methodology: MDS plus
-density ridge are used in a cool way to estimate (consistently?, let's
-hope..) pairwise geodesic distances.
+\section{Proposed method for estimating geodesic
+distances}\label{proposed-method-for-estimating-geodesic-distances}
 
-\hypertarget{related-work-susan}{%
-\subsection{Related Work {[}Susan{]}}\label{related-work-susan}}
+Suppose that each curve \(X_i \in (\M,g)\) is observed with measurement
+error on a time grid
+\(T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b\),
+i.e.~we observe a sample of \(K\)-dimensional vectors \(Y_1,\ldots,Y_n\)
+with \(Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}\), where the random
+variables \(\epsilon_{ij}\) have mean zero and are uncorrelated with
+each other. We assume that each observation grid \(T_1,\ldots,T_n\) is
+dense, i.e.~\(K\) is large.
 
-(Chen and Muller 2012)
+We begin by converting the discretly-sampled functional observations
+into continuous ones. Let \(\tilde X_1,\ldots,\tilde X_n\) denote the
+functional versions of the raw data obtained by some smoothing method.
+For example, we may employ spline smoothing ???add reference for spline
+smoothing??? to recover the functional versions of the raw data, i.e.\\
 
-(Lin and Yao 2017)
+\begin{equation}\label{eq_spline_smoothing} 
+\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
+\end{equation}
 
-(Dimeglio et al. 2014)
+where \(\lambda>0\) is a tuning parameter controlling the smoothness of
+\(\tilde X_i\). The proposed methodology for estimating the pairwise
+geodesic distances \(\{ d_\M (X_i,X_j)\}_{i>j}\) is agnostic to the
+smoothing method employed.
 
-\hypertarget{setting-susan}{%
-\subsection{Setting {[}Susan{]}}\label{setting-susan}}
+Our method is based on the idea that the underlying functional manifold
+\(\M\) can be sufficiently well-recovered by the subspace-constrained
+mean-shift (SCMS) algorithm (Ozertem and Erdogmus 2011). Let
+\(\hat M \subset L^2([a,b],\mathbb R)\) denote the collection of basins
+of attraction of the estimated probability density function of \(X\).
+The SCMS algorithm sends each point
+\(\tilde X_i \in L^2([a,b],\mathbb R)\) to a point in \(\hat M\). This
+destination is unique and we denote it \(\tilde X_i^{\hat \M}\). We
+expect the points
+\(\tilde X^{\hat \M}_1,\ldots,\tilde X^{\hat \M}_n \in \hat M\) to lie
+close to the real manifold \(\M\). Theoretical justificaton that
+\(\hat M\) as estimated by SCMS is a reasonable surrogate for the true
+manifold \(\M\) can be found in (Genovese et al. 2014).
 
-Let \(X_1,\ldots,X_n\) be independent realizations of a random function
-\(X\in\mathcal{M}\subset L^2([a,b],\mathbb{R})\), where \(\mathcal{M}\)
-is an unknown manifold. Suppose that each curve \(X_i\) is observed with
-measurements errors on a grid \(T_i=(t_{i1},\ldots,t_{iK})\), i.e.~we
-observe a sample of \(K\)-dimensional vectors \(Y_1,\ldots,Y_n\) with
-\(Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}\), where the random variables
-\(\epsilon_{ij}\) are of mean zero and uncorrelated with each other. We
-assume that the grids \(T_1,\ldots,T_n\) are dense.
-
-Let \(\tilde X_1,\ldots,\tilde X_n\) denote the functional versions of
-the raw data by some smoothing method. Our proposed methodology is
-unspecific to the smoothing method employed. For simplicity however,
-let's say we employ spline smoothing to recover the functional versions
-of the raw data, i.e.\\
-\[ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}\]
-where \(\lambda>0\) is a tuning parameter controlling the smoothness of
-\(\tilde X_i\).
-
-\hypertarget{relevant-differential-geometry-concepts-susan}{%
-\subsection{Relevant differential geometry concepts
-{[}Susan{]}}\label{relevant-differential-geometry-concepts-susan}}
-
-We first rigorously define the concepts of geodesic and geodesic
-distance. Following (Lin et al. 2014), let \((\mathcal{M},g)\) be a
-\(d\)-dimensional Riemannian manifold, where \(g\) is a Riemannian
-metric tensor on \(\mathcal{M}\) which can be used to assign a metric on
-the manifold. Specifically, for each point \(p\) on the manifold, the
-Riemannian metric tensor \(g\) has an inner product \(g_p\) on the
-tangent space \(T_p \mathcal{M}\). The norm of a tangent vector
-\(v \in T_p \mathcal{M}\) is defined as \[||v|| = \sqrt{g_p(v,v)}.\] If
-\(\gamma: [a,b] \subset \mathbb{R}\to \mathcal{M}\) is a smooth curve,
-its length is then defined as
-\[ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt .\] The
-geodesic distance between two points \(p,q\) on the manifold
-\(\mathcal{M}\), based on this metric tensor \(g\), is defined as
-\[ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \mathcal{M}\text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.\]
-
-\hypertarget{proposed-method-for-estimating-geodesic-distances-marie}{%
-\section{Proposed method for estimating geodesic distances
-{[}Marie{]}}\label{proposed-method-for-estimating-geodesic-distances-marie}}
-
-We first describe an important helper algorithm which we call IsoGeo. It
-is a subroutine of the classic Isomap procedure. Isomap is a three steps
-procedure that takes as input a set of points
-\(x_1,\ldots,x_n\in \mathbb{R}^D\) and produces an embedding of the
-input data in the space \(\mathbb{R}^d\) with \(d<D\), that preserves
-pairwise geodesic distances.
+Before presenting our procedure, we need to first review the ISOMAP
+algorithm (Tenenbaum, Silva, and Langford 2000), a three-step procedure
+that takes as input a set of points \(x_1,\ldots,x_n\) in a submanifold
+\(\M\) of \(\R^D\) and produces an embedding in the space \(\R^d\) with
+\(d<D\) that preserves pairwise geodesic distances. The procedure is as
+follows.
 
 \begin{enumerate}
-\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\mathbb{R}^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-\item In the second step, the pairwise geodesic distances $d_\mathcal{M}(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Warshall1962]
-<!-- or Dijkstra’s  which one do we actually use in our implementation, the FW or D? -->
-\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\mathbb{R}^d$.
+\item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
+\item Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
+Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+\item Use multidimensional scaling ???add reference??? to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
 \end{enumerate}
 
-In what follows, we call IsoGeo the procedure of estimating pairwise
-geodesic distances with the two first steps of Isomap. Note that Floyd
-or Dijkstra are methods to find the smallest path given a weigthed graph
-but they are not related to how to calculate that graph (which is the
-difficult part, depend on number of neighbors, many method out there to
-calculate that graph in a robust way, P-ISOMAP is one of these).
+Note that shortest-path algorithms such as the Floyd-Warshall algorithm
+are used to find the smallest path \textit{given} a weigthed graph but
+they do not produce the weighted graphs themselves. Both P-ISOMAP and
+robust-ISOMAP are modifications of ISOMAP in the sense that they change
+the construction of the weighted graph in step 1.\\
 
-Our method is described by the following steps:
+In what follows, we call IsoGeo the two-step procedure which performs a
+modified version of the first step of ISOMAP followed by the original
+second step of ISOMAP. The first step of IsoGeo constructs the weighted
+graph \(G\) accoding to Dimeglio et al. (2014) so that graph is
+independent of the tuning parameter \(\epsilon\). Specifically, we begin
+by constructing the complete ???explain what complete means??? weighted
+graph \(G_c\) with nodes corresponding to the observations
+\(x_1,\ldots,x_n\in \R^D\). Next, we obtain the minimal spanning tree
+associated with \(G_c\) and denote its set of edges by \(E_s\). The
+graph \(G\) is obtained by adding all edges \(e_{ij}\) to \(E_s\) for
+which the following condition is true:
+\[ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),\] where
+\(B(x_i,\epsilon_i)\) is the open ball of center \(x_i\) and radius
+\(\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}\), and
+\(\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}\).
+Let the distance estimated by IsoGeo be denoted
+\(\text{IsoGeo}(x_i,x_j)\). Note that IsoGeo has no tuning parameters.
+
+We are now ready to describe our procedure. Since SCMS is based on
+kernel density estimation, we first reduce the dimension of our data
+with multidimensional scaling before applying SCMS to avoid the curse of
+dimensionality.
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Obtain a \(s\)-dimensional representation
-  \(\tilde X^s_1,\ldots,\tilde X^s_n\) of the functions
-  \(\tilde X_1,\ldots,\tilde X_n\) that ``preserves'' the pairwise
-  \(L^2\) distances by using MDS.
-\item
-  Obtain
-  \(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
-  a ``projection'' of \(\tilde X^s_1,\ldots,\tilde X^s_n\) onto a ridge
-  \(\hat \mathcal{M}\) which is computed with the subspace constrained
-  mean-shift algorithm (Ozertem and Erdogmus 2011).
-\item
-  Use IsoGeo to approximate the pairwise geodesic distances
-  \(\{d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j)\}_{i>j}\)
-  and define the \(n\) by \(n\) matrix \[
+\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$.
+Let $s$ be a positive integer, much smaller than $D$.
+Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ are preserved.
+\item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
+\item Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
+\end{enumerate}
+
+Our geodesic distance estimator is the \(n{\times}n\) matrix \(\hat G\)
+whose elements are given by \[
 \hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
-   d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j) & \textrm{if $i\neq j$,}\\
+ \text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
  0 & \textrm{otherwise.}
@@
 \]
-\end{enumerate}
 
-Since the ridge estimation obtained from noisy measurements of a
-manifold should approximate well the manifold (Genovese et al. 2014), we
-expect the points
-\(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
-to lie close to the real manifold \(\mathcal{M}\) and then
-\(d_{\hat \mathcal{M}}\) to be close to \(d_\mathcal{M}\). Ridge
-estimation suffers from the curse of dimensionality, this is why we
-first reduce the dimension of our data with MDS and then apply the
-shift-mean algorithm to estimate the ridge.
+There are two tuning parameters to our procedure. First, there is the
+dimension \(s\) in step 1. In our simulations we try
+\(s \in \{1,2,3\}\). The other parameter is the bandwidth \(h\) in the
+subspace constrained mean-shift. In our simulations we set \(h\)
+according to the heuristic in Equation (A1) of (Chen et al. 2015). In
+reality, depending on the downstream task, both \(s\) and \(h\) can be
+tuned in a data-adaptive way, say using cross-validation.
 
-\hypertarget{selection-of-tuning-parameters-marie}{%
-\subsection{Selection of tuning parameters
-{[}Marie{]}}\label{selection-of-tuning-parameters-marie}}
-
-Describe heuristics. We pick the bandwidth \(h\) in the subspace
-constrained mean-shift using Equation (A1) of (Chen et al. 2015).
-
-\hypertarget{simulation-study}{%
-\section{Simulation study}\label{simulation-study}}
+\section{Simulation study}\label{simulation-study}
 
 We perform a simulation study to ascertain the efficacy of our method
-for estimating pairwise geodesic distances for discretely-observed noisy
-functional data.
+for estimating pairwise geodesic distances. Three different metrics are
+used to assess the quality of a pairwise geodesic distance estimator.
 
-\hypertarget{alternative-estimators-of-geodesic-distance-marie}{%
+The first metric assesses near-\(\epsilon\) isometry, which we define as
+the percentage of estimated pairwise geodesic distances between
+\(1-\epsilon\) and \(1+\epsilon\) of the corresponding true pairwise
+geodesic distance. We form the receiver operating curve with
+\(\epsilon\) on the \(x\)-axis and near-\(\epsilon\) isometry on the
+\(y\)-axis. The metric is the area under the receiver operating curve.
+
+The second metric \(||d-\hat d||/||d||\) is the Frobenius norm of the
+estimation error matrix \(d-\hat d\), relative to the Frobenius norm of
+the true distance matrix \(d\). The third metric is the Pearson
+correlation coefficient between the upper diagonal of \(d\) and
+\(\hat d\).
+
 \subsection{Alternative estimators of geodesic distance
-{[}Marie{]}}\label{alternative-estimators-of-geodesic-distance-marie}}
+{[}Marie{]}}\label{alternative-estimators-of-geodesic-distance-marie}
 
-\begin{itemize}
-\tightlist
-\item
-  \textbf{(Raw Data) RD} Apply IsoGeo on the raw data \(Y_1,\ldots,Y_n\)
-  to obtain an estimator \(\hat G_{\textrm{RD}}\). Note that this
-  procedure can only be used if the grid \(T_i\) is the same for each
-  \(i\).
-\item
-  \textbf{(Spline Smoothing) SS} Transform each vector \(Y_i\) into a
-  function \(\tilde X_i\) by spline smoothing. Apply IsoGeo on the
-  vectors \(\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n\), where
-  \(t_1,\ldots,t_K\) is a regular grid of \([a,b]\), to obtain an
-  estimator \(\hat G_{\textrm{SS}}\).
-\item
-  \textbf{(P-ISOMAP) pI} Transform each vector \(Y_i\) into a function
-  \(\tilde X_i\) by spline smoothing. Calculate the weighted graph of
-  step I of Isomap with a penalty (describe the penalty of Chen and
-  Muller). Apply step II of Isomap to obtain \(\hat G_{\textrm{pI}}\).
-\item
-  \textbf{(Random Projection) RP} same method as our but change step 2
-  by : obtain \(s\)-dimensional representation by random projection and
-  setp 4 by obtain \(\hat G\) using a ensemble method.
-\end{itemize}
+We compare the performance of our method to each of the following four
+methods. The first method, denoted RD for raw data, takes the naive
+approach of applying IsoGeo directly on the raw vectors
+\(Y_1,\ldots,Y_n \in R^K\). Note that this procedure is only sensible if
+all the grids \(T_1,\ldots,T_n\) are the same. In the second method, we
+first smooth the discretely-sampled noisy functional data and then
+estimate their pairwise distance using the \(L^2\) distance.
 
-We ran Chen and Muller's P-ISOMAP separately in Matlab with their
-automatic penalty selection and discovered that ???.
+The third alternative, denoted SS for spline smoothing, applies IsoGeo
+on the smoothed version \(\tilde X_1,\ldots,\tilde X_n\) of the raw data
+obtained by spline smoothing as described in
+(\ref{eq_spline_smoothing}). The final alternative we consider is
+P-ISOMAP, denoted pI, which is the two-step procedure developed in Chen
+and Muller (2012) designed to handle noisy functional data. In the first
+step, a penalty is incorporated to robustify the construction of the
+weighted graph. This is followed by the original second step of ISOMAP.
 
-\hypertarget{performance-metrics-susan}{%
-\subsection{Performance metrics
-{[}Susan{]}}\label{performance-metrics-susan}}
+We ran Chen and Muller's P-ISOMAP separately in Matlab and tuned for the
+number of neighbors and the penalty parameter using the relative
+Frobenius assessment metric. We found that the penalty parameter chosen
+in this way was always zero, thus reducing hte P-ISOMAP procedure to
+standard ISOMAP. Given this, we did not make further comparison to
+P-ISOMAP in our simulation study.
 
-We describe three different metrics we use to assess the quality of a
-pairwise geodesic distance estimator. This is the ROC curve with
-\(\epsilon\) on the \(x\)-axis and degree to which near-\(\epsilon\)
-isometry holds, i.e.~the percentage of estimated pairwise distances
-between \(1-\epsilon\) and \(1+\epsilon\) of the truth pairwise
-distance.
+\subsection{Simulation scenarios}\label{simulation-scenarios}
 
-\hypertarget{simulation-scenarios-susan}{%
-\subsection{Simulation scenarios
-{[}Susan{]}}\label{simulation-scenarios-susan}}
+We shall consider three functional manifolds in our simulation study. As
+of yet, there is little work as to how to sample from a functional
+manifold. Even in the Euclidean case, it is not obvious how sampling
+should be done (Diaconis, Holmes, and Shahshahani 2013). How to properly
+sample from a functional manifold could be interesting future work. For
+now, to safeguard against sampling unevenly on the functional manifold,
+we sample on a very concentrated measure for the intrinsic parameters.
 
-Acknowledge that we are sampling on a very concentrated measure to make
-sure functional manifold is well sampled. Cite Diaconis et. al's
-``Sampling from a manifold''. How to sample properly from a functional
-manifold could be interesting future work.
+\textbf{Isometric functional manifold of normal density functions} This
+scenario is modified from what is referred to as Manifold 2 in (Chen and
+Muller 2012) by fixing the variance of the normal density to be \(1\).
+We have
+\[\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}\] with
+the \(L_2\) inner product as the metric tensor of \(\M\), where
+\(X_\beta: [a,b] \to \R\) is given by
+\(X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}\).
+First, note that the ``straight'' line connecting \(X_{\beta_1}\) and
+\(X_{\beta_2}\) in \(\M\) does not always stay inside of \(\M\). Thus
+the geodesic distance will not coincide with the \(L^2\) distance.
 
-\begin{itemize}
-\tightlist
-\item
-  warping manifold (scenario 5 in sim\_functional\_data.R)
-\item
-  pdf manifold (scenario 2 in sim\_functional\_data.R)
-\item
-  square root pdf manifold (scenario 4 )
-\end{itemize}
+We set \(a=-4\) and \(b=4\) and sample \(\beta\) according to a
+truncated standard normal with support on \([-1,1]\). The geodesic
+distance between the curves \(X_{\beta_1}\) and \(X_{\beta_2}\) is given
+by
 
-\hypertarget{results}{%
-\subsubsection{Results}\label{results}}
+\begin{eqnarray*}
+d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
+&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
+&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
+&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
+&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
+\end{eqnarray*}
 
-Submit compare\_methods.sh as slurm job using sbatch.
+where the approximation comes from the fact that we are integrating on
+\([a,b]=[-4,4]\) and not on \(\R\). We can see this manifold is
+isometric, since the geodesic distance between \(X_{\beta_1}\) and
+\(X_{\beta_2}\) in \(\M\) is proportional to \(\beta_2-\beta_1\), the
+Euclidean distance between the \(\beta\)'s.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\ExtensionTok{sbatch}\NormalTok{ compare_methods.sh}
-\end{Highlighting}
-\end{Shaded}
+\textbf{Functional manifold of square root velocity functions} It was
+shown in Joshi, Srivastava, and Jermyn (2007) that the square root
+representation of probability density functions gives rise to a simple
+closed-form geodesic distance. They consider the manifold
+\[ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}\]
+with the metric tensor given by the Fisher-Rao metric tensor
+\[ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds \] for two tangent vectors
+\(v_1,v_2 \in T_\psi(\M)\). Note that this concides with the
+\(L_2[0,1]\) inner product. Joshi, Srivastava, and Jermyn (2007) showed
+that the geodesic distance between any two \(\psi_1\) and \(\psi_2\) in
+\(\M\) is simply \[d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.\] We
+will specifically examine the square root of \(Beta(\alpha,\beta)\)
+distributions which is supported on \([0,1]\). That is,
+\[ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} \]
+where \(\psi_{\alpha,\beta}: [0,1] \to \R\) is the pdf of
+\(Beta(\alpha,\beta)\). We sample \(\alpha\) according to a truncated
+normal with mean 3 and variance 0.09 with support on \([1,5]\). We
+sample \(\beta\) according to a truncated normal with mean 3.5 and
+variance 0.09 with support on \([2,5]\).
 
-Alternatively, we can run the R script compare\_methods.R locally by
-removing the slurm\_arrayid chunk
+\textbf{Functional manifold of warping functions} This is based on
+Equations 17 and 18 in Kneip and Ramsay (2008) but with
+\(z_{i1}, z_{i2}\) both set to \(1\). (Note that Equation 17 has a typo
+where the exponentials are missing negative signs.) Let
+\(X_\alpha(t) = \mu(h_\alpha(t))\) be defined on \([-3,3]\) where
+\[ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}\] and
+\[ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}\]
+if \(\alpha \ne 0\) and \(h_\alpha(t) = t\) otherwise.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{R compare_methods.R}
-\end{Highlighting}
-\end{Shaded}
+Consider the manifold \[ M = \{X_\alpha: -1 \le \alpha \le 1\}.\] We
+sample \(\alpha\) according to a truncated standard normal with support
+on \([-1,1]\). The geodesic distance is then
+\[ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha\]
+which is estimated using numerical integration.
 
-Simulation results are then visualized by invoking cluster\_results.R.
-There's some unfortunate hardcoding still in cluster\_results.R
+\subsection{Results}\label{results}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{R cluster_results.R}
-\end{Highlighting}
-\end{Shaded}
+???Still need to describe parameters under study: SNR, denseness of
+observations (\(K_{observed}\)). Describe parameters that are fixed:
+\(K_smooth\) and com\_grid. Describe tuning parameters for our method:
+\(s\) and \(h\).
 
-\hypertarget{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}{%
-\subsection{The cost of employing proposed method when manifold is flat
-{[}Susan/Marie?{]}}\label{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}}
+\begin{figure}
+\includegraphics{sim_results/sce=5_SNR=high_Kobs=30}
+\caption{blah}
+\end{figure}
 
-Suppose we take scenario 1 where the geodesic distance coincides with
-the \(L_2\) distance. Then what is the cost of employing our distance
-estimator compared to smoothing the data and doing numerical integration
-to find the \(L_2\) distance?
-
-\hypertarget{distance-based-functional-classification-marie}{%
 \section{Distance-based functional classification
-{[}Marie{]}}\label{distance-based-functional-classification-marie}}
+{[}Marie{]}}\label{distance-based-functional-classification-marie}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ChenMuller2012}{}%
+\hypertarget{ref-ChenMuller2012}{}
 Chen, Dong, and Hans-Georg Muller. 2012. ``NONLINEAR Manifold
@@
 
-\leavevmode\hypertarget{ref-ChenHo2015}{}%
+\hypertarget{ref-ChenHo2015}{}
 Chen, Y., S. Ho, P. E. Freeman, C. R. Genovese, and L. Wasserman. 2015.
@@
 Algorithm.'' \emph{Monthly Notices of the Royal Astronomical Society}
-454 (1): 1140--56. \url{https://doi.org/10.1093/mnras/stv1996}.
+454 (1): 1140--56.
+doi:\href{https://doi.org/10.1093/mnras/stv1996}{10.1093/mnras/stv1996}.
 
-\leavevmode\hypertarget{ref-Dimeglio2014}{}%
+\hypertarget{ref-Diaconis2013}{}
+Diaconis, Persi, Susan Holmes, and Mehrdad Shahshahani. 2013. ``Sampling
+from a Manifold.'' In \emph{Advances in Modern Statistical Theory and
+Applications: A Festschrift in Honor of Morris L. Eaton}, Volume
+10:102--25. Collections. Beachwood, Ohio, USA: Institute of Mathematical
+Statistics.
+doi:\href{https://doi.org/10.1214/12-IMSCOLL1006}{10.1214/12-IMSCOLL1006}.
+
+\hypertarget{ref-Dimeglio2014}{}
 Dimeglio, Chloe, Santiago Gallon, Jean-Michel Loubes, and Elie Maza.
@@
 Amsterdam, The Netherlands, The Netherlands: Elsevier Science Publishers
-B. V.: 373--86. \url{https://doi.org/10.1016/j.csda.2013.09.030}.
+B. V.: 373--86.
+doi:\href{https://doi.org/10.1016/j.csda.2013.09.030}{10.1016/j.csda.2013.09.030}.
 
-\leavevmode\hypertarget{ref-Genovese2014}{}%
+\hypertarget{ref-Floyd1962}{}
+Floyd, Robert W. 1962. ``Algorithm 97: Shortest Path.'' \emph{Commun.
+ACM} 5 (6). New York, NY, USA: ACM: 345.
+doi:\href{https://doi.org/10.1145/367766.368168}{10.1145/367766.368168}.
+
+\hypertarget{ref-Genovese2014}{}
 Genovese, Christopher R., Marco Perone-Pacifico, Isabella Verdinelli,
@@
 
-\leavevmode\hypertarget{ref-Lin2014}{}%
+\hypertarget{ref-Joshi2007}{}
+Joshi, S., A. Srivastava, and I.H. Jermyn. 2007. ``Riemannian Analysis
+of Probability Density Functions with Applications in Vision.'' In
+\emph{2007 Ieee Conference on Computer Vision and Pattern Recognition ;
+Proceedings}, 1664--71. Piscataway, NJ: IEEE.
+\url{http://dro.dur.ac.uk/18442/}.
+
+\hypertarget{ref-Kneip2008}{}
+Kneip, Alois, and James O Ramsay. 2008. ``Combining Registration and
+Fitting for Functional Models.'' \emph{Journal of the American
+Statistical Association} 103 (483). Taylor \& Francis: 1155--65.
+doi:\href{https://doi.org/10.1198/016214508000000517}{10.1198/016214508000000517}.
+
+\hypertarget{ref-Lin2014}{}
 Lin, Binbin, Ji Yang, Xiaofei He, and Jieping Ye. 2014. ``Geodesic
@@
 
-\leavevmode\hypertarget{ref-LinYao2017}{}%
-Lin, Zhenhua, and Fang Yao. 2017. ``Functional Regression on Manifold
-with Contamination.'' \emph{arXiv E-Prints}, April, arXiv:1704.03005.
-
-\leavevmode\hypertarget{ref-Ozertem2011}{}%
+\hypertarget{ref-Ozertem2011}{}
 Ozertem, Umut, and Deniz Erdogmus. 2011. ``Locally Defined Principal
@@
 
-\end{document}
+\hypertarget{ref-Tenenbaum2000}{}
+Tenenbaum, Joshua B., Vin de Silva, and John C. Langford. 2000. ``A
+Global Geometric Framework for Nonlinear Dimensionality Reduction.''
+\emph{Science} 290 (5500). American Association for the Advancement of
+Science: 2319--23.
+doi:\href{https://doi.org/10.1126/science.290.5500.2319}{10.1126/science.290.5500.2319}.

### 2019-08-22 paperAAAI.tex
rmarkdown outputs to body

@@
 
-\newcommand {\To}{\rightarrow}
-\newcommand {\TO}{\Rightarrow}
-\newcommand {\R}{\mathbb{R}}
-\newcommand {\Prob}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand {\cov}{\textrm{Cov}}
-\newcommand {\var}{\textrm{Var}}
-\newcommand {\1}{\textrm{\textbf{1}}}
-\newcommand{\M}{\mathcal{M}}
 %\nocopyright
@@
 
-\begin{abstract}
-Functional data analysis is the statistical analysis of smooth infinite-dimensional curves. 
-The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the dimensionality of curves is only artifically high if they are smooth.
-There is also another sense in which the dimensionality of the curves under study may be low. 
-This is the much less explored idea that the curves may actually live in a submanifold with low intrinsic dimension.
-Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
-%Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
-In this work we address the estimation of the pairwise geodesic distance between functional manifold data when we only have access to their noisy realizations which live near, rather than exactly on, the manifold.
-% This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
-%We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
-The proposed methodology first projects the functional data on to the underlying hidden manifold and then performs operations on this hidden manifold. 
-Good estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.
-
-\end{abstract}
-
-\section{Introduction}
-
-Many statistical methods depend on a measure of distance.
-Methods in functional data analysis, the study of infinite-dimensional smooth curves, are no exception.
-Certain classes of functional data may naturally live on manifold, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
-In situations where the manifold hypothesis might be plausible, one might consider using a geodesic distance that takes into account the intrinsic structure of the manifold rather than the standard $L^2$ distance.
-The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.
-
-Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a low-dimensional nonlinear Riemannian manifold $\M \subset L^2([a,b],\R)$ where $g$ is a Riemannian metric tensor on $\M$ which can be used to assign a metric on the manifold as follows \cite{Lin2014}.
-For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
-The norm of a tangent vector $V \in T_X \M$ is defined as $$\|V\| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-\begin{eqnarray*} 
-d_g(X_i,X_j)&:=&\inf \big\{l(\gamma): \gamma:[a,b] \to \M \text{ piecewise smooth},\\
-&&\quad \quad  \gamma(a) = X_i, \gamma(b) = X_j] \big \},
-\end{eqnarray*}
-where $$ l(\gamma) := \int_a^b \left\| \frac{\,d\gamma}{\,d t}(t) \right\| \,dt$$
-is the length of a smooth curve $\gamma: [a,b] \subset \R \to \M$.
-
-%<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->
-
-%<!-- (gives examples and references for functional manifolds). -->
-%<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->
-
-%<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision.
-However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio.
-(is it true?
-ref?)  -->
-
-Given perfect functional observations $X_i$ and $X_j$, i.e.\ with no measurement error and on a very dense domain grid, the estimation of the geodesic distance $d_g(X_i,X_j)$ can be accomplished straightforwardly using a shortest-path algorithm, e.g.
-the Floyd-Warhsall algorithm \cite{Floyd1962}.
-However, when functional data are observed with noise, this approach is likely to fail. This is because the error may be such that the noisy functional manifold data no longer live on a manifold. 
-A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
-This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
-%<!-- This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
-
-%<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data).
-Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.
--->
-
-In this work, we put forth a technique for the specific task of recovering the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
-$$
-G(i,j)=G(j,i) = \left\{ \begin{array}{ll}
- d_g (X_i,X_j) & \textrm{if $i\neq j$,}\\
- 0 & \textrm{otherwise,}
-  \end{array} \right.
-$$ 
-when we have, in lieu of $X_i$ and $X_j$, access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.
-This is accomplished by first projecting the functional data on to the underlying hidden manifold and then performing operations on this hidden manifold.
-
-% Pairwise distances are important for many downstream tasks, e.g.
-classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression).
-Certainly, pairwise geodesic distances are legitimate objects of interest in themselves but our motivation is more general in nature.
-Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.
--->
-
-\subsection{Related work}
-
-The work of Chen and Muller \shortcite{ChenMuller2012} was among the first in functional data analysis to consider the manifold hypothesis for functional data.
-A notion of the mean and variation of functional manifold data was introduced.
-Of particular relevance to our work is the modified ISOMAP procedure, called P-ISOMAP, introduced in Chen and Muller \shortcite{ChenMuller2012}, which adds a data-adaptive penalty to allow for noisy functional observations whereas in the classic ISOMAP algorithm, observations are assumed to lie exactly on the manifold.
-Another existing work for estimating the pairwise geodesic distance for functional manifold data can be found in Dimeglio et al.
-\shortcite{Dimeglio2014}.
-The estimator is called Robust-ISOMAP for the fact that it is motivated by modifying the ISOMAP algorithm to be less sensitive to outliers.
-We will discuss both P-ISOMAP and robust ISOMAP in more details in the simulation section.
-
-%<!-- [@LinYao2017]  -->
-
-\section{Proposed method for estimating geodesic distances}
-
-Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e.
-we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
-We assume that each observation grid $T_1,\ldots,T_n$ is dense.
-
-We begin by converting the discrete functional observations into continuous ones.
-Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
-Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
-For example we may employ spline smoothing to recover the functional versions of the raw data, i.e.\
-\begin{equation}\label{eq_spline_smoothing} 
-\tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
-\end{equation}
-where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
-
-Before introducing our procedure, let us first review the ISOMAP algorithm \cite{Tenenbaum2000}, a three-step procedure that takes a set of points $x_1,\ldots,x_n$ in $(\M,g)$, a submanifold of $\R^D$ as input and produces an embedding of the input data in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
-\begin{enumerate}
-\item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
-Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ if the distance $d_{ij}=||x_i-x_j ||_{2}$ is smaller than a given value $\epsilon$, and the weight associated to an edge $e_{ij}$ is $d_{ij}$.
-\item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms.
-Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e./ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm \cite{Floyd1962} or with the Dijkstra algorithm \cite{Dijkstra59anote}.
-%<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
-\item Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
-\end{enumerate}
-Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves.
-Formulating the graph in step 1 of ISOMAP is in fact the most challenging asepct of the algorithm since it depends heavily on the choice of the tuning parameter $\epsilon$.
-Both P-ISOMAP and robust ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.
-%<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
-
-In what follows, we call IsoGeo the procedure which performs a modified version of the first step of ISOMAP, proposed in @Dimeglio2014, followed by the original second step of ISOMAP.
-The first step of IsoGeo modifies Step 1 of ISOMAP so that the constructed graph $G$ is independent of the tuning parameter $\epsilon$:
-1'.
-Construct the complete weigthed graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
-Obtain the minimal spanning tree $G_s$ associated with $G_c$ and denote its set of edges by $E_s$.
-The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
-$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
-where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
-Note that IsoGeo has no tuning parameters.
-
-Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm \cite{Ozertem2011}.
-Theoretical justificaton for this can be found in Genovese et al.
-\shortcite{Genovese2014}.
-We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$.
-Our procedure is described by the following steps:
-
-%<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
-%<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
-%<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
-\begin{enumerate}
-\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a common grid $\tilde T=(t_{1},\ldots,t_{K})$.
-Let $s$ be a positive integer, smaller than $K$.
-Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^K$ Euclidean distances on the discretised functions $\tilde X_i$ are preserved.
-\item Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011} to each of $\tilde X^s_i$ to obtain  $\tilde X^{s,\hat \M}_i$.
-\item Use IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
-\end{enumerate}
-Thus our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
-$$
-\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
- d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
- 0 & \textrm{otherwise.}
-  \end{array} \right.
-$$
-Since ridge estimation suffers from the curse of dimensionality, we first reduce the dimension of our data with multidimensional scaling in Step 1 before applying the mean-shift algorithm.
-
-\subsection{Selection of tuning parameters}
-Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of \cite{ChenHo2015}.
-
-\section{Simulation study}
-
-We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances when one only has access to discretely-observed noisy functional data.
-Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.
-
-\begin{itemize}
-\item the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis.
-Near-$\epsilon$ isometry is a relaxation of isometry measured by the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.
-\item  the relative Frobenius metric: this is given by
-$||d-\hat d||/||d||$
-where $d$ and $\hat d$ are the true and estimated $n{\times}n$ geodesic distance matrices, respectively. 
-\item  the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
-\end{itemize}
-
-\subsection{Alternative estimators of geodesic distance [Marie]}
-
-We compare the performance of our method to the one of the three following methods:
-\begin{enumerate}
-\item \emph{Raw Data (RD)} The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
-<!-- to obtain an estimator $\hat G_{\textrm{RD}}$.--> Note that this procedure only makes sense if all the grids $T_1,\ldots,T_n$ are the same.
-\item   \emph{Spline Smoothing (SS)} The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
-%<!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
-Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
-\item  \emph{P-ISOMAP (pI)} The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
-%<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->
-\end{enumerate}
-We ran Chen and Muller's P-ISOMAP separately in Matlab with their automatic penalty selection and discovered that ???.
-
-\subsection{Simulation scenarios }
-%<!-- Susan  -->
-
-\subsubsection{Isometric functional manifold of normal density functions}
-%<!-- Scenario 5 -->
-This scenario is modified from what is referred to as Manifold 2 in \cite{ChenMuller2012} by fixing the variance of the normal density to be $1$.
-We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
-We set $a=-4$ and $b=4$.The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
-\begin{eqnarray*}
-d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
-&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
-&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
-&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
-&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
-\end{eqnarray*}
-where the approximation comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
-We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is the Euclidan distance between the $\beta$'s, up to some scaling factor.
-Note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$, so we cannot employ the calculation technique of Scenario 1.
-
-\subsubsection{Functional manifold of square root velocity functions}
-It was shown in \cite{Joshi2007} that the square root representation of probability density functions has a nice closed form geodesic.
-They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
-Note that this concides with the $L_2[0,1]$ inner product.
-\cite{Joshi2007} showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
-We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$.
-That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$.
-
-\subsubsection{Functional manifold of warping functions}
-%<!-- Scenario 5 -->
-This is based on Equations (17) and (18) of\cite{Kneip2008} but with $z_{i1}, z_{i2}$ set to $1$.
-(Equation 17 has a typo where the exponentials are missing negative signs).Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$ and $h_\alpha(t) = t$ if $\alpha = 0$.
-Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\} $$.
-The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$
-
-\subsection{Sampling from a functional manifold}
-As of yet, there is little work as to how to sample from a functional manifold.
-Even in the Euclidean case, it is not obvious how sampling should be done \cite{Diaconis2013}.
-To safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
-For both the manifold of normal densities and the manifold of warping functions, we sample $\alpha$ according to ???%rtruncnorm(samplesize,alphamin,alphamax,(alphamin+alphamax)/2,1)???
-For the manifold of square root beta densities, we sample as follows       
-%alpha<-$ rtruncnorm(nb_alpha,alphamin,alphamax,(alphamax+alphamin)/2,0.3)$
-%beta<- $rtruncnorm(nb_beta,betamin,betamax,(betamax+betamin)/2,0.3)$
-How to sample properly from a functional manifold could be interesting future work.
-
-\subsection{Results}
-
-\subsection{The cost of employing proposed method when manifold is flat}
-
-\section{Distance-based functional classification [Marie]}
-
-In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task.
-There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification.
-It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.
-For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
-Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
-$$
-\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
-$$
-We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.
-
-The bandwidth in the classifier should be tuned individually for each method.
-Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.
-
-Datasets used by functional classification papers
-\begin{itemize}
-\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
-\item Berkeley growth curves in \cite{ChenReiss2014}.
-\item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper.
-\item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data"
-\end{itemize}
-
-Datasets used in functional manifold papers
-\begin{itemize}
-\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012}
-\item Tecator in \cite{LinYao2017} contamination paper
-\item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper
-\end{itemize}
-
-\bibliographystyle{aaai} 
-\bibliography{references}
+\input{index}
 

### 2019-08-22 cluster_results.R

@@ -1,5 +1,3 @@
-#setwd('./Users/UQAM/Dropbox/Marie-Moi/Susan_project/manifold learning/Geo_dist_calculation/Cluster_results/spartanSim_clusterOutput')
-
 choose_el<- function(list_in,ind1,ind2){
   list_in[[ind1]][ind2]
 }
@@ -10,8 +8,16 @@ choose_el2<-function(ind_vec,list_in1){
 
 create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
 
+  base_pattern = do.call(sprintf, c(list("sce=%d_SNR=%.01f_Kobs=%d"), sce_details))
+  
   # TODO: fix hardcoding
-  base_pattern = do.call(sprintf, c(list("sce=%d_SNR=%.01f_regsamp=%d_Kobs=%d_Ksmooth=%d"), sce_details))
+  if(sce_details[2]==0.1){
+    base_pattern_out = do.call(sprintf, c(list("sce=%d_SNR=low_Kobs=%d"), sce_details[1],sce_details[3]))
+  }
+  else{
+    base_pattern_out = do.call(sprintf, c(list("sce=%d_SNR=high_Kobs=%d"), sce_details[1],sce_details[3]))
+  }
+  
 
   data_sce = list.files(pattern=paste("_", base_pattern,"_mc=*",sep=""))
   if (length(data_sce) == 0){
@@ -25,29 +31,30 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=3*nr_methods)
     
     ### We plot one boxplot per method and per assesment measure
-    pdf(paste(base_pattern,".pdf",sep=""),width=15,height=5)
+    pdf(paste(base_pattern_out,".pdf",sep=""),width=4,height=2)
     
     par(mfrow=c(1,3))
-    par(oma=c(3,3,3,3))
+    # par(oma=c(3,3,3,3))
+    boxplot(results[,1:nr_methods],main="MSE",names=method_names)
+    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="isometry",names=method_names)
+    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson",names=method_names)
     
-    boxplot(results[,1:nr_methods],main="relative MSE",names=method_names)
[…]

### 2019-08-22 compare_methods.R

@@ -69,7 +69,7 @@ meth <- list("NN" = FALSE,
              "SS" = TRUE,
              "pI" = FALSE,
              "OUR" = FALSE,
-             "OUR2" = TRUE,
+             "OUR2" = FALSE,
              "OUR3"=TRUE,
              "RP" = FALSE,
              "L2" = TRUE, 

### 2019-08-22 run_this.R

@@ -1,0 +1,2 @@
+knitr::knit("index.Rmd")
+rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex", citeproc = TRUE)

### 2019-08-22 cluster_results.R
putting sim pdfs

@@ -45,16 +45,4 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
 }
 
 
-setwd("./sim_results")
-
-method_names=c("SS","s=2","s=3","L2")
-nr_methods = length(method_names)
-
-combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
-# TODO: fix hardcoding
-combination_para <- expand.grid(c(5,2,4),c(0.5),c(30)) # combinations of the parameters, copied from compare_methods.R
-
-apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)
-
-setwd("../")
 

### 2019-08-23 index.Rmd
turned off cluster_results chunk
formatted some sim result pictures

@@
 # Abstract
-The clustering and classification of functional data, discretely-sampled data varying over a continuum, is frequently encountered in machine learning.
-While one can naively treat the data as vectors in Euclidean space, one may also regard each data point as a function.
-The challenge of analyzing infinite-dimensional functions can be ameliorated if the curves are smooth since the dimensionality is then only artificially high. 
+The clustering and classification of functional data, discretely-sampled curves varying over a continuum, is frequently encountered in machine learning.
+While one can naively treat the data as vectors in Euclidean space, one may also regard the data as functions.
+The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high. 
 This insight underpins functional data analysis, a subfield of statistics that studies functional data.
-For certain types of functional data such as classes of probability density functions and classes of warped curves of a common template function, there is further structure to exploit.
-Namely, when functional data lie on a manifold, the dimensionality at play might be even lower if the manifold is itself of low intrinsic dimension.
-<!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
-This work addresses the estimation of pairwise geodesic distances between functional manifold data that are observed with noise and which may live near, rather than on, a manifold.
-<!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
+For certain types of functional data such as probability density functions and warped curves of a common template function, there is further structure to exploit.
+Namely, certain functional data may lie on a manifold of low intrinsic dimension.
+For nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
+This work addresses the estimation of pairwise geodesic distances between noisy functional manifold data that live near, rather than on, a manifold.
+This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
 <!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
@@
 Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold. 
-Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distance-based functional classification.
+Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distance-based functional classification/clustering.
 
@@
 \tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
-\end{equation}
-where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
+\end{equation} where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
 The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
@@
 
-???Still need to describe parameters under study: SNR, denseness of observations ($K_{observed}$).
-Describe parameters that are fixed:  $K_smooth$ and com_grid.
-Describe tuning parameters for our method: $s$ and $h$.
+The tuning parameters for the proposed method were chosen as described at the end of Section \ref{proposed-method-for-estimating-geodesic-distances}.
+In step 1 of our procedure, we take $D = 100$ for the size of the common time grid.
+We will investigate how two parameters affect the performance of the various methods in consideration.
+The first one is the signal-to-noise ratio ???describe???
+which we fix at either $0.1$ (low) or $0.5$ (high).
+The second parameter is $K$, the dimension of the dscrietly-observed noisy $Y_i$'s.
+We set $K$ to either $30$ (sparse) or $100$ (abundant).
+
+We show the performance of the SS and L2 method versus our method for three different values of $s$.
+We omit the result of the square root velocity scenario because all methods perform comparably.
+We can see from Figures \ref{fig:normal} and \ref{fig:warping} that ???.
 
@@
 
-```{r, echo=FALSE, eval=TRUE, results='hide'}
+```{r, echo=FALSE, eval=FALSE, results='hide'}
 
@@
 # TODO: fix hardcoding
-combination_para <- expand.grid(c(5,2,4),c(0.5),c(30)) # combinations of the parameters, copied from compare_methods.R
+combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(30,100)) # combinations of the parameters, copied from compare_methods.R
 
@@
 \begin{figure}
-\includegraphics{sim_results/sce=5_SNR=high_Kobs=30}
-\caption{blah}
+\centering
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
+\caption{Normal density}
+\label{fig:normal}
+\end{figure}
+
+<!-- \begin{figure} -->
+<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_MSE} -->
+<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_isometry} -->
+<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_Pearson} -->
+<!-- \caption{Square root velocity} -->
+<!-- \end{figure} -->
+
+\begin{figure}
+\centering
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
+\caption{Warping function}
+\label{fig:warping}
 \end{figure}

### 2019-08-23 cluster_results.R

@@ -31,18 +31,30 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     results<- matrix(unlist(temp_res2,use.names=FALSE),ncol=3*nr_methods)
     
     ### We plot one boxplot per method and per assesment measure
-    pdf(paste(base_pattern_out,".pdf",sep=""),width=4,height=2)
     
-    par(mfrow=c(1,3))
-    # par(oma=c(3,3,3,3))
-    boxplot(results[,1:nr_methods],main="MSE",names=method_names)
-    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="isometry",names=method_names)
-    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson",names=method_names)
+    pdf(paste(base_pattern_out,"_MSE",".pdf",sep=""),width=16/4,height=3)
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
+    # par(
+    #   # Change the colors
+    #   col.main="red", col.lab="blue", col.sub="black",
+    #   # Titles in italic and bold
+    #   font.main=4, font.lab=4, font.sub=4,
+    #   # Change font size
+    #   cex.main=0.5, cex.lab=0.5, cex.sub=0.5
+    # )
+    boxplot(results[,1:nr_methods],main="MSE",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
+    dev.off()
+    
+    pdf(paste(base_pattern_out,"_isometry",".pdf",sep=""),width=16/4,height=3)
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
+    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="isometry",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
+    dev.off()
+    
+    pdf(paste(base_pattern_out,"_Pearson",".pdf",sep="") ,width=16/4,height=3)
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
+    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
+    dev.off()
     
[…]

### 2019-08-23 index.tex
.gitignore fix

@@
 The clustering and classification of functional data, discretely-sampled
-data varying over a continuum, is frequently encountered in machine
+curves varying over a continuum, is frequently encountered in machine
 learning. While one can naively treat the data as vectors in Euclidean
-space, one may also regard each data point as a function. The challenge
-of analyzing infinite-dimensional functions can be ameliorated if the
-curves are smooth since the dimensionality is then only artificially
-high. This insight underpins functional data analysis, a subfield of
-statistics that studies functional data. For certain types of functional
-data such as classes of probability density functions and classes of
-warped curves of a common template function, there is further structure
-to exploit. Namely, when functional data lie on a manifold, the
-dimensionality at play might be even lower if the manifold is itself of
-low intrinsic dimension. This work addresses the estimation of pairwise
-geodesic distances between functional manifold data that are observed
-with noise and which may live near, rather than on, a manifold. The
-proposed methodology first sends the observed functional data to the
-hidden manifold, estimated using subspace-constrained mean-shift.
-Geodesic distances are subsequently calculated by employing
-shortest-path algorithms on this estimated manifold. Improved estimation
-of the pairwise geodesic distance has beneficial implications for many
-downstream tasks in functional data analysis which we illustrate in the
-case of distance-based functional classification.
+space, one may also regard the data as functions. The challenge of
+analyzing infinite-dimensional objects can be ameliorated if the curves
+are smooth since the dimensionality is then only artificially high. This
+insight underpins functional data analysis, a subfield of statistics
+that studies functional data. For certain types of functional data such
+as probability density functions and warped curves of a common template
+function, there is further structure to exploit. Namely, certain
+functional data may lie on a manifold of low intrinsic dimension. For
+nonlinear manifolds, even basic operations such as addition and
+subtraction require special consideration. This work addresses the
+estimation of pairwise geodesic distances between noisy functional
+manifold data that live near, rather than on, a manifold. This setting
+cannot be tackled by classic manifold learning techniques which require
+data to live exactly on a manifold. The proposed methodology first sends
+the observed functional data to the hidden manifold, estimated using
+subspace-constrained mean-shift. Geodesic distances are subsequently
+calculated by employing shortest-path algorithms on this estimated
+manifold. Improved estimation of the pairwise geodesic distance has
+beneficial implications for many downstream tasks in functional data
+analysis which we illustrate in the case of distance-based functional
+classification/clustering.
 
@@
 
-???Still need to describe parameters under study: SNR, denseness of
-observations (\(K_{observed}\)). Describe parameters that are fixed:
-\(K_smooth\) and com\_grid. Describe tuning parameters for our method:
-\(s\) and \(h\).
+The tuning parameters for the proposed method were chosen as described
+at the end of Section
+\ref{proposed-method-for-estimating-geodesic-distances}. In step 1 of
+our procedure, we take \(D = 100\) for the size of the common time grid.
+We will investigate how two parameters affect the performance of the
+various methods in consideration. The first one is the signal-to-noise
+ratio ???describe??? which we fix at either \(0.1\) (low) or \(0.5\)
+(high). The second parameter is \(K\), the dimension of the
+dscrietly-observed noisy \(Y_i\)'s. We set \(K\) to either \(30\)
+(sparse) or \(100\) (abundant).
+
+We show the performance of the SS and L2 method versus our method for
+three different values of \(s\). We omit the result of the square root
+velocity scenario because all methods perform comparably. We can see
+from Figures \ref{fig:normal} and \ref{fig:warping} that ???.
 
 \begin{figure}
-\includegraphics{sim_results/sce=5_SNR=high_Kobs=30}
-\caption{blah}
+\centering
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
+\caption{Normal density}
+\label{fig:normal}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
+\caption{Warping function}
+\label{fig:warping}
 \end{figure}

### 2019-08-23 Analysis_phoneme.R
add analysis of phoneme data

@@ -1,0 +1,100 @@
+# Classification phoneme data, comparison with results from Delaigle and Hall
+
+library(reticulate)
+library(fda)
+library(scatterplot3d)
+library(igraph)
+library(ROCR)
+library(caret)
+library(vows)
+library(kmed)
+library(ElemStatLearn)
+library(KernSmooth)
+
+# this may need to be set at the beginning of the session 
+# Susan's
+# use_python('/usr/bin/python',required=TRUE)
+# use_python('/usr/local/bin/python3.7',required=TRUE)
+# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
+# Marie's
+# use_python('/anaconda3/envs/r-reticulate/bin/python',required=TRUE)
+use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+
+
+source('robust_isomap.R')
+source('pairwise_geo_estimation.R')
+source("weight_L2.R")
+source("processRealData.R")
+source("NP_FDA_Ferraty.R")
+
+PhonData_aa<-phoneme[which(phoneme$g=="aa"),]
+PhonData_ao<-phoneme[which(phoneme$g=="ao"),]
+Phon_aa_ao_temp<- rbind(PhonData_aa,PhonData_ao)
+n_aa<-nrow(PhonData_aa)
+n_ao<-nrow(PhonData_ao)
+n<- n_aa+n_ao
+true_group<- rep(FALSE,n)
+true_group[which(Phon_aa_ao_temp$g=="aa")] <-TRUE
+true_class<- rep(0,n)
+true_class[true_group]<- 1
+Phon_aa_ao<- as.matrix(Phon_aa_ao_temp[,-c(257,258)])
+grid<- 1:256
+
+# We smooth the data with local linear smoother
+Smooth_Phon<-matrix(nrow=n,ncol=256)
+for(i in 1:n){
+  Smooth_Phon[i,]<-locpoly(grid,Phon_aa_ao[i,],degree=1,kernel="normal",bandwidth=4,gridsize=256)$y
+}
+matplot(grid,t(Smooth_Phon[1:n_aa,]),type='l')
+matplot(grid,t(Smooth_Phon[(n_aa+1):n,]),type='l')
+
+# we just considerd grid = [1,50] as in Delaigle and Hall and we take a subset of each group (dataset too big)
[…]

### 2019-08-24 index.Rmd
cleaned up intro a bit

@@
 The clustering and classification of functional data, discretely-sampled curves varying over a continuum, is frequently encountered in machine learning.
-While one can naively treat the data as vectors in Euclidean space, one may also regard the data as functions.
-The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high. 
-This insight underpins functional data analysis, a subfield of statistics that studies functional data.
-For certain types of functional data such as probability density functions and warped curves of a common template function, there is further structure to exploit.
-Namely, certain functional data may lie on a manifold of low intrinsic dimension.
-For nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
+While one can naively treat the data as vectors in Euclidean space, one may do better to regard the data as functions.
+For certain types of functional data such as probability density functions or warped curves of a common template function, there is further structure to exploit.
+Namely, the curves may lie on a manifold of low intrinsic dimension.
 This work addresses the estimation of pairwise geodesic distances between noisy functional manifold data that live near, rather than on, a manifold.
@@
 Many statistical and machine learning methods rely on some measure of distance.
-For example, clustering and classification of functional data, discretely-sampled data varying over a continuum, is a common machine learning task in which distance plays a crucial role.
-The successful development of functional data analysis, a branch of statistics dedicated to the analysis of infinite-dimensional smooth curves, demonstrates that employing the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between functional observations treated as vectors.
+For example, clustering and classification of functional data, discretely-sampled curves varying over a continuum, is a common machine learning task in which distance plays a crucial role.
+In a naive treatment, the data may be simply processed as vectors in Euclidean space.
+Doing so however risks losing the richness of functional data; indeed, one may do better to regard the data as functions.
+
+The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high.
+This insight underpins functional data analysis, a subfield of statistics that studies functional data. This has many implications for the processing of functional data. 
+For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between functional observations treated as vectors.
+
 In this work, we advocate that for functional manifold data, i.e.
-functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification of functional manifold data.
-In other words, when the manifold hypothesis is plausible, e.g.\ classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L_2$ distance.
+functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
+In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L_2$ distance.
+The manifold setting presents its own challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
 
@@
 The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_{\M}(X_i,X_j):=\inf \{l(\gamma): \gamma \text{ is piecewise-smooth function from } [a,b] \text{ to } \M, \gamma(a) = X_i, \gamma(b) = X_j \}$$
+$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[a,b] \to \M, \gamma(a) = X_i, \gamma(b) = X_j} l(\gamma)$$
 where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
-is the length of the curve $\gamma$. ???Is it clear how $l(\gamma)$ depends on $g$???
+is the length of the piecewise-smooth curve $\gamma$. 
+<!-- ???Is it clear how $l(\gamma)$ depends on $g$??? -->
 
@@
 Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
-However, most shortest-path algorithms critically assume observations live on the manifold ???add reference???
-and are thus likely to fail for functional data which are almost certainly observed with noise.
+However, most shortest-path algorithms critically assume observations live on the manifold [@Yeh2005] and are thus likely to fail for functional data which are almost certainly observed with noise.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
@@
 Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
-For example, we may employ spline smoothing ???add reference for spline smoothing???
-to recover the functional versions of the raw data, i.e.\
+For example, we may employ spline smoothing [@Ramsay2005] to recover the functional versions of the raw data, i.e.\
 \begin{equation}\label{eq_spline_smoothing} 
@@
 The procedure is as follows.
-\begin{enumerate}
-\item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+
+1.
+Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
 Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
-\item Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
+2.
+Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
 Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
-\item Use multidimensional scaling ???add reference??? to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
-\end{enumerate}
+3. Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
+
 Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weigthed graph but they do not produce the weighted graphs themselves.
@@
 <!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
-\begin{enumerate}
-\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$.
+1.
+Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$.
 Let $s$ be a positive integer, much smaller than $D$.
 Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ are preserved.
-\item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
-\item Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
-\end{enumerate}
+2. Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
+3. Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
+
 Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
@@
 
-\textbf{Isometric functional manifold of normal density functions}
+### Isometric functional manifold of normal density functions
 <!-- Scenario 2 in sim_functional_data -->
@@
 d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
-&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
-&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
+% &=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\ 
+&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta   \\
 &\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
-&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
+&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}}
 \end{eqnarray*} 
-where the approximation comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
+where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
 We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$, the Euclidean distance between the $\beta$'s.
 
-\textbf{Functional manifold of square root velocity functions}
+### Functional manifold of square root velocity functions
 <!-- Scenario 4 in sim_functional_data -->
@@
 
-\textbf{Functional manifold of warping functions}
+### Functional manifold of warping functions
 <!-- Scenario 5 in sim_functional_data -->
@@
 
-```{r, echo=FALSE, eval=FALSE, results='hide'}
+```{r, echo=FALSE, eval=FaALSE, results='hide'}
 

### 2019-08-24 cluster_results.R

@@ -33,7 +33,7 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     ### We plot one boxplot per method and per assesment measure
     
     pdf(paste(base_pattern_out,"_MSE",".pdf",sep=""),width=16/4,height=3)
-    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
     # par(
     #   # Change the colors
     #   col.main="red", col.lab="blue", col.sub="black",
@@ -42,17 +42,17 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     #   # Change font size
     #   cex.main=0.5, cex.lab=0.5, cex.sub=0.5
     # )
-    boxplot(results[,1:nr_methods],main="MSE",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
+    boxplot(results[,1:nr_methods],main="MSE",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
     dev.off()
     
     pdf(paste(base_pattern_out,"_isometry",".pdf",sep=""),width=16/4,height=3)
-    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
-    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="isometry",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
+    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="Isometry",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
     dev.off()
     
     pdf(paste(base_pattern_out,"_Pearson",".pdf",sep="") ,width=16/4,height=3)
-    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
-    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson",names=method_names,cex.main=0.5,cex.names=0.5,cex.axis=0.5)
[…]

### 2019-08-25 index.Rmd
finished sim sec

@@
   html_document:
-    fig_caption: true
-bibliography: references.bib
+    fig_caption: yes
 ---
@@
 functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
-In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L_2$ distance.
+In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L^2$ distance.
 The manifold setting presents its own challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
@@
 Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014].
+Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows \citep{Lin2014}.
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
@@
 
-<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->
+<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L^2$ distance.??? -->
 
@@
 Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error.
-Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
-However, most shortest-path algorithms critically assume observations live on the manifold [@Yeh2005] and are thus likely to fail for functional data which are almost certainly observed with noise.
+Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall \citep{Floyd1962} may be used.
+However, most shortest-path algorithms critically assume observations live on the manifold \citep{Yeh2005} and are thus likely to fail for functional data which are almost certainly observed with noise.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
-This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
+This is because smoothing techniques in functional data analysis are designed for the $L^2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$.
 This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs.
@@
 The first step of IsoGeo constructs the weighted graph $G$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$.
-Specifically, we begin by constructing the complete ???explain what complete means???
-weighted graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Specifically, the complete weighted graph $G_c$ is constructed with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
 Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$.
@@
 
-We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances.
-Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.
+We perform simulations to study various estimators of pairwise geodesic distances.
+We compare the performance of our method to each of the following four alternatives.
+The first takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+Note that this procedure is only sensible if all the time grids $T_1,\ldots,T_n$ are the same.
 
-The first metric assesses near-$\epsilon$ isometry, which we define as the percentage of estimated pairwise geodesic distances between $1-\epsilon$ and $1+\epsilon$ of the corresponding true pairwise geodesic distance.
-We form the receiver operating curve with $\epsilon$ on the $x$-axis and near-$\epsilon$ isometry on the $y$-axis.
-The metric is the area under the receiver operating curve.
+For the next two alternatives, we first smooth the discretely-sampled noisy functional data using splines as described in (\ref{eq_spline_smoothing}).
+Then, we either estimate pairwise distance between the smoothed $\tilde X_1,\ldots,\tilde X_n$ evaluated on a common time grid of $D$ points using IsoGeo (denoted Spline IsoGeo) or the $L^2$ distance (denoted Spline L2).
 
-The second metric $||d-\hat d||/||d||$ is the Frobenius norm of the estimation error matrix $d-\hat d$, relative to the Frobenius norm of the true distance matrix $d$.
-The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
-
-## Alternative estimators of geodesic distance [Marie]
-We compare the performance of our method to each of the following four methods.
-The first method, denoted RD for raw data, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
-Note that this procedure is only sensible if all the grids $T_1,\ldots,T_n$ are the same.
-In the second method, we first smooth the discretely-sampled noisy functional data and then estimate their pairwise distance using the $L^2$ distance.
-
-The third alternative, denoted SS for spline smoothing, applies IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}).
-<!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing.
-Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
-The final alternative we consider is P-ISOMAP, denoted pI, which is the two-step procedure developed in @ChenMuller2012 designed to handle noisy functional data.
+The final alternative we consider is P-ISOMAP which is the two-step procedure developed in @ChenMuller2012 for handling noisy functional data.
 In the first step, a penalty is incorporated to robustify the construction of the weighted graph.
@@
 
-We ran Chen and Muller's P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric.
+We ran P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric.
 We found that the penalty parameter chosen in this way was always zero, thus reducing hte P-ISOMAP procedure to standard ISOMAP.
 Given this, we did not make further comparison to P-ISOMAP in our simulation study.
+We also do not present the results for the RD method which always produced very bad estimates of the geodesic distance.
 
-## Simulation scenarios 
+Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.
+The first metric assesses near-$\epsilon$ isometry, which we define as the percentage of estimated pairwise geodesic distances between $1-\epsilon$ and $1+\epsilon$ of the corresponding true pairwise geodesic distance.
+We form the receiver operating curve with $\epsilon$ on the $x$-axis and near-$\epsilon$ isometry on the $y$-axis.
+The metric is the area under this receiver operating curve.
+Thus a good estimator has high values for the near-isometry metric.
+
+The second metric $||d-\hat d||/||d||$ is the Frobenius norm of the estimation error matrix $d-\hat d$, relative to the Frobenius norm of the true distance matrix $d$.
+The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
+A good geodesic distance estimator has low Frobenius norma nd high Pearson correlation coefficient.
+
 We shall consider three functional manifolds in our simulation study.
-As of yet, there is little work as to how to sample from a functional manifold.
+Throughout, we make use of a helpful property of one-dimensional manifolds.
+If $X \in \M$ with the $L^2$ metric tensor depends on a parameter $\theta \in \mathbb R$, then
+$$
+d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2}. 
+$$
+
+We also need to make a disclaimer about our sampling technique below.
+Proper sampling from a manifold remains an open problem.
 Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
-How to properly sample from a functional manifold could be interesting future work.
-For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.
+For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very narrow support of the manifold parameters.
 
-### Isometric functional manifold of normal density functions
+## Isometric functional manifold of normal density functions
 <!-- Scenario 2 in sim_functional_data -->
-This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$.
-We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
-First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$.
+We take Manifold 2 in @ChenMuller2012 and set the variance of the normal density to be $1$.
+Let $X_\beta: [a,b] \to \R$ be given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
+Consider $$\M = \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L^2$ inner product as the metric tensor.
+First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ does not always stay inside of $\M$.
 Thus the geodesic distance will not coincide with the $L^2$ distance.
@@
 \begin{eqnarray*}
-d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
+d_\M(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
 % &=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\ 
@@
 \end{eqnarray*}
-where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
+\noindent where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from integrating on $[a,b]=[-4,4]$ rather than $\R$. 
 We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$, the Euclidean distance between the $\beta$'s.
 
-### Functional manifold of square root velocity functions
+## Functional manifold of square root velocity functions
 <!-- Scenario 4 in sim_functional_data -->
 It was shown in @Joshi2007 that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
-They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
-Note that this concides with the $L_2[0,1]$ inner product.
-@Joshi2007 showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
-We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$.
-That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$.
+They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
+Note that this concides with the $L^2[0,1]$ inner product.
+It was shown in [@Joshi2007] that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$.
+That is consider $$ M = \{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$.
 We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$.
@@
 
-### Functional manifold of warping functions
+## Functional manifold of warped functions
 <!-- Scenario 5 in sim_functional_data -->
-This is based on Equations 17 and 18 in @Kneip2008 but with $z_{i1}, z_{i2}$ both set to $1$.
-(Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
+This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$.
+(Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
 
@@
 We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
-The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha$$ which is estimated using numerical integration.
+The geodesic distance is then $$ d_\M(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$ We estimate this using numerical integration.
 
@@
 We will investigate how two parameters affect the performance of the various methods in consideration.
-The first one is the signal-to-noise ratio ???describe???
+The first one is the signal-to-noise ratio ???[Marie] describe???
 which we fix at either $0.1$ (low) or $0.5$ (high).
@@
 
-We show the performance of the SS and L2 method versus our method for three different values of $s$.
-We omit the result of the square root velocity scenario because all methods perform comparably.
-We can see from Figures \ref{fig:normal} and \ref{fig:warping} that ???.
+We show the performance of the Spline IsoGeo and Spline L2 method versus our method for three dimensions $s=1,2,3$.
+We present the result of the square root velocity manifold scenario in the Appendix.
+We can see from Figures \ref{fig:normal} and \ref{fig:warping} that spline smoothing followed by either IsoGeo or the $L^2$ distance calculation estimates the pairwise geodesic distances reasonably well though not as well as the proposed method for $s=2$ and $s=3$.
+Interestingly, when $s$ in our method is not selected properly, for instance when $s=1$, our estimator performs poorly in terms of the relative Frobenius norm although still better than Spline IsoGeo and Spline L2 in the other two metrics.
+It may be possible that for certain downstream tasks such as clustering or classification, precise estimation of the geodesic distance is less important than an estimation that respects the ordering of the distances.
+In such cases, the near-isometry and Pearson correlation coefficients may give a better indication of downstream performance of the geodesic distance estimator.
 
@@
 
-```{r, echo=FALSE, eval=FaALSE, results='hide'}
+```{r, echo=FALSE, eval=TRUE, results='hide'}
 
@@
 
-method_names=c("SS","s=1","s=2","s=3","L2")
+method_names=c("Spline IsoGeo","s=1","s=2","s=3","Spline L2")
 nr_methods = length(method_names)
@@
 \includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
-\caption{Normal density}
+\caption{Manifold of $N(\beta,1)$ probability density functions. Across all three metrics, our method with $s=2$ performs the best, followed by our method with $s=3$.}
 \label{fig:normal}
 \end{figure}
-
-<!-- \begin{figure} -->
-<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_MSE} -->
-<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_isometry} -->
-<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_Pearson} -->
-<!-- \caption{Square root velocity} -->
-<!-- \end{figure} -->
 
@@
 \includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
-\caption{Warping function}
+\caption{Manifold of warped functions. The best performing estimators are given by our method with $s=2$ and then $s=3$.}
 \label{fig:warping}
@@
 
-Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance.
-Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?
+Suppose we take scenario 1 where the geodesic distance coincides with the $L^2$ distance.
+Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L^2$ distance?
 
@@
 $$
-We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.
+We shall compare our method to using $L^2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.
 
@@
 
-# References
+# Acknowledgement
+The authors would like to acknowledge support from the ARC Centre of Excellence for Mathematical & Statistical Frontiers (ACEMS) which enabled the two authors to work on the project at the University of Melbourne.
+
+# Appendices
+\begin{figure}
+\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_MSE}
+\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_isometry}
+\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
+\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
+\end{figure}
+
+\bibliography{references.bib}
+\bibliographystyle{aaai.bst}

### 2019-08-25 cluster_results.R

@@ -42,17 +42,17 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     #   # Change font size
     #   cex.main=0.5, cex.lab=0.5, cex.sub=0.5
     # )
-    boxplot(results[,1:nr_methods],main="MSE",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,1:nr_methods],main="relative Frobenius",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
     dev.off()
     
     pdf(paste(base_pattern_out,"_isometry",".pdf",sep=""),width=16/4,height=3)
     par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
-    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="Isometry",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="near-isometry",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
     dev.off()
     
     pdf(paste(base_pattern_out,"_Pearson",".pdf",sep="") ,width=16/4,height=3)
     par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
-    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
     dev.off()
     
     # mtext(base_pattern,outer=TRUE,cex=1.5)

### 2019-08-25 run_this.R

@@ -1,2 +1,3 @@
 knitr::knit("index.Rmd")
-rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex", citeproc = TRUE)
+rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex")
+system("pdflatex paperAAAI.tex")

### 2019-08-28 index.Rmd
fixed paperAAAI preamble to comply

@@
     fig_caption: yes
+bibliography: references.bib
 ---
@@
 
-# Abstract
-The clustering and classification of functional data, discretely-sampled curves varying over a continuum, is frequently encountered in machine learning.
-While one can naively treat the data as vectors in Euclidean space, one may do better to regard the data as functions.
-For certain types of functional data such as probability density functions or warped curves of a common template function, there is further structure to exploit.
-Namely, the curves may lie on a manifold of low intrinsic dimension.
-This work addresses the estimation of pairwise geodesic distances between noisy functional manifold data that live near, rather than on, a manifold.
-This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
-<!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
-The proposed methodology first sends the observed functional data to the hidden manifold, estimated using subspace-constrained mean-shift.
-Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold. 
-Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distance-based functional classification/clustering.
-
 # Introduction 
@@
 This insight underpins functional data analysis, a subfield of statistics that studies functional data. This has many implications for the processing of functional data. 
-For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between functional observations treated as vectors.
+For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between noisy functional observations treated as vectors.
 
 In this work, we advocate that for functional manifold data, i.e.
-functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
+functional data that lie on a manifold, the geodesic distance is more appropriate than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
 In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L^2$ distance.
@@
 Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows \citep{Lin2014}.
+Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014].
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
 The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[a,b] \to \M, \gamma(a) = X_i, \gamma(b) = X_j} l(\gamma)$$
-where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
+$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[0,1] \to \M, \gamma(0) = X_i, \gamma(1) = X_j} l(\gamma)$$
+where $$ l(\gamma) := \int_0^1 || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
 is the length of the piecewise-smooth curve $\gamma$. 
@@
 Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error.
-Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall \citep{Floyd1962} may be used.
-However, most shortest-path algorithms critically assume observations live on the manifold \citep{Yeh2005} and are thus likely to fail for functional data which are almost certainly observed with noise.
+Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
+However, most shortest-path algorithms critically assume observations live on the manifold [@Yeh2005] and are thus likely to fail for functional data which are almost certainly observed with noise.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
@@
 <!-- Scenario 5 in sim_functional_data -->
-This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$.
+This is based on Equations 17 and 18 in [@Kneip2008] but with $z_{i1}, z_{i2}$ both set to $1$.
 (Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
@@
 
-\bibliography{references.bib}
-\bibliographystyle{aaai.bst}
+# References

### 2019-08-28 paperAAAI.tex
finished sim sec
fixed paperAAAI preamble to comply

@@
 
-\usepackage{hyperref}       % hyperlinks
+\def\tightlist{}
+
+% 
+% 
 \usepackage{url}            % simple URL typesetting
@@
 \usepackage{microtype}      % microtypography
-
+% 
+% 
+% 
 \usepackage{amsmath}
@@
 \usepackage{epsfig}
-\usepackage[usenames]{color}
-\usepackage{graphicx}
 
@@
 % \natbib} -- This package is specifically forbidden -- use the following workaround:
+%\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}} 
+%\newcommand{\citep}{\cite} 
+%\newcommand{\citealp}[1]{\citeauthor{#1} \citeyear{#1}}
 % \usepackage{savetrees} -- This package is specifically forbidden
@@
 % command to correct it. You may not alter the value below 2.5 in
-\title{The hidden manifold distance for functional data}
+\title{The Hidden Manifold Distance for Functional ata}
 %Your title must be in mixed case, not sentence case. 
@@
 
+\begin{abstract}
+In many machine learning applications, the observed data is a
+discretely-sampled realization of a continuous process. While one can
+naively treat such data as vectors in Euclidean space, one may do better
+to regard them as functional data. For certain types of functional data,
+there is further structure to exploit, e.g.~probability density
+functions or warped curves of a common template function. Namely, the
+underlying curves of interest may lie on a manifold of low intrinsic
+dimension. This work addresses the estimation of pairwise geodesic
+distances between functional manifold data that are observed with noise,
+causing them to lie off the manifold. This setting falls outside of
+classic manifold learning techniques which require data to live exactly
+on or very near a manifold. The proposed methodology first sends the
+observed functional data to the hidden manifold, estimated using
+subspace-constrained mean-shift. Geodesic distances are subsequently
+calculated by employing shortest-path algorithms on this estimated
+manifold. Improved estimation of the pairwise geodesic distance has
+beneficial implications for many downstream tasks in functional data
+analysis which we illustrate in the case of distance-based functional
+classification/clustering.
+\end{abstract}
+
 \input{index}

### 2019-08-28 run_this.R

@@ -1,3 +1,3 @@
 knitr::knit("index.Rmd")
-rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex")
-system("pdflatex paperAAAI.tex")
+rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex",citeproc=FALSE)
+# system("pdflatex paperAAAI.tex")

### 2019-08-28 index.Rmd
some more writing

@@
 Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error.
-Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
-However, most shortest-path algorithms critically assume observations live on the manifold [@Yeh2005] and are thus likely to fail for functional data which are almost certainly observed with noise.
+Then the geodesic distance $d_\M(X_i,X_j)$ can be estimated using any of a variety of shortest-path algorithms, e.g.\ the Floyd-Warhsall [@Floyd1962] algorithm.
+However, most shortest-path algorithms critically assume observations live near the manifold with little noise[@Yeh2005] and are thus likely to fail for functional data which can be observed with low signal-to-noise ratio.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
@@
 [@Dimeglio2014] also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP that is less sensitive to outliers.
-We will discuss P-ISOMAP and robust-ISOMAP in more details in the simulation section.
+We will discuss P-ISOMAP and robust-ISOMAP in more details below.
 
@@
 # Proposed method for estimating geodesic distances 
-Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
+Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$.
+Let the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
 We assume that each observation grid $T_1,\ldots,T_n$ is dense, i.e.\ $K$ is large.
@@
 Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm [@Ozertem2011].
-Let $\hat M \subset L^2([a,b],\mathbb R)$ denote the collection of basins of attraction of the estimated probability density function of $X$.
+Let the set $\hat M \subset L^2([a,b],\mathbb R)$ be comprised of the basins of attraction of the estimated probability density function of $X$.
 The SCMS algorithm sends each point $\tilde X_i \in L^2([a,b],\mathbb R)$ to a point in $\hat M$.
-This destination is unique and we denote it $\tilde X_i^{\hat \M}$.
+We denote this destination as $\tilde X_i^{\hat \M}$.
 We expect the points $\tilde X^{\hat \M}_1,\ldots,\tilde X^{\hat \M}_n \in \hat M$ to lie close to the real manifold $\M$.
@@
 Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
-Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+Specifically, this is the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
 3. Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
@@
 In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP.
-The first step of IsoGeo constructs the weighted graph $G$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$.
-Specifically, the complete weighted graph $G_c$ is constructed with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+The first step of IsoGeo constructs the complete weighted ???Marie: does this mean all weights are nonzero???
+graph $G_c$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$.
 Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$.
@@
 The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
-A good geodesic distance estimator has low Frobenius norma nd high Pearson correlation coefficient.
+A good geodesic distance estimator has low Frobenius norm and high Pearson correlation coefficient.
 
@@
 $$
-d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2}. 
+d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2} \,d\theta
 $$

### 2019-08-30 MarieDataAnalysis.tex
final writing touches

--- 2019-07-29 index.tex
@@
-\documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={The hidden manifold distance for functional data},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{The hidden manifold distance for functional data}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
-\begin{document}
-\maketitle
-
-\newcommand {\To}{\rightarrow}
-\newcommand {\TO}{\Rightarrow}
-\newcommand {\R}{\mathbb{R}}
-\newcommand {\Prob}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand {\cov}{\textrm{Cov}}
-\newcommand {\var}{\textrm{Var}}
-\newcommand {\1}{\textrm{\textbf{1}}}
-\newcommand{\M}{\mathcal{M}}
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
-Functional data analysis is the statistical analysis of smooth
-infinite-dimensional curves. The challenge of analyzing
-infinite-dimensional objects is ameliorated by the fact that, since
-observations are smooth over their domain, the dimensionality of the
-curves is only artifically high. Less studied in functional data
-analysis is the idea that the curves may actually live in a submanifold
-with low intrinsic dimension. Interesting classes of functional manifold
-data include the class of probability density functions and the class of
-warped curves of a common template function. Since these spaces are not
-vector spaces, even basic operations such as addition and subtraction
-require special consideration. In this work we set ourselves the task of
-calculating the pairwise geodesic distance between functional
-observations that, due to noise, live near a non-linear manifold rather
-than exactly on the manifold. We will show that good estimation of the
-geodesic distance cannot be accomplished by naive methods such as
-smoothing followed by a shortest-path algorithm. Instead, we propose a
-dedicated technique for pairwise geodesic distance estimation by first
-projecting the functional data to the underlying hidden manifold and
-then perofrming operations on this hidden manifold. We conclude the work
-by showing the implications for many downstream tasks in functional data
-analysis, in particular distanced-based functional classification.
-
-\hypertarget{introduction-susan}{%
-\section{Introduction {[}Susan{]}}\label{introduction-susan}}
-
-Let \(X_1,\ldots,X_n\) be a sample of \(n\) independant realizations of
-a random variable \(X\) that takes value in the Hilbert space
-\(L^2([a,b],\mathbb{R})\). We additionally assume that the function
-\(X\) belongs to a low-dimensional nonlinear manifold
-\(\mathcal{M}\subset L^2([a,b],\mathbb{R})\), in which case a geodesic
-distance taking into account the intrinsic structure of \(\mathcal{M}\)
-rather than the \(L^2\) distance should be used as a measure of distance
-between two functions in \(\mathcal{M}\). As many statistical methods
-depend on some measure of distance, downstream benefits of using the
-geodesic distance can be expected.
-
-???insert example that shows pairwise geodesic distances give visually
-more interesting results than pairwise \(L_2\) distance.???
-
-Given perfect functional observations, i.e.~with no measurement error
-and on a very dense domain grid, the estimation of the geodesic distance
-can be accomplished straightforwardly using algorithms such as Floyd's.
-However, functional data are observed with noise in such a way that the
-observed functional data no longer live on a manifold. Though powerful
-techniques exist for smoothing discretely observed noisy functional
-data, the smoothing is designed for the \(L_2\) recovery of the original
-curve, not to ensure the recovered functional versions of the data lie
-on or close to the functional manifold \(\mathcal{M}\).
-
-With this challenge in mind, we put forth a technique for the specific
-task of estimating pairwise geodesic distances
-\(\{ d_\mathcal{M}(X_i,X_j\}_{i>j}\), to be defined when we have access
-only to discretely-observed noisy functional observations that possibly
-live off the true manifold. Briefly outline the methodology: MDS plus
-density ridge are used in a cool way to estimate (consistently?, let's
-hope..) pairwise geodesic distances.
-
-\hypertarget{related-work-susan}{%
-\subsection{Related Work {[}Susan{]}}\label{related-work-susan}}
-
-(Chen and Muller 2012)
-
-(Lin and Yao 2017)
-
-(Dimeglio et al. 2014)
-
-\hypertarget{setting-susan}{%
-\subsection{Setting {[}Susan{]}}\label{setting-susan}}
-
-Let \(X_1,\ldots,X_n\) be independent realizations of a random function
-\(X\in\mathcal{M}\subset L^2([a,b],\mathbb{R})\), where \(\mathcal{M}\)
-is an unknown manifold. Suppose that each curve \(X_i\) is observed with
-measurements errors on a grid \(T_i=(t_{i1},\ldots,t_{iK})\), i.e.~we
-observe a sample of \(K\)-dimensional vectors \(Y_1,\ldots,Y_n\) with
-\(Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}\), where the random variables
-\(\epsilon_{ij}\) are of mean zero and uncorrelated with each other. We
-assume that the grids \(T_1,\ldots,T_n\) are dense.
-
-Let \(\tilde X_1,\ldots,\tilde X_n\) denote the functional versions of
-the raw data by some smoothing method. Our proposed methodology is
-unspecific to the smoothing method employed. For simplicity however,
-let's say we employ spline smoothing to recover the functional versions
-of the raw data, i.e.\\
-\[ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}\]
-where \(\lambda>0\) is a tuning parameter controlling the smoothness of
-\(\tilde X_i\).
-
-\hypertarget{relevant-differential-geometry-concepts-susan}{%
-\subsection{Relevant differential geometry concepts
-{[}Susan{]}}\label{relevant-differential-geometry-concepts-susan}}
-
-We first rigorously define the concepts of geodesic and geodesic
-distance. Following (Lin et al. 2014), let \((\mathcal{M},g)\) be a
-\(d\)-dimensional Riemannian manifold, where \(g\) is a Riemannian
-metric tensor on \(\mathcal{M}\) which can be used to assign a metric on
-the manifold. Specifically, for each point \(p\) on the manifold, the
-Riemannian metric tensor \(g\) has an inner product \(g_p\) on the
-tangent space \(T_p \mathcal{M}\). The norm of a tangent vector
-\(v \in T_p \mathcal{M}\) is defined as \[||v|| = \sqrt{g_p(v,v)}.\] If
-\(\gamma: [a,b] \subset \mathbb{R}\to \mathcal{M}\) is a smooth curve,
-its length is then defined as
-\[ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt .\] The
-geodesic distance between two points \(p,q\) on the manifold
-\(\mathcal{M}\), based on this metric tensor \(g\), is defined as
-\[ d_g(p,q):=\inf \{l(\gamma): \gamma:[a,b] \to \mathcal{M}\text{ piecewise smooth}, \gamma(a) = p, \gamma(b) = q]  \}.\]
-
-\hypertarget{proposed-method-for-estimating-geodesic-distances-marie}{%
-\section{Proposed method for estimating geodesic distances
-{[}Marie{]}}\label{proposed-method-for-estimating-geodesic-distances-marie}}
-
-We first describe an important helper algorithm which we call IsoGeo. It
-is a subroutine of the classic Isomap procedure. Isomap is a three steps
-procedure that takes as input a set of points
-\(x_1,\ldots,x_n\in \mathbb{R}^D\) and produces an embedding of the
-input data in the space \(\mathbb{R}^d\) with \(d<D\), that preserves
-pairwise geodesic distances.
-
-\begin{enumerate}
-\item The first step consists in constructing a weighted graph $G$ with nodes corresponding to the input data and for which the weight of an edge between two nodes $x_i$ and $x_j$ is equal to $||x_i-x_j ||_{\mathbb{R}^D}$.
-The graph $G$ is constructed such that a node is connected to a fix number $N$ of neighbors or to all nodes that are at a distance smaller than a given value $\epsilon$.
-\item In the second step, the pairwise geodesic distances $d_\mathcal{M}(x_i,x_j)$
-are estimated based on $G$.
-The geodesic distance between two nodes is simply the length of the shortest path in $G$ between these two nodes, which we calculate with the Floyd-Warshall algorithm [@Warshall1962]
-<!-- or Dijkstra’s  which one do we actually use in our implementation, the FW or D? -->
-\item The last step consists in using MDS on the matrix of pairwise geodesic distances obtained in the first step in order to obtain an embedding in $\mathbb{R}^d$.
-\end{enumerate}
-
-In what follows, we call IsoGeo the procedure of estimating pairwise
-geodesic distances with the two first steps of Isomap. Note that Floyd
-or Dijkstra are methods to find the smallest path given a weigthed graph
-but they are not related to how to calculate that graph (which is the
-difficult part, depend on number of neighbors, many method out there to
-calculate that graph in a robust way, P-ISOMAP is one of these).
-
-Our method is described by the following steps:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Obtain a \(s\)-dimensional representation
-  \(\tilde X^s_1,\ldots,\tilde X^s_n\) of the functions
-  \(\tilde X_1,\ldots,\tilde X_n\) that ``preserves'' the pairwise
-  \(L^2\) distances by using MDS.
-\item
-  Obtain
-  \(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
-  a ``projection'' of \(\tilde X^s_1,\ldots,\tilde X^s_n\) onto a ridge
-  \(\hat \mathcal{M}\) which is computed with the subspace constrained
-  mean-shift algorithm (Ozertem and Erdogmus 2011).
-\item
-  Use IsoGeo to approximate the pairwise geodesic distances
-  \(\{d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j)\}_{i>j}\)
-  and define the \(n\) by \(n\) matrix \[
-  \hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
-   d_{\hat \mathcal{M}}(\tilde X^{s,\hat \mathcal{M}}_i,\tilde X^{s,\hat \mathcal{M}}_j) & \textrm{if $i\neq j$,}\\
-   0 & \textrm{otherwise.}
-    \end{array} \right.
-  \]
-\end{enumerate}
-
-Since the ridge estimation obtained from noisy measurements of a
-manifold should approximate well the manifold (Genovese et al. 2014), we
-expect the points
-\(\tilde X^{s,\hat \mathcal{M}}_1,\ldots,\tilde X^{s,\hat \mathcal{M}}_n\)
-to lie close to the real manifold \(\mathcal{M}\) and then
-\(d_{\hat \mathcal{M}}\) to be close to \(d_\mathcal{M}\). Ridge
-estimation suffers from the curse of dimensionality, this is why we
-first reduce the dimension of our data with MDS and then apply the
-shift-mean algorithm to estimate the ridge.
-
-\hypertarget{selection-of-tuning-parameters-marie}{%
-\subsection{Selection of tuning parameters
-{[}Marie{]}}\label{selection-of-tuning-parameters-marie}}
-
-Describe heuristics. We pick the bandwidth \(h\) in the subspace
-constrained mean-shift using Equation (A1) of (Chen et al. 2015).
-
-\hypertarget{simulation-study}{%
-\section{Simulation study}\label{simulation-study}}
-
-We perform a simulation study to ascertain the efficacy of our method
-for estimating pairwise geodesic distances for discretely-observed noisy
-functional data.
-
-\hypertarget{alternative-estimators-of-geodesic-distance-marie}{%
-\subsection{Alternative estimators of geodesic distance
-{[}Marie{]}}\label{alternative-estimators-of-geodesic-distance-marie}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{(Raw Data) RD} Apply IsoGeo on the raw data \(Y_1,\ldots,Y_n\)
-  to obtain an estimator \(\hat G_{\textrm{RD}}\). Note that this
-  procedure can only be used if the grid \(T_i\) is the same for each
-  \(i\).
-\item
-  \textbf{(Spline Smoothing) SS} Transform each vector \(Y_i\) into a
-  function \(\tilde X_i\) by spline smoothing. Apply IsoGeo on the
-  vectors \(\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n\), where
-  \(t_1,\ldots,t_K\) is a regular grid of \([a,b]\), to obtain an
-  estimator \(\hat G_{\textrm{SS}}\).
-\item
-  \textbf{(P-ISOMAP) pI} Transform each vector \(Y_i\) into a function
-  \(\tilde X_i\) by spline smoothing. Calculate the weighted graph of
-  step I of Isomap with a penalty (describe the penalty of Chen and
-  Muller). Apply step II of Isomap to obtain \(\hat G_{\textrm{pI}}\).
-\item
-  \textbf{(Random Projection) RP} same method as our but change step 2
-  by : obtain \(s\)-dimensional representation by random projection and
-  setp 4 by obtain \(\hat G\) using a ensemble method.
-\end{itemize}
-
-We ran Chen and Muller's P-ISOMAP separately in Matlab with their
-automatic penalty selection and discovered that ???.
-
-\hypertarget{performance-metrics-susan}{%
-\subsection{Performance metrics
-{[}Susan{]}}\label{performance-metrics-susan}}
-
-We describe three different metrics we use to assess the quality of a
-pairwise geodesic distance estimator. This is the ROC curve with
-\(\epsilon\) on the \(x\)-axis and degree to which near-\(\epsilon\)
-isometry holds, i.e.~the percentage of estimated pairwise distances
-between \(1-\epsilon\) and \(1+\epsilon\) of the truth pairwise
-distance.
-
-\hypertarget{simulation-scenarios-susan}{%
-\subsection{Simulation scenarios
-{[}Susan{]}}\label{simulation-scenarios-susan}}
-
-Acknowledge that we are sampling on a very concentrated measure to make
-sure functional manifold is well sampled. Cite Diaconis et. al's
-``Sampling from a manifold''. How to sample properly from a functional
-manifold could be interesting future work.
-
-\begin{itemize}
-\tightlist
-\item
-  warping manifold (scenario 5 in sim\_functional\_data.R)
-\item
-  pdf manifold (scenario 2 in sim\_functional\_data.R)
-\item
-  square root pdf manifold (scenario 4 )
-\end{itemize}
-
-\hypertarget{results}{%
-\subsubsection{Results}\label{results}}
-
-Submit compare\_methods.sh as slurm job using sbatch.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\ExtensionTok{sbatch}\NormalTok{ compare_methods.sh}
-\end{Highlighting}
-\end{Shaded}
-
-Alternatively, we can run the R script compare\_methods.R locally by
-removing the slurm\_arrayid chunk
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{R compare_methods.R}
-\end{Highlighting}
-\end{Shaded}
-
-Simulation results are then visualized by invoking cluster\_results.R.
-There's some unfortunate hardcoding still in cluster\_results.R
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{R cluster_results.R}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}{%
-\subsection{The cost of employing proposed method when manifold is flat
-{[}Susan/Marie?{]}}\label{the-cost-of-employing-proposed-method-when-manifold-is-flat-susanmarie}}
-
-Suppose we take scenario 1 where the geodesic distance coincides with
-the \(L_2\) distance. Then what is the cost of employing our distance
-estimator compared to smoothing the data and doing numerical integration
-to find the \(L_2\) distance?
-
-\hypertarget{distance-based-functional-classification-marie}{%
 \section{Distance-based functional classification
-{[}Marie{]}}\label{distance-based-functional-classification-marie}}
+{[}Marie{]}}\label{distance-based-functional-classification-marie}
 
@@
 each functional object \(x\) is a binary \(y\) indicating class
-membership. Consider the classifier proposed in Ferraty and Vieu
-(2003,2006) which is a functional version of the Nadaraya-Watson kernel
-estimator of class membership probabilities: \[
+membership. Consider the classifier proposed in \cite{Ferraty2006} which
+is a functional version of the Nadaraya-Watson kernel estimator of class
+membership probabilities: \[
 \hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
-\] We shall compare our method to using \(L_2\) distance, possibly
+\] We shall compare our method to using \(L^2\) distance, possibly
 weighted, and with curve registration already accomplished. Describe
@@
 \begin{itemize}
-\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
-\item Tecator in [@LinYao2017] contamination paper
-\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
+\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012}
+\item Tecator in \cite{LinYao2017} contamination paper
+\item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper
 \end{itemize}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ChenMuller2012}{}%
-Chen, Dong, and Hans-Georg Muller. 2012. ``NONLINEAR Manifold
-Representations for Functional Data.'' \emph{The Annals of Statistics}
-40 (1). Institute of Mathematical Statistics: 1--29.
-\url{http://www.jstor.org/stable/41713625}.
-
-\leavevmode\hypertarget{ref-ChenHo2015}{}%
-Chen, Y., S. Ho, P. E. Freeman, C. R. Genovese, and L. Wasserman. 2015.
-``Cosmic Web Reconstruction Through Density Ridges: Method and
-Algorithm.'' \emph{Monthly Notices of the Royal Astronomical Society}
-454 (1): 1140--56. \url{https://doi.org/10.1093/mnras/stv1996}.
-
-\leavevmode\hypertarget{ref-Dimeglio2014}{}%
-Dimeglio, Chloe, Santiago Gallon, Jean-Michel Loubes, and Elie Maza.
-2014. ``A Robust Algorithm for Template Curve Estimation Based on
-Manifold Embedding.'' \emph{Comput. Stat. Data Anal.} 70 (February).
-Amsterdam, The Netherlands, The Netherlands: Elsevier Science Publishers
-B. V.: 373--86. \url{https://doi.org/10.1016/j.csda.2013.09.030}.
-
-\leavevmode\hypertarget{ref-Genovese2014}{}%
-Genovese, Christopher R., Marco Perone-Pacifico, Isabella Verdinelli,
-and Larry Wasserman. 2014. ``NONPARAMETRIC Ridge Estimation.'' \emph{The
-Annals of Statistics} 42 (4). Institute of Mathematical Statistics:
-1511--45. \url{http://www.jstor.org/stable/43556332}.
-
-\leavevmode\hypertarget{ref-Lin2014}{}%
-Lin, Binbin, Ji Yang, Xiaofei He, and Jieping Ye. 2014. ``Geodesic
-Distance Function Learning via Heat Flow on Vector Fields.'' In
-\emph{Proceedings of the 31st International Conference on International
-Conference on Machine Learning - Volume 32}, II--145--II--153. ICML'14.
-Beijing, China: JMLR.org.
-\url{http://dl.acm.org/citation.cfm?id=3044805.3044909}.
-
-\leavevmode\hypertarget{ref-LinYao2017}{}%
-Lin, Zhenhua, and Fang Yao. 2017. ``Functional Regression on Manifold
-with Contamination.'' \emph{arXiv E-Prints}, April, arXiv:1704.03005.
-
-\leavevmode\hypertarget{ref-Ozertem2011}{}%
-Ozertem, Umut, and Deniz Erdogmus. 2011. ``Locally Defined Principal
-Curves and Surfaces.'' \emph{J. Mach. Learn. Res.} 12 (July). JMLR.org:
-1249--86. \url{http://dl.acm.org/citation.cfm?id=1953048.2021041}.
-
-\end{document}

### 2019-08-30 index.Rmd
final writing touches

@@
     fig_caption: yes
-bibliography: references.bib
 ---
@@
 Many statistical and machine learning methods rely on some measure of distance.
-For example, clustering and classification of functional data, discretely-sampled curves varying over a continuum, is a common machine learning task in which distance plays a crucial role.
+For example, clustering and classification of functional data, discretely-sampled curves varying over a continuum, is a common learning task in which distance plays a crucial role.
 In a naive treatment, the data may be simply processed as vectors in Euclidean space.
 Doing so however risks losing the richness of functional data; indeed, one may do better to regard the data as functions.
+The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high.
+This insight underpins functional data analysis, a subfield of statistics that studies functional data.
 
-The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high.
-This insight underpins functional data analysis, a subfield of statistics that studies functional data. This has many implications for the processing of functional data. 
-For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between noisy functional observations treated as vectors.
+<!-- This has many implications for the processing of functional data.  -->
+<!-- For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between noisy functional observations treated as vectors.  -->
 
-In this work, we advocate that for functional manifold data, i.e.
-functional data that lie on a manifold, the geodesic distance is more appropriate than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
-In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L^2$ distance.
-The manifold setting presents its own challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
+In this work, we advocate that for functional manifold data, i.e. functional data that lie on a manifold, the geodesic distance is more appropriate than the $L^2$ distance.
+Indeed, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the geodesic distance may give better results than using the $L^2$ distance.
 
-Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014].
+The manifold setting presents certainly challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
+We will carefully lay out a technique for estimating the geodesic distance between noisy functional manifold data.
+
+Let $X_1,\ldots,X_n$ be a sample of $n$ independent realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
+Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows \cite{Lin2014}.
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
@@
 Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error.
-Then the geodesic distance $d_\M(X_i,X_j)$ can be estimated using any of a variety of shortest-path algorithms, e.g.\ the Floyd-Warhsall [@Floyd1962] algorithm.
-However, most shortest-path algorithms critically assume observations live near the manifold with little noise[@Yeh2005] and are thus likely to fail for functional data which can be observed with low signal-to-noise ratio.
+Then the geodesic distance $d_\M(X_i,X_j)$ can be estimated using any of a variety of shortest-path algorithms, e.g.\ the Floyd-Warhsall \cite{Floyd1962} algorithm.
+However, most shortest-path algorithms critically assume observations live near the manifold with little noise \cite{Yeh2005} and are thus likely to fail for functional data which can be observed with low signal-to-noise ratio.
 <!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail.
@@
 
-The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data.
+The work of \cite{ChenMuller2012} was among the first in functional data analysis to consider the manifold hypothesis for functional data.
 A notion of the mean and variation of functional manifold data was introduced there.
-Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold.
-[@Dimeglio2014] also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP that is less sensitive to outliers.
+Of particular relevance to this work is their proposed P-ISOMAP procedure which allows for noisy functional observations.
+This is in contrast to the classic ISOMAP algorithm \cite{Tenenbaum2000} which assumes observations lie exactly on the manifold.
+\cite{Dimeglio2014} also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP for functional data that is less sensitive to outliers.
 We will discuss P-ISOMAP and robust-ISOMAP in more details below.
 
-<!-- [@LinYao2017]  -->
+<!-- \cite{LinYao2017]  -->
 
@@
 Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$.
-Let the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other.
+Let the random variables $\epsilon_{ij}$ have mean zero and be uncorrelated with each other.
 We assume that each observation grid $T_1,\ldots,T_n$ is dense, i.e.\ $K$ is large.
 
-We begin by converting the discretly-sampled functional observations into continuous ones.
+We begin by converting the discretely-sampled functional observations into continuous ones.
 Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method.
-For example, we may employ spline smoothing [@Ramsay2005] to recover the functional versions of the raw data, i.e.\
+For example, we may employ spline smoothing \cite{Ramsay2005} to recover the functional versions of the raw data, i.e.\
 \begin{equation}\label{eq_spline_smoothing} 
 \tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
-\end{equation} where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.
-The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.
+\end{equation} 
+where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
+<!-- The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed. -->
+Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm \cite{Ozertem2011} which we now review briefly in the Euclidean case.
+Suppose $Z_1,\ldots,Z_n$ is a sample of realisations of a random vector $Z \in \M$.
+We can use a kernel density estimator $\hat f$ based on this data to estimate the probability density function of $Z$.
+Let the set $\hat M$ be comprised of the basins of attraction of $\hat f$.
+The SCMS algorithm sends each point $Z_i$ to its destination in $\hat M$.
+Theoretical justification that $\hat M$ as estimated by SCMS is a reasonable surrogate for the true manifold $\M$ can be found in \cite{Genovese2014}.
 
-Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm [@Ozertem2011].
-Let the set $\hat M \subset L^2([a,b],\mathbb R)$ be comprised of the basins of attraction of the estimated probability density function of $X$.
-The SCMS algorithm sends each point $\tilde X_i \in L^2([a,b],\mathbb R)$ to a point in $\hat M$.
-We denote this destination as $\tilde X_i^{\hat \M}$.
-We expect the points $\tilde X^{\hat \M}_1,\ldots,\tilde X^{\hat \M}_n \in \hat M$ to lie close to the real manifold $\M$.
-Theoretical justificaton that $\hat M$ as estimated by SCMS is a reasonable surrogate for the true manifold $\M$ can be found in [@Genovese2014].
-
-Before presenting our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold $\M$ of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
+Before presenting our procedure, we also need to review the ISOMAP algorithm \cite{Tenenbaum2000}, a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold $\M \subset \R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances.
 The procedure is as follows.
@@
 Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
-Specifically, this is the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
+Specifically, this is the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm \cite{Floyd1962} or with the Dijkstra algorithm \cite{Dijkstra59anote}.
 3. Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
 
-Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weigthed graph but they do not produce the weighted graphs themselves.
-Both P-ISOMAP and robust-ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in step 1.
+Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weighted graph but they do not produce the weighted graphs themselves.
+Both P-ISOMAP and robust-ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in step 1 above.
 <!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->
@@
 The first step of IsoGeo constructs the complete weighted ???Marie: does this mean all weights are nonzero???
-graph $G_c$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$.
+graph $G_c$ according to \cite{Dimeglio2014} so that the graph is independent of the tuning parameter $\epsilon$.
 Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$.
@@
 Let $s$ be a positive integer, much smaller than $D$.
-Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ are preserved.
-2. Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
-3. Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
-
+Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling so that the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ can be preserved as much as possible.
+2. Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011} to each of $\tilde X^s_i$ to obtain its destination $\tilde X^{s,\hat \M}_i$.
+3. Use IsoGeo on the inputs $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n \in \R^s$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
 Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
@@
 The other parameter is the bandwidth $h$ in the subspace constrained mean-shift.
-In our simulations we set $h$ according to the heuristic in Equation (A1) of [@ChenHo2015].
+In our simulations we set $h$ according to the heuristic in Equation (A1) of \cite{ChenHo2015}.
 In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.
@@
 We compare the performance of our method to each of the following four alternatives.
-The first takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+The first alternative, denoted RD, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
 Note that this procedure is only sensible if all the time grids $T_1,\ldots,T_n$ are the same.
@@
 
-The final alternative we consider is P-ISOMAP which is the two-step procedure developed in @ChenMuller2012 for handling noisy functional data.
+The final alternative we consider is P-ISOMAP which is the two-step procedure developed in \cite{ChenMuller2012} for handling noisy functional data.
 In the first step, a penalty is incorporated to robustify the construction of the weighted graph.
@@
 We ran P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric.
-We found that the penalty parameter chosen in this way was always zero, thus reducing hte P-ISOMAP procedure to standard ISOMAP.
+We found that the penalty parameter chosen in this oracle fashion was always zero, thus reducing the P-ISOMAP procedure to standard ISOMAP.
 Given this, we did not make further comparison to P-ISOMAP in our simulation study.
-We also do not present the results for the RD method which always produced very bad estimates of the geodesic distance.
+We also do not present the results for the RD method which always produced poor estimates of the geodesic distance.
 
@@
 
-The second metric $||d-\hat d||/||d||$ is the Frobenius norm of the estimation error matrix $d-\hat d$, relative to the Frobenius norm of the true distance matrix $d$.
-The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.
+The second metric $||G-\hat G||/||G||$ is the Frobenius norm of the estimation error matrix $G-\hat G$, relative to the Frobenius norm of the true distance matrix $G$.
+The third metric is the Pearson correlation coefficient between the upper diagonal of $G$ and $\hat G$.
 A good geodesic distance estimator has low Frobenius norm and high Pearson correlation coefficient.
@@
 Throughout, we make use of a helpful property of one-dimensional manifolds.
-If $X \in \M$ with the $L^2$ metric tensor depends on a parameter $\theta \in \mathbb R$, then
+If $X \in \M$, with the $L^2$ metric tensor, depends on a parameter $\theta \in \mathbb R$, then
 $$
-d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2} \,d\theta
+d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2} \,d\theta.
 $$
@@
 Proper sampling from a manifold remains an open problem.
-Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013].
+Even in the Euclidean case, it is not obvious how sampling should be done \cite{Diaconis2013}.
 For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very narrow support of the manifold parameters.
@@
 <!-- Scenario 2 in sim_functional_data -->
-We take Manifold 2 in @ChenMuller2012 and set the variance of the normal density to be $1$.
+We take Manifold 2 in \cite{ChenMuller2012} and set the variance of the normal density to be $1$.
 Let $X_\beta: [a,b] \to \R$ be given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
@@
 \noindent where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from integrating on $[a,b]=[-4,4]$ rather than $\R$. 
-We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$, the Euclidean distance between the $\beta$'s.
+We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$.
 
@@
 <!-- Scenario 4 in sim_functional_data -->
-It was shown in @Joshi2007 that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
+It was shown in \cite{Joshi2007} that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
 They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
-Note that this concides with the $L^2[0,1]$ inner product.
-It was shown in [@Joshi2007] that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+Note that this coincides with the $L^2[0,1]$ inner product.
+It was shown in \cite{Joshi2007} that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
 We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$.
-That is consider $$ M = \{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$.
+That is, consider $$ M = \{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$.
 We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$.
@@
 <!-- Scenario 5 in sim_functional_data -->
-This is based on Equations 17 and 18 in [@Kneip2008] but with $z_{i1}, z_{i2}$ both set to $1$.
-(Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
+This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$.
+(Note that Equation 17 therein has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
 
@@
 
-The tuning parameters for the proposed method were chosen as described at the end of Section \ref{proposed-method-for-estimating-geodesic-distances}.
+The tuning parameters for the proposed method were chosen as described at the end of the methodology section.
 In step 1 of our procedure, we take $D = 100$ for the size of the common time grid.
@@
 which we fix at either $0.1$ (low) or $0.5$ (high).
-The second parameter is $K$, the dimension of the dscrietly-observed noisy $Y_i$'s.
+The second parameter is $K$, the dimension of the discretely-observed noisy $Y_i$'s.
 We set $K$ to either $30$ (sparse) or $100$ (abundant).
 
+We will present the results for the high signal-to-noise and sparse $K$ setting.
 We show the performance of the Spline IsoGeo and Spline L2 method versus our method for three dimensions $s=1,2,3$.
@@
 Interestingly, when $s$ in our method is not selected properly, for instance when $s=1$, our estimator performs poorly in terms of the relative Frobenius norm although still better than Spline IsoGeo and Spline L2 in the other two metrics.
-It may be possible that for certain downstream tasks such as clustering or classification, precise estimation of the geodesic distance is less important than an estimation that respects the ordering of the distances.
-In such cases, the near-isometry and Pearson correlation coefficients may give a better indication of downstream performance of the geodesic distance estimator.
+It may be possible that for certain downstream tasks such as clustering or classification, precise estimation of the geodesic distance is less important than one that respects the ordering of the distances.
+In such cases, the near-isometry and Pearson correlation metrics may give a better indication of downstream performance of the geodesic distance estimator.
 
@@
 
-# Distance-based functional classification [Marie]
+<!-- # Distance-based functional classification [Marie] -->
 
-In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task.
+<!-- In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task.
 There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification.
 It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.
+-->
 
-For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
-Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
-$$
-\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
-$$
-We shall compare our method to using $L^2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.
+<!-- For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership. -->
+<!-- Consider the classifier proposed in \cite{Ferraty2006} which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities: -->
+<!-- $$ -->
+<!-- \hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] } -->
+<!-- $$ -->
+<!-- We shall compare our method to using $L^2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.  -->
 
-The bandwidth in the classifier should be tuned individually for each method.
+<!-- The bandwidth in the classifier should be tuned individually for each method.
 Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.
+-->
 
-Datasets used by functional classification papers
-\begin{itemize}
-\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
-\item Berkeley growth curves in [@ChenReiss2014].
-\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
-\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
-\end{itemize}
+<!-- Datasets used by functional classification papers -->
+<!-- \begin{itemize} -->
+<!-- \item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data" -->
+<!-- \item Berkeley growth curves in \cite{ChenReiss2014}. -->
+<!-- \item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper. -->
+<!-- \item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data" -->
+<!-- \end{itemize} -->
 
-Datasets used in functional manifold papers
-\begin{itemize}
-\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
-\item Tecator in [@LinYao2017] contamination paper
-\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
-\end{itemize}
+<!-- Datasets used in functional manifold papers -->
+<!-- \begin{itemize} -->
+<!-- \item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012} -->
+<!-- \item Tecator in \cite{LinYao2017} contamination paper -->
+<!-- \item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper -->
+<!-- \end{itemize} -->
 
@@
 
-# Acknowledgement
-The authors would like to acknowledge support from the ARC Centre of Excellence for Mathematical & Statistical Frontiers (ACEMS) which enabled the two authors to work on the project at the University of Melbourne.
+<!-- # Acknowledgement -->
+<!-- The authors would like to acknowledge support from the ARC Centre of Excellence for Mathematical & Statistical Frontiers (ACEMS) which enabled the two authors to work on the project at the University of Melbourne.
+-->
 
-# Appendices
-\begin{figure}
-\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_MSE}
-\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_isometry}
-\includegraphics{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
-\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
-\end{figure}
+<!-- # Appendices -->
+<!-- The result of the square root velocity manifold scenario is presented in Figure \ref{fig:betas}.
+The Spline L2 method has similar performance to our proposed method.
+This could be due to the fact that we did not sample the manifold broadly enough so that the sample lives in a locally flat region where the $L^2$ distance holds.
+-->
 
-# References
+<!-- \begin{figure} -->
+<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE} -->
+<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry} -->
+<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson} -->
+<!-- \label{fig:betas} -->
+<!-- \caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.} -->
+<!-- \end{figure} -->

### 2019-08-30 paperAAAI.tex
final writing touches

@@
 % command to correct it. You may not alter the value below 2.5 in
-\title{The Hidden Manifold Distance for Functional ata}
+\title{The Hidden Manifold Distance for Functional Data}
 %Your title must be in mixed case, not sentence case. 
@@
 In many machine learning applications, the observed data is a
-discretely-sampled realization of a continuous process. While one can
+discretely-sampled realisation of a continuous process. While one can
 naively treat such data as vectors in Euclidean space, one may do better
-to regard them as functional data. For certain types of functional data,
-there is further structure to exploit, e.g.~probability density
-functions or warped curves of a common template function. Namely, the
-underlying curves of interest may lie on a manifold of low intrinsic
-dimension. This work addresses the estimation of pairwise geodesic
+to regard them as functional data.  
+When the underlying curves of interest lie on a manifold of low intrinsic
+dimension, e.g.~probability density
+functions or warped curves of a common template function, there is further structure to exploit. 
+This work addresses the estimation of pairwise geodesic
 distances between functional manifold data that are observed with noise,
@@
 manifold. Improved estimation of the pairwise geodesic distance has
-beneficial implications for many downstream tasks in functional data
-analysis which we illustrate in the case of distance-based functional
-classification/clustering.
+potential benefits for downstream tasks such as distance-based functional
+???Marie: update classification/clustering???.
 \end{abstract}
@@
 
+\input{MarieDataAnalysis}
+
+\section{Acknowledgement}\label{acknowledgement}
+
+The authors would like to acknowledge support from the ARC Centre of
+Excellence for Mathematical and Statistical Frontiers (ACEMS) which
+enabled the two authors to work on the project at the University of
+Melbourne.
+
+\section{Appendices}\label{appendices}
+
+The result of the square root velocity manifold scenario is presented in
+Figure \ref{fig:betas}. The Spline L2 method has similar performance to
+our proposed method. This could be due to the fact that we did not
+sample the manifold broadly enough so that the sample lives in a locally
+flat region where the \(L^2\) distance holds.
+
+\begin{figure}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
+\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
+\label{fig:betas}
+\end{figure}
+
+\bibliography{references}
+\bibliographystyle{aaai}
+
 \end{document}

### 2019-08-30 run_this.R

@@ -1,3 +1,2 @@
 knitr::knit("index.Rmd")
 rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex",citeproc=FALSE)
-# system("pdflatex paperAAAI.tex")

### 2019-09-03 MarieDataAnalysis.tex
wrote section classification of real data

@@
-\section{Distance-based functional classification
-{[}Marie{]}}\label{distance-based-functional-classification-marie}
+\section{Distance-based functional classification}\label{distance-based-functional-classification-marie}
 
@@
 distanced-based functional clustering, but we will focus on
-distance-based functional classification. It must be noted that while
-curve alignment, also known as curve registration, is necessarily
-performed as a preprocessing technique prior to clustering and
-classification, our geodesic distance estimator allows one to forsake
-this step.
+distance-based functional classification.
 
-For simplicity, assume the task is binary classification. Associated to
-each functional object \(x\) is a binary \(y\) indicating class
-membership. Consider the classifier proposed in \cite{Ferraty2006} which
-is a functional version of the Nadaraya-Watson kernel estimator of class
-membership probabilities: \[
-\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
-\] We shall compare our method to using \(L^2\) distance, possibly
-weighted, and with curve registration already accomplished. Describe
-alternative methods in detail.
+We work with the non-parametric classifier proposed in \cite{Ferraty2006} which is a functional version of the Nadaraya-Watson kernel estimator.
+Given a sample of curves $X_1,\ldots,X_n$ where each function $X_i$ is associated to a class $Y_i$, the estimator of the probability that a new observation $X^\star$ is in class $Y^\star=y$ is defined as
+\begin{equation}\label{func_NW}
+\hat P(Y^\star = y | X^\star)= \frac{ \sum_{i=1}^n k[h^{-1} d(X^\star,X_i)] {\bf{1}}(Y_i = y) }{ \sum_{i=1}^n k[h^{-1} d(X^\star,X_i)] },
+\end{equation}
+where $d$ is a distance, $k$ a kernel, $h$ a bandwidth and $\bf{1}$ the indicator function.
+The curve $X^\star$ is then classified in the class $Y^\star =y$ for which the conditional probability (\ref{func_NW}) is larger.
+We compare the classification performance obtained with the distance $d$ being equal to our proposed geodesic distance with $s=1,2$ and $3$, and to the one obtained with Spline IsoGeo and Spline L2 described in the simulation study section.
+We set $k$ to be a quadratic kernel and the bandwidth $h$ is chosen by cross-validation.
 
-The bandwidth in the classifier should be tuned individually for each
-method. Also we might need to tune MDS dimension \(s\) since in real
-data, the dimension of the manifold might be much higher than
-encountered in the simulation scenarios where it never goes above 2.
+The classification task is perform on a famous dataset in Functional Data Analysis which is called the Berkeley growth curves dataset and is available in the \emph{fda} package \cite{Ram-Hoo-Gra} for R \cite{Rproject}.
+We chose this dataset because we expect the data to contain nonlinear features as it is hinted in \cite{ChenMuller2012}.
+The data consist in the height of $n=93$ individuals, 39 boys and 54 girls, measured on a common grid $t_1,\ldots,t_K$ of $K=31$ points taken between the ages of 1 and 18 years.
+The raw data $Y_1,\ldots,Y_n$ are transformed to continuous functions by smoothing with B-spline bases with a roughness penalty on the fourth derivative :
+$$\tilde X_i(t) = \sum_{l=1}^{b} c_{il}B_l(t), $$
+where
+ \begin{eqnarray*}
+ (c_{i1},\ldots,c_{ib})&=&\arg \min_{c\in \R^b}\Bigg\{\sum_{j=1}^K\left(\sum_{l=1}^{b} c_{il}B_l(t_j)-Y_{ij}\right)^2\\
+ &&\quad +\lambda \left\|\sum_{l=1}^{b} c_{il}\partial^4_tB_l\right\|^2_{L^2}\Bigg\}.
+ \end{eqnarray*} 
+We use $b=35$ B-splines bases of order 6 and the parameter $\lambda$ is chosen by generalized cross-validation.
+We then differentiate the resulting fonctions to obtain velocity curves; the data are illustrated for boys and girls on Figure \ref{fig:data_illu}.
 
-Datasets used by functional classification papers
+To assess the classification performance of the different distances, we randomly split $200$ times the dataset into a training set of $50$ curves and a test set of $43$ curves.
+For each split, we calculate the misclassification error for the 5 different distances, and we average these errors over the splits.
+The results are presented in Figure \ref{fig:class_err_velo}, we can see that our method with $s=3$ performs a little bit better than the others and that our method with $s=2$ is equivalent to Spline IsoGeo and Spline L2.
 
-\begin{itemize}
-\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
-\item Berkeley growth curves in \cite{ChenReiss2014}.
-\item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper.
-\item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data"
-\end{itemize}
+\begin{figure}
+\centering
+\includegraphics[height=0.55\textheight]{./Velocity_curves_data_illustration.pdf}
+\caption{Velocity curves of 39 boys (top) and 54 girls (bottom).}
+\label{fig:data_illu}
+\end{figure}
 
-Datasets used in functional manifold papers
+\begin{figure}
+\centering
+\includegraphics[height=0.3\textheight]{./Erreur_classification_velocity.pdf}
+\caption{Classification errors for the velocity curves, our method with $s=3$ performs better while our method with $s=2$ is equivalent to Spline IsoGeo and Spline L2.}
+\label{fig:class_err_velo}
+\end{figure}
 
-\begin{itemize}
-\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012}
-\item Tecator in \cite{LinYao2017} contamination paper
-\item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper
-\end{itemize}
+% It must be noted that while
+% curve alignment, also known as curve registration, is necessarily
+% performed as a preprocessing technique prior to clustering and
+% classification, our geodesic distance estimator allows one to forsake
+% this step.
+
+% For simplicity, assume the task is binary classification. Associated to
+% each functional object \(x\) is a binary \(y\) indicating class
+% membership. Consider the classifier proposed in \cite{Ferraty2006} which
+% is a functional version of the Nadaraya-Watson kernel estimator of class
+% membership probabilities: \[
+% \hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
+% \] We shall compare our method to using \(L^2\) distance, possibly
+% weighted, and with curve registration already accomplished. Describe
+% alternative methods in detail.
+% 
+% The bandwidth in the classifier should be tuned individually for each
+% method. Also we might need to tune MDS dimension \(s\) since in real
+% data, the dimension of the manifold might be much higher than
+% encountered in the simulation scenarios where it never goes above 2.
+
+% Datasets used by functional classification papers
+% 
+% \begin{itemize}
+% \item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
+% \item Berkeley growth curves in \cite{ChenReiss2014}.
+% \item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper.
+% \item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data"
+% \end{itemize}
+% 
+% Datasets used in functional manifold papers
+% 
+% \begin{itemize}
+% \item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012}
+% \item Tecator in \cite{LinYao2017} contamination paper
+% \item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper
+% \end{itemize}

### 2019-09-03 index.Rmd
wrote section classification of real data

@@
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
-The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
-$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[0,1] \to \M, \gamma(0) = X_i, \gamma(1) = X_j} l(\gamma)$$
-where $$ l(\gamma) := \int_0^1 || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
+The norm of a tangent vector $V \in T_X \M$ is defined as $$\|V\| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
+$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[0,1] \to \M, \gamma(0) = X_i, \gamma(1) = X_j} l(\gamma),$$
+where $$ l(\gamma) := \int_0^1 \left\| \frac{\,d\gamma}{\,d t}(t) \right\| \,dt$$
 is the length of the piecewise-smooth curve $\gamma$. 
@@
 1.
-Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
-Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
+Construct a weighted graph $\mathcal{G}$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$.
+Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \varepsilon \}$ where $d_{ij}=\|x_i-x_j \|_{2}$ and $\varepsilon > 0$.
 2.
-Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms.
+Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $\mathcal{G}$ using shortest-path algorithms.
 Specifically, this is the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm \cite{Floyd1962} or with the Dijkstra algorithm \cite{Dijkstra59anote}.
@@
 In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP.
-The first step of IsoGeo constructs the complete weighted ???Marie: does this mean all weights are nonzero???
-graph $G_c$ according to \cite{Dimeglio2014} so that the graph is independent of the tuning parameter $\epsilon$.
-Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$.
-The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
-$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
-where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
+The first step of IsoGeo constructs a weighted graph $\mathcal{G}$ according to \cite{Dimeglio2014}.
+First we construct a complete weighted graph $\mathcal{G}_c$, meaning that every pair of nodes $x_i$ and $x_j$ are connected by an edge of weight $d_{ij}$, so that the graph is independent of the tuning parameter $\varepsilon$.
+Next, we obtain the minimal spanning tree associated with $\mathcal{G}_c$ and denote its set of edges by $E_s$.
+The graph $\mathcal{G}$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
+$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\varepsilon_i),$$
+where $B(x_i,\varepsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
 Let the distance estimated by IsoGeo be denoted $\text{IsoGeo}(x_i,x_j)$.
@@
 The other parameter is the bandwidth $h$ in the subspace constrained mean-shift.
-In our simulations we set $h$ according to the heuristic in Equation (A1) of \cite{ChenHo2015}.
+In our simulations we set $h$ according to the heuristic in Equation (A1) of \cite{ChenHo2015}:
+$$h = A_0\left(\frac{1}{s+2}\right)^{\frac{1}{s+4}}n^{\frac{-1}{s+4}}\sigma_{\min}, $$ where $A_0 = ?$ and $\sigma_{\min}$ is the minimum over $j=1,\ldots,s$ of the standard deviation of $\tilde X^s_{1,j},\ldots, \tilde X^s_{n,j}$, with $\tilde X^s_{i,j}$ being the $j$th entry of the vector $\tilde X^s_{i}$.
 In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.
@@
 
-We ran P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric.
+We ran P-ISOMAP separately in Matlab and tuned for the parameter $\varepsilon$ and the penalty parameter using the relative Frobenius assessment metric.
 We found that the penalty parameter chosen in this oracle fashion was always zero, thus reducing the P-ISOMAP procedure to standard ISOMAP.
@@
 
-The second metric $||G-\hat G||/||G||$ is the Frobenius norm of the estimation error matrix $G-\hat G$, relative to the Frobenius norm of the true distance matrix $G$.
+The second metric $\|G-\hat G\|/\|G\|$ is the Frobenius norm of the estimation error matrix $G-\hat G$, relative to the Frobenius norm of the true distance matrix $G$.
 The third metric is the Pearson correlation coefficient between the upper diagonal of $G$ and $\hat G$.
@@
 $$
-d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2} \,d\theta.
+d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} \left\| \frac{\partial X}{\partial \theta}(t)\right\|_{L^2} \,d\theta.
 $$
@@
 Let $X_\beta: [a,b] \to \R$ be given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$.
-Consider $$\M = \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L^2$ inner product as the metric tensor.
+Consider $$\M = \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \},$$ with the $L^2$ inner product as the metric tensor.
 First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ does not always stay inside of $\M$.
@@
 &\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
-&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}}
+&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
 \end{eqnarray*}
@@
 It was shown in \cite{Joshi2007} that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
-They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
+They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \},$$ with the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds ,$$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$.
 Note that this coincides with the $L^2[0,1]$ inner product.
@@
 We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$.
-That is, consider $$ M = \{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$.
+That is, consider $$ \M = \big\{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\big\}, $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$.
 We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$.
@@
 This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$.
-(Note that Equation 17 therein has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
+(Note that Equation 17 therein has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\},$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1},$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
 
-Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\}.$$ 
+Consider the manifold $$ \M = \{X_\alpha: -1 \le \alpha \le 1\}.$$ 
 We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
@@
 
+For each manifold $\M$ described above, we simulate ???Susan how many?
+1000??
+samples of $200$ functions $X_i\in\M$.
+We then construct vectors $Y_i\in \R^K$ such that $Y_{ij}=X_i(t_j) + \epsilon_{ij}$ with $T=(t_1,\ldots,t_K)$ a common grid of equally spaced points and $\epsilon_{ij}\sim N(0,\sigma_\epsilon^2)$.
 The tuning parameters for the proposed method were chosen as described at the end of the methodology section.
@@
 We will investigate how two parameters affect the performance of the various methods in consideration.
-The first one is the signal-to-noise ratio ???[Marie] describe???
-which we fix at either $0.1$ (low) or $0.5$ (high).
+The first one is the signal-to-noise ratio
+$$SNR = 10 log_{10}\left(\frac{\sigma^2_{sig}}{\sigma^2_\epsilon}\right),$$
+where $\sigma^2_{sig}$ is the variance of the signal.
+We fix SNR at either $0.1$ (low) or $0.5$ (high).
 The second parameter is $K$, the dimension of the discretely-observed noisy $Y_i$'s.
@@
 We show the performance of the Spline IsoGeo and Spline L2 method versus our method for three dimensions $s=1,2,3$.
-We present the result of the square root velocity manifold scenario in the Appendix.
 We can see from Figures \ref{fig:normal} and \ref{fig:warping} that spline smoothing followed by either IsoGeo or the $L^2$ distance calculation estimates the pairwise geodesic distances reasonably well though not as well as the proposed method for $s=2$ and $s=3$.
@@
 In such cases, the near-isometry and Pearson correlation metrics may give a better indication of downstream performance of the geodesic distance estimator.
+The result of the square root velocity manifold scenario is presented in
+Figure \ref{fig:betas}. The Spline L2 method has similar performance to
+our proposed method. This could be due to the fact that we did not
+sample the manifold broadly enough so that the sample lives in a locally
+flat region where the \(L^2\) distance holds.
 
@@
 \label{fig:warping}
+\end{figure}
+
+\begin{figure}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
+\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
+\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
+\label{fig:betas}
 \end{figure}

### 2019-09-03 paperAAAI.tex
wrote section classification of real data

--- 2019-08-28 paperAAAI.tex
@@
 % directly follow a colon or long dash
-\author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be in the same font size and format.
-Use \Large and \textbf to achieve this result when breaking a line
-\textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ %If you have multiple authors and multiple affiliations
+
+% \author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be
+% \textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ }
+
+%in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
+
+%If you have multiple authors and multiple affiliations
 % use superscripts in text and roman font to identify them.
@@
  % email address must be in roman text type, not monospace or sans serif
-}
+
 \begin{document}
@@
 manifold. Improved estimation of the pairwise geodesic distance has
-beneficial implications for many downstream tasks in functional data
-analysis which we illustrate in the case of distance-based functional
-classification/clustering.
+potential benefits for downstream tasks such as distance-based functional classification.
 \end{abstract}
@@
 
+\input{MarieDataAnalysis}
+
+%\section{Acknowledgement}\label{acknowledgement}
+
+%The authors would like to acknowledge support from the ARC Centre of
+%Excellence for Mathematical and Statistical Frontiers (ACEMS) which
+%enabled the two authors to work on the project at the University of
+%Melbourne.
+
+% \section{Appendices}\label{appendices}
+% 
+% The result of the square root velocity manifold scenario is presented in
+% Figure \ref{fig:betas}. The Spline L2 method has similar performance to
+% our proposed method. This could be due to the fact that we did not
+% sample the manifold broadly enough so that the sample lives in a locally
+% flat region where the \(L^2\) distance holds.
+% 
+% \begin{figure}
+% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
+% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
+% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
+% \caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
+% \label{fig:betas}
+% \end{figure}
+
+\newpage
+
+\bibliography{references}
+\bibliographystyle{aaai}
+
 \end{document}

### 2019-09-03 NP_FDA_Ferraty.R

@@ -927,10 +927,10 @@ funopadi.knn.lcv <- function(ind_train,data, kind.of.kernel = "quadratic", pairw
     Classes.predicted <- as.vector((PROB.PREDICTED == apply(
       PROB.PREDICTED, 1, max)) %*% (1:nbclass))
     Misclas<- sum(Classes.predicted != class_test)/n2
-    return(Misclas.estimated)
+    return(Misclas)
     #return(list(Estimated.classnumber = Classes.estimated, 
-                #Predicted.classnumber = Classes.predicted, Bandwidths
-                #= Bandwidth.opt, Misclas = Misclas.estimated))
+    #Predicted.classnumber = Classes.predicted, Bandwidths
+    #= Bandwidth.opt, Misclas = Misclas.estimated))
     
   }else {
     return(list(Estimated.classnumber = Classes.estimated, 
@@ -3259,6 +3259,141 @@ unbal2equibal <- function(DATA, INSTANTS, Range.grid, lnewgrid)
   return(list(NEWDATA = NEWDATA, Newgrid = Newgrid))
 }
 
+
+#####################################################################
+funopadi.knn.lcv_ORIGINAL_DISTANCE <- function(ind_train,data, ...,kind.of.kernel = "quadratic",semimetric,true_class)
+{
+  Classes<- true_class[ind_train]
+  CURVES<- data[ind_train,]
+  PRED<- data[-ind_train,]
+  class_test<- true_class[-ind_train]
+  ################################################################
+  # Performs functional discrimination of a sample of curves when 
+  # a categorical response is observed (supervised classification). 
+  # A local bandwidth (i.e. local number of neighbours) is selected 
+  # by a cross-validation procedure.
+  #    "Classes" vector containing the categorical responses
+  #              giving the group number for each curve in 
+  #              the matrix CURVES (if nbclass is the number of 
+  #              groups, "Classes" contains numbers 1,2,...,nbclass)
+  #    "CURVES" matrix containing the curves dataset (row by row) 
+  #             used for the estimating stage
[…]

### 2019-09-03 analysis_growth_data.R

@@ -1,0 +1,179 @@
+
+library(reticulate)
+library(fda)
+library(igraph)
+library(caret)
+library(vows)
+library(KernSmooth)
+library(truncnorm)
+
+# this may need to be set at the beginning of the session 
+# Susan's
+# use_python('/usr/bin/python',required=TRUE)
+# use_python('/usr/local/bin/python3.7',required=TRUE)
+# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
+
+# Marie's
+# use_python('/anaconda3/envs/r-reticulate/bin/python',required=TRUE)
+use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
+
+source('NP_FDA_Ferraty.R')
+source('robust_isomap.R')
+source('pairwise_geo_estimation.R')
+source('pairwise_geo_OUR_tuned.R')
+
+attach(growth)
+true_class<- c(rep(1,39),rep(2,54))
+true_group<- c(rep(TRUE,39),rep(FALSE,54))
+grid<- seq(1,18,length.out=101)
+raw_data <- cbind(hgtm,hgtf)
+smooth_growth<- smooth_FD_bspline(t(raw_data),age,grid)
+detach(growth)
+
+
+pdf("./Velocity_curves_data_illustration.pdf",width=5,height=8)
+par(mar=c(3,4,1,1),mfrow=c(2,1))
+matplot(grid,t(smooth_growth[1:39,]),type='l',col="black",lty=1,ylab='Velocity in cm per year')
+par(mar=c(4,4,0,1))
+matplot(grid,t(smooth_growth[40:93,]),type='l',col="black",lty=2,xlab='Age',ylab='Velocity in cm per year')
+dev.off()
+
+
+# meth <- list("NN" = FALSE,"RD_o" = FALSE,"RD" = FALSE,"SS_o" = FALSE,"SS" = TRUE,"pI" = FALSE,"OUR" = FALSE,"OUR2" = FALSE,"OUR3"=TRUE,"RP" = FALSE , "L2"=TRUE, "w_L2"= FALSE)
+# 
+# # estimation of geo distance with methods TRUE in meth
+# geo_estim_growth = pairwise_geo_estimation(method=meth,
+#                                               true_data=NA,
+#                                               discrete_data=smooth_growth,
+#                                               true_geo=NA,
+#                                               plot_true=FALSE,
[…]

### 2019-09-03 knn_dist_matrix.R

@@ -1,0 +1,105 @@
+
+
+
+
+
+#############################################################################################
+# knn provides with the classification with knn method for a given distance matrix
+#############################################################################################
+
+knn_dist_matrix <- function(ind_train,group,ksNN,lksNN, dist_matrix,N){
+ 
+  ind_train<- sort(ind_train)
+  groups<- group[ind_train]
+  ind_seq<- 1:N
+  ind_test<- ind_seq[-ind_train]
+  n <- length(ind_train)
+  n.new<- length(ind_test)
+  good.class.l2 <- matrix(NA,nrow=lksNN,ncol=1)
+  
+  kkNN <- 1
+  for (k in ksNN){
+    
+    groups.l2 <- matrix(NA,nrow=n,ncol=1)
+    
+    for (i in 1 : n){
+      ind_temp<- ind_train[i]
+      dist.l2 <- matrix(dist_matrix[ind_temp,ind_train],nrow=n,ncol=1)
+      dist.l2[i] <- max(dist.l2)
+      
+      dist.l2.s <- sort(dist.l2,index.return = TRUE)
+      g1 <- sum(groups[dist.l2.s$ix[1:k]]==1)
+      g2 <- sum(groups[dist.l2.s$ix[1:k]]==2)
+      if (g1 > g2){groups.l2[i] <- 1}else{groups.l2[i] <- 2}
+      
+    }
+    good.class.l2[kkNN] <- sum(groups.l2==groups)
+    kkNN <- kkNN + 1
+    
+  }
+  
+  k.knn.l2 <- min(which(good.class.l2==max(good.class.l2)))
+  groups.knn.l2 <- matrix(NA,nrow=n.new,ncol=1)
+  
+  for (i in 1 : n.new){
+    ind_temp<- ind_test[i]
+    dist.l2 <- matrix(dist_matrix[ind_temp,ind_train],nrow=n,ncol=1)
+
+    dist.l2.s <- sort(dist.l2,index.return = TRUE)
+    g1 <- sum(groups[dist.l2.s$ix[1:k.knn.l2]]==1)
+    g2 <- sum(groups[dist.l2.s$ix[1:k.knn.l2]]==2)
+    if (g1 > g2){groups.knn.l2[i] <- 1}else{groups.knn.l2[i] <- 2}
+  }
+  
[…]

### 2019-09-05 MarieDataAnalysis.tex
final touches

@@
 
+There are many downstream analysis tasks in functional data analysis that are based on pairwise distances, including distance-based nonparametric regression and distanced-based functional clustering.
 In this section, we explore whether our geodesic distance estimator has
-benefits for downstream analysis task. There are many tasks we could
-consider here such as distance-based nonparametric regression and
-distanced-based functional clustering, but we will focus on
-distance-based functional classification.
+benefits for the downstream analysis task of functional classification.
 
 We work with the non-parametric classifier proposed in \cite{Ferraty2006} which is a functional version of the Nadaraya-Watson kernel estimator.
-Given a sample of curves $X_1,\ldots,X_n$ where each function $X_i$ is associated to a class $Y_i$, the estimator of the probability that a new observation $X^\star$ is in class $Y^\star=y$ is defined as
+Given a sample of curves $X_1,\ldots,X_n$ where each function $X_i$ is associated to a class label $Y_i$, the estimator of the probability that a new observation $X^\star$ is in class $Y^\star=y$ is defined as
 \begin{equation}\label{func_NW}
@@
 where $d$ is a distance, $k$ a kernel, $h$ a bandwidth and $\bf{1}$ the indicator function.
-The curve $X^\star$ is then classified in the class $Y^\star =y$ for which the conditional probability (\ref{func_NW}) is larger.
-We compare the classification performance obtained with the distance $d$ being equal to our proposed geodesic distance with $s=1,2$ and $3$, and to the one obtained with Spline IsoGeo and Spline L2 described in the simulation study section.
+The curve $X^\star$ is then classified in the class $Y^\star =y$ for which the conditional probability (\ref{func_NW}) is largest.
+We compare the classification performance obtained with the distance $d$ given by our proposed geodesic distance with $s=1,2$ and $3$, as well as that given by Spline IsoGeo and Spline L2 described in the simulation study section.
 We set $k$ to be a quadratic kernel and the bandwidth $h$ is chosen by cross-validation.
 
-The classification task is perform on a famous dataset in Functional Data Analysis which is called the Berkeley growth curves dataset and is available in the \emph{fda} package \cite{Ram-Hoo-Gra} for R \cite{Rproject}.
-We chose this dataset because we expect the data to contain nonlinear features as it is hinted in \cite{ChenMuller2012}.
-The data consist in the height of $n=93$ individuals, 39 boys and 54 girls, measured on a common grid $t_1,\ldots,t_K$ of $K=31$ points taken between the ages of 1 and 18 years.
+The classification task is performed on the well-known Berkeley growth curves dataset in functional data analysis, available in the \emph{fda} package \cite{Ram-Hoo-Gra} for R \cite{Rproject}.
+We chose this dataset because we expect the data to contain nonlinear features as suggested in \cite{ChenMuller2012}.
+The data consist of the height of $n=93$ individuals, 39 boys and 54 girls, measured on a common grid $t_1,\ldots,t_K$ of $K=31$ points taken between the ages of 1 and 18 years.
 The raw data $Y_1,\ldots,Y_n$ are transformed to continuous functions by smoothing with B-spline bases with a roughness penalty on the fourth derivative :
@@
 
-To assess the classification performance of the different distances, we randomly split $200$ times the dataset into a training set of $50$ curves and a test set of $43$ curves.
-For each split, we calculate the misclassification error for the 5 different distances, and we average these errors over the splits.
+To assess the classification performance of the the proposed geodesic distance, we randomly split $200$ times the dataset into a training set of $50$ curves and a test set of $43$ curves.
+For each split, we calculate the misclassification error, and we average these errors over the splits.
 The results are presented in Figure \ref{fig:class_err_velo}, we can see that our method with $s=3$ performs a little bit better than the others and that our method with $s=2$ is equivalent to Spline IsoGeo and Spline L2.
 
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[height=0.55\textheight]{./Velocity_curves_data_illustration.pdf}
+\includegraphics[height=0.5\textheight]{./Velocity_curves_data_illustration.pdf}
 \caption{Velocity curves of 39 boys (top) and 54 girls (bottom).}
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \centering

### 2019-09-05 index.Rmd
final touches

@@
 source("cluster_results.R",echo=FALSE)
+source('sim_functional_data.R')
 ```
@@
 
-The manifold setting presents certainly challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
+The manifold setting presents certain challenges since, for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.
 We will carefully lay out a technique for estimating the geodesic distance between noisy functional manifold data.
@@
 Let $X_1,\ldots,X_n$ be a sample of $n$ independent realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$.
-Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows \cite{Lin2014}.
+Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which are often used to assign a metric on the manifold as follows \cite{Lin2014}.
 For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$.
@@
 -->
-In this work, we put forth a technique for the recovery of the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
+In this work, we put forth a technique for the recovery of the $n{\times}n$ matrix $G$ of pairwise geodesic distances:
 $$
@@
 This is in contrast to the classic ISOMAP algorithm \cite{Tenenbaum2000} which assumes observations lie exactly on the manifold.
-\cite{Dimeglio2014} also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP for functional data that is less sensitive to outliers.
+This drawback to ISOMAP was also realised in \cite{Dimeglio2014} who proposed a procedure we will call robust-ISOMAP for functional data that is less sensitive to outliers.
 We will discuss P-ISOMAP and robust-ISOMAP in more details below.
@@
 In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP.
-The first step of IsoGeo constructs a weighted graph $\mathcal{G}$ according to \cite{Dimeglio2014}.
-First we construct a complete weighted graph $\mathcal{G}_c$, meaning that every pair of nodes $x_i$ and $x_j$ are connected by an edge of weight $d_{ij}$, so that the graph is independent of the tuning parameter $\varepsilon$.
+The first step of IsoGeo constructs a weighted graph $\mathcal{G}$ according to \cite{Dimeglio2014} as follows.
+First we construct a complete weighted graph $\mathcal{G}_c$, meaning that every pair of nodes $x_i$ and $x_j$ is connected by an edge of weight $d_{ij}$, so that the graph is independent of the tuning parameter $\varepsilon$.
 Next, we obtain the minimal spanning tree associated with $\mathcal{G}_c$ and denote its set of edges by $E_s$.
@@
 2. Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011} to each of $\tilde X^s_i$ to obtain its destination $\tilde X^{s,\hat \M}_i$.
-3. Use IsoGeo on the inputs $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n \in \R^s$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
-Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
+3. Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
 $$
@@
 In our simulations we set $h$ according to the heuristic in Equation (A1) of \cite{ChenHo2015}:
-$$h = A_0\left(\frac{1}{s+2}\right)^{\frac{1}{s+4}}n^{\frac{-1}{s+4}}\sigma_{\min}, $$ where $A_0 = ?$ and $\sigma_{\min}$ is the minimum over $j=1,\ldots,s$ of the standard deviation of $\tilde X^s_{1,j},\ldots, \tilde X^s_{n,j}$, with $\tilde X^s_{i,j}$ being the $j$th entry of the vector $\tilde X^s_{i}$.
+$$h = \left(\frac{4}{s+2}\right)^{\frac{1}{s+4}}n^{\frac{-1}{s+4}}\sigma_{\min}, $$ where $\sigma_{\min}$ is the minimum over $j=1,\ldots,s$ of the standard deviation of $\tilde X^s_{1,j},\ldots, \tilde X^s_{n,j}$, with $\tilde X^s_{i,j}$ being the $j$th entry of the vector $\tilde X^s_{i}$.
 In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.
@@
 We compare the performance of our method to each of the following four alternatives.
-The first alternative, denoted RD, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
+The first alternative, denoted **RD**, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$.
 Note that this procedure is only sensible if all the time grids $T_1,\ldots,T_n$ are the same.
@@
 For the next two alternatives, we first smooth the discretely-sampled noisy functional data using splines as described in (\ref{eq_spline_smoothing}).
-Then, we either estimate pairwise distance between the smoothed $\tilde X_1,\ldots,\tilde X_n$ evaluated on a common time grid of $D$ points using IsoGeo (denoted Spline IsoGeo) or the $L^2$ distance (denoted Spline L2).
+Then, we either estimate pairwise distance between the smoothed $\tilde X_1,\ldots,\tilde X_n$ evaluated on a common time grid of $D$ points using IsoGeo or the $L^2$ distance.
+The former method will be denoted **Spline IsoGeo** and the latter **Spline L2**.
 
@@
 
-We ran P-ISOMAP separately in Matlab and tuned for the parameter $\varepsilon$ and the penalty parameter using the relative Frobenius assessment metric.
+The code for P-ISOMAP was provided by \cite{ChenMuller2012} in MATLAB.
+We ran P-ISOMAP and tuned for the parameter $\varepsilon$ and the penalty parameter $\delta$ using the relative Frobenius assessment metric, described below.
 We found that the penalty parameter chosen in this oracle fashion was always zero, thus reducing the P-ISOMAP procedure to standard ISOMAP.
@@
 Throughout, we make use of a helpful property of one-dimensional manifolds.
-If $X \in \M$, with the $L^2$ metric tensor, depends on a parameter $\theta \in \mathbb R$, then
+If $X \in \M$, with the $L^2$ metric tensor, depends on a scalar parameter $\theta \in \mathbb R$, then
 $$
@@
 <!-- Scenario 2 in sim_functional_data -->
+```{r, echo=FALSE, eval=TRUE, cache=TRUE, results='hide'}
+# Generate data
+sim <- sim_functional_data(sce=2,samplesize=100,K=30,plot_true=2,reg_sampling=1)
+```
 We take Manifold 2 in \cite{ChenMuller2012} and set the variance of the normal density to be $1$.
@@
 Thus the geodesic distance will not coincide with the $L^2$ distance.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{sim_sce=2}
+\caption{One hundred noisy curves in the functional manifold of normal probability density.}
+\label{fig:sce2}
+\end{figure}
 
@@
 <!-- Scenario 4 in sim_functional_data -->
+```{r, echo=FALSE, eval=TRUE, cache=TRUE, results='hide'}
+# Generate data
+sim <- sim_functional_data(sce=4,samplesize=100,K=30,plot_true=2,reg_sampling=1)
+```
 It was shown in \cite{Joshi2007} that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
@@
 It was shown in \cite{Joshi2007} that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{sim_sce=4}
+\caption{One hundred noisy curves in the functional manifold of square root probability density functions.}
+\label{fig:sce4}
+\end{figure}
+
 We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$.
@@
 <!-- Scenario 5 in sim_functional_data -->
-This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$.
-(Note that Equation 17 therein has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\},$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1},$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
+```{r, echo=FALSE, eval=TRUE, cache=TRUE, results='hide'}
+# Generate data
+sim <- sim_functional_data(sce=5,samplesize=100,K=30,plot_true=2,reg_sampling=1)
+```
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{sim_sce=5}
+\caption{One hundred noisy curves in the functional manifold of warped functions.}
+\label{fig:sce5}
+\end{figure}
+
+Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\},$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1},$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.
 
@@
 We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
-The geodesic distance is then $$ d_\M(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$ We estimate this using numerical integration.
+This manifold is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ therein both set to $1$.
+(Note that Equation 17 therein has a typo where the exponentials are missing negative signs.)
+
+The geodesic distance is $$ d_\M(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$ We estimate this using numerical integration.
 
@@
 
-For each manifold $\M$ described above, we simulate ???Susan how many?
-1000??
-samples of $200$ functions $X_i\in\M$.
+For each manifold $\M$ described above, we simulate 100 Monte Carlo samples of $100$ functions $X_i\in\M$.
 We then construct vectors $Y_i\in \R^K$ such that $Y_{ij}=X_i(t_j) + \epsilon_{ij}$ with $T=(t_1,\ldots,t_K)$ a common grid of equally spaced points and $\epsilon_{ij}\sim N(0,\sigma_\epsilon^2)$.
-The tuning parameters for the proposed method were chosen as described at the end of the methodology section.
+The tuning parameters for the proposed method, $s$ and $h$, were chosen as described at the end of the methodology section.
 In step 1 of our procedure, we take $D = 100$ for the size of the common time grid.
@@
 We show the performance of the Spline IsoGeo and Spline L2 method versus our method for three dimensions $s=1,2,3$.
-We can see from Figures \ref{fig:normal} and \ref{fig:warping} that spline smoothing followed by either IsoGeo or the $L^2$ distance calculation estimates the pairwise geodesic distances reasonably well though not as well as the proposed method for $s=2$ and $s=3$.
-Interestingly, when $s$ in our method is not selected properly, for instance when $s=1$, our estimator performs poorly in terms of the relative Frobenius norm although still better than Spline IsoGeo and Spline L2 in the other two metrics.
+We can see from Figures \ref{fig:normal} and \ref{fig:warping} that spline smoothing followed by either IsoGeo or the $L^2$ distance estimates the pairwise geodesic distances reasonably well though not as well as the proposed method for $s=2$ and $s=3$.
+Notably, when $s$ in our method is not selected properly, for instance when $s=1$, our estimator performs poorly in terms of the relative Frobenius norm although still better than Spline IsoGeo and Spline L2 in the other two metrics.
 It may be possible that for certain downstream tasks such as clustering or classification, precise estimation of the geodesic distance is less important than one that respects the ordering of the distances.
 In such cases, the near-isometry and Pearson correlation metrics may give a better indication of downstream performance of the geodesic distance estimator.
+
 The result of the square root velocity manifold scenario is presented in
@@
 sample the manifold broadly enough so that the sample lives in a locally
-flat region where the \(L^2\) distance holds.
+flat region where the $L^2$ distance holds.
 
@@
 \includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
-\caption{Manifold of $N(\beta,1)$ probability density functions. Across all three metrics, our method with $s=2$ performs the best, followed by our method with $s=3$.}
+\caption{Manifold of $N(\beta,1)$ probability density functions.
+The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples.
+Across all three metrics, our method with $s=2$ performs the best, followed by our method with $s=3$.}
 \label{fig:normal}
@@
 \includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
-\caption{Manifold of warped functions. The best performing estimators are given by our method with $s=2$ and then $s=3$.}
+\caption{Manifold of warped functions.
+The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples.
+The best performing estimators are given by our method with $s=2$ and then $s=3$.}
 \label{fig:warping}
@@
 \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
-\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
+\caption{Manifold of square root Beta densities.
+The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples.
+Besides Spline IsoGeo, all methods perform comparably well.}
 \label{fig:betas}

### 2019-09-05 paperAAAI.tex
final touches

@@
 to regard them as functional data.  
-When the underlying curves of interest lie on a manifold of low intrinsic
-dimension, e.g.~probability density
+When the underlying curves of interest lie on a manifold, e.g.~probability density
 functions or warped curves of a common template function, there is further structure to exploit. 

### 2019-09-05 cluster_results.R

@@ -42,17 +42,20 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
     #   # Change font size
     #   cex.main=0.5, cex.lab=0.5, cex.sub=0.5
     # )
-    boxplot(results[,1:nr_methods],main="relative Frobenius",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,1:nr_methods],main="relative Frobenius"
+            ,names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75,font.main=1)
     dev.off()
     
     pdf(paste(base_pattern_out,"_isometry",".pdf",sep=""),width=16/4,height=3)
     par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
-    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="near-isometry",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,(1+nr_methods):(2*nr_methods)],main="near-isometry"
+            ,names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75,font.main=1)
     dev.off()
     
     pdf(paste(base_pattern_out,"_Pearson",".pdf",sep="") ,width=16/4,height=3)
     par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
-    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation",names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75)
+    boxplot(results[,(2*nr_methods+1):(3*nr_methods)],main="Pearson correlation"
+            ,names=method_names,cex.main=0.75,cex.names=0.75,cex.axis=0.75,cex.lab=0.75,font.main=1)
     dev.off()
     
     # mtext(base_pattern,outer=TRUE,cex=1.5)

### 2019-09-05 sim_functional_data.R

@@ -79,7 +79,7 @@
 # SNR (signal to noise ratio)
 # reg_sampling : 0 if manifold parameters are drawn uniformly and 1 if manifold parameters are drawn from a concentrated measure.
 # com_grid : if 1 each curve is observed on the same grid and if 0 the grid for each curve is randomly genarated from unif[a,b]
-# plot_true : if 1 plot the true data, the observed data and the true geodesic matrix
+# plot_true : if 1 plot the true data, the observed data and the true geodesic matrix, 2 if plots are for publication
 
 ## Output
 # noiseless_data : samplesize x K matrix containing the original data (no noise)
@@ -381,6 +381,18 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
     matplot(t(grid),t(noisy_data),main="Observed data",type='l', col=rainbow(samplesize))
     image.plot(analytic_geo,main='analytic geodesic')
   }
+  if (plot_true==2){
+    pdf(paste(sprintf("sim_sce=%d",sce),".pdf",sep=""),width=2,height=2)
+    par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
+    par(cex.axis=0.5,cex.lab=0.5)
+    # matplot(t(grid),t(noiseless_data),main="True data",type='l', col=rainbow(samplesize))
+    matplot(t(grid),t(noisy_data),main="",type='l', col=rainbow(samplesize))
+    title(main = list("Observed data", cex = 0.5, font=1))
+    # image.plot(analytic_geo,main='analytic geodesic')
+    dev.off()    
+    
+  }
+    
   return(list('noiseless_data'=noiseless_data,'noisy_data'=noisy_data,'analytic_geo'=analytic_geo,'grid'=grid,'reg_grid'=reg_grid))
 }
 

### 2019-11-20 ecai.tex
ecai files

--- 2019-09-05 paperAAAI.tex
@@
-\def\year{2020}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
-\usepackage{aaai20}  % DO NOT CHANGE THIS
-\usepackage{times}  % DO NOT CHANGE THIS
-\usepackage{helvet} % DO NOT CHANGE THIS
-\usepackage{courier}  % DO NOT CHANGE THIS
-\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
-\usepackage{graphicx} % DO NOT CHANGE THIS
-\urlstyle{rm} % DO NOT CHANGE THIS
-\def\UrlFont{\rm}  % DO NOT CHANGE THIS
-\usepackage{graphicx}  % DO NOT CHANGE THIS
-\frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
+\documentclass{ecai}
+\usepackage{times}
+\usepackage{graphicx}
+\usepackage{latexsym}
 
-\def\tightlist{}
-
-% 
 % 
@@
 
-%\nocopyright
-%PDF Info Is REQUIRED.
-% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
-% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
- \pdfinfo{
-/Title (The hidden manifold distance for functional data)
-/Author (Susan Wei and Marie-H\'el\`{e}ne Descary)
-} %Leave this	
-% /Title ()
-% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
-% Leave the space between \Title and the beginning parenthesis alone
-% /Author ()
-% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case. 
-% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands, 
-% remove them.
+\def\tightlist{}
 
-% DISALLOWED PACKAGES
-% \usepackage{authblk} -- This package is specifically forbidden
-% \usepackage{balance} -- This package is specifically forbidden
-% \usepackage{caption} -- This package is specifically forbidden
-% \usepackage{color (if used in text)
-% \usepackage{CJK} -- This package is specifically forbidden
-% \usepackage{float} -- This package is specifically forbidden
-% \usepackage{flushend} -- This package is specifically forbidden
-% \usepackage{fontenc} -- This package is specifically forbidden
-% \usepackage{fullpage} -- This package is specifically forbidden
-% \usepackage{geometry} -- This package is specifically forbidden
-% \usepackage{grffile} -- This package is specifically forbidden
-% \usepackage{hyperref} -- This package is specifically forbidden
-% \usepackage{navigator} -- This package is specifically forbidden
-% (or any other package that embeds links such as navigator or hyperref)
-% \indentfirst} -- This package is specifically forbidden
-% \layout} -- This package is specifically forbidden
-% \multicol} -- This package is specifically forbidden
-% \nameref} -- This package is specifically forbidden
-% \natbib} -- This package is specifically forbidden -- use the following workaround:
-%\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}} 
-%\newcommand{\citep}{\cite} 
-%\newcommand{\citealp}[1]{\citeauthor{#1} \citeyear{#1}}
-% \usepackage{savetrees} -- This package is specifically forbidden
-% \usepackage{setspace} -- This package is specifically forbidden
-% \usepackage{stfloats} -- This package is specifically forbidden
-% \usepackage{tabu} -- This package is specifically forbidden
-% \usepackage{titlesec} -- This package is specifically forbidden
-% \usepackage{tocbibind} -- This package is specifically forbidden
-% \usepackage{ulem} -- This package is specifically forbidden
-% \usepackage{wrapfig} -- This package is specifically forbidden
-% DISALLOWED COMMANDS
-% \nocopyright -- Your paper will not be published if you use this command
-% \addtolength -- This command may not be used
-% \balance -- This command may not be used
-% \baselinestretch -- Your paper will not be published if you use this command
-% \clearpage -- No page breaks of any kind may be used for the final version of your paper
-% \columnsep -- This command may not be used
-% \newpage -- No page breaks of any kind may be used for the final version of your paper
-% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
-% \pagestyle -- This command may not be used
-% \tiny -- This is not an acceptable font size.
-% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
-% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
-
-\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
-
-% The file aaai20.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
-% command to correct it. You may not alter the value below 2.5 in
-\title{The Hidden Manifold Distance for Functional Data}
-%Your title must be in mixed case, not sentence case. 
-% That means all verbs (including short verbs like be, is, using,and go), 
-% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
-% articles, conjunctions, and prepositions are lower case unless they
-% directly follow a colon or long dash
-
-% \author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be
-% \textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ }
-
-%in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
-
-%If you have multiple authors and multiple affiliations
-% use superscripts in text and roman font to identify them.
-For example, Sunil Issar,\textsuperscript{\rm 2} J.
-Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}.
-Note that the comma should be placed BEFORE the superscript for optimum readability
- % email address must be in roman text type, not monospace or sans serif
+%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
+                  % Do NOT use for camera-ready version of paper.
 
@@
 
+\title{The hidden manifold distance for functional data}
+
+\author{Susan Wei\institute{University of Melbourne,
+Australia, email: susan.wei@@unimelb.edu.edu} \and Marie-H\'el\`{e}ne Descary\institute{Universit\'{e} du Qu\'{e}bec \`{a} Montr\'{e}al,
+Canada, email:descary.marie-helene@uqam.ca}}
+
 \maketitle
+\bibliographystyle{ecai}
 
@@
 
-%\section{Acknowledgement}\label{acknowledgement}
+\input{Conclusion}
 
-%The authors would like to acknowledge support from the ARC Centre of
-%Excellence for Mathematical and Statistical Frontiers (ACEMS) which
-%enabled the two authors to work on the project at the University of
-%Melbourne.
-
-% \section{Appendices}\label{appendices}
-% 
-% The result of the square root velocity manifold scenario is presented in
-% Figure \ref{fig:betas}. The Spline L2 method has similar performance to
-% our proposed method. This could be due to the fact that we did not
-% sample the manifold broadly enough so that the sample lives in a locally
-% flat region where the \(L^2\) distance holds.
-% 
-% \begin{figure}
-% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
-% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
-% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
-% \caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
-% \label{fig:betas}
-% \end{figure}
+\ack We would like to thank the referees for their comments, which
+helped improve this paper considerably
 
@@
 \bibliography{references}
-\bibliographystyle{aaai}
-
 \end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-11-20 index.Rmd
ecai files

@@
 
-# Introduction 
+# INTRODUCTION 
 Many statistical and machine learning methods rely on some measure of distance.
@@
 
-# Proposed method for estimating geodesic distances 
+# PROPOSED METHOD FOR ESTIMATING GEODESIC DISTANCES
 Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$.
@@
 
-# Simulation study
+# SIMULATION STUDY
 

### 2020-04-29 writing/abstract.tex
created writing folder, moved paper graphics, put paper into sadm template

--- 2019-11-20 ecai.tex
@@
-\documentclass{ecai}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{latexsym}
-
-% 
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-% 
-% 
-% 
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{amsbsy}
-\usepackage{graphics}
-\usepackage{epsfig}
-
-\def\tightlist{}
-
-%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
-                  % Do NOT use for camera-ready version of paper.
-
-\begin{document}
-
-\title{The hidden manifold distance for functional data}
-
-\author{Susan Wei\institute{University of Melbourne,
-Australia, email: susan.wei@@unimelb.edu.edu} \and Marie-H\'el\`{e}ne Descary\institute{Universit\'{e} du Qu\'{e}bec \`{a} Montr\'{e}al,
-Canada, email:descary.marie-helene@uqam.ca}}
-
-\maketitle
-\bibliographystyle{ecai}
-
-\begin{abstract}
 In many machine learning applications, the observed data is a
@@
 potential benefits for downstream tasks such as distance-based functional classification.
-\end{abstract}
-
-\input{index}
-
-\input{MarieDataAnalysis}
-
-\input{Conclusion}
-
-\ack We would like to thank the referees for their comments, which
-helped improve this paper considerably
-
-\newpage
-
-\bibliography{references}
-\end{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-29 writing/index.tex
created writing folder, moved paper graphics, put paper into sadm template

\newcommand {\To}{\rightarrow}
\newcommand {\TO}{\Rightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\Prob}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand {\cov}{\textrm{Cov}}
\newcommand {\var}{\textrm{Var}}
\newcommand {\1}{\textrm{\textbf{1}}}
\newcommand{\M}{\mathcal{M}}

\section{INTRODUCTION}\label{introduction}

Many statistical and machine learning methods rely on some measure of
distance. For example, clustering and classification of functional data,
discretely-sampled curves varying over a continuum, is a common learning
task in which distance plays a crucial role. In a naive treatment, the
data may be simply processed as vectors in Euclidean space. Doing so
however risks losing the richness of functional data; indeed, one may do
better to regard the data as functions. The challenge of analyzing
infinite-dimensional objects can be ameliorated if the curves are smooth
since the dimensionality is then only artificially high. This insight
underpins functional data analysis, a subfield of statistics that
studies functional data.

In this work, we advocate that for functional manifold data,
i.e.~functional data that lie on a manifold, the geodesic distance is
more appropriate than the \(L^2\) distance. Indeed, when the manifold
hypothesis is plausible, e.g.~for classes of probability density
functions and classes of warped curves of a common template function,
clustering and classification using the geodesic distance may give
better results than using the \(L^2\) distance.

The manifold setting presents certain challenges since, for nonlinear
manifolds, even basic operations such as addition and subtraction
require special consideration. We will carefully lay out a technique for
estimating the geodesic distance between noisy functional manifold data.

Let \(X_1,\ldots,X_n\) be a sample of \(n\) independent realizations of
a random variable \(X\) that takes value in the Hilbert space
\(L^2([a,b],\R)\). Suppose additionally that the function \(X\) belongs
to a Riemannian manifold \(\M \subset L^2([a,b],\R)\) with Riemannian
metric tensor \(g\) which are often used to assign a metric on the
manifold as follows \cite{Lin2014}. For each point \(X\) on the
manifold, the Riemannian metric tensor \(g\) has an inner product
\(g_X\) on the tangent space \(T_X \M\). The norm of a tangent vector
\(V \in T_X \M\) is defined as \[\|V\| = \sqrt{g_X(V,V)}.\] The geodesic
distance between two functions \(X_i,X_j\) on the manifold \(\M\), based
on this metric tensor \(g\), is defined as
\[ d_{\M}(X_i,X_j):=\inf_{\gamma:[0,1] \to \M, \gamma(0) = X_i, \gamma(1) = X_j} l(\gamma),\]
where
\[ l(\gamma) := \int_0^1 \left\| \frac{\,d\gamma}{\,d t}(t) \right\| \,dt\]
is the length of the piecewise-smooth curve \(\gamma\).

Ideally, the functions \(X_i\) and \(X_j\) are observed on a very dense
domain grid with no measurement error. Then the geodesic distance
\(d_\M(X_i,X_j)\) can be estimated using any of a variety of
shortest-path algorithms, e.g.~the Floyd-Warhsall \cite{Floyd1962}
algorithm. However, most shortest-path algorithms critically assume
observations live near the manifold with little noise \cite{Yeh2005} and
are thus likely to fail for functional data which can be observed with
low signal-to-noise ratio. In this work, we put forth a technique for
the recovery of the \(n{\times}n\) matrix \(G\) of pairwise geodesic
distances: \[
G(i,j)=G(j,i) = \begin{cases} 
      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
      0 & \textrm{otherwise.}
   \end{cases}
\] In lieu of \(X_i\) and \(X_j\), we have access only to
discretely-sampled noisy functional observations \(Y_i\) and \(Y_j\)
that possibly lie off the true manifold.

The work of \cite{ChenMuller2012} was among the first in functional data
analysis to consider the manifold hypothesis for functional data. A
notion of the mean and variation of functional manifold data was
introduced there. Of particular relevance to this work is their proposed
P-ISOMAP procedure which allows for noisy functional observations. This
is in contrast to the classic ISOMAP algorithm \cite{Tenenbaum2000}
which assumes observations lie exactly on the manifold. This drawback to
ISOMAP was also realised in \cite{Dimeglio2014} who proposed a procedure
we will call robust-ISOMAP for functional data that is less sensitive to
outliers. We will discuss P-ISOMAP and robust-ISOMAP in more details
below.

\section{RELATED WORK}

\section{PROPOSED METHOD FOR ESTIMATING GEODESIC
DISTANCES}\label{proposed-method-for-estimating-geodesic-distances}

Suppose that each curve \(X_i \in (\M,g)\) is observed with measurement
error on a time grid
\(T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b\),
i.e.~we observe a sample of \(K\)-dimensional vectors \(Y_1,\ldots,Y_n\)
with \(Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}\). Let the random variables
\(\epsilon_{ij}\) have mean zero and be uncorrelated with each other. We
assume that each observation grid \(T_1,\ldots,T_n\) is dense,
i.e.~\(K\) is large.

We begin by converting the discretely-sampled functional observations
into continuous ones. Let \(\tilde X_1,\ldots,\tilde X_n\) denote the
functional versions of the raw data obtained by some smoothing method.
For example, we may employ spline smoothing \cite{Ramsay2005} to recover
the functional versions of the raw data, i.e.\\

\begin{equation}\label{eq_spline_smoothing} 
\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
\end{equation}

where \(\lambda>0\) is a tuning parameter controlling the smoothness of
\(\tilde X_i\). Our method is based on the idea that the underlying
functional manifold \(\M\) can be sufficiently well-recovered by the
subspace-constrained mean-shift (SCMS) algorithm \cite{Ozertem2011}
which we now review briefly in the Euclidean case. Suppose
\(Z_1,\ldots,Z_n\) is a sample of realisations of a random vector
\(Z \in \M\). We can use a kernel density estimator \(\hat f\) based on
this data to estimate the probability density function of \(Z\). Let the
set \(\hat M\) be comprised of the basins of attraction of \(\hat f\).
The SCMS algorithm sends each point \(Z_i\) to its destination in
\(\hat M\). Theoretical justification that \(\hat M\) as estimated by
SCMS is a reasonable surrogate for the true manifold \(\M\) can be found
in \cite{Genovese2014}.

Before presenting our procedure, we also need to review the ISOMAP
algorithm \cite{Tenenbaum2000}, a three-step procedure that takes as
input a set of points \(x_1,\ldots,x_n\) in a submanifold
\(\M \subset \R^D\) and produces an embedding in the space \(\R^d\) with
\(d<D\) that preserves pairwise geodesic distances. The procedure is as
follows.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Construct a weighted graph \(\mathcal{G}\) with nodes corresponding to
  the observations \(x_1,\ldots,x_n\in \R^D\). Two nodes \(x_i\) and
  \(x_j\) are connected by an edge \(e_{ij}\) with weight
  \(d_{ij} 1\{d_{ij} \le \varepsilon \}\) where
  \(d_{ij}=\|x_i-x_j \|_{2}\) and \(\varepsilon > 0\).
\item
  Estimate the pairwise geodesic distances \(d_\M(x_i,x_j)\) based on
  \(\mathcal{G}\) using shortest-path algorithms. Specifically, this is
  the sum of the weights of the edges forming the shortest path, which
  is calculated either with the Floyd-Warshall algorithm
  \cite{Floyd1962} or with the Dijkstra algorithm
  \cite{Dijkstra59anote}.
\item
  Use multidimensional scaling to obtain an embedding in \(\R^d\) that
  preserves the pairwise geodesic distances estimated above.
\end{enumerate}

Note that shortest-path algorithms such as the Floyd-Warshall algorithm
are used to find the smallest path \textit{given} a weighted graph but
they do not produce the weighted graphs themselves. Both P-ISOMAP and
robust-ISOMAP are modifications of ISOMAP in the sense that they change
the construction of the weighted graph in step 1 above.

In what follows, we call IsoGeo the two-step procedure which performs a
modified version of the first step of ISOMAP followed by the original
second step of ISOMAP. The first step of IsoGeo constructs a weighted
graph \(\mathcal{G}\) according to \cite{Dimeglio2014} as follows. First
we construct a complete weighted graph \(\mathcal{G}_c\), meaning that
every pair of nodes \(x_i\) and \(x_j\) is connected by an edge of
weight \(d_{ij}\), so that the graph is independent of the tuning
parameter \(\varepsilon\). Next, we obtain the minimal spanning tree
associated with \(\mathcal{G}_c\) and denote its set of edges by
\(E_s\). The graph \(\mathcal{G}\) is obtained by adding all edges
\(e_{ij}\) to \(E_s\) for which the following condition is true:
\[ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\varepsilon_i),\]
where \(B(x_i,\varepsilon_i)\) is the open ball of center \(x_i\) and
radius \(\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}\), and
\(\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}\).
Let the distance estimated by IsoGeo be denoted
\(\text{IsoGeo}(x_i,x_j)\). Note that IsoGeo has no tuning parameters.

We are now ready to describe our procedure. Since SCMS is based on
kernel density estimation, we first reduce the dimension of our data
with multidimensional scaling before applying SCMS to avoid the curse of
dimensionality.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Discretise the functions \(\tilde X_1,\ldots,\tilde X_n\) by
  evaluating them on a dense common grid
  \(\tilde T=(t_{1},\ldots,t_{D})\). Let \(s\) be a positive integer,
  much smaller than \(D\). Obtain
  \(\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s\) using
  multidimensional scaling so that the pairwise \(\mathbb R^D\)
  Euclidean distances of the discretised functions \(\tilde X_i\) can be
  preserved as much as possible.
\item
  Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011}
  to each of \(\tilde X^s_i\) to obtain its destination
  \(\tilde X^{s,\hat \M}_i\).
\item
  Our geodesic distance estimator is the \(n{\times}n\) matrix
  \(\hat G\) whose elements are given by \[
  \hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
   \text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
   0 & \textrm{otherwise.}
    \end{array} \right.
  \]
\end{enumerate}

There are two tuning parameters to our procedure. First, there is the
dimension \(s\) in step 1. In our simulations we try
\(s \in \{1,2,3\}\). The other parameter is the bandwidth \(h\) in the
subspace constrained mean-shift. In our simulations we set \(h\)
according to the heuristic in Equation (A1) of \cite{ChenHo2015}:
\[h = \left(\frac{4}{s+2}\right)^{\frac{1}{s+4}}n^{\frac{-1}{s+4}}\sigma_{\min}, \]
where \(\sigma_{\min}\) is the minimum over \(j=1,\ldots,s\) of the
standard deviation of \(\tilde X^s_{1,j},\ldots, \tilde X^s_{n,j}\),
with \(\tilde X^s_{i,j}\) being the \(j\)th entry of the vector
\(\tilde X^s_{i}\). In reality, depending on the downstream task, both
\(s\) and \(h\) can be tuned in a data-adaptive way, say using
cross-validation.

\section{SIMULATION STUDY}\label{simulation-study}

We perform simulations to study various estimators of pairwise geodesic
distances. We compare the performance of our method to each of the
following four alternatives. The first alternative, denoted \textbf{RD},
takes the naive approach of applying IsoGeo directly on the raw vectors
\(Y_1,\ldots,Y_n \in R^K\). Note that this procedure is only sensible if
all the time grids \(T_1,\ldots,T_n\) are the same.

For the next two alternatives, we first smooth the discretely-sampled
noisy functional data using splines as described in
(\ref{eq_spline_smoothing}). Then, we either estimate pairwise distance
between the smoothed \(\tilde X_1,\ldots,\tilde X_n\) evaluated on a
common time grid of \(D\) points using IsoGeo or the \(L^2\) distance.
The former method will be denoted \textbf{Spline IsoGeo} and the latter
\textbf{Spline L2}.

The final alternative we consider is P-ISOMAP which is the two-step
procedure developed in \cite{ChenMuller2012} for handling noisy
functional data. In the first step, a penalty is incorporated to
robustify the construction of the weighted graph. This is followed by
the original second step of ISOMAP.

The code for P-ISOMAP was provided by \cite{ChenMuller2012} in MATLAB.
We ran P-ISOMAP and tuned for the parameter \(\varepsilon\) and the
penalty parameter \(\delta\) using the relative Frobenius assessment
metric, described below. We found that the penalty parameter chosen in
this oracle fashion was always zero, thus reducing the P-ISOMAP
procedure to standard ISOMAP. Given this, we did not make further
comparison to P-ISOMAP in our simulation study. We also do not present
the results for the RD method which always produced poor estimates of
the geodesic distance.

Three different metrics are used to assess the quality of a pairwise
geodesic distance estimator. The first metric assesses near-\(\epsilon\)
isometry, which we define as the percentage of estimated pairwise
geodesic distances between \(1-\epsilon\) and \(1+\epsilon\) of the
corresponding true pairwise geodesic distance. We form the receiver
operating curve with \(\epsilon\) on the \(x\)-axis and
near-\(\epsilon\) isometry on the \(y\)-axis. The metric is the area
under this receiver operating curve. Thus a good estimator has high
values for the near-isometry metric.

The second metric \(\|G-\hat G\|/\|G\|\) is the Frobenius norm of the
estimation error matrix \(G-\hat G\), relative to the Frobenius norm of
the true distance matrix \(G\). The third metric is the Pearson
correlation coefficient between the upper diagonal of \(G\) and
\(\hat G\). A good geodesic distance estimator has low Frobenius norm
and high Pearson correlation coefficient.

We shall consider three functional manifolds in our simulation study.
Throughout, we make use of a helpful property of one-dimensional
manifolds. If \(X \in \M\), with the \(L^2\) metric tensor, depends on a
scalar parameter \(\theta \in \mathbb R\), then \[
d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} \left\| \frac{\partial X}{\partial \theta}(t)\right\|_{L^2} \,d\theta.
\]

We also need to make a disclaimer about our sampling technique below.
Proper sampling from a manifold remains an open problem. Even in the
Euclidean case, it is not obvious how sampling should be done
\cite{Diaconis2013}. For now, to safeguard against sampling unevenly on
the functional manifold, we sample on a very narrow support of the
manifold parameters.

\subsection{Isometric functional manifold of normal density
functions}\label{isometric-functional-manifold-of-normal-density-functions}

We take Manifold 2 in \cite{ChenMuller2012} and set the variance of the
normal density to be \(1\). Let \(X_\beta: [a,b] \to \R\) be given by
\(X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}\).
Consider
\[\M = \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \},\] with
the \(L^2\) inner product as the metric tensor. First, note that the
``straight'' line connecting \(X_{\beta_1}\) and \(X_{\beta_2}\) does
not always stay inside of \(\M\). Thus the geodesic distance will not
coincide with the \(L^2\) distance.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=2}
\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=4}
\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=5}
\caption{One hundred noisy curves in the functional manifold of normal probability density.}
\label{fig:sce2} %previous labels {fig:sce4} and {fig:sce5} need to be updated
\end{figure}

We set \(a=-4\) and \(b=4\) and sample \(\beta\) according to a
truncated standard normal with support on \([-1,1]\). The geodesic
distance between the curves \(X_{\beta_1}\) and \(X_{\beta_2}\) is given
by

\begin{eqnarray*}
d_\M(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
% &=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\ 
&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta   \\
&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
\end{eqnarray*}

\noindent where \(f\) is the density of a N\((\beta,1/2)\) and the
approximation in the next to last step comes from integrating on
\([a,b]=[-4,4]\) rather than \(\R\). We can see this manifold is
isometric, since the geodesic distance between \(X_{\beta_1}\) and
\(X_{\beta_2}\) in \(\M\) is proportional to \(\beta_2-\beta_1\).

\subsection{Functional manifold of square root velocity
functions}\label{functional-manifold-of-square-root-velocity-functions}

It was shown in \cite{Joshi2007} that the square root representation of
probability density functions gives rise to a simple closed-form
geodesic distance. They consider the manifold
\[ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \},\]
with the Fisher-Rao metric tensor
\[ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds ,\] for two tangent vectors
\(v_1,v_2 \in T_\psi(\M)\). Note that this coincides with the
\(L^2[0,1]\) inner product. It was shown in \cite{Joshi2007} that the
geodesic distance between any two \(\psi_1\) and \(\psi_2\) in \(\M\) is
simply \[d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.\]

We will specifically examine the square root of \(Beta(\alpha,\beta)\)
densities, which are of course supported on \([0,1]\). That is, consider
\[ \M = \big\{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\big\}, \]
where \(\psi_{\alpha,\beta}: [0,1] \to \R\) is the probability density
function of a Beta random variable with shape parameters \(\alpha\) and
\(\beta\). We sample \(\alpha\) according to a truncated normal with
mean 3 and variance 0.09 with support on \([1,5]\). We sample \(\beta\)
according to a truncated normal with mean 3.5 and variance 0.09 with
support on \([2,5]\).

\subsection{Functional manifold of warped
functions}\label{functional-manifold-of-warped-functions}

Let \(X_\alpha(t) = \mu(h_\alpha(t))\) be defined on \([-3,3]\) where
\[ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\},\] and
\[ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1},\] if
\(\alpha \ne 0\) and \(h_\alpha(t) = t\) otherwise.

Consider the manifold \[ \M = \{X_\alpha: -1 \le \alpha \le 1\}.\] We
sample \(\alpha\) according to a truncated standard normal with support
on \([-1,1]\). This manifold is based on Equations 17 and 18 in
\cite{Kneip2008} but with \(z_{i1}, z_{i2}\) therein both set to \(1\).
(Note that Equation 17 therein has a typo where the exponentials are
missing negative signs.)

The geodesic distance is
\[ d_\M(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.\]
We estimate this using numerical integration.

\subsection{Results}\label{results}

For each manifold \(\M\) described above, we simulate 100 Monte Carlo
samples of \(100\) functions \(X_i\in\M\). We then construct vectors
\(Y_i\in \R^K\) such that \(Y_{ij}=X_i(t_j) + \epsilon_{ij}\) with
\(T=(t_1,\ldots,t_K)\) a common grid of equally spaced points and
\(\epsilon_{ij}\sim N(0,\sigma_\epsilon^2)\). The tuning parameters for
the proposed method, \(s\) and \(h\), were chosen as described at the
end of the methodology section. In step 1 of our procedure, we take
\(D = 100\) for the size of the common time grid. We will investigate
how two parameters affect the performance of the various methods in
consideration. The first one is the signal-to-noise ratio
\[SNR = 10 log_{10}\left(\frac{\sigma^2_{sig}}{\sigma^2_\epsilon}\right),\]
where \(\sigma^2_{sig}\) is the variance of the signal. We fix SNR at
either \(0.1\) (low) or \(0.5\) (high). The second parameter is \(K\),
the dimension of the discretely-observed noisy \(Y_i\)'s. We set \(K\)
to either \(30\) (sparse) or \(100\) (abundant).

We will present the results for the high signal-to-noise and sparse
\(K\) setting. We show the performance of the Spline IsoGeo and Spline
L2 method versus our method for three dimensions \(s=1,2,3\). We can see
from Figures \ref{fig:normal} and \ref{fig:warping} that spline
smoothing followed by either IsoGeo or the \(L^2\) distance estimates
the pairwise geodesic distances reasonably well though not as well as
the proposed method for \(s=2\) and \(s=3\). Notably, when \(s\) in our
method is not selected properly, for instance when \(s=1\), our
estimator performs poorly in terms of the relative Frobenius norm
although still better than Spline IsoGeo and Spline L2 in the other two
metrics. It may be possible that for certain downstream tasks such as
clustering or classification, precise estimation of the geodesic
distance is less important than one that respects the ordering of the
distances. In such cases, the near-isometry and Pearson correlation
metrics may give a better indication of downstream performance of the
geodesic distance estimator.

The result of the square root velocity manifold scenario is presented in
Figure \ref{fig:betas}. The Spline L2 method has similar performance to
our proposed method. This could be due to the fact that we did not
sample the manifold broadly enough so that the sample lives in a locally
flat region where the \(L^2\) distance holds.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_MSE}
\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_isometry}
\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_Pearson}
\caption{Manifold of $N(\beta,1)$ probability density functions. The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples. Across all three metrics, our method with $s=2$ performs the best, followed by our method with $s=3$.}
\label{fig:normal}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_MSE}
\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_isometry}
\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_Pearson}
\caption{Manifold of warped functions. The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples. The best performing estimators are given by our method with $s=2$ and then $s=3$.}
\label{fig:warping}
\end{figure}

\begin{figure}
\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_MSE}
\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_isometry}
\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_Pearson}
\caption{Manifold of square root Beta densities. The boxplots indicate the distributions of each of the three metrics over 100 Monte Carlo samples. Besides Spline IsoGeo, all methods perform comparably well.}
\label{fig:betas}
\end{figure}

\section{Conclusion}

### 2020-04-29 writing/wrapper_sadm.tex
created writing folder, moved paper graphics, put paper into sadm template

\documentclass[AMS,STIX1COL]{WileyNJD-v2}

\def\tightlist{}

%\usepackage{amsmath,amsfonts}

\articletype{Article Type}%

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\raggedbottom

\begin{document}

\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}

\author[1]{Author One*}

\author[2,3]{Author Two}

\author[3]{Author Three}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[2]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\address[3]{\orgdiv{Org Division}, \orgname{Org Name}, \orgaddress{\state{State name}, \country{Country name}}}

\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}

\presentaddress{This is sample for present address text this is sample for present address text}

\abstract[Summary]{
\include{abstract}
}

\keywords{keyword1, keyword2, keyword3, keyword4}

\jnlcitation{\cname{%
\author{Williams K.}, 
\author{B. Hoskins}, 
\author{R. Lee}, 
\author{G. Masato}, and 
\author{T. Woollings}} (\cyear{2016}), 
\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}

\maketitle

\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}

\include{index}

\include{MarieDataAnalysis.tex}

%\backmatter

\section*{Acknowledgments}
This is acknowledgment text~\cite{Elbaum2002}. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here.

\subsection*{Author contributions}

This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.

\subsection*{Financial disclosure}

None reported.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

\section*{Supporting information}

The following supporting information is available as part of the online article:

\noindent
\textbf{Figure S1.}
{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}

\noindent
\textbf{Figure S2.}
{The SST anomalies for GC2C calculated against the observations (OIsst).}

\appendix

\section{Section title of first appendix\label{app1}}

%\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
\bibliography{references.bib}%

\section*{Author Biography}

\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2020-04-29 index.Rmd
changed graphics paths, updated index.Rmd

@@
 <!-- \cite{LinYao2017]  -->
+
+# RELATED WORK
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=2}
+\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=4}
+\includegraphics[width=0.3\textwidth]{./graphics/sim_sce=5}
+\caption{Left: One hundred noisy curves in the functional manifold of normal probability density.
+Middle: One hundred noisy curves in the functional manifold of square root probability density functions.
+Right: One hundred noisy curves in the functional manifold of warped functions.}
+\label{fig:sce2} 
+\end{figure}
+<!-- previous labels {fig:sce4} and {fig:sce5} need to be updated -->
+
 ## Isometric functional manifold of normal density functions
@@
 Thus the geodesic distance will not coincide with the $L^2$ distance.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{sim_sce=2}
-\caption{One hundred noisy curves in the functional manifold of normal probability density.}
-\label{fig:sce2}
-\end{figure}
 
@@
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{sim_sce=4}
-\caption{One hundred noisy curves in the functional manifold of square root probability density functions.}
-\label{fig:sce4}
-\end{figure}
-
 We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$.
@@
 ```
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{sim_sce=5}
-\caption{One hundred noisy curves in the functional manifold of warped functions.}
-\label{fig:sce5}
-\end{figure}
 
@@
 \centering
-\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
-\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
-\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_MSE}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_isometry}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=2_SNR=high_Kobs=30_Pearson}
 \caption{Manifold of $N(\beta,1)$ probability density functions.
@@
 \centering
-\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
-\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
-\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_MSE}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_isometry}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=5_SNR=high_Kobs=30_Pearson}
 \caption{Manifold of warped functions.
@@
 \begin{figure}
-\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
-\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
-\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_MSE}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_isometry}
+\includegraphics[width=0.3\textwidth]{./graphics/sce=4_SNR=high_Kobs=30_Pearson}
 \caption{Manifold of square root Beta densities.

### 2020-04-29 writing/wrapper_sadm.tex
changed graphics paths, updated index.Rmd

@@
 
-\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
+\begin{biography}{\includegraphics[width=60pt,height=70pt,draft]{graphics/empty}}{\textbf{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
 \end{biography}

### 2020-04-29 cluster_results.R

@@ -12,10 +12,10 @@ create_boxplot<- function(sce_details,mat_ind,method_names,nr_methods){
   
   # TODO: fix hardcoding
   if(sce_details[2]==0.1){
-    base_pattern_out = do.call(sprintf, c(list("sce=%d_SNR=low_Kobs=%d"), sce_details[1],sce_details[3]))
+    base_pattern_out = do.call(sprintf, c(list("../writing/graphics/sce=%d_SNR=low_Kobs=%d"), sce_details[1],sce_details[3]))
   }
   else{
-    base_pattern_out = do.call(sprintf, c(list("sce=%d_SNR=high_Kobs=%d"), sce_details[1],sce_details[3]))
+    base_pattern_out = do.call(sprintf, c(list("../writing/graphics/sce=%d_SNR=high_Kobs=%d"), sce_details[1],sce_details[3]))
   }
   
 

### 2020-04-29 run_this.R

@@ -1,2 +1,4 @@
 knitr::knit("index.Rmd")
-rmarkdown::pandoc_convert("index.md", to = "latex", output = "index.tex",citeproc=FALSE)
+rmarkdown::pandoc_convert("index.md", to = "latex", output = "./writing/index.tex",citeproc=FALSE)
+
+# if index.Rmd is only changed in terms of writing, set cache=True to avoid long running times producing certain figures

### 2020-04-29 sim_functional_data.R

@@ -382,7 +382,7 @@ sim_functional_data<-function(sce,samplesize=100,K=30,SNR=1,reg_sampling=1,com_g
     image.plot(analytic_geo,main='analytic geodesic')
   }
   if (plot_true==2){
-    pdf(paste(sprintf("sim_sce=%d",sce),".pdf",sep=""),width=2,height=2)
+    pdf(paste(sprintf("./writing/graphics/sim_sce=%d",sce),".pdf",sep=""),width=2,height=2)
     par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.5,1.5,1.5,1.5))
     par(cex.axis=0.5,cex.lab=0.5)
     # matplot(t(grid),t(noiseless_data),main="True data",type='l', col=rainbow(samplesize))

### 2020-04-29 index.Rmd
starting conclusion section

@@
 
+<!-- \cite{LinYao2017]  -->
+
+# RELATED WORK
 The work of \cite{ChenMuller2012} was among the first in functional data analysis to consider the manifold hypothesis for functional data.
@@
 We will discuss P-ISOMAP and robust-ISOMAP in more details below.
-
-<!-- \cite{LinYao2017]  -->
-
-# RELATED WORK
 
@@
 
+# CONCLUSION
+
+In this work, we firstly challenge the existing paradigm of measuring distance for functional data.
+We advocated for the manifold perspective.
+We develop a method for estimating geodesic distances between functional manifold data.
+The approach relies on integrating ideas across several disciplines including manifold learning, functional data analysis, and signal processing.
+
 <!--
